### 2016-10-30 docs/statement_of_research/mlinb_sor.tex
Barebones of the variational LDA. Referencing to the Blei's online LDA implementation.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\setlength{\parskip}{1em}

\title{Statement of Research}
\author{Arijus Pleska}

\begin{document}

\maketitle

\par The project's domain lies in applying machine learning techniques in order to determine structures in metabolomics. Some of the recent research \cite{Rogers} has shown that structures of metabolites can be discovered using topic modelling methodology. 
Thereby, we will investigate this branch of machine learning to push the current results even further.

\par We have chosen Latent Dirichlet Allocation (LDA) \cite{Blei} as our main subject of research. However, we will take into account other techniques such as Probabilistic Latent Semantic Indexing (PLSI) \cite{Hofmann} as well. The objective of our research is to optimise the performance of the latter models in order to be compatible with metabolomics. Most importantly, we are going to address the issue of the impact of noise, which is caused by the limitation of the data collection technology. Also, we are going to enhance the currently used mathematical underlying, and suggest appropriate software engineering practises. Further, we will use robust tools to deliver and display results in high quality.

\par Effectively, the research is expected to increase performance of the discussed models, and to mitigate the previously mentioned obstacles. First of all, we will familiarise with the current research in topic modelling. This preliminary step will allow to familiarise with the terms used in machine learning, and, also, it might suggest alternative approaches in tackling the problem. As a part of the familiarisation process, the models studied will be implemented concurrently with the research. Finally, the results will be documented using adaptive tools. That is, suggested ideas will be written in Jupyter notebook; also, the results will be processed using the Plotly platform to produce interactive plots of high quality.

\bibliography{mlinb_sor}{}
\bibliographystyle{plain}
\end{document}

### 2016-10-30 code/variational_lda.py

@@ -1,0 +1,121 @@
+import os
+import pandas as pd
+import numpy as np
+from scipy.special import psi
+np.set_printoptions(suppress=True)
+HOME_PATH = os.path.expanduser('~')
+
+meanchangethresh = 0.001
+
+class Variational_LDA(object):
+
+    def __init__(self, corpus, K, alpha, eta, tau0, kappa):
+        self.corpus = corpus
+        self.vocabulary = create_vocabulary(corpus)
+        self._K = K
+        self._W = len(self.vocabulary)
+        self._alpha = alpha
+        self._eta = eta
+        self._tau0 = tau0 + 1
+        self._kappa = kappa
+        self._updatect = 0
+
+
+        self._lambda = np.random.gamma(100., 1./100., (self._K, self._W))
+        self._Elogbeta = dirichlet_expectation(self._lambda)
+        self._expElogbeta = np.exp(self._Elogbeta)
+
+    def update_lambda_docs(self):
+        self._rhot = pow(self._tau0 + self._updatect, -self._kappa)
+
+        (gamma, sstats) = self.do_e_step(self.corpus, self.vocabulary)
+
+        # INVESTIGATE D
+        self._lambda = self._lambda * (1-self._rhot) + \
+            self._rhot * (self._eta + sstats / len(self.corpus))
+        self._Elogbeta = dirichlet_expectation(self._lambda)
+        self._expElogbeta = np.exp(self._Elogbeta)
+        self._updatect += 1
+
+        return gamma
+
+
+    def do_e_step(self, corpus, vocabulary):
+
+        (wordids, wordcts) = parse_doc_list(corpus, vocabulary)
+
+        gamma = 1*np.random.gamma(100., 1./100., (len(corpus), self._K))
+        Elogtheta = dirichlet_expectation(gamma)
+        expElogtheta = np.exp(Elogtheta)
+
+        sstats = np.zeros(self._lambda.shape)
+
+        meanchange = 0
+        for d in range(0, len(corpus)):
+            ids = wordids[d]
+            cts = wordcts[d]
+            gammad = gamma[d, :]
+            Elogthetad = Elogtheta[d, :]
+            expElogthetad = expElogtheta[d, :]
+            expElogbetad = self._expElogbeta[:, ids]
+            phinorm = np.dot(expElogthetad, expElogbetad) + 1e-100
+            for it in range(0, 100):
+                lastgamma = gammad
+                gammad = self._alpha + expElogthetad * \
[…]

### 2016-11-01 code/variational_lda.py
Implementation of a Variational LDA. Referrencing to Dr Rogers code.

@@ -1,120 +1,98 @@
+# Implementation of a variational LDA.
+# 
+# Code refers to the variational LDA implementation of Dr Simon Rogers.
+# Comments refer to the LDA paper (Blei et al. 2003)  
+# and the the online LDA implementation (Hoffman et al. 2010).
+
 import os
 import pandas as pd
 import numpy as np
 from scipy.special import psi
+
 np.set_printoptions(suppress=True)
 HOME_PATH = os.path.expanduser('~')
 
-meanchangethresh = 0.001
-
 class Variational_LDA(object):
 
-    def __init__(self, corpus, K, alpha, eta, tau0, kappa):
+    def __init__(self, corpus, K, alpha, eta):
+        """
+        Arguments:
+        corpus: Collection of (mass, intensity) metabolite tuples;
+        K: Number of topics;
+        alpha: Hyperparameter for prior on weight vectors theta;
+        eta: Hyperparameter for prior on topics beta.
+        """
         self.corpus = corpus
         self.vocabulary = create_vocabulary(corpus)
         self._K = K
-        self._W = len(self.vocabulary)
         self._alpha = alpha
         self._eta = eta
-        self._tau0 = tau0 + 1
-        self._kappa = kappa
-        self._updatect = 0
+        self._V = len(self.vocabulary)
+        self._M = len(self.corpus)
 
+    def init_vb(self):
+        # Initialising Beta.
+        self._beta = np.random.rand(self._K, self._V) 
+        self._beta /= self._beta.sum(axis=1)[:, None]
 
-        self._lambda = np.random.gamma(100., 1./100., (self._K, self._W))
-        self._Elogbeta = dirichlet_expectation(self._lambda)
-        self._expElogbeta = np.exp(self._Elogbeta)
+        # Initialising Gamma and Phi.
+        self._phi = {}
+        self._gamma = np.zeros((self._M, self._K))
+        for d, doc in enumerate(self.corpus):
+            self._phi[doc] = {}
+            for word in self.corpus[doc]:
+                w = self.vocabulary[word]
+                # Line (1) in Figure 6.
+                self._phi[doc][w] = np.zeros(self._K)
+            total_intensity = sum(self.corpus[doc].values())
+            # Line (1) in Figure 6 .
+            self._gamma[d, :] = self._alpha + total_intensity / self._K
 
-    def update_lambda_docs(self):
[…]

### 2016-11-01 code/lda.py
Comments' syntax fixes.

@@ -1,100 +1,0 @@
-import os
-import numpy as np
-import pandas as pd
-
-from scipy.special import polygamma as pg
-from scipy.special import psi as psi
-
-HOME_PATH = os.path.expanduser('~')
-
-class LDA(object):
-    def __init__(self, corpus, topic_count):
-        self.corpus = corpus
-
-        self.k = topic_count
-        self.doc_count = len(self.corpus)
-        self.word_count = count_words(self.corpus)
-
-        self.eta = 0.1
-        self.alpha = 1
-        self.init_beta()
-        self.init_phi()
-        self.gamma = np.zeros((self.doc_count, self.k))
-
-    def init_beta(self):
-        self.beta = np.random.rand(self.k, self.word_count)
-        self.beta /= self.beta.sum(axis=1)[:,None]
-
-    def init_phi(self):
-        self.phi = {}
-        for doc in self.corpus:
-            self.phi[doc] = {}
-            for word in self.corpus[doc]:
-                self.phi[doc][word] = np.zeros(self.k) 
-
-    def run(self, iteration_count):
-        # Initialisisation.
-        for i, doc in enumerate(self.corpus):
-            N = 0
-            for word in self.corpus[doc]:
-                N += self.corpus[doc][word]
-            self.gamma[i, :] = self.alpha + N / self.k
-
-        # The main loop.
-        for i in range(iteration_count):
-            temp_beta = np.zeros((self.k, self.word_count))
-            # e step
-            for doc_id, doc in enumerate(self.corpus):
-                temp_gamma = np.zeros(self.k) + self.alpha
-                for word_id, word in enumerate(self.corpus[doc]):
-                    self.phi[doc][word] = self.beta[:, word_id] * np.exp(psi(self.gamma[doc_id, :])).T
-                    self.phi[doc][word] /= self.phi[doc][word].sum()
-                    temp_gamma += self.phi[doc][word] * self.corpus[doc][word]
-                    temp_beta[:, word_id] += self.phi[doc][word] * self.corpus[doc][word]
-                self.gamma[doc_id, :] = temp_gamma
-            temp_beta += self.eta
-            temp_beta /= temp_beta.sum(axis=1)[:,None]
-            self.beta = temp_beta
-            self.alpha = self.update_alpha()
-
-    def update_alpha(self, iteration_count=20, init_alpha=[]):
-        M,K = self.gamma.shape
[…]

### 2016-11-01 code/variational_lda.py

@@ -22,6 +22,8 @@ class Variational_LDA(object):
         alpha: Hyperparameter for prior on weight vectors theta;
         eta: Hyperparameter for prior on topics beta.
         """
+        print corpus.keys()
+        exit()
         self.corpus = corpus
         self.vocabulary = create_vocabulary(corpus)
         self._K = K
@@ -45,7 +47,7 @@ class Variational_LDA(object):
                 # Line (1) in Figure 6.
                 self._phi[doc][w] = np.zeros(self._K)
             total_intensity = sum(self.corpus[doc].values())
-            # Line (1) in Figure 6 .
+            # Line (2) in Figure 6 .
             self._gamma[d, :] = self._alpha + total_intensity / self._K
 
     def run(self, iterations):

### 2016-12-01 code/gibbs_lda.py
The review on the original LDA paper; document generation and plotting.

@@ -1,68 +1,45 @@
 import os
-import logging
 import pandas as pd
 import numpy as np
 
 HOME_PATH = os.path.expanduser('~')
+np.set_printoptions(edgeitems=5)
+
 
 class Gibbs_LDA(object):
 
-    def __init__(self, K, iter_count, alpha, eta):
-        """
-        Arguments:
-        corpus: Collection of (mass, intensity) metabolite tuples;
-        K: Number of topics;
-        alpha: Hyperparameter for prior on weight vectors theta;
-        eta: Hyperparameter for prior on topics beta.
-        """
+    def __init__(self, K, iter_count):
         self.iter_count = iter_count
-        self._K = K
-        self._alpha = alpha
-        self._eta = eta
+        self.K = K
+        self.thetas_ = []
 
     def fit(self, corpus):
         self._fit(corpus)
 
     def dump_thetas(self, path):
-        normalised_thetas = self._theta / self._sampling_count
-        normalised_thetas.dump(path)
-
-    def dump_phis(self, path):
-        normalised_phis = self._phi / self._sampling_count
-        normalised_phis.dump(path)
+        pickle = np.array(self.thetas_)
+        pickle.dump(path)
 
     def _fit(self, corpus):
         self._initialise(corpus)
 
-        thetas_over_time = []
-
         for it in range(self.iter_count):
-            print("Iteration %s." % it) 
+            print("Iteration %s:" % it) 
             self._sample_topics(corpus)
-
-            self._sampling_count += 1.0
-            for d, _ in enumerate(corpus):
-                z_counts = self._dz_counts[d, :]
-                self._theta[d, :] += np.random.dirichlet(z_counts)
-            for z in range(self._K):
-                w_counts = self._zw_counts[z, :]
-                self._phi[z, :] += np.random.dirichlet(w_counts)
-            current_theta = self._theta / self._sampling_count
-            print("Current theta:\n %s." % current_theta) 
-
-            thetas_over_time.append(current_theta)
-
-        pickle = np.array(thetas_over_time)
-        pickle.dump(HOME_PATH + '/Projects/tminm/heavy_pickles/thetas_over_time.pickle')
-
+            
+            self._store_theta()
+            print("Current theta:")
+            print(self.thetas_[it]) 
 
     def _initialise(self, corpus):
-        self._D, self._W = corpus.shape
[…]

### 2016-12-01 code/preprocessing.py

@@ -12,6 +12,8 @@ def create_vocab(corpus):
                 vocab.append(word)
     vocab_panda = pd.Series(vocab)
     vocab_panda.to_pickle(HOME_PATH + '/Projects/tminm/heavy_pickles/' + 'vocab.pickle')
+    print(vocab)
+    exit()
     return vocab
 
 def preprocess_corpus(corpus, vocab, factor):

### 2016-12-01 code/variational_lda.py

@@ -9,12 +9,12 @@ import pandas as pd
 import numpy as np
 from scipy.special import psi
 
-np.set_printoptions(suppress=True)
 HOME_PATH = os.path.expanduser('~')
+np.set_printoptions(edgeitems=5)
 
 class Variational_LDA(object):
 
-    def __init__(self, corpus, K, alpha, eta):
+    def __init__(self, iter_count, K, alpha, eta):
         """
         Arguments:
         corpus: Collection of (mass, intensity) metabolite tuples;
@@ -22,77 +22,87 @@ class Variational_LDA(object):
         alpha: Hyperparameter for prior on weight vectors theta;
         eta: Hyperparameter for prior on topics beta.
         """
-        self.corpus = corpus
-        self.vocabulary = create_vocabulary(corpus)
-        self._K = K
-        self._alpha = alpha
-        self._eta = eta
-        self._V = len(self.vocabulary)
-        self._M = len(self.corpus)
+        self.iter_count = iter_count
+        self.K = K
+        self.alpha = alpha
+        self.eta = eta
 
-    def init_vb(self):
-        # Initialising Beta.
-        self._beta = np.random.rand(self._K, self._V) 
+    def fit(self, corpus, vocab):
+        self._fit(corpus, vocab)
+
+    def get_theta(self):
+        return self._gamma / self._gamma.sum(axis=1)[:, None]
+
+    def _fit(self, corpus, vocab):
+        self._init_vb(corpus, vocab)
+
+        for it in range(self.iter_count):
+            print("Iteration: %s" % it)
+            self._vb_step()
+
+    def _init_vb(self, corpus, vocab):
+        self.corpus = corpus
+        self.vocab = vocab
+        self.V = len(vocab)
+        self.D = len(corpus)
+
+        self._beta = np.random.rand(self.K, self.V) 
         self._beta /= self._beta.sum(axis=1)[:, None]
 
         # Initialising Gamma and Phi.
         self._phi = {}
-        self._gamma = np.zeros((self._M, self._K))
+        self._gamma = np.zeros((self.D, self.K))
         for d, doc in enumerate(self.corpus):
             self._phi[doc] = {}
             for word in self.corpus[doc]:
-                w = self.vocabulary[word]
                 # Line (1) in Figure 6.
-                self._phi[doc][w] = np.zeros(self._K)
[…]

### 2016-12-01 docs/MSci_proposal/mprop.tex
Adding dynamic topic modelling.

\documentclass{mprop}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{siunitx}
\usepackage{xspace}
\usepackage{cite}
\graphicspath{ {images/} }

% alternative font if you prefer
% \usepackage{times}

% for alternative page numbering use the following package
% and see documentation for commands
\usepackage{fancyheadings}

% other potentially useful packages
\usepackage{amssymb,amsmath}
%\usepackage{url}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Smooth Latent Dirichlet Allocation in Metabolomics}
\author{Arijus Pleska}
\date{Date of submission placed here}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{intro}

% Hook intro
\par This research proposal is directed towards applying machine learning in biology. We intend to study the scans of low level biological entities in order to derive their patterns. To be more specific, we aim to tackle the issue of noise, which arises from the hardware limitation of the data collection devices.

% Metabolomics
\par The branch of biology we are expecting to carry out the research is metabolomics. In brief, metabolomics is the scientific study of chemical processes which involve small molecules -- metamobilites. These chemical processes can be predicted by the features of metabolites; and metabolites are these low level biological entities. By using mass spectrometers, we can obtain the masses and intensities of metabolites. Note that the mass suggests different types of metabolites, whereas the intensity displays the level of concentration. Thereby, a mass spectrum can be used to distinguish and visualise the observed metabolites; and combinations of specific metabolites types   may suggest a specific pattern for the chemical process leading to a particular formation.

% machine learning in vision
\par In the terms of machine learning applicability, we are looking into a problem of vision. The metabolites of some biological tissue are sequentially captured within regions of fixed size. Effectively, this process is generating an image: each scan is a pixel, therefore, the collections of scans produce an image. However, each pixel contains a large amount of metabolites, so the visualisation of these in a single image would be rather complex. For this reason, we can choose a specific metabolite type and produce pixels reflecting such metabolite activity. Further, we can choose some specific combination (pattern) of metabolites and produce an image of such pattern concentrations. The problem arises upon finding such valid patterns of metabolites, therefore, we are applying machine learning to find the perspective patterns for us.

% Unsupervised learning > Topic modelling > probabilistic machine learning > LDA > Noise
\par The objective to discover new patterns makes it a problem of unsupervised machine learning. That is, we have no prior settings in how these patterns would look like thus we are utilising machine learning in order to structure the data. The problems of this flavour can be tackled using topic modelling. This branch of unsupervised machine learning is a statistical process. In other words, we are looking into probabilistic models -- the hidden structure of the data is inferred by the frequencies of its instances. To be more specific, we will be focusing on enhancing a particular probabilistic machine learning model -- Latent Dirichlet Allocation (LDA). This model dictates the state-of-the-art of topic modeling and is widely applied in the current research on metabolomics [insert citation]. Since the process of the data collection in metabolomics is prone to noise, we will carry a research on whether Latent Dirichlet Allocation could be applied as a method of error correction.

\par The structure of this proposal is set-up as follows: in Section 2 we present the issue of noise in Metabolomics; in section 3 we cover the literature in topic modeling which focuses on the variations of Latent Dirichlet Allocation; in Section 4 we provide the metric of success of the proposed research; finally, in Section 5, we go into details on how the project will be executed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Statement of Problem}

% Problem: noise > annotating the data
\par The current mass spectrometry (MS) equipment display a significant accuracy error upon reflecting the exact metabolite construction in a given sample. Rather than relying upon the advances of such equipment, we can investigate the prospect of enhancing the data processing techniques. As described in Section 1, topic modelling is capable to distinguish the hidden structures of metabolites. However, at this point of time, there is no well set generative approach on how to configure Latent Dirichlet Allocation to reduce the impact of noise in a given dataset.

% Objectives: utilise LDA to correct the errors
\par The primary goal of this research is to expand the consensus on whether topic modelling is a valid of approach to smooth datasets. To be more specific, the objectives we will be looking into are listed as follows:
\begin{enumerate}
    \item Adapt and utilise the generative Latent Dirichlet Allocation models in metabolomics;
    \item Reduce the impact of noise in the datasets of metabolomics;
    \item Give a basis to the generative Latent Dirichlet Allocation model in error correction.
\end{enumerate}
The first objective is expected to investigate the perspective applications of topic modelling which are successfully utilised in other domains. Effectively, we want to assess the general models (or models displaying significant results in other fields). The second objective would adapt, and thus enhance, these general models for the particular use of error correction in metabolomics. Finally, the third objective can be treated as a general contribution to the field of machine learning.

% Hypothesis: topic modelling can be utilised in the error correction
\par We are setting a hypothesis that topic modelling, or more specifically Latent Dirichlet Allocation, can be utilised to correct the errors in the datasets of metabolomics. Both cases of proving and disproving this hypothesis would be a contribution in metabolomics. The successful outcome would lead to the potential applications provided in the following paragraph, whereas the unsuccessful outcome would set-up additional requirements to tackle this issue.

% Results: tools to annotate the metabolites > basis for generative error correction 
\par The potential applications of this research can be induced directly induced from the set-up objectives. First of all, we would show that models from other domains can be applied in metabolomics and, thus, generalised. Further, the developed error correction methodology for the datasets of Metabolomics would improve the accuracy of the chemical processes prediction. Finally, the generalisation of the developed error correction methodologies would provide an alternative tool-kit for other domains as well as expand the potential applications of the results derived from metabolomics.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background Survey}

% intro
\par The provided literature review on latent Dirichlet allocation assesses its underlying methodology and proposes several different implementation variations. The review is structured in progressing order: to start with, we look into the original latent Dirichlet allocation paper and define the terminology which will be used throughout this proposal; then, we familiarise with two different approaches used for the inference of latent Dirichlet allocation;  finally, we review the perspective advances which could be used for error correction.

\subsection{Terminology}

% Topic modelling in a nutshell > Example
\par Recall that in topic modelling we are trying to induce some hidden structure within the studied data. For instance, if we were analysing a book, it would be likely to address several different topics. Further, the chapters of the book itself would have unique concentrations of such topics. In order to cluster these topics, a probabilistic topic modelling method would look into the structure of the words used in the book. Also, the performance of the method would depend on its ability to optimise this clustering, for example, we could dismiss the commonly used words such as: and, or, the, etcetera. Note that in we will assess only the topic modelling where we do not know the naming of topics. That is, we will be looking into a \textit{unsupervised learning} problem.

% Terminology
\par The terminology used in this proposal is equivalent to the one used in the survey on probabilistic topic models carried by Blei et al. [citation]. Figure \ref{fig:lda} below represents the latent Dirichlet allocation model in plate notation. We will use it as a guide to familiarise with the terminology.   
\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{lda}
  \caption{Latent Dirichlet allocation. COPYRIGHT. CHANGE M TO D}
  \label{fig:lda}
\end{figure}

% High level terminology.
\par To start with, we will familiarise with the domain of the model in a high level. The grey circle represents a single word, whereas a collection of words (document) is represented by the inner plate, and the outer plate corresponds to a collection of documents (corpus). In the model, words are the only \textit{observable variables} -- this is indicated by the grey colour of the circle, whereas the white coloured circles indicate the \textit{hidden variables}. As suggested by the figure, a \textit{word} is denoted by $w$; a \textit{document} (sequence of words) is denoted by $\mathbf{w} = \{w_1, w_2, \dots, w_N\}$, where $N$ indicates the number of words in the document; and a \textit{corpus} is denoted by $\mathcal{\mathbf{D}} = \{\mathbf{w_1}, \mathbf{w_2}, \dots, \mathbf{w_M}\}$, where $D$ is the number of documents in the corpus.

% Low Level terminology.
\par Now, we will introduce the lower level entities -- the hidden variables. Recall that latent Dirichlet allocation attempts to induce an arbitrary mixture of topics. The corpus and each document is expected to vary in the proportions of the induced topics; and each document is likely to have a different number of word. Let's denote the number of such arbitrary topics by $K$, and the number of unique words in the corpus by $V$. This notation will simplify the explanation on the remaining entities in Figure \ref{fig:lda}. To start with, $\beta$ is the hyper-parameter (a matrix of dimension $V \times K$) referring to the prior distribution of the topics over the vocabulary. The hyper-parameter $\alpha$ (a vector of length $K$) refers to the prior distribution of the topics over the documents. Further, the latent variable $\theta$ (a matrix of dimension $D \times K$) is the topic distribution over documents. Finally, $z$ is the latent variable on the topic assignments of each word in every document, effectively, it is a matrix of dimension $M \times V$.

\subsection{Preliminaries}
% Relationship between the variables.
% To-do: COPYRIGHT for the equation.
\par The relationship between observable and hidden variables, effectively, describes the inference of the underlying topic structure. We will go through an arbitrary example in order to give intuition in the inference process and provide the notation of the variable instances. By looking into Figure \ref{fig:lda}, we can see that $\alpha_k$ (the prior weight of the $k$th topic in a document) influences $\theta_{:, k}$ (the latent $k$th topic distribution over $M$ documents). It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in the document $d$) influences $z_{d, :}$ (the topic assignment over $V$ vocabulary terms in document $d$). Note that $z_{d, v}$ would refer to the assignment of some specific vocabulary term $v$. Now, by taking $z_{d, v}$ and $\beta_{v, k}$ (the prior weight of the $k$th topic on the vocabulary $v$) we describe the origins of word $w_{d, v}$. The relationship of the terms, effectively, induces the generation of the corpus, this can be expressed as the following joint probability distribution
\begin{equation}
p(\beta, \theta, z, w) = \prod_{k=1}^Kp(\beta_{:, k})\prod_{d=1}^Dp(\theta_{d, :})\bigg(\prod_{v=1}^Vp(z_{d, v} | \theta_{d, :})p(w_{d, v} | \beta, z_{d, v})\bigg).
\end{equation}
By using Bayes Theorem, we derive the computation of the conditional distribution
\begin{equation}
p(\beta, \theta, z | w) = \frac{p(\beta, \theta, z, w)}{p(w)}. 
\end{equation}
This equation expresses the computation of the posterior -- the derivation of the hidden corpus structure given observable variable $w$. The problem arises from inefficient analytical computation of evidence $p(w)$. By reviewing the literature on latent Dirichlet allocation [CITATION], we will also familiarise with the numerical methods used in approximating $p(w)$.

% Intro > methodology > experiments
\subsection{Latent Dirichlet allocation}

% Intro:
% Attention to detail giving a basis fo the following model interpretations
\par Latent Dirichlet allocation is a probabilistic topic modelling technique  proposed by Blei et al [citation]. Since this subsection is dedicated for the origins of the technique, it will be broader in detail. This will cover aspects addressed in the following subsections covering the advances on latent Dirichlet Allocation.

% Assumptions: 
% - Exchengable docs and words > bag of words
% - Discrete data
% - Dimensionality of topic variable is assumed to be known and fixed\
\par The initial latent Dirichlet allocation model sets some assumptions for the data which will be processed by the model. First of all, it is assumed that both documents and words in the documents are \textit{exchangeable}. That is, the order in which these will be processed does not matter. More formally, this assumption means that the model follows the \textit{bag-of-words} principle. The next assumption is that the model works on discrete data. Effectively, the original paper does not cover the treatment of continuous features. Another assumption is on the dimensionality of the topic variables: the number of topics is set to be fixed during the complete run of inference.

% Document Generation:
% - Algorithm
% - Alpha and Beta are sampled once, whereas w and z are sampled once per each word in a document
% - Three level bayesian hierarchy~~
\par As discussed in the preliminaries subsection, the process of latent Dirichlet allocation assumes that the corpus can be generated by a probabilistic process. The generation of the corpus is made of three levels: the parameters $\alpha$ and $\beta$ are sampled once; the variable $\theta$ is sampled once per document; and the variables $z$ and $w$ are sampled once per word. The generative algorithm given in the original paper is equivalent to Algorithm 1 below.
\begin{algorithm}[H]
\caption{Document Generation.}
\label{alg:document_generation}
\begin{algorithmic}[1]
% \Procedure {generate_document}{$\xi, \alpha, \beta$}
\State $N \sim \mbox{Poisson}(\xi)$
\State $\theta \sim \mbox{Dir}(\alpha)$
\For {$n \leftarrow 1, N$}
\State $z_n \sim \mbox{Multinomial}(\theta)$
\State $w_n \sim \mbox{p}(w_n | z_n, \beta)$
\EndFor
% \EndProcedure
\end{algorithmic}
\end{algorithm}
Note that the algorithm describes the generation of a single document. In order to generate the corpus, we would run the algorithm several times. To start with, in the line 1 we draw the number of words in the document $N$, where $\xi$ is an ancillary variable suggesting the mean for the Poisson Distribution. In the line 2 we draw the topic distribution $\theta$ from the Dirichlet distribution on $\alpha$. Now in the lines 3--6 we generate the words: in the line 4 we draw topic $z_n$ from the multinomial distribution on $\theta$; and, finally, in the line 5 we obtain the word $w_n$ from the conditional probability on $z_n$ and the distribution over the words in the vocabulary $\beta$.

% Inference
% - Setting alpha and beta as hypterparameters, i.e. reforming the posterior formula. (untouched)
% - Variational inference: variational parameters gamma and fi
\par The inference is, effectively, the computation of the evidence for the posterior distribution. The original latent Dirichlet allocation paper uses the method of variational inference. Basically, the authors introduce the variational parameters $\gamma$ and $\phi$ (computation of these set-ups a problem of optimisation). Note that the original paper provides pseudo code for solving the optimisation problem.

% Parmeter estimation
% - M Step update for beta
% - Update for alpha using newton-raphson method
% Problem with smoothing: Large datasets > sparsity, training corpus blocks the possibility of the new words coming from new documents
\par Additionally, the authors emphasise the estimations of the parameters $\alpha$ and $\beta$. They suggest a two-step EM procedure utilising the variational parameters $\gamma$ and $\phi$: E-step is the optimisation problem addressed in the previous paragraph; and M-step updates the parameters $\alpha$ and $\beta$. Note that the authors provide an analytical update for $\beta$ and a Newton--Raphson method for updating $\alpha$. Apart from that, the authors also introduce the issue of a sparse corpus. Such corpus would be expected upon having a large vocabulary or a significant amount of documents. The authors suggest addressing this problem by introducing smoothing. They introduce an additional parameter $\eta$ which impacts the estimation of $\beta$.

% Experiments:
% - Document modelling: issue with pLSI and mixture unigrams~; perplexity (UNTOUCHED)
% - Document classification: dimensionality reduction; comparison with SVM
\par The conducted experiments display the improvement performance over the predecessor models and suggest the application domain of latent Dirichlet allocation. The first experiment addresses document modelling. The latent Dirichlet allocation model is reported to display balanced hidden topic proportions in the set of test documents. That is, the test set has replicated the proportions of the training set. The second experiment addresses document classification. In order to display better performance compared to a support vector machine (SVM) model, the authors address the dimensionality reduction of the tested documents. This means that the documents would possess a lower amount of features. It follows that the latent Dirichlet allocation model has successfully reduced the number of features and displayed an improvement in accuracy compared to the support vector machine model.

% Discussion and Conclusion
\par The review on the initial latent Dirichlet allocation model has set-up a basis for introducing the successor variations. These will be provided in the following sections. We have introduced the initial methodology for the following reasons: (1) the novel models will display the relaxation of the initial assumptions; (2) the provided document generation methodology will be compared to the document generation in batches; (3) we will cover different inference and parameter estimation approaches. Also, note that the paper is relatively old in time, and the experiments do not display the state-of-the-art performance. Rather than that, they were introduced to address the application domains of the model.

\subsection{Dynamic Topic Models}

% Intro:
% - Time series
% - Both qualitative and quantitative results
\par

% Assumptions:
% - Data is divided in time slices
\par

% Generative Process:
\par

% Inference:
% Variational methods are a deterministic alternavive for stochastic simulation
% Two variations: Kalman filter and wavelet regression
\par

% Experiments:
% - Evolution of Science
\par

% Concluison
\par

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed Approach}

state how you propose to solve the software development problem. Show that your proposed approach is feasible, but identify any risks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Work Plan}

show how you plan to organize your work, identifying intermediate deliverables and dates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% it is fine to change the bibliography style if you want
\bibliographystyle{plain}
\bibliography{mprop}
\end{document}

% To-do: relaxing the matrix dimension assumptions, i.e., introducing the dictionary variation

### 2016-12-01 docs/reports/report_generative-process/report_generative-process.tex
Adding dynamic topic modelling.

\documentclass{article}
\usepackage{graphicx}
\graphicspath{ {images/} }
\usepackage{float}

% alternative font if you prefer
% \usepackage{times}

% for alternative page numbering use the following package
% and see documentation for commands
\usepackage{fancyheadings}

% other potentially useful packages
\usepackage{amssymb,amsmath}
%\usepackage{url}
%\usepackage{fancyvrb}
%\usepackage[final]{pdfpages}
\setlength{\parindent}{0pt}
\setlength{\parskip}{11pt plus 2pt}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Report: Generative Processes of Static and Dynamic LDA Models}
\author{Arijus Pleska}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% - Comparing the static and dynamics models.
% - Questions 
\par During this week I have familiarised with the generative processes of static and dynamic LDA models. In this report I will present the results by plotting the generated documents, address the estimation of the parameters $\alpha$ and $\beta$, and raise questions to be considered for the next meeting.

% Generative process
\par I have managed to generate documents using both static and dynamic variations. The static approach in generating corpus is given in Figure 1, and the dynamic approach is given in Figure 2.

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{static}
  \caption{The distribution of a word in a corpus produced by the static generative process.}
  \label{fig:static}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=\textwidth]{dynamic}
  \caption{The distribution of a word in a corpus produced by the dynamic generative process.}
  \label{fig:dynamic}
\end{figure}

In Figure 1 we can see that the word is present throughout all the corpus, whereas Figure 2 displays a smoother distribution of the word. That is, the dynamic generative process induces a structure of the word's existence in the corpus. Also, note that both corpus contain the same number of documents and the number of words in a document is drawn from the Poisson distribution on the same mean $\xi = 200$.

The implementation of the generative processes follow directly from the original LDA paper \cite{Blei:2003:LDA:944919.944937} and the paper on dynamic topic models \cite{Blei:2006:DTM:1143844.1143859}.

\bibliographystyle{unsrt}
\bibliography{report_generative-process}

\end{document}

### 2016-12-01 code/d_gibbs_lda.py

@@ -1,0 +1,126 @@
+import os
+import pandas as pd
+import numpy as np
+
+HOME_PATH = os.path.expanduser('~')
+np.set_printoptions(edgeitems=5)
+
+
+class Dynamic_Gibbs_LDA(object):
+
+    def __init__(self, K, iter_count, n_seq):
+        self.iter_count = iter_count
+        self.K = K
+        self.n_seq = n_seq
+        self.thetas_ = []
+        self.phis_ = []
+        self.alphas_ = []
+
+    def fit(self, corpus):
+        self._fit(corpus)
+
+    def dump(self, array, path):
+        pickle = np.array(array)
+        pickle.dump(path)
+
+    def _fit(self, corpus):
+        list_of_seq_corpus = np.array_split(corpus, self.n_seq)
+
+        for seq_c, seq_corpus in enumerate(list_of_seq_corpus):
+
+            self._initialise(seq_corpus, seq_c)
+            print(self.alpha.shape)
+            print(self.alpha)
+            print(self.alpha.sum())
+            print(self.beta.shape)
+            print(self.beta[0, :])
+            print(self.beta[0, :].sum())
+
+            for it in range(self.iter_count):
+                
+                print("Sequence: %s, Iteration: %s." % (seq_c, it)) 
+                self._sample_topics(seq_corpus)
+                self._store_theta()
+                self._store_phi()
+                print("Current theta:")
+                print(self.thetas_[it]) 
+
+            self._dynamic_step(seq_c)
+
+
+    def _initialise(self, corpus, iteration):
+        self.D, self.W = corpus.shape
+        self._total_frequencies = corpus.sum()
+        if iteration == 0:
+            self.alpha = np.full(self.K, 1./self.K)
+            self.beta = np.full((self.K, self.W), 1./self.W)
+
+        self._zw_counts = np.zeros((self.K, self.W), dtype=np.int)
+        self._dz_counts = np.zeros((self.D, self.K), dtype=np.int)
+        self._z_counts = np.zeros(self.K, dtype=np.int)
+        self._z_history = []
+
+        for d, doc in enumerate(corpus):
+            self._z_history.append([])
+            for w, word in enumerate(doc):
+                self._z_history[d].append([])
+                word_frequency = int(word)
+                for _ in range(word_frequency):
+                    z = np.random.randint(self.K)
[…]

### 2016-12-02 docs/MSci_proposal/mprop.tex
Dynamic topic model.

@@
 We will use it as a guide to familiarise with the terminology.
-\begin{figure}[h]
+\begin{figure}[H]
   \centering
@@
 % - Both qualitative and quantitative results
-\par
+\par Now we review the paper on Dynamic Topic Models [CITATION].
+Effectively, the dynamic topic model can be treated as an extension of static topic models (such as latent Dirichlet allocation).
+The main idea behind a dynamic topic model is that it induces the topic evolution the processed corpus.
+That is, the documents in the corpus evolve over time.
+In this subsection, we review the key assumptions of dynamic topic models, go over the generative process of documents, review the suggested inference methods, and familiarise with the results of the carried experiment.
 
@@
 % - Data is divided in time slices
-\par
+\par We will emphasise the assumptions given on the dynamic topic models.
+First of all, even though we are discussing a model of time series suggesting the use of continuous variables, the data remains categorical.
+The distinction between continuous and categorical data is that continuous data is strictly numeric and infinite, whereas categorical data can take only particular values.
+The second assumption relaxes the ex-changeability on documents.
+That is, the documents are expected to form a sequence.
+Or in other words, the corpus is divided into time-slices.
+Note that in such time-slice documents are exchangeable.
 
 % Generative Process:
-\par
+\par The generative process of the models updates the parameters $\alpha$ and $\beta$.
+This means that the topic distributions over documents and words change over time.
+The visualisation of the process is given in Figure 2 below.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{dynamic_topic_model}
+  \caption{Dynamic topic model. COPYRIGHT.}
+  \label{fig:dtm}
+\end{figure}
+Notice that each time-slice is equivalent to the model shown in Figure 1.
+Also, the parameters $\alpha$ and $\beta$ are induced by their predecessors in the previous time-slice.
+More formal representation of the generative process is given in Algorithm 2.
+\begin{algorithm}[H]
+\caption{Dynamic document generation.}
+\label{alg:dynamic_document_generation}
+\begin{algorithmic}[2]
+\State $\beta_t | \beta_{t-1} \sim \mathcal{N}(\beta_{t-1}, \sigma^2)$
+\State $\alpha_t | \alpha_{t-1} \sim \mathcal{N}(\alpha_{t-1}, \delta^2)$
+\For {$d \leftarrow 1, D$}
+\State $\eta_d \sim \mathcal{N}(\alpha_{t}, a^2)$
+\For {$n \leftarrow 1, N$}
+\State $z_{d, n} \sim \mbox{Multinomial}(\pi(\eta_d))$
+\State $w_{t, d, n} \sim \mbox{Multinomial}(\pi(\beta_t | z_{d, n}))$
+\EndFor
+\EndFor
+% \EndProcedure
+\end{algorithmic}
+\end{algorithm}
+The given algorithm describes the generative process for the time-scale $t$.
+It starts by drawing the parameters $\alpha_t$ and $\beta_t$ from the Gaussian distribution with the means given by the previous parameter values.
+For now, we will not discuss the meaning of the variance values.
+Further, for each document in the time-scale, we carry the following procedure: we draw the topic distribution in a document $\eta_d$ from the Gaussian distribution with the mean given by $\alpha_t$; then, for each word in the document, we draw the topic $z_{d, n}$ from the Multinomial distribution on the parametrised $\eta_d$, and, finally, we draw the word $w_{t, d, n}$ from the Multinomial distribution on the parametrised $\beta_t$ which is conditioned on $z_{d, n}$.
+By the use of \textit{parametrisation}, we are expressing the result of the Gaussian distributions into a format suitable for Multinomial distribution.
+This mapping is expressed as
+\begin{equation}
+\pi(x)_n = \dfrac{\exp{(x_n)}}{\sum^n_{i=1}\exp{(x_i)}}.
+\end{equation}
 
@@
 % Two variations: Kalman filter and wavelet regression
-\par
+\par The authors suggest using variational methods to approximate the inference posterior.
+It is claimed that stochastic simulation would be unable to scale with large data sets.
+Two techniques of variational methods are provided: Kalman Filtering and Wavelet Regression.
+The implementation details of these are provided in a brief manner suggesting the computation of the parameter $\beta_t$.
 
@@
 % - Evolution of Science
-\par
+\par The dynamic topic model is evaluated by conducting an experiment on the journals of \textit{Science}.
+The experiment is expected to establish the time series on the topic development with the articles.
+Indeed, the experiment is able to produce results displaying the rises and falls of the topics indicating scientific fields.
+Note that the experiment is performed using both variational inference methods.
 
 % Concluison
-\par
+\par The review on the dynamic topic models has familiarised us with more realistic generative process: the documents are induced by time series.
+Further, we have provided the visualisation of the model and introduced the algorithm for the generative process.
+Also, we have captured the techniques used for inference.
+Finally, we have familiarised with the experiment settings.
 

### 2016-12-03 docs/MSci_proposal/mprop.tex
Gibbs sampling review.

@@
 
+\subsection{Finding scientific topics}
+
+% Intro:
+% - The paper
+% - Sampling as a method of inference
+% - Experiments
+\par The paper on "Finding scientific topics" [CITATION] suggests an alternative approach to the inference in latent Dirichlet allocation.
+To be more specific, the paper provides a readable introduction on how to apply a Markov chain Monte Carlo (MCMC) algorithm as a sampling-based method for the inference in topic modelling.
+Further, the authors discuss the method's application settings in vision and conduct an experiment on deriving topics from the abstracts of papers published in "Proceedings of the National Academy of Sciences" (PNAS).
+Finally, the authors compare the performance of the suggested sampling-based inference method to variational inference methods.
+
+% The method (Gibbs Sampling)
+% - Introduce the meaning of phi and theta.
+% - Introduce the inference method
+\par The inference method using a Markov chain Monte Carlo algorithm can be referred by \textit{Gibbs sampling}.
+The authors claim that the method provides a first-order approximation in deriving the topics on the given corpus.
+That is, we can establish quantitative reasoning on the document correlation in terms of content and content's change over time.
+Recall that by $\theta$ we denote the topic distribution over documents and by $\phi$ we denote the topic distribution over words.
+In Gibbs sampling, these values can be obtained by examining the posterior distribution.
+Essentially, the variables are sequentially sampled from their respective distributions.
+The process draws topic assignment $z_i$ and keeps a track of this assignment with respect to each word and document.
+Note that $z_i$ is drawn from the following distribution:
+\begin{equation}
+P(z_i = j | z_{-i}, w) \propto \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + W\beta}\dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + T\alpha}.
+\end{equation}
+It follows that this distribution is utilise the previously mentioned topic assignments.
+For a brief familiarisation with the notation used in the previous expression, note the following: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$; $W$ is the number of words; and $T$ is the number of topics.
+Ultimately, the left side of the expression is $\phi$ and the right side is $\theta$.
+That is,
+\begin{align*}
+\phi_j &= \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)}},\\
+\theta_j &= \dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + T\alpha}.
+\end{align*}
+This process of topic assignment would be executed for every word in a document and for all documents in the corpus.
+The procedure would be repeated until the distributions displayed fluctuations below some set threshold.
+Notice that the procedure does not require direct assignment of $\theta$ and $\phi$.
+If needed, the values of these can be sampled upon converge by using the previously provided representations.
+
+% Experiments
+% - Time series
+% - PNAS
+\par The authors have executed Gibbs sampling in the domain of vision.
+They have represented an image as a document and each pixel as a word.
+This means that they were able to recover the underlying basis from which the images could be constructed.
+The performance was compared to the latent Dirichlet allocations models which use variational inference.
+Note that we define the performance in terms of perplexity, which indicates the uncertainty of the word assignment to a topic.
+It appears that, with relatively small data sets, Gibbs sampling displays superior performance.
+Further, the authors have investigated the performance upon varying the number of topics.
+The results indicate that there exists a trade-off between the model being able to cover the significant underlying topics (low number of topics) and start deriving irrelevant branches (high number of topics).
+
+\par The main experiment was carried in the PNAS data set, note that the processed papers were published from 1991 to 2001.
+By having a varying notion of time, the authors were able to investigate the prevailing topics.
+Thus, establish a sense in dynamic topic modelling.
+Even though the model was not adjusted to smooth the time-series, from the derived results we can deduce the varying relevance of underlying science domains.
+
+% Conclusion
+\par We have reviewed an alternative method for the inference in latent Dirichlet allocation.
+The idea behind the inference process was explained by considering the primary aspects: the notion of counts and the implicit use of hidden variables.
+Also, we have introduced the experiment settings its results.
+This allowed to familiarise with an example of topic modelling application and vision.
+Further, the experiment on PNAS has emphasised the relevance of time series.
+
 \subsection{Dynamic Topic Models}

### 2016-12-05 docs/MSci_proposal/mprop.tex
Fixing the outline.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Smooth Latent Dirichlet Allocation in Metabolomics}
+\title{Probabilistic Topic Modelling for Reducing the Impact of Noise in Metabolomics}
 \author{Arijus Pleska}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}\label{intro}
 
-% Hook intro
-\par This research proposal is directed towards applying machine learning in biology.
+% Metabolomics
+% Vision
+
+\section{Introduction}
+
+% Intro
+% - ML in biology
+% - Data patterns
+% - Noise
+\par This research proposal suggests an application of machine learning in biology.
 We intend to study the scans of low level biological entities in order to derive their patterns.
@@
 % Metabolomics
+% - Definition
+% - Data gatheting technology
 \par The branch of biology we are expecting to carry out the research is metabolomics.
@@
 
-% machine learning in vision
+% The domain of vision
+% - Expalanation how the scans of metabolites can be treated as images.
+% - Visualisation of particular metabolites or metabolite patterns.
 \par In the terms of machine learning applicability, we are looking into a problem of vision.
@@
 
-% Unsupervised learning > Topic modelling > probabilistic machine learning > LDA > Noise
+% Unsupervisder learning
+% - Topic modelling
+% - Probabilistic processes
+% - LDA variations
 \par The objective to discover new patterns makes it a problem of unsupervised machine learning.
-That is, we have no prior settings in how these patterns would look like thus we are utilising machine learning in order to structure the data.
+That is, we have no initial settings in how these patterns would look like thus we are utilising machine learning in order to structure the data.
 The problems of this flavour can be tackled using topic modelling.
@@
 
+% Outline
+% - Sections
 \par The structure of this proposal is set-up as follows: in Section 2 we present the issue of noise in Metabolomics; in section 3 we cover the literature in topic modeling which focuses on the variations of Latent Dirichlet Allocation; in Section 4 we provide the metric of success of the proposed research; finally, in Section 5, we go into details on how the project will be executed.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Statement of Problem}
 
-% Problem: noise > annotating the data
-\par The current mass spectrometry (MS) equipment display a significant accuracy error upon reflecting the exact metabolite construction in a given sample.
+% Intro
+% - Suggesting an advance in processing rather than equipment.
+\par Current mass spectrometry (MS) equipment displays a significant accuracy error upon reflecting the exact metabolite construction in a given sample.
 Rather than relying upon the advances of such equipment, we can investigate the prospect of enhancing the data processing techniques.
@@
 
-% Objectives: utilise LDA to correct the errors
+% Objectives
+% - Research in the state-of-the-art methodology.
+% - Reducing the impact of noise in metabolomics.
+% - Providing a general-purpose machine learning model.
 \par The primary goal of this research is to expand the consensus on whether topic modelling is a valid of approach to smooth datasets.
@@
 
-% Hypothesis: topic modelling can be utilised in the error correction
+% Hypothesis
+% - Topic modelling can improve the quality of metabolomics data sets.
+% - The negation of the hypothesis would contribute in terms of reassessing the complexity of the problem.
 \par We are setting a hypothesis that topic modelling, or more specifically Latent Dirichlet Allocation, can be utilised to correct the errors in the datasets of metabolomics.
@@
 % Results: tools to annotate the metabolites > basis for generative error correction 
+% - Expanding the application of existing machine learning models.
+% - Advancing the current state-of-the-art in metabolomics.
+% - Contributing to machine learning in general.
 \par The potential applications of this research can be induced directly induced from the set-up objectives.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Background Survey}
 
-% intro
+% Intro
+% - The reason why the literature review is conducted.
+% - Outline: preliminaries, notation, literature.
+% - A notification the the results will be discussed in the following section.
 \par The provided literature review on latent Dirichlet allocation assesses its underlying methodology and proposes several different implementation variations.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Terminology}
 
-% Topic modelling in a nutshell > Example
+% General intuition on topic modelling
+% - Example
 \par Recall that in topic modelling we are trying to induce some hidden structure within the studied data.
@@
 In order to cluster these topics, a probabilistic topic modelling method would look into the structure of the words used in the book.
-Also, the performance of the method would depend on its ability to optimise this clustering, for example, we could dismiss the commonly used words such as: and, or, the, etcetera.
+Also, the performance of the method would depend on its ability to optimise this clustering, for example, we could dismiss the commonly used words (such as and, or, the, etc.).
 Note that in we will assess only the topic modelling where we do not know the naming of topics.
@@
 
-% Terminology
+% LDA intro
+% - the model
+% - A notification that the terminology and the notation will be useful throughout the proposal.
 \par The terminology used in this proposal is equivalent to the one used in the survey on probabilistic topic models carried by Blei et al.
@@
 
-% High level terminology.
+% High level LDA terminology
+% - Intro to observable and hidden variables
+% - Notation:
+% -- List the definitions
+% -- Explain the definitions
 \par To start with, we will familiarise with the domain of the model in a high level.
@@
 
-% Low Level terminology.
+% Low level LDA terminology
+% - An explanation on the hidden variables
+% - Notation:
+% -- List the definitions
+% -- Explain the definitions
 \par Now, we will introduce the lower level entities -- the hidden variables.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Preliminaries}
-% Relationship between the variables.
-% To-do: COPYRIGHT for the equation.
+
+% Basic LDA
+% - Generative process
+% - The derivation of the posterior
 \par The relationship between observable and hidden variables, effectively, describes the inference of the underlying topic structure.
@@
 By reviewing the literature on latent Dirichlet allocation [CITATION], we will also familiarise with the numerical methods used in approximating $p(w)$.
+% To-do: COPYRIGHT for the equation
+% To-do: reconsider the name of the subsection
 
-% Intro > methodology > experiments
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Latent Dirichlet allocation}
 
-% Intro:
-% Attention to detail giving a basis fo the following model interpretations
+% Intro
+% - Detailed approach in covering implementations details
 \par Latent Dirichlet allocation is a probabilistic topic modelling technique  proposed by Blei et al [citation].
@@
 
-% Assumptions: 
-% - Exchengable docs and words > bag of words
-% - Discrete data
-% - Dimensionality of topic variable is assumed to be known and fixed\
+% Assumptions 
+% - Bag of words principle
+% - Discrete data sets
+% - Topic dimensionality
 \par The initial latent Dirichlet allocation model sets some assumptions for the data which will be processed by the model.
@@
 
-% Document Generation:
-% - Algorithm
-% - Alpha and Beta are sampled once, whereas w and z are sampled once per each word in a document
-% - Three level bayesian hierarchy~~
+% Generative process
+% - Three level hierarchy
+% - The generative algorithm
 \par As discussed in the preliminaries subsection, the process of latent Dirichlet allocation assumes that the corpus can be generated by a probabilistic process.
@@
 \begin{algorithmic}[1]
-% \Procedure {generate_document}{$\xi, \alpha, \beta$}
 \State $N \sim \mbox{Poisson}(\xi)$
@@
 \EndFor
-% \EndProcedure
 \end{algorithmic}
@@
 
-% Inference
-% - Setting alpha and beta as hypterparameters, i.e. reforming the posterior formula. (untouched)
-% - Variational inference: variational parameters gamma and fi
+% Variational inference 
+% - Variational parameters gamma and fi
 \par The inference is, effectively, the computation of the evidence for the posterior distribution.
@@
 Note that the original paper provides pseudo code for solving the optimisation problem.
+% To-do: connect with the parameter estimation paragraph
+% To-do: reforming the posterior formula
 
-% Parmeter estimation
-% - M Step update for beta
-% - Update for alpha using newton-raphson method
-% Problem with smoothing: Large datasets > sparsity, training corpus blocks the possibility of the new words coming from new documents
+% Parameter estimation
+% - M-step is an update of beta
+% - Newton--Raphson is used to update alpha
+% - The smoothing issue: 
+% -- Sparsity from large data sets
+% -- The incoorperation of documents which were not used for the training
 \par Additionally, the authors emphasise the estimations of the parameters $\alpha$ and $\beta$.
@@
 
-% Experiments:
-% - Document modelling: issue with pLSI and mixture unigrams~; perplexity (UNTOUCHED)
-% - Document classification: dimensionality reduction; comparison with SVM
+% Experiments
+% - Document modelling: issue with pLSI and mixture unigrams~ 
+% - Document classification: 
+% -- Dimensionality reduction
+% -- Comparison with SVM
 \par The conducted experiments display the improvement performance over the predecessor models and suggest the application domain of latent Dirichlet allocation.
@@
 It follows that the latent Dirichlet allocation model has successfully reduced the number of features and displayed an improvement in accuracy compared to the support vector machine model.
+% To-do: explain perplexity (UNTOUCHED)
 
-% Discussion and Conclusion
+% Conclusion
+% - Summary of the results established by the paper
 \par The review on the initial latent Dirichlet allocation model has set-up a basis for introducing the successor variations.
@@
 Rather than that, they were introduced to address the application domains of the model.
+% To-do: rework the paragraph to reflect its purpose: remove the discussion
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Finding scientific topics}
 
-% Intro:
-% - The paper
+% Intro
 % - Sampling as a method of inference
-% - Experiments
 \par The paper on "Finding scientific topics" [CITATION] suggests an alternative approach to the inference in latent Dirichlet allocation.
@@
 
-% The method (Gibbs Sampling)
-% - Introduce the meaning of phi and theta.
+% Gibbs Sampling
+% - Recall the meaning of phi and theta.
 % - Introduce the inference method
@@
 
-% Experiments
+% Auxilliary xperiments
+% - Vision
 % - Time series
-% - PNAS
 \par The authors have executed Gibbs sampling in the domain of vision.
@@
 
+% The main experiment
+% - PNAS
 \par The main experiment was carried in the PNAS data set, note that the processed papers were published from 1991 to 2001.
@@
 This allowed to familiarise with an example of topic modelling application and vision.
-Further, the experiment on PNAS has emphasised the relevance of time series.
+Further, the experiment on PNAS has emphasised the relevance of time series
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Dynamic Topic Models}
@@
 % - Time series
-% - Both qualitative and quantitative results
+% - Qualitative and quantitative improvement
 \par Now we review the paper on Dynamic Topic Models [CITATION].
@@
 
-% Assumptions:
+% Assumptions
 % - Data is divided in time slices
@@
 
-% Generative Process:
+% Generative Process
+% - Algorithm
+% - An explanation of the algorithm
 \par The generative process of the models updates the parameters $\alpha$ and $\beta$.
@@
 
-% Inference:
+% Inference
 % Variational methods are a deterministic alternavive for stochastic simulation
-% Two variations: Kalman filter and wavelet regression
+% Two variations: Kalman filtering and wavelet regression
 \par The authors suggest using variational methods to approximate the inference posterior.
@@
 
-% Experiments:
+% Experiments
 % - Evolution of Science
@@
 
-% Concluison
+% Conclusion
 \par The review on the dynamic topic models has familiarised us with more realistic generative process: the documents are induced by time series.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Section Conclusion
+\par
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Proposed Approach}
 
-state how you propose to solve the software development problem. Show that your proposed approach is feasible, but identify any risks.
+% Section intro
+\par
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Overview}
+
+% Discussion over the literature
+\par
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Approach}
+
+% Design
+\par
+
+% Implementation
+\par
+
+% Evaluation/Experiments
+\par
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Risk Management}
+
+% Section conclusion
+\par
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Work Plan}
 
-show how you plan to organize your work, identifying intermediate deliverables and dates.
+% Intro
+\par
+
+% Schedule
+% - Dates
+% - Tasks
+\subsection{Schedule}
+
+% Code
+% Dissertation
+\subsection{Deliverables}
+
+% Conclusion
+\par
 

### 2016-12-06 docs/MSci_proposal/mprop.tex
First two paragraphs review.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Probabilistic Topic Modelling for Reducing the Impact of Noise in Metabolomics}
+\title{Probabilistic Topic Modelling to Reduce the Noise in Metabolomics}
 \author{Arijus Pleska}
@@
 % Vision
-
 \section{Introduction}
@@
 We intend to study the scans of low level biological entities in order to derive their patterns.
-To be more specific, we aim to tackle the issue of noise, which arises from the hardware limitation of the data collection devices.
+To be more specific, we will be tackling a problem of noise.
+This issue of noisiness arise from the hardware limitation of the high precision data collection devices.
 
@@
 In brief, metabolomics is the scientific study of chemical processes which involve small molecules -- metamobilites.
-These chemical processes can be predicted by the features of metabolites; and metabolites are these low level biological entities.
-By using mass spectrometers, we can obtain the masses and intensities of metabolites.
-Note that the mass suggests different types of metabolites, whereas the intensity displays the level of concentration.
-Thereby, a mass spectrum can be used to distinguish and visualise the observed metabolites; and combinations of specific metabolites types   may suggest a specific pattern for the chemical process leading to a particular formation.
+Note that these chemical processes can be predicted by the features of metabolites.
+More specifically, mass spectrometry (MS) is often used to gather metabolite data sets.
+By this technique, we have information on the masses and intensities of the metabolites within a sample, i.e., we have the mass spectrum of a sample.
+Note that a different mass suggests different types of metabolites, whereas intensity suggests the level of concentration.
+Thereby, a mass spectrum is used to distinguish and visualise the observed metabolites.
+By deriving the formulae of specific metabolites types, we can induce patterns leading to the prediction of chemical processes.
 

### 2016-12-08 docs/MSci_proposal/mprop.tex
Review until the middle of preliminaries.

@@
 \section{Introduction}
+% To-do:
+% - Mention Rogers' paper for the first attempt to use LDA in Metabolomics datasets
 
@@
 % Metabolomics
-% - Definition
-% - Data gatheting technology
-\par The branch of biology we are expecting to carry out the research is metabolomics.
+% - Mass spectrometry
+\par The application domain of this project is metabolomics -- a branch of computional biology.
 In brief, metabolomics is the scientific study of chemical processes which involve small molecules -- metamobilites.
-Note that these chemical processes can be predicted by the features of metabolites.
-More specifically, mass spectrometry (MS) is often used to gather metabolite data sets.
-By this technique, we have information on the masses and intensities of the metabolites within a sample, i.e., we have the mass spectrum of a sample.
-Note that a different mass suggests different types of metabolites, whereas intensity suggests the level of concentration.
+Ultimately, these chemical processes can be predicted by the features of metabolites.
+We intend to set the basis of our study on data sets acquired by \textit{mass spectrometry} (MS).
+By using this technique, we obtain the masses and intensities of metabolites within a sample of interest.
+This data representation is known as mass spectrum.
+Note that a different mass indicates different types of metabolites, whereas intensity suggests the level of concentration.
 Thereby, a mass spectrum is used to distinguish and visualise the observed metabolites.
@@
 However, each pixel contains a large amount of metabolites, so the visualisation of these in a single image would be rather complex.
-For this reason, we can choose a specific metabolite type and produce pixels reflecting such metabolite activity.
-Further, we can choose some specific combination (pattern) of metabolites and produce an image of such pattern concentrations.
-The problem arises upon finding such valid patterns of metabolites, therefore, we are applying machine learning to find the perspective patterns for us.
+Thereby, we can choose a specific metabolite type and produce pixels reflecting such metabolite activity.
+For a rigorous discussion on machine learning applications in metabolomics, the reader can consult the recent survey conducted by Alonso et al.
+\cite{Alonso-et-al}.
+Going back to the visualisation of metabolites, we can choose a specific combination (structure) and produce an image where activated pixels would suggest the concentration of the chosen metabolite structure.
+The problem arises upon discovering valid matabolite structures.
+Therefore, the techniques of machine learning, or more specifically topic modelling, are used to induce the prospective structures.
 
 % Unsupervisder learning
-% - Topic modelling
-% - Probabilistic processes
-% - LDA variations
-\par The objective to discover new patterns makes it a problem of unsupervised machine learning.
-That is, we have no initial settings in how these patterns would look like thus we are utilising machine learning in order to structure the data.
-The problems of this flavour can be tackled using topic modelling.
-This branch of unsupervised machine learning is a statistical process.
-In other words, we are looking into probabilistic models -- the hidden structure of the data is inferred by the frequencies of its instances.
-To be more specific, we will be focusing on enhancing a particular probabilistic machine learning model -- Latent Dirichlet Allocation (LDA).
-This model dictates the state-of-the-art of topic modeling and is widely applied in the current research on metabolomics [insert citation].
-Since the process of the data collection in metabolomics is prone to noise, we will carry a research on whether Latent Dirichlet Allocation could be applied as a method of error correction.
+% - Topic moels
+% - Probabilistic models
+% - LDA:
+% -- General
+% -- Metabolomics
+\par The objective to discover novel structures is treated as a problem of \textit{unsupervised} machine learning.
+That is, we have no initial settings in how these structures are expected to look like: the structures will be inferred by machine learning methodology.
+The inference of unknown structures is studied within \textit{topic modelling}.
+Further, note that we are concentrating on statistical topic modelling.
+In other words, we are putting the emphasis on the \textit{probabilistic} models: the structures are inferred from the data instance distributions.
+To be more specific, we are focusing to enhance a particular probabilistic machine learning model -- \textit{latent Dirichlet allocation} (LDA).
+First of all, the successor LDA models are treated as a state-of-the-art method in tackling semantic-analysis type problems.
+For more detailed review on the LDA applications, the reader can consult the survey \cite{blei_2012} conducted by Blei.
+Further, Hooft et al.
+\cite{hooft} has recently addressed the applicability of LDA in metabolomics.
+Note that this proposal is directly influenced by the latter prospect: we will attempt to utilise the LDA-like models to reduce the impact of noise in the data sets of the metabolomics domain.
 
@@
 % - Sections
-\par The structure of this proposal is set-up as follows: in Section 2 we present the issue of noise in Metabolomics; in section 3 we cover the literature in topic modeling which focuses on the variations of Latent Dirichlet Allocation; in Section 4 we provide the metric of success of the proposed research; finally, in Section 5, we go into details on how the project will be executed.
+\par The structure of this proposal is given as follows: in Section 2 we present the issue of noise in Metabolomics; in section 3 we cover the literature in topic modeling which focuses on the variations of Latent Dirichlet Allocation; in Section 4 we provide the metric of success of the proposed research; finally, in Section 5, we go into details on how the project will be executed.
+% To-do: rewrite the outline
 
@@
 % - Suggesting an advance in processing rather than equipment.
-\par Current mass spectrometry (MS) equipment displays a significant accuracy error upon reflecting the exact metabolite construction in a given sample.
-Rather than relying upon the advances of such equipment, we can investigate the prospect of enhancing the data processing techniques.
-As described in Section 1, topic modelling is capable to distinguish the hidden structures of metabolites.
-However, at this point of time, there is no well set generative approach on how to configure Latent Dirichlet Allocation to reduce the impact of noise in a given dataset.
+\par Currently, MS equipment is incapable to fully represent the metabolite construction.
+As discussed by Palmer et al.
+\cite{palmer}, even though the current machinery differentiates metabolites in millidalton precision, we still obtain a significant amount of false positives.
+Thereby, rather than relying upon the advances of MS equipment, we can investigate the prospect of enhancing the data processing techniques.
+As described in Section 1, LDA derivatives are promising methods in discovering the hidden structures in metabolomics data sets.
+For this reason, we raise a hypothesis that the degeneracies of metabolomics data sets can be detected using LDA-like models.
+Note that currently there is no set approach on how to configure LDA-like models to reduce the impact of noise in metabolomics data sets.
 
@@
 % - Providing a general-purpose machine learning model.
-\par The primary goal of this research is to expand the consensus on whether topic modelling is a valid of approach to smooth datasets.
-To be more specific, the objectives we will be looking into are listed as follows:
+\par The primary goal of this research is to expand the consensus on whether topic modelling is a valid approach to smooth metabolomics data sets.
+To be more specific, we set our objectives to the following:
 \begin{enumerate}
-    \item Adapt and utilise the generative Latent Dirichlet Allocation models in metabolomics;
-    \item Reduce the impact of noise in the datasets of metabolomics;
-    \item Give a basis to the generative Latent Dirichlet Allocation model in error correction.
+    \item Utilise general topic modelling methods in metabolomics;
+    \item Reduce the impact of noise in metabolomics data sets;
+    \item Give a basis to a general topic modelling method for error correction.
 \end{enumerate}
-The first objective is expected to investigate the perspective applications of topic modelling which are successfully utilised in other domains.
-Effectively, we want to assess the general models (or models displaying significant results in other fields).
-The second objective would adapt, and thus enhance, these general models for the particular use of error correction in metabolomics.
-Finally, the third objective can be treated as a general contribution to the field of machine learning.
+Briefly, the first objective is expected to utilise the general state-of-the-art topic modelling methodology to distinguish the techniques displaying better performance in metobolomics applications.
+In other words, we will conduct a survey on LDA-like topic modelling methodology.
+The second objective is focusing on enhancing the general models for the particular task of error correction in metabolomics.
+Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered techniques might be applicable to other domains.
 
@@
 % - The negation of the hypothesis would contribute in terms of reassessing the complexity of the problem.
-\par We are setting a hypothesis that topic modelling, or more specifically Latent Dirichlet Allocation, can be utilised to correct the errors in the datasets of metabolomics.
-Both cases of proving and disproving this hypothesis would be a contribution in metabolomics.
-The successful outcome would lead to the potential applications provided in the following paragraph, whereas the unsuccessful outcome would set-up additional requirements to tackle this issue.
+\par We are setting a hypothesis that LDA-like topic modelling can be successfully utilised to correct errors in metabolomics data sets.
+Both cases of proving and disproving this hypothesis would contribute to metabolomics.
+The successful outcome would lead to the potential applications discussed in the following paragraph, whereas the unsuccessful outcome would set-up additional requirements to tackle the issue of noisiness in metabolomics data sets.
 
@@
 % - Contributing to machine learning in general.
-\par The potential applications of this research can be induced directly induced from the set-up objectives.
-First of all, we would show that models from other domains can be applied in metabolomics and, thus, generalised.
-Further, the developed error correction methodology for the datasets of Metabolomics would improve the accuracy of the chemical processes prediction.
-Finally, the generalisation of the developed error correction methodologies would provide an alternative tool-kit for other domains as well as expand the potential applications of the results derived from metabolomics.
+\par The potential applications of this research can be directly induced from the previously listed objectives.
+First of all, we would show that the general models or the models applied in other domains can be utilised in metabolomics.
+Further, the developed error correction methodology would improve the accuracy of the chemical processes prediction.
+Finally, the generalisation of the developed error correction model might suggest different approaches in tackling error correction in general; also, it would impact the awareness in metabolomics and the field's potential in developing machine learning generalisations.
 
@@
 % - A notification the the results will be discussed in the following section.
-\par The provided literature review on latent Dirichlet allocation assesses its underlying methodology and proposes several different implementation variations.
-The review is structured in progressing order: to start with, we look into the original latent Dirichlet allocation paper and define the terminology which will be used throughout this proposal; then, we familiarise with two different approaches used for the inference of latent Dirichlet allocation;  finally, we review the perspective advances which could be used for error correction.
+\par The provided literature review on topic modelling is directed towards familiarising with LDA derivatives.
+Also, the review assesses the methodology of LDA-like models and proposes several implementation variations.
+The review is structured in progressing order: at the start, we look into the original LDA paper and define the terminology (it will be used throughout the proposal); then, we familiarise with two different techniques used for the inference of LDA-like models; finally, we review LDA derivatives which shows the potential to be utilised in designing the error correction technique.
 
@@
 % - Example
-\par Recall that in topic modelling we are trying to induce some hidden structure within the studied data.
-For instance, if we were analysing a book, it would be likely to address several different topics.
-Further, the chapters of the book itself would have unique concentrations of such topics.
-In order to cluster these topics, a probabilistic topic modelling method would look into the structure of the words used in the book.
-Also, the performance of the method would depend on its ability to optimise this clustering, for example, we could dismiss the commonly used words (such as and, or, the, etc.).
-Note that in we will assess only the topic modelling where we do not know the naming of topics.
-That is, we will be looking into a \textit{unsupervised learning} problem.
+\par Recall that in topic modelling induces a hidden structure within the studied data sets.
+We will provide an example providing an intuition in this process.
+Say, if we were analysing a book, its chapters might address different topics or address the topics at a different extent.
+Further, the chapters itself would have unique distributions of the topics.
+Note that we are assuming that we do not know what the topics addressed in the book.
+In order to derive these hidden topics, a probabilistic topic modelling method would analyse the structure of the words used in the book.
+In other words, the model would induce patterns which would refer to a particular topic.
+Regarding the method's performance, it would depend on the method's ability to optimise the inference of the patterns.
+For example, commonly used words (such as and, or, the, etc.) would be relevant in all topics.
+Therefore, their semantic impact in defining the topics would be negligible.
+For this reason, we could dismiss these to improve the time performance of the method.
 
@@
 % - A notification that the terminology and the notation will be useful throughout the proposal.
-\par The terminology used in this proposal is equivalent to the one used in the survey on probabilistic topic models carried by Blei et al.
-[citation].
-Figure \ref{fig:lda} below represents the latent Dirichlet allocation model in plate notation.
+\par The terminology used in this proposal is equivalent to the terminology used by Blei \cite{blei_2012}.
+Figure \ref{fig:lda} below represents the original LDA model in plate notation.
 We will use it as a guide to familiarise with the terminology.
@@
   \centering
-  \includegraphics[width=0.7\textwidth]{lda}
-  \caption{Latent Dirichlet allocation. COPYRIGHT. CHANGE M TO D}
+  \includegraphics[width=0.7\textwidth]{lda_model}
+  \caption{The design of the original LDA model}
   \label{fig:lda}
@@
 % High level LDA terminology
+% - Notation:
+% -- List LDA definitions
+% -- Explain the definitions
+\par To start with, we will familiarise with the higher level entities of the model.
+The circles represent an entity, whereas the plates indicate the number of entities.
+For example, there would $N \times M$ entities of $z$ and $M$ entities of $\theta$.
+Further, LDA is a three-level hierarchical structure: the smallest entity (grey circle) is defined as a \textit{word}; then a collection of words (inner plate) is defined as a \textit{document}; and a collection of documents (outer plate) is a \textit{corpus}.
+Note that Figure 1 represents $M$ documents and $N$ words per document, it follows that in the corpus there would be $M \times N$ words.
+Also, as suggested by Figure 1, let's denote a word by $w$.
+Then, let's denote the sequence of words in document $d$ by
+\begin{equation*}
+\mathbf{w}_d = \{w_1, w_2, \dots, w_N\},
+\end{equation*}
+then the sequence of words in the corpus would be denoted by
+\begin{equation*}
+\mathcal{\mathbf{D}} = \{\mathbf{w_1}, \mathbf{w_2}, \dots, \mathbf{w_M}\}.
+\end{equation*}
+
+% Low level LDA terminology
 % - Intro to observable and hidden variables
@@
 % -- Explain the definitions
-\par To start with, we will familiarise with the domain of the model in a high level.
-The grey circle represents a single word, whereas a collection of words (document) is represented by the inner plate, and the outer plate corresponds to a collection of documents (corpus).
-In the model, words are the only \textit{observable variables} -- this is indicated by the grey colour of the circle, whereas the white coloured circles indicate the \textit{hidden variables}.
-As suggested by the figure, a \textit{word} is denoted by $w$; a \textit{document} (sequence of words) is denoted by $\mathbf{w} = \{w_1, w_2, \dots, w_N\}$, where $N$ indicates the number of words in the document; and a \textit{corpus} is denoted by $\mathcal{\mathbf{D}} = \{\mathbf{w_1}, \mathbf{w_2}, \dots, \mathbf{w_M}\}$, where $D$ is the number of documents in the corpus.
+\par Now, we will introduce the lower level entities.
+The grey circle indicates that an \textit{observable variables}, whereas the white coloured circles indicate the \textit{hidden variables}.
+Note that the words are the only observable variables in the model; all other entities are within the underlying structure of the model.
+Further, recall that topic modelling attempts to discover a mixture of topics.
+To be more specific, a corpus and each document could be represented as an arbitrary mixture of topics.
 
-% Low level LDA terminology
-% - An explanation on the hidden variables
-% - Notation:
-% -- List the definitions
-% -- Explain the definitions
-\par Now, we will introduce the lower level entities -- the hidden variables.
 Recall that latent Dirichlet allocation attempts to induce an arbitrary mixture of topics.
@@
 % it is fine to change the bibliography style if you want
+\bibliography{mprop}
 \bibliographystyle{plain}
-\bibliography{mprop}
 \end{document}

### 2016-12-09 docs/MSci_proposal/mprop.tex
Fixing the section on the basic LDA model.@

@@
 \par Now, we will introduce the lower level entities.
-The grey circle indicates that an \textit{observable variables}, whereas the white coloured circles indicate the \textit{hidden variables}.
-Note that the words are the only observable variables in the model; all other entities are within the underlying structure of the model.
-Further, recall that topic modelling attempts to discover a mixture of topics.
-To be more specific, a corpus and each document could be represented as an arbitrary mixture of topics.
-
-Recall that latent Dirichlet allocation attempts to induce an arbitrary mixture of topics.
-The corpus and each document is expected to vary in the proportions of the induced topics; and each document is likely to have a different number of word.
-Let's denote the number of such arbitrary topics by $K$, and the number of unique words in the corpus by $V$.
-This notation will simplify the explanation on the remaining entities in Figure \ref{fig:lda}.
-To start with, $\beta$ is the hyper-parameter (a matrix of dimension $V \times K$) referring to the prior distribution of the topics over the vocabulary.
-The hyper-parameter $\alpha$ (a vector of length $K$) refers to the prior distribution of the topics over the documents.
-Further, the latent variable $\theta$ (a matrix of dimension $D \times K$) is the topic distribution over documents.
-Finally, $z$ is the latent variable on the topic assignments of each word in every document, effectively, it is a matrix of dimension $M \times V$.
+The grey circle indicates that the \textit{observable} variables, whereas the white coloured circles indicate the \textit{hidden} variables.
+Note that in Figure 1 the words are the only observable variables; all other entities are hidden, i.e., they describe the underlying structure of the model.
+Further, recall that topic modelling is applied in order to discover a mixture of topics; or in other words, topic modelling could describe a corpus and its documents in terms of topic distributions.
+This intuition will simplify the definitions of the remaining entities in Figure 1, these are given as follows:
+\begin{itemize}
+\item $K$ is the number of topics;
+\item $V$ is the size of the vocabulary, i.e. the number of unique words in the corpus;
+\item $\alpha$ is the parameter referring to the prior distribution of the topics over documents;
+\item $\beta$ is the parameter referring to the prior distribution of the vocabulary over topics;
+\item $\theta$ is the latent variable referring to the topic distribution over documents;
+\item $\phi$ is the latent variable referring to the vocabulary over topics
+\item $w$ is the observable variable referring to the words in documents;
+\item and $z$ is the latent variable of topic assignments over the words.
+\end{itemize}
+Further, note that the listed entities can be perceived as matrices of the following dimensions: $\alpha$ is $1 \times K$; $\beta$ is $K \times V$; $\theta$ is $M \times K$; $\phi$ is $M \times K$; $w$ is $M \times N$; and $z$ is $M \times V$.
+% To-do: check the dimensions
 
@@
 % - The derivation of the posterior
-\par The relationship between observable and hidden variables, effectively, describes the inference of the underlying topic structure.
-We will go through an arbitrary example in order to give intuition in the inference process and provide the notation of the variable instances.
-By looking into Figure \ref{fig:lda}, we can see that $\alpha_k$ (the prior weight of the $k$th topic in a document) influences $\theta_{:, k}$ (the latent $k$th topic distribution over $M$ documents).
-It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in the document $d$) influences $z_{d, :}$ (the topic assignment over $V$ vocabulary terms in document $d$).
-Note that $z_{d, v}$ would refer to the assignment of some specific vocabulary term $v$.
-Now, by taking $z_{d, v}$ and $\beta_{v, k}$ (the prior weight of the $k$th topic on the vocabulary $v$) we describe the origins of word $w_{d, v}$.
-The relationship of the terms, effectively, induces the generation of the corpus, this can be expressed as the following joint probability distribution
+\par The relationship the between observable and hidden variables, effectively, describes the inference of the underlying topic structure.
+We will go through an arbitrary example in order to give an intuition in the inference process and familiarise with the lower level notation of the previous listed variables.
+Figure \ref{fig:lda} shows see that $\alpha_k$ (the prior for the $k$th topic in a document) influences $\theta_{:, k}$ (the latent variable describing the distribution of topic $k$ over all documents).
+It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in document $d$) influences $z_{d, :}$ (the topic assignment over all vocabulary terms for document $d$).
+Further, $z_{d, v}$ refers to the topic assignment of vocabulary term $v$.
+By taking $z_{d, v}$ and $\beta_{v, k}$ (the prior of topic $k$ on vocabulary term $v$), we obtain word $w_{d, v}$.
+In the complete perspective, the generative process of the corpus is given by the following joint probability distribution
 \begin{equation}
@@
 The problem arises from inefficient analytical computation of evidence $p(w)$.
-By reviewing the literature on latent Dirichlet allocation [CITATION], we will also familiarise with the numerical methods used in approximating $p(w)$.
-% To-do: COPYRIGHT for the equation
-% To-do: reconsider the name of the subsection
+The approximation of it will be considered during the literature review on LDA-like models.
+The naming of the following subsections corresponds to the reviewed papers.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Latent Dirichlet allocation}
+\subsection{Latent Dirichlet Allocation}
 
@@
 % - Detailed approach in covering implementations details
-\par Latent Dirichlet allocation is a probabilistic topic modelling technique  proposed by Blei et al [citation].
-Since this subsection is dedicated for the origins of the technique, it will be broader in detail.
-This will cover aspects addressed in the following subsections covering the advances on latent Dirichlet Allocation.
+\par Latent Dirichlet allocation is a probabilistic topic modelling technique proposed by Blei et al.
+\cite{blei_2003}.
+Since this paper is the origin of LDA-like models, it will be discussed in more detail to introduce the basis of LDA-like methodology.
+As a result, this coverage will simplify the review on the LDA derivatives discussed in the upcoming subsections.
 
@@
 % - Topic dimensionality
-\par The initial latent Dirichlet allocation model sets some assumptions for the data which will be processed by the model.
-First of all, it is assumed that both documents and words in the documents are \textit{exchangeable}.
-That is, the order in which these will be processed does not matter.
-More formally, this assumption means that the model follows the \textit{bag-of-words} principle.
-The next assumption is that the model works on discrete data.
-Effectively, the original paper does not cover the treatment of continuous features.
-Another assumption is on the dimensionality of the topic variables: the number of topics is set to be fixed during the complete run of inference.
+\par The original LDA model sets some assumptions for the data processed by the model.
+First of all, it is assumed that documents and words are \textit{exchangeable}.
+That is, the order in which these entities are processed does not matter.
+In other words, we can say the model follows the \textit{bag-of-words} principle.
+The next assumption states that we should use discrete data.
+Effectively, the basic LDA does not cover  continuous features.
+Another assumption is on the number of topics: it is set to be fixed throughout the complete run of inference.
 
@@
 % - The generative algorithm
-\par As discussed in the preliminaries subsection, the process of latent Dirichlet allocation assumes that the corpus can be generated by a probabilistic process.
-The generation of the corpus is made of three levels: the parameters $\alpha$ and $\beta$ are sampled once; the variable $\theta$ is sampled once per document; and the variables $z$ and $w$ are sampled once per word.
-The generative algorithm given in the original paper is equivalent to Algorithm 1 below.
+\par As discussed in the preliminaries subsection, we are also taking an assumption that the corpus is induced by a probabilistic process.
+The generation of the corpus has three levels: the parameters $\alpha$ and $\beta$ are sampled once; the variable $\theta$ is sampled once per document; and the variables $z$ and $w$ are sampled once per word.
+This hierarchical process relates to the generative algorithm given in the original LDA paper, it is equivalent to Algorithm 1 below.
 \begin{algorithm}[H]
-\caption{Document Generation.}
+\caption{Document Generation}
 \label{alg:document_generation}
@@
 Note that the algorithm describes the generation of a single document.
-In order to generate the corpus, we would run the algorithm several times.
-To start with, in the line 1 we draw the number of words in the document $N$, where $\xi$ is an ancillary variable suggesting the mean for the Poisson Distribution.
-In the line 2 we draw the topic distribution $\theta$ from the Dirichlet distribution on $\alpha$.
-Now in the lines 3--6 we generate the words: in the line 4 we draw topic $z_n$ from the multinomial distribution on $\theta$; and, finally, in the line 5 we obtain the word $w_n$ from the conditional probability on $z_n$ and the distribution over the words in the vocabulary $\beta$.
+In order to generate the whole corpus, we would run the algorithm for each document.
+Going back to Algorithm 1, a brief intuition behind it is given as follows: in line 1 we draw the number of words $N$ in a document ($\xi$ is an ancillary variable used as the mean for the Poisson Distribution); in line 2 we draw topic distribution $\theta$ from the Dirichlet distribution on parameter $\alpha$; further, in lines 3--6 we generate the words: in line 4 we draw topic $z_n$ from the multinomial distribution on $\theta$ and, finally, in line 5 we obtain  word $w_n$ from the conditional probability on $z_n$ and $\beta$, recall that $\beta$ is the prior on the vocabulary terms correspondence to topics.
 
@@
 % - Variational parameters gamma and fi
-\par The inference is, effectively, the computation of the evidence for the posterior distribution.
-The original latent Dirichlet allocation paper uses the method of variational inference.
-Basically, the authors introduce the variational parameters $\gamma$ and $\phi$ (computation of these set-ups a problem of optimisation).
-Note that the original paper provides pseudo code for solving the optimisation problem.
-% To-do: connect with the parameter estimation paragraph
-% To-do: reforming the posterior formula
+\par As suggested in the ending paragraph of the preliminaries subsection, the inference relates to the computation of evidence $p(w)$ (it is the denominator term given in Equation 2).
+Even though the original paper mentions the inference methods such as Markov chain Monte Carlo, it suggests using the method of variational inference.
+In order to develop an intuition over the method, the authors introduce the variational parameters $\gamma$ and $\phi$.
+It is also mentioned that the computation of $\gamma$ and $\phi$ is an optimisation problem.
+That is, we are assessing the minimum bound over the latter parameters.
+Further, recall that in Figure 1 we explicitly express the parameter $\phi$ in describing the basic LDA model.
+However, the authors have chosen to show it separately -- upon describing the variational method.
+Their expression is given in Figure \ref{fig:var_par} below.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.4\textwidth]{variational_parameters}
+  \caption{Variational parameters}
+  \label{fig:var_par}
+\end{figure}
+Note that the authors provide pseudo code to build an intuition in tackling this optimisation problem.
 
@@
 % -- The incoorperation of documents which were not used for the training
-\par Additionally, the authors emphasise the estimations of the parameters $\alpha$ and $\beta$.
-They suggest a two-step EM procedure utilising the variational parameters $\gamma$ and $\phi$: E-step is the optimisation problem addressed in the previous paragraph; and M-step updates the parameters $\alpha$ and $\beta$.
-Note that the authors provide an analytical update for $\beta$ and a Newton--Raphson method for updating $\alpha$.
-Apart from that, the authors also introduce the issue of a sparse corpus.
-Such corpus would be expected upon having a large vocabulary or a significant amount of documents.
-The authors suggest addressing this problem by introducing smoothing.
-They introduce an additional parameter $\eta$ which impacts the estimation of $\beta$.
+\par Additionally, the authors emphasise the estimation of the parameters $\alpha$ and $\beta$.
+They use a two-step EM procedure which involves the previously introduced variational parameters $\gamma$ and $\phi$.
+E-step is the solution of the variational parameters optimisation problem; and M-step is an update of the parameters $\alpha$ and $\beta$.
+Note that the authors provide an analytical approach to estimate $\beta$ which is proportional to the changes of $\phi$; and for $\alpha$, they provide an approach utilising the Newton--Raphson method.
+Apart from that, the authors introduce the an issue of \textit{sparsity} in a corpus structure.
+A sparse corpus is expected upon using a large vocabulary or a large amount of documents.
+This issue is addressed by introducing smoothing: an additional parameter $\eta$ is utilised to smooth the estimation of $\beta$.
 
@@
 % -- Comparison with SVM
-\par The conducted experiments display the improvement performance over the predecessor models and suggest the application domain of latent Dirichlet allocation.
+\par The conducted experiments display an improvement in performance over the predecessor models and suggest the application domains of LDA-like models.
 The first experiment addresses document modelling.
-The latent Dirichlet allocation model is reported to display balanced hidden topic proportions in the set of test documents.
-That is, the test set has replicated the proportions of the training set.
-The second experiment addresses document classification.
-In order to display better performance compared to a support vector machine (SVM) model, the authors address the dimensionality reduction of the tested documents.
-This means that the documents would possess a lower amount of features.
-It follows that the latent Dirichlet allocation model has successfully reduced the number of features and displayed an improvement in accuracy compared to the support vector machine model.
+LDA  has displayed balanced hidden topic proportions in the set of tested documents.
+That is, the tested documents has replicated the topic proportions of the training documents.
+The second experiment is directed towards the problem of document classification.
+They have addressed the \textit{dimensionality reduction} in order to surpass the performance of the support-vector-machine-like models.
+To be more specific, the documents with reduced dimensionality would possess a lower amount of features.
+The results of the experiment indicate that the basic LDA model has successfully reduced the number of features.
+Also, it has displayed an improvement in accuracy compared to the support-vector machine-like model.
 % To-do: explain perplexity (UNTOUCHED)
@@
 % - Summary of the results established by the paper
-\par The review on the initial latent Dirichlet allocation model has set-up a basis for introducing the successor variations.
-These will be provided in the following sections.
-We have introduced the initial methodology for the following reasons: (1) the novel models will display the relaxation of the initial assumptions; (2) the provided document generation methodology will be compared to the document generation in batches; (3) we will cover different inference and parameter estimation approaches.
-Also, note that the paper is relatively old in time, and the experiments do not display the state-of-the-art performance.
-Rather than that, they were introduced to address the application domains of the model.
-% To-do: rework the paragraph to reflect its purpose: remove the discussion
+\par The review on the original LDA model has set up a basis required to understand the model's derivatives.
+That is, the introduced methodology will be useful in considering the following improvements: (1) the  relaxation of the exchangeability assumption; (2) the adaptation of time-series suggesting word correspondence to topic over time; (3) the discussion over alternative inference methods.
+Apart from that, the experiments introduced in this section has reflected the LDA performance upon the model's introduction to the public.
+The current state-of-the-hard topic modelling techniques overcome the shown results.
+The experiments were addressed in order to suggest the application fields of LDA-like models.
+% To-do: rework the paragraph name to reflect its purpose: remove the discussion (you might think it's done, but it's not)
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Finding scientific topics}
+\subsection{Finding Scientific Topics}
 

### 2016-12-10 docs/MSci_proposal/mprop.tex
Final review on the literature review.

@@
 \end{itemize}
-Further, note that the listed entities can be perceived as matrices of the following dimensions: $\alpha$ is $1 \times K$; $\beta$ is $K \times V$; $\theta$ is $M \times K$; $\phi$ is $M \times K$; $w$ is $M \times N$; and $z$ is $M \times V$.
+Further, note that the listed entities can be perceived as matrices of the following dimensions: $\alpha$ is $1 \times K$; $\beta$ is $K \times V$; $\theta$ is $M \times K$; $\phi$ is $K \times V$; $w$ is $M \times N$; and $z$ is $M \times V$.
 % To-do: check the dimensions
@@
 Also, it has displayed an improvement in accuracy compared to the support-vector machine-like model.
-% To-do: explain perplexity (UNTOUCHED)
 
@@
 % - Sampling as a method of inference
-\par The paper on "Finding scientific topics" [CITATION] suggests an alternative approach to the inference in latent Dirichlet allocation.
-To be more specific, the paper provides a readable introduction on how to apply a Markov chain Monte Carlo (MCMC) algorithm as a sampling-based method for the inference in topic modelling.
-Further, the authors discuss the method's application settings in vision and conduct an experiment on deriving topics from the abstracts of papers published in "Proceedings of the National Academy of Sciences" (PNAS).
-Finally, the authors compare the performance of the suggested sampling-based inference method to variational inference methods.
+\par The article titled as `Finding Scientific Topics' \cite{griffiths} suggests an alternative approach to the inference which is suitable for LDA-like models.
+To be more specific, the article provides a readable introduction on the application of the Markov chain Monte Carlo (MCMC) algorithm.
+That is, the MCMC algorithm is applied as a sampling-based method to infer an underlying topic structure.
+Further, the authors discuss the method's application settings for the machine learning problems in vision.
+Also, the authors conduct an experiment on inferring the structure of topics in the abstracts of papers published in `Proceedings of the National Academy of Sciences' (PNAS).
+Finally, the paper is concluded by a discussion on assessing the sampling-based method usability.
+To be more specific, the performance of the sampled-based method is compared to the performance of variational inference methods.
 
@@
 % - Introduce the inference method
-\par The inference method using a Markov chain Monte Carlo algorithm can be referred by \textit{Gibbs sampling}.
-The authors claim that the method provides a first-order approximation in deriving the topics on the given corpus.
-That is, we can establish quantitative reasoning on the document correlation in terms of content and content's change over time.
-Recall that by $\theta$ we denote the topic distribution over documents and by $\phi$ we denote the topic distribution over words.
-In Gibbs sampling, these values can be obtained by examining the posterior distribution.
-Essentially, the variables are sequentially sampled from their respective distributions.
-The process draws topic assignment $z_i$ and keeps a track of this assignment with respect to each word and document.
-Note that $z_i$ is drawn from the following distribution:
+\par The MCMC-based inference method is often referred by \textit{Gibbs sampling}.
+The authors claim that the method provides a first-order approximation upon inferring the hidden structure of the data.
+That is, the performance is sufficient to establish quantitative reasoning in assessing the correlating sections and semantic structure of the data.
+Recall that by $\theta$ we denote the topic distribution over the documents and by $\phi$ we denote the vocabulary terms distribution over the topics.
+In Gibbs sampling, the latter variables are obtained by examining the posterior distribution, whereas topic assignments $z$ are sampled sequentially.
+That is, we sample topic assignment $z$ and update the distribution depending on the obtained value (this will impact the next sample of $z$); we repeat this process until the topic distribution converges.
+To be more specific, each drawn $z_i$ ($i$ denotes the iteration) is recorded as a count.
+That is, we keep a track of topic assignments for particular words and documents.
+The value of drawn topic $z_i$ is proportional to the distribution given below:
 \begin{equation}
-P(z_i = j | z_{-i}, w) \propto \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + W\beta}\dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + T\alpha}.
+P(z_i = j | z_{-i}, w) \propto \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + V\beta}\dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + K\alpha}.
 \end{equation}
-It follows that this distribution is utilise the previously mentioned topic assignments.
-For a brief familiarisation with the notation used in the previous expression, note the following: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$; $W$ is the number of words; and $T$ is the number of topics.
+For a brief familiarisation with the notation used in the expression, note the following: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$; recall that $V$ is the number of words in the vocabulary and $K$ is the number of topics.
+Also, note that the counts do not include the last assignment of the topic, this is indicated by $-i$.
 Ultimately, the left side of the expression is $\phi$ and the right side is $\theta$.
 That is,
-\begin{align*}
-\phi_j &= \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)}},\\
-\theta_j &= \dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + T\alpha}.
-\end{align*}
-This process of topic assignment would be executed for every word in a document and for all documents in the corpus.
-The procedure would be repeated until the distributions displayed fluctuations below some set threshold.
-Notice that the procedure does not require direct assignment of $\theta$ and $\phi$.
-If needed, the values of these can be sampled upon converge by using the previously provided representations.
+\begin{align}
+\phi_j &= \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + W\beta},\\
+\theta_j &= \dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + K\alpha}.
+\end{align}
+This process of the topic assignment is executed for every word in the document and for all documents in the corpus.
+The procedure is repeated until convergence or, more pragmatically, until the distribution displays the assigned topic fluctuations below the set threshold.
+In the Gibbs sampling procedure, there is no need to explicitly assign $\theta$ and $\phi$ values.
+If needed, these values are obtained directly from the topic distribution, as shown by the two previous equations.
 
-% Auxilliary xperiments
+% Auxilliary experiments
 % - Vision
 % - Time series
-\par The authors have executed Gibbs sampling in the domain of vision.
-They have represented an image as a document and each pixel as a word.
-This means that they were able to recover the underlying basis from which the images could be constructed.
-The performance was compared to the latent Dirichlet allocations models which use variational inference.
-Note that we define the performance in terms of perplexity, which indicates the uncertainty of the word assignment to a topic.
-It appears that, with relatively small data sets, Gibbs sampling displays superior performance.
-Further, the authors have investigated the performance upon varying the number of topics.
-The results indicate that there exists a trade-off between the model being able to cover the significant underlying topics (low number of topics) and start deriving irrelevant branches (high number of topics).
+\par The authors has shown a use of Gibbs sampling on a problem of vision.
+In their experiment settings, it is assumed that a collection of images represents a corpus; an image represent a document; and a pixel of an image is a word.
+The experiment has shown that the model is able to infer the underlying basis (topics) from which the images could be constructed.
+The performance of the model was compared to the variational inference based LDA models.
+Note that the performance was assessed in terms of \textit{perplexity} -- the uncertainty in assigning a word assignment to a topic.
+It has been shown that Gibbs sampling displays superior performance in the small data sets, whereas the performance in larger data tends to fall off.
+Furthermore, the authors have assessed the performance upon varying the number of topics.
+The results of this experiment indicate a trade-off in deriving high-level and low-level structures.
+That is, the low number of topics might result in mashing several structures, whereas the high number of topics is likely to infer irrelevant structures.
 
@@
 % - PNAS
-\par The main experiment was carried in the PNAS data set, note that the processed papers were published from 1991 to 2001.
-By having a varying notion of time, the authors were able to investigate the prevailing topics.
-Thus, establish a sense in dynamic topic modelling.
-Even though the model was not adjusted to smooth the time-series, from the derived results we can deduce the varying relevance of underlying science domains.
+\par The main experiment of the article is conducted on the PNAS data set containing abstracts of the articles published from 1991 to 2001.
+The authors have noticed that the topic distribution over the vocabulary terms change over time.
+However, the model had not yet been optimised to take this notion of time-series into account upon the inference.
+Nevertheless, the authors have emphasised the necessity to assess the dynamic topic modelling.
 
 % Conclusion
-\par We have reviewed an alternative method for the inference in latent Dirichlet allocation.
-The idea behind the inference process was explained by considering the primary aspects: the notion of counts and the implicit use of hidden variables.
-Also, we have introduced the experiment settings its results.
-This allowed to familiarise with an example of topic modelling application and vision.
-Further, the experiment on PNAS has emphasised the relevance of time series
+\par The review has introduced an alternative method used in the process of inference of LDA-like models.
+We have familiarised with the intuition in the implementation requirements of the sampling-based inference method.
+Also, we have reviewed experiments suggesting an approach in tackling problems of vision.
+Further, we have introduced the relevance of dynamic topic modelling (topic modelling with the notion of time-series).
 
@@
 % - Qualitative and quantitative improvement
-\par Now we review the paper on Dynamic Topic Models [CITATION].
-Effectively, the dynamic topic model can be treated as an extension of static topic models (such as latent Dirichlet allocation).
-The main idea behind a dynamic topic model is that it induces the topic evolution the processed corpus.
-That is, the documents in the corpus evolve over time.
-In this subsection, we review the key assumptions of dynamic topic models, go over the generative process of documents, review the suggested inference methods, and familiarise with the results of the carried experiment.
+\par At this point, we have set up the preliminaries to cover the article on utilising the notion of time-series in LDA-like models.
+We review the initial article on utilising time-series -- `Dynamic Topic Models' \cite{blei_2006}.
+Note that by a \textit{static} model we mean that the assumed generative process does not take the development/change of the vocabulary terms into account, whereas in \textit{dynamic} topic modelling we assume the notion of that the documents were generated in a sequential process.
+That is, in dynamic topic modelling, the topics evolve in terms of their distribution over the vocabulary terms.
+Upon reviewing the dynamic topic modelling article, we emphasise the following aspects: the fundamental dynamic topic modelling assumptions and their differences compared to the basic LDA model; the generative process; the suggested inference methods; and the results displayed by the experiments on the initial dynamic topic model.
+% To-do: qualitative and quantitative improvements
 
@@
 % - Data is divided in time slices
-\par We will emphasise the assumptions given on the dynamic topic models.
-First of all, even though we are discussing a model of time series suggesting the use of continuous variables, the data remains categorical.
-The distinction between continuous and categorical data is that continuous data is strictly numeric and infinite, whereas categorical data can take only particular values.
-The second assumption relaxes the ex-changeability on documents.
-That is, the documents are expected to form a sequence.
-Or in other words, the corpus is divided into time-slices.
-Note that in such time-slice documents are exchangeable.
+\par The assumptions set for the dynamic topic model emphasises its strengths over the basic LDA model.
+Most importantly, the dynamic topic model relaxes the assumption of the document exchangeability.
+Since the documents are generated over time, we assign them to specific time-slices.
+The combination of these time-slices would represent the corpus.
+Note that the documents in each time-slice are exchangeable.
+This treatment of time set-ups a basis to tackle the problems which involve the sequential development of the data.
+On the other hand, note that another assumption of discrete data remains the same as in the basic LDA model.
+Even though we are inducing time series (which would suggest the use of \textit{continuous} variables), the authors use only \textit{categorical} (discrete) data.
+The distinction between continuous and categorical data is the following: continuous data is strictly numeric and infinite, whereas categorical data takes a fixed value from a finite set of values.
 
@@
 % - An explanation of the algorithm
-\par The generative process of the models updates the parameters $\alpha$ and $\beta$.
-This means that the topic distributions over documents and words change over time.
-The visualisation of the process is given in Figure 2 below.
+\par In the dynamic topic modelling, the generative process updates the parameters $\alpha$ and $\beta$ for each new time-slice.
+This means that the topic distribution over documents and the vocabulary terms distribution over topics change over time.
+The visualisation of this process is given in Figure 3 below.
 \begin{figure}[H]
@@
 \end{figure}
-Notice that each time-slice is equivalent to the model shown in Figure 1.
-Also, the parameters $\alpha$ and $\beta$ are induced by their predecessors in the previous time-slice.
-More formal representation of the generative process is given in Algorithm 2.
+Effectively, Figure 3 is a sequential combination of the basic LDA model illustrated in Figure 1.
+In Figure 3, one basic LDA model represents one time-slice.
+Also, the parameters $\alpha$ and $\beta$ are induced by their predecessors.
+More formal representation of the generative process is given in Algorithm 2 below.
 \begin{algorithm}[H]
@@
 \begin{algorithmic}[2]
-\State $\beta_t | \beta_{t-1} \sim \mathcal{N}(\beta_{t-1}, \sigma^2)$
-\State $\alpha_t | \alpha_{t-1} \sim \mathcal{N}(\alpha_{t-1}, \delta^2)$
-\For {$d \leftarrow 1, D$}
-\State $\eta_d \sim \mathcal{N}(\alpha_{t}, a^2)$
+\State $\beta_t | \beta_{t-1} \sim \mathcal{N}(\beta_{t-1}, \sigma^2I)$
+\State $\alpha_t | \alpha_{t-1} \sim \mathcal{N}(\alpha_{t-1}, \delta^2I)$
+\For {$d \leftarrow 1, M$}
+\State $\eta_d \sim \mathcal{N}(\alpha_{t}, a^2I)$
 \For {$n \leftarrow 1, N$}
@@
 \EndFor
-% \EndProcedure
 \end{algorithmic}
 \end{algorithm}
-The given algorithm describes the generative process for the time-scale $t$.
-It starts by drawing the parameters $\alpha_t$ and $\beta_t$ from the Gaussian distribution with the means given by the previous parameter values.
-For now, we will not discuss the meaning of the variance values.
-Further, for each document in the time-scale, we carry the following procedure: we draw the topic distribution in a document $\eta_d$ from the Gaussian distribution with the mean given by $\alpha_t$; then, for each word in the document, we draw the topic $z_{d, n}$ from the Multinomial distribution on the parametrised $\eta_d$, and, finally, we draw the word $w_{t, d, n}$ from the Multinomial distribution on the parametrised $\beta_t$ which is conditioned on $z_{d, n}$.
-By the use of \textit{parametrisation}, we are expressing the result of the Gaussian distributions into a format suitable for Multinomial distribution.
-This mapping is expressed as
+
+Algorithm 2 describe the generative process of the documents in time-slice $t$.
+The algorithm is initialised by drawing the parameters $\alpha_t$ and $\beta_t$ from the Gaussian distribution with the means given by the previous parameter values.
+Note that the parameters $\alpha_0$ and $\beta_0$ are the the special case.
+They can be initialised randomly, i.e.
+the mean of the Gaussian distribution is random.
+However, for the parameters $\alpha_0$ and $\beta_0$ we take larger variance values compared to $\alpha_t$ and $\beta_t$.
+The larger variance values  induce the initial time-slice with larger fluctuations in the topic and vocabulary term distributions; and the smaller variance values in the following time-slices induce the smoothness.
+Further, in lines 3--9, for every document in time-slice $t$ we execute the following procedure: for document $d$, we draw the topic distribution $\eta_d$ (the meaning of the parameter is equivalent to the previously introduced $\theta_d$); then, in lines 5--7, for every word in document $d$ we draw topic assignment $z_{d, n}$ and, finally, we draw word $w_{t, d, n}$.
+Note that the means for the Multinomial distributions are normalised by the use of \textit{parametrisation}.
+Parametrisation mapping $\pi$ is expressed as
 \begin{equation}
-\pi(x)_n = \dfrac{\exp{(x_n)}}{\sum^n_{i=1}\exp{(x_i)}}.
+\pi(x)_n = \dfrac{\exp{(x_n)}}{\sum^N_{i=1}\exp{(x_i)}}.
 \end{equation}
@@
 % Two variations: Kalman filtering and wavelet regression
-\par The authors suggest using variational methods to approximate the inference posterior.
-It is claimed that stochastic simulation would be unable to scale with large data sets.
-Two techniques of variational methods are provided: Kalman Filtering and Wavelet Regression.
-The implementation details of these are provided in a brief manner suggesting the computation of the parameter $\beta_t$.
+\par The authors have used variational methods to approximate the inference of the posterior.
+It is claimed that stochastic simulation (sampling-based variance methods) would be unable to scale with larger data sets.
+Further, the authors discuss two methods used for the variational inference: Kalman Filtering and Wavelet Regression.
+The implementation details of these methods can be found in the `Dynamic Topic Modelling' article \cite{blei_2006}.
+Note that the implementation details are provided in a brief manner suggesting the use of additional literature, these are also provided in the article.
 
@@
 % - Evolution of Science
-\par The dynamic topic model is evaluated by conducting an experiment on the journals of \textit{Science}.
-The experiment is expected to establish the time series on the topic development with the articles.
-Indeed, the experiment is able to produce results displaying the rises and falls of the topics indicating scientific fields.
-Note that the experiment is performed using both variational inference methods.
+\par The dynamic topic model is evaluated by conducting an experiment on the journals of `Science'.
+The experiment is expected to infer the prevalent science fields and the prevalent vocabulary used in these fields.
+Note that the relevance of the science fields can be respectively obtained from the latent variables $\theta$ and $\phi$.
+That is, the authors would compare these variables throughout all time-slices in order to induce the reasoning over the prevalent science fields (by using $\theta$) and the prevalent vocabulary terms (by using $\phi$).
+The experiment has successfully established this notion of time-series.
+This is proved by comparing the inferred rises and falls of the scientific fields to the history of science.
+Note that the experiment has been performed both Kalman Filtering and Wavelet Regression methods.
 
 % Conclusion
-\par The review on the dynamic topic models has familiarised us with more realistic generative process: the documents are induced by time series.
-Further, we have provided the visualisation of the model and introduced the algorithm for the generative process.
-Also, we have captured the techniques used for inference.
-Finally, we have familiarised with the experiment settings.
+\par By reviewing the original dynamic topic model, we have familiarised with the generative process which takes time-series into account.
+We have provided the visualisation of the model and introduced the algorithm for the generative process.
+Further, we have outlined the inference methods suggests be the authors.
+Finally, we have familiarised with the experiment settings which were used to infer time series in the articles of `Science'.
 

### 2016-12-11 docs/MSci_proposal/mprop.tex
The initial write-up of the dicussion over the literature.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Probabilistic Topic Modelling to Reduce the Noise in Metabolomics}
+\title{Probabilistic Topic Modelling for Spatial Smoothing in Mass Spectrometry Imaging}
 \author{Arijus Pleska}
@@
 % Metabolomics
+% Mass Spectrometry
 % Vision
 \section{Introduction}
+
 % To-do:
-% - Mention Rogers' paper for the first attempt to use LDA in Metabolomics datasets
+% - Ionisation
+% - Mass Spectrum
+
+% Notes:
+% -- Mass spectrum measures the masses of the ions entering mass spectrometer. For this reason, the sample has to be ionised beforehand.
+% -- The ionisation process takes the molecules of a metabolite and a adds to something like a `proton'. Then, the resulting measure of mass would be the metabolite plus the proton.
+
+% Structure:
+% 1. General statement of the proposal (novel techniques for the analysis in Mass Spectrometry Imaging (MSI))
+% 2. Mass Spectrometry Imaging. What is it used for?
+% 3. Challenges in the analysis of MSI data.
+% 4. Your contributions to the domain.
+% 5. Report outline.
 
@@
 % - ML in biology
-% - Data patterns
+% - MSI
+% - Spatial Smoothing
+\par This proposal is focused on applying machine learning techniques to improve the analysis of biology data sets.
+We intend to enhance the quality of \textit{mass spectrometry imaging} (MSI) data.
+Briefly, by the use of MSI, the researchers are able to interpret the formation of chemical processes.
+However, as suggested by a recent overview by Smith et al.
+\cite{smith_2004}, the pre-processing of molecular-level entity data sets is lagging in terms of the current noise reduction algorithm effectiveness.
+We will attempt to tackle this issue of noise by utilising spatial smoothing.
+
+% MSI
+% - Connection to metabolomics
+% -- Metabolomics definition
+% - Vision
+% - Chemical Process occurrence
+
+\par In this project, the application domain of MSI is set to be \textit{metabolomics} -- it is the scientific study of chemical processes which involve metabolites (small molecules) and metabolomes (sets of small molecules).
+In the settings of metabolomics, we utilise \text{mass spectrometry} (MS) to obtain the distribution of chemical entities within a sample.
+By utilising MSI, we can visualise the distributions of the ions (or ion sets) of different masses.
+For instance, Figure \ref{fig:b_and_e} below represents the whole sample as an image, where each picture's pixel is a small region within the sample.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{b_and_e}
+  \caption{MSI application on a sample with the letters b and e.}
+  \label{fig:b_and_e}
+\end{figure}
+Note that each pixel contains a distribution of chemical entities as well.
+In Figure \ref{fig:b_and_e} we have chosen the range of mass to reflect the ink used to inscribe the letters.
+That is, in each pixel we have summed the intensities of the chemical entities meeting the mass criteria.
+Then, each pixel has been assigned an intensity value.
+Therefore, the sample could be expressed as the latter image.
+An open problem in metabolomics is to find the sets of chemical entities  inducing valid visual patterns.
+Ultimately, the discovered biological patterns could be used to suggest and measure the likelihood of particular chemical process occurrence.
+For a rigorous discussion on machine learning applications in metabolomics, the reader can consult the recent survey conducted by Alonso et al.
+\cite{alonso-et-al}.
+
+% -- The ionisation process might be executed in different means. This results in a different sum of masses.
+% -- Metabolite might be fragmentised during the ionisation.
+% -- Atoms of the same thing might have a different mass. This would result in the different ion mass.
+
+% Challenges
+% - Explanation of the data capturing process
+% - Ionisation
+% -- Fragmentation
+% -- Different means to ionise the metabolite
+% -- Atoms might have a different mass
 % - Noise
-\par This research proposal suggests an application of machine learning in biology.
-We intend to study the scans of low level biological entities in order to derive their patterns.
-To be more specific, we will be tackling a problem of noise.
-This issue of noisiness arise from the hardware limitation of the high precision data collection devices.
+% -- Spatial Smoothing
 
-% Metabolomics
-% - Mass spectrometry
-\par The application domain of this project is metabolomics -- a branch of computional biology.
-In brief, metabolomics is the scientific study of chemical processes which involve small molecules -- metamobilites.
-Ultimately, these chemical processes can be predicted by the features of metabolites.
-We intend to set the basis of our study on data sets acquired by \textit{mass spectrometry} (MS).
-By using this technique, we obtain the masses and intensities of metabolites within a sample of interest.
-This data representation is known as mass spectrum.
-Note that a different mass indicates different types of metabolites, whereas intensity suggests the level of concentration.
-Thereby, a mass spectrum is used to distinguish and visualise the observed metabolites.
-By deriving the formulae of specific metabolites types, we can induce patterns leading to the prediction of chemical processes.
+\par The challenges on working with MSI data arise from the data capturing process.
+In metabolomics, the MS equipment requires the ionisation of metabolites.
+Effectively, a metabolite is compounded with an appropriate chemical entity to produce an ion.
+It follows that the mass spectrometers are capable to capture some properties of the ionised metabolites.
+Since the mass of an ion belongs to these properties, the collection of captured ions can be represented in terms of mass spectra.
+However, the following issues have to be considered upon reasoning from the MSI data: (1) the ionisation of a metabolite can be performed using different chemical entities, therefore the ions of same metabolite type would have different masses; (2) a metabolite might be fragmented during the ionisation; (3) the metabolites of same type might have different atom structure, which suggests a different mass value.
+The latter issues, effectively, corresponds to a problem of error correction.
 
@@
 % - Visualisation of particular metabolites or metabolite patterns.
-\par In the terms of machine learning applicability, we are looking into a problem of vision.
+
+% \par In the terms of machine learning applicability, we are looking into a problem of vision.
 The metabolites of some biological tissue are sequentially captured within regions of fixed size.
@@
 Thereby, we can choose a specific metabolite type and produce pixels reflecting such metabolite activity.
-For a rigorous discussion on machine learning applications in metabolomics, the reader can consult the recent survey conducted by Alonso et al.
-\cite{Alonso-et-al}.
 Going back to the visualisation of metabolites, we can choose a specific combination (structure) and produce an image where activated pixels would suggest the concentration of the chosen metabolite structure.
@@
 % -- Metabolomics
-\par The objective to discover novel structures is treated as a problem of \textit{unsupervised} machine learning.
-That is, we have no initial settings in how these structures are expected to look like: the structures will be inferred by machine learning methodology.
-The inference of unknown structures is studied within \textit{topic modelling}.
-Further, note that we are concentrating on statistical topic modelling.
-In other words, we are putting the emphasis on the \textit{probabilistic} models: the structures are inferred from the data instance distributions.
-To be more specific, we are focusing to enhance a particular probabilistic machine learning model -- \textit{latent Dirichlet allocation} (LDA).
-First of all, the successor LDA models are treated as a state-of-the-art method in tackling semantic-analysis type problems.
-For more detailed review on the LDA applications, the reader can consult the survey \cite{blei_2012} conducted by Blei.
-Further, Hooft et al.
-\cite{hooft} has recently addressed the applicability of LDA in metabolomics.
-Note that this proposal is directly influenced by the latter prospect: we will attempt to utilise the LDA-like models to reduce the impact of noise in the data sets of the metabolomics domain.
+\par Since this proposal is focused on utilising spatial smoothing to reduce the impact of noise in MSI data sets, we will familiarise with the machine learning techniques and problems tackled in the MSI domain.
+The objective to discover novel patterns can be treated as a problem of \textit{unsupervised} machine learning.
+That is, we have no initial settings in how the patterns are expected to look like.
+However, upon appropriate optimisation, the unsupervised machine learning techniques are capable to learn the patterns on their own.
+To be more specific, we will utilise the techniques of \textit{topic modelling} (in this case, a pattern is referred by a topic).
+Also, we are focusing on the \textit{probabilistic} (statistical) models, more specifically, \textit{latent Dirichlet allocation} (LDA).
+First of all, the successor LDA-like models are treated as a state-of-the-art method in tackling semantic-analysis type problems.
+For more detailed review on the LDA applications, the reader can consult the survey conducted by Blei \cite{blei_2012}.
+Also, Hooft et al.
+\cite{hooft} has recently addressed the applicability of LDA in MSI data sets.
+This proposal is directly influenced by the latter study: we will utilise the LDA-like models to reduce the impact of noise in the MSI data sets.
 
@@
 % - Sections
-\par The structure of this proposal is given as follows: in Section 2 we present the issue of noise in Metabolomics; in section 3 we cover the literature in topic modeling which focuses on the variations of Latent Dirichlet Allocation; in Section 4 we provide the metric of success of the proposed research; finally, in Section 5, we go into details on how the project will be executed.
+\par The structure of this proposal is given as follows: in Section 2 we present the issue of noise in MSI data sets; in section 3 we cover the literature in topic modelling which focuses on the variations of LDA-like models; in Section 4 we provide the metric of success of the proposed research; finally, in Section 5, we go into details on how the project will be executed.
 % To-do: rewrite the outline
@@
 As discussed by Palmer et al.
-\cite{palmer}, even though the current machinery differentiates metabolites in millidalton precision, we still obtain a significant amount of false positives.
-Thereby, rather than relying upon the advances of MS equipment, we can investigate the prospect of enhancing the data processing techniques.
-As described in Section 1, LDA derivatives are promising methods in discovering the hidden structures in metabolomics data sets.
-For this reason, we raise a hypothesis that the degeneracies of metabolomics data sets can be detected using LDA-like models.
-Note that currently there is no set approach on how to configure LDA-like models to reduce the impact of noise in metabolomics data sets.
+\cite{palmer},  the current hardware limit in differentiating between the ions in millidalton precision does not mitigate the issue of the false positive data.
+Thereby, rather than relying upon the advances in MS equipment, we can investigate the prospect of enhancing the data processing techniques.
+As described in Section 1, LDA derivatives are promising methods in discovering the hidden patterns in MSI data sets.
+For this reason, we raise a hypothesis that the degeneracies of MSI data sets can be detected using LDA-like models.
 
@@
 % - Providing a general-purpose machine learning model.
-\par The primary goal of this research is to expand the consensus on whether topic modelling is a valid approach to smooth metabolomics data sets.
+\par The primary goal of this research is to expand the consensus on whether topic modelling is a valid approach to apply spatial smoothing in MSI data sets.
 To be more specific, we set our objectives to the following:
 \begin{enumerate}
-    \item Utilise general topic modelling methods in metabolomics;
-    \item Reduce the impact of noise in metabolomics data sets;
+    \item Optimise the general topic modelling methods for MSI data sets;
+    \item Successfully apply spatial smoothing in MSI data sets;
     \item Give a basis to a general topic modelling method for error correction.
 \end{enumerate}
-Briefly, the first objective is expected to utilise the general state-of-the-art topic modelling methodology to distinguish the techniques displaying better performance in metobolomics applications.
-In other words, we will conduct a survey on LDA-like topic modelling methodology.
-The second objective is focusing on enhancing the general models for the particular task of error correction in metabolomics.
-Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered techniques might be applicable to other domains.
+Briefly, the first objective is expected to utilise the general state-of-the-art topic modelling methodology to distinguish the techniques displaying better performance in MSI applications -- we will conduct a survey on LDA-like topic modelling methodology.
+The second objective focuses on optimising those general models for the particular task of spatial smoothing in MSI data sets.
+Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered spatial smoothing techniques might be applicable to other domains.
 
@@
 % - The negation of the hypothesis would contribute in terms of reassessing the complexity of the problem.
-\par We are setting a hypothesis that LDA-like topic modelling can be successfully utilised to correct errors in metabolomics data sets.
-Both cases of proving and disproving this hypothesis would contribute to metabolomics.
-The successful outcome would lead to the potential applications discussed in the following paragraph, whereas the unsuccessful outcome would set-up additional requirements to tackle the issue of noisiness in metabolomics data sets.
+\par The main project hypothesis -- the LDA-like models can detect and correct the degeneracies of MSI data sets -- can be interpreted as the spatial smoothing of MSI data sets.
+Both cases of proving and disproving this hypothesis would contribute to the research in MSI.
+The successful outcome would lead to the potential applications discussed in the following paragraph, whereas the unsuccessful outcome would set up additional requirements to tackle the issue of noisiness in MSI data sets.
 
@@
 % - Contributing to machine learning in general.
-\par The potential applications of this research can be directly induced from the previously listed objectives.
-First of all, we would show that the general models or the models applied in other domains can be utilised in metabolomics.
-Further, the developed error correction methodology would improve the accuracy of the chemical processes prediction.
-Finally, the generalisation of the developed error correction model might suggest different approaches in tackling error correction in general; also, it would impact the awareness in metabolomics and the field's potential in developing machine learning generalisations.
+\par The potential applications of the successful research outcome can be directly induced from the previously listed objectives.
+First of all, we would show that the general models (or the models applied in other domains) can be utilised for MSI data analysis.
+Further, the developed spatial smoothing methodology would improve the accuracy of the chemical processes prediction.
+Finally, the generalisation of the developed spatial smoothing model might suggest different approaches in tackling error correction in general; finally, the successful research outcome would suggest the applicability prospect of topic modelling in MSI data sets; reversely, it would also impact the prospect to derive machine learning generalisations from the research in MSI data sets.
 
@@
 Further, LDA is a three-level hierarchical structure: the smallest entity (grey circle) is defined as a \textit{word}; then a collection of words (inner plate) is defined as a \textit{document}; and a collection of documents (outer plate) is a \textit{corpus}.
-Note that Figure 1 represents $M$ documents and $N$ words per document, it follows that in the corpus there would be $M \times N$ words.
-Also, as suggested by Figure 1, let's denote a word by $w$.
+Note that Figure \ref{fig:lda} represents $M$ documents and $N$ words per document, it follows that in the corpus there would be $M \times N$ words.
+Also, as suggested by Figure \ref{fig:lda}, let's denote a word by $w$.
 Then, let's denote the sequence of words in document $d$ by
@@
 \item $w$ is the observable variable referring to the words in documents;
-\item and $z$ is the latent variable of topic assignments over the words.
+\item $z$ is the latent variable of topic assignments over the words.
 \end{itemize}
@@
 That is, we are assessing the minimum bound over the latter parameters.
-Further, recall that in Figure 1 we explicitly express the parameter $\phi$ in describing the basic LDA model.
+Further, recall that in Figure \ref{fig:lda} we explicitly express the parameter $\phi$ in describing the basic LDA model.
 However, the authors have chosen to show it separately -- upon describing the variational method.
@@
 Note that the documents in each time-slice are exchangeable.
-This treatment of time set-ups a basis to tackle the problems which involve the sequential development of the data.
+This treatment of time set ups a basis to tackle the problems which involve the sequential development of the data.
 On the other hand, note that another assumption of discrete data remains the same as in the basic LDA model.
@@
 This means that the topic distribution over documents and the vocabulary terms distribution over topics change over time.
-The visualisation of this process is given in Figure 3 below.
+The visualisation of this process is given in Figure \ref{fig:dtm} below.
 \begin{figure}[H]
@@
 \end{figure}
-Effectively, Figure 3 is a sequential combination of the basic LDA model illustrated in Figure 1.
-In Figure 3, one basic LDA model represents one time-slice.
+Effectively, Figure \ref{fig:dtm} is a sequential combination of the basic LDA model illustrated in Figure \ref{fig:lda}.
+In Figure \ref{fig:dtm}, one basic LDA model represents one time-slice.
 Also, the parameters $\alpha$ and $\beta$ are induced by their predecessors.
@@
 
-% Section Conclusion
-\par
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Section intro
+% - Discussion over literature
+% - Design of the model
+% - Evaluation of the model
+\par In this section, we will set up the approach of the project's execution.
+To start with, we will go over the literature reviewed in Section 3 in order to derive the promising techniques in terms of applicability to MSI data sets.
+Then, we will go over the high-level design and implementation requirements.
+Finally, we will set the basis for the model's evaluation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Discussion over the literature
+\subsection{Discussion}
+
+% Intro
+% - Literature review shows the mathematical underlying
+% - Extract the relevant bits
+\par Throughout the literature review, we have looked into the papers which are considered to have a significant influence on the LDA-like modelling.
+For this reason, we have set up the basis of the LDA-like model mathematical underlying.
+However, the approach in utilising those models in MSI like data sets have not been reviewed.
+Nevertheless, two sources of relevant research have been mentioned (the paper by Hooft et al.
+\cite{hooft} and the survey by Alonso et al.
+\cite{alonso-et-al}).
+Ultimately, this section is focused to emphasise the superior approaches of the previously reviewed papers.
+
+% Generative processess
+% - Comparison between static and dynamic data sets
+\par The reviewed generative processes sets an assumption on the formulation of the data.
+We have reviewed two generative processes: static and dynamic.
+The static generative process is given by the initial LDA paper -- `Latent Dirichlet Allocation', and the dynamic generative processes is given by the 'Dynamic Topic Models' paper.
+These generative processes are expressed respectively in Figure \ref{fig:static} and Figure \ref{fig:dynamic} below.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{static}
+  \caption{Static generative process.}
+  \label{fig:static}
+\end{figure}
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{dynamic}
+  \caption{Dynamic generative process.}
+  \label{fig:dynamic}
+\end{figure}
+In the previous plots, we have visualised the occurrences of a particular entity within a corpus.
+The static generative process suggests a random distribution of the entities, whereas the dynamic generative process induces a particular pattern.
+Since in metabolomics we expect the metabolites of same type to form clusters (rather than to be randomly distributed across the sample), the static generative process is more appropriate to reflect the nature of a biological entity development.
+
+% Generative process
+% - Rather than sequential development, we induce a relationship between the documents
+
+The initial dynamic topic modelling motivation suggests the application of time-series into the generative process of data sets.
+However, the way we capture the ions is done simultaneously for the whole corpus.
+Therefore, we are relaxing the assumption of time-series.
+Instead of that, we will use the dynamic topic modelling techniques to establish the relationship between the time-slices (even though we are not taking time into account, we will stick to the same terminology to preserve the consistency).
+For example, if we assume that the data is generated in a 1-dimensional manner, then time-slice $t$ would depends on its adjacent time-slices $t-1$ and $t+1$.
+However, since the data sets we are working on are 2-dimensional, the induced relationship considers more adjacent time-slices.
+This relationship is shown in Figure \ref{fig:2-d_generative_process} below.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\textwidth]{2-d_generative_process}
+  \caption{2-dimensional generative process.}
+  \label{fig:2-d_generative_process}
+\end{figure}
+In the boundary case shown on the left, time-slice $t_{0,0}$ relates to  three surrounding time-slices, whereas, in the general case indicated on the right side, time-slice $t_{1,1}$ relates to eight surrounding time-slices.
+
+time-slice $t$ would be dependant on it's adjacent time-slices as shown  
+Even though the dynamic topic modelling suggests
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Design}
+
+% Design
+% - The required model supplements
+% -- Dictionary
+\par
+% To-do: induce time-series in terms of z axis
+
+% Implementation
 \par
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Overview}
-
-% Discussion over the literature
-\par
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Approach}
-
-% Design
-\par
-
-% Implementation
-\par
+\subsection{Evaluation}
 
 % Evaluation/Experiments
-\par
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Risk Management}
-
-% Section conclusion
 \par

### 2016-12-11 docs/MSci_proposal/mprop.tex
The start of the initial design write-up.

@@
 % - Extract the relevant bits
-\par Throughout the literature review, we have looked into the papers which are considered to have a significant influence on the LDA-like modelling.
-For this reason, we have set up the basis of the LDA-like model mathematical underlying.
+\par Throughout the literature review, we have looked into the papers establishing the mathematical underlying of the LDA-like models.
 However, the approach in utilising those models in MSI like data sets have not been reviewed.
-Nevertheless, two sources of relevant research have been mentioned (the paper by Hooft et al.
+Nevertheless, two sources of relevant research have been provided (the paper by Hooft et al.
 \cite{hooft} and the survey by Alonso et al.
 \cite{alonso-et-al}).
-Ultimately, this section is focused to emphasise the superior approaches of the previously reviewed papers.
+Ultimately, this section is focused to emphasise the superior approaches of the previously reviewed papers and establish the appropriate techniques in the generative and inference processes.
 
@@
 We have reviewed two generative processes: static and dynamic.
-The static generative process is given by the initial LDA paper -- `Latent Dirichlet Allocation', and the dynamic generative processes is given by the 'Dynamic Topic Models' paper.
+The static generative process is given by the initial LDA paper -- `Latent Dirichlet Allocation', and the dynamic generative processes is given by the `Dynamic Topic Models' paper.
 These generative processes are expressed respectively in Figure \ref{fig:static} and Figure \ref{fig:dynamic} below.
@@
 In the boundary case shown on the left, time-slice $t_{0,0}$ relates to  three surrounding time-slices, whereas, in the general case indicated on the right side, time-slice $t_{1,1}$ relates to eight surrounding time-slices.
+Thereby, the notion of dynamic topic modelling is preserved even upon relaxing the assumption of time-series.
 
-time-slice $t$ would be dependant on it's adjacent time-slices as shown  
-Even though the dynamic topic modelling suggests
+% Inference
+% - Stochastic and optimisation based methods
+% - Evaluation
+\par The reviewed inference techniques have covered the mathematical underlying of two approaches: stochastic (sampling-based) and optimisation (variational).
+The variational methods are discussed in the `Latent Dirichlet Allocation' and `Dynamic Topic Models' papers, and the sampling-based method is discussed in the `Finding Scientific Topics' paper.
+Since both methods have been utilised in the state-of-the-art topic modelling applications, we will implement both of them in order to set up a discussion for the developed model's evaluation.
+However, since the authors of `Dynamic Topic Models' has utilised only the variational methods, we have to address the compatibility of the sampling-based methods in terms of applicability to the dynamic topic modelling.
 
@@
 
-% Design
+% Intro
 % - The required model supplements
-% -- Dictionary
+\par By covering the design of the model, we will set up the requirements for the project's execution and propose methods which would supplement the techniques derived from the reviewed literature.
+To start with, we will address the dimensionality reduction of the MSI data sets.
+Then, we will introduce the additional data capturing requirements shifting the set-up of the generative process.
+Finally, we will suggest an approach for utilising sampling-based methodology in the dynamic topic modelling.
+
+% Dimensionality Reduction
+% - Dictionary
+% - Preprocessing
+% -- Dismissing low intensity values
+% -- Storing the ranges of ions
+\par In order to address the scalability of the model in terms of the large MSI data sets, we will address two techniques for the dimensionality reduction: utilisation of a dictionary and pre-processing of the data.
+Both of these techniques reduces the sparsity of the data sets.
+To start with, the use of dictionary will eliminate the irrelevant use of memory in storing the words which do not occur in particular documents.
+For example, let's assume that the masses of the ions in our data can be represented as the set $\{1, 2, 3, 4, 5\}$, also assume that the ions can be represented in the form of (mass, intensity) tupple.
+Then, if our sample contained two following documents: $\mathbf{w_0} = \{(1, 100), (2, 200)\}$ and $\mathbf{w_1} = \{(3, 300)\}$, the sample representation in matrix and dictionary forms would be interchangeable with the expression given in Table \ref{tab:m_and_d} below.
+
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{|c||c|c|c|c|c|}
+\hline
+&1&2&3&4&5\\
+\hline
+$\mathbf{w_0}$&100&200&0&0&0\\
+$\mathbf{w_1}$&0&0&300&0&0\\
+\hline
+\end{tabular}
+\quad
+\begin{tabular}{|c||c|c|c|c|c|}
+\hline
+&1&2&3&4&5\\
+\hline
+$\mathbf{w_0}$&100&200&$\O$&$\O$&$\O$\\
+$\mathbf{w_1}$&$\O$&$$\O$$&300&$\O$&$\O$\\
+\hline
+\end{tabular}
+\end{center}
+\caption{Matrix and dictionary representation of the data set.}
+\label{tab:m_and_d}
+\end{table}
+
+% Generative process
+% - induce time-series in terms of z axis
 \par
-% To-do: induce time-series in terms of z axis
 
-% Implementation
+% Inference
 \par

### 2016-12-11 docs/MSci_proposal/mprop.tex
The design subsection: 3-d generative processes.

@@
 To start with, the use of dictionary will eliminate the irrelevant use of memory in storing the words which do not occur in particular documents.
-For example, let's assume that the masses of the ions in our data can be represented as the set $\{1, 2, 3, 4, 5\}$, also assume that the ions can be represented in the form of (mass, intensity) tupple.
-Then, if our sample contained two following documents: $\mathbf{w_0} = \{(1, 100), (2, 200)\}$ and $\mathbf{w_1} = \{(3, 300)\}$, the sample representation in matrix and dictionary forms would be interchangeable with the expression given in Table \ref{tab:m_and_d} below.
-
+For example, let's assume that the masses of the ions in our data can be represented as the set $\{1, 2, 3, 4, 5\}$, also assume that the ions can be represented in the form of (mass, intensity) tuple.
+Then, if our sample contained two following documents: $\mathbf{w_0} = \{(1, 0.1), (2, 200), (3, 500)\}$ and $\mathbf{w_1} = \{(5, 300)\}$, the sample representation in matrix and dictionary forms would be interchangeable with the expression given in Table \ref{tab:m_and_d} below.
 \begin{table}[H]
@@
 \hline
-$\mathbf{w_0}$&100&200&0&0&0\\
-$\mathbf{w_1}$&0&0&300&0&0\\
+$\mathbf{w_0}$&0.1&200&500&0&0\\
+$\mathbf{w_1}$&0&0&0&0&300\\
 \hline
@@
 \hline
-$\mathbf{w_0}$&100&200&$\O$&$\O$&$\O$\\
-$\mathbf{w_1}$&$\O$&$$\O$$&300&$\O$&$\O$\\
+$\mathbf{w_0}$&0.1&200&500&$\O$&$\O$\\
+$\mathbf{w_1}$&$\O$&$\O$&$\O$&$\O$&300\\
 \hline
@@
 \end{table}
+Note that, in these settings, empty set symbol $\O$ represents a value which is not assigned to the memory; and the matrix and dictionary forms are respectively given on the left and right hand sides.
+It follows that the use of dictionary requires about as half as much memory.
+Another dimensionality reduction technique can be introduced by setting the minimum threshold value.
+That is, we can limit the data set by dismissing values which do not reach the sufficient significance criteria.
+If we set the threshold of intensity to be $1$, then the dictionary could be represented in Table \ref{tab:d_threshhold} below.
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{|c||c|c|c|c|c|}
+\hline
+&1&2&3&4&5\\
+\hline
+$\mathbf{w_0}$&\O&200&500&$\O$&$\O$\\
+$\mathbf{w_1}$&$\O$&$\O$&$\O$&$\O$&300\\
+\hline
+\end{tabular}
+\end{center}
+\caption{The dictionary of the data set with the minimum intensity threshold of 1.}
+\label{tab:d_threshhold}
+\end{table}
+
+% Ion grouping
+\par As discussion in Section 1, the captured ions of same time are likely to have fluctuations in their masses.
+For this reason, we are applying another pre-processing step.
+By setting the threshold which would indicate the maximum fluctuation between ions masses, we would group the ions which are likely to represent the same time, also, we would reduce the data set even further.
+By taking the example of previous paragraph and setting the fluctuation tolerance to $1$, we obtain the data structure given in Table \ref{tab:d_f_threshhold} below.
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{|c||c|c|}
+\hline
+&[2,3]&[5,5]\\
+\hline
+$\mathbf{w_0}$&700&$\O$\\
+$\mathbf{w_1}$&$\O$&300\\
+\hline
+\end{tabular}
+\end{center}
+\caption{The dictionary of the data set with the fluctuation threshold of 1.}
+\label{tab:d_f_threshhold}
+\end{table}
+Notice that the masses are now represented by intervals, and the intensities below the threshold are summed together.
+Ultimately, by utilising the latter dimensionality reduction and ion grouping techniques, we establish data structure which would scale better compared to the initial data capturing settings.
 
 % Generative process
-% - induce time-series in terms of z axis
-\par
+% 3-D data sets
+\par The generative process of the MSI data set discussed in the discussion over the reviewed literature has utilised 2-dimensional data set.
+However, we are expecting to work on 3-dimensional data sets.
+Therefore, the established relationship between time-slices is illustrated by Figure \ref{fig:3-d_generative_process} below.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{3-d_generative_process}
+  \caption{3-dimensional generative process.}
+  \label{fig:3-d_generative_process}
+\end{figure}
+The time-slice indicated by the green colour, which is one of the boundary cases, is related to eleven other documents indicated by the red colour.
+Note that in the general case, the time-slice would have twenty-six relating documents.
+Effectively, we are intending to establish a 3-dimensional topic modelling.
+% - To-ask: 4-dimensions (time-series)
 
 % Inference
+% - Metropolis--Hasting
 \par

### 2016-12-11 docs/MSci_proposal/mprop.tex
Design section. (one time read).

@@
 \par In this section, we will set up the approach of the project's execution.
-To start with, we will go over the literature reviewed in Section 3 in order to derive the promising techniques in terms of applicability to MSI data sets.
+To start with, we will go over the literature reviewed in Section 3 to derive the promising techniques in terms of applicability to MSI data sets.
 Then, we will go over the high-level design and implementation requirements.
@@
 \cite{alonso-et-al}).
-Ultimately, this section is focused to emphasise the superior approaches of the previously reviewed papers and establish the appropriate techniques in the generative and inference processes.
+Ultimately, this section is focused to emphasise the superior approaches of the previously reviewed papers; also, we will establish the appropriate techniques for the generative and inference processes.
 
@@
 % - Comparison between static and dynamic data sets
-\par The reviewed generative processes sets an assumption on the formulation of the data.
-We have reviewed two generative processes: static and dynamic.
+\par The reviewed generative processes describe how the initial data is expected to be formed.
+Recall that we have reviewed two generative processes: static and dynamic.
 The static generative process is given by the initial LDA paper -- `Latent Dirichlet Allocation', and the dynamic generative processes is given by the `Dynamic Topic Models' paper.
-These generative processes are expressed respectively in Figure \ref{fig:static} and Figure \ref{fig:dynamic} below.
+The applications of these generative processes are expressed respectively in Figure \ref{fig:static} and Figure \ref{fig:dynamic} below.
 \begin{figure}[H]
@@
 The static generative process suggests a random distribution of the entities, whereas the dynamic generative process induces a particular pattern.
-Since in metabolomics we expect the metabolites of same type to form clusters (rather than to be randomly distributed across the sample), the static generative process is more appropriate to reflect the nature of a biological entity development.
+Since we expect the ions of same type be in clusters (rather than to be randomly distributed across the sample), the dynamic generative process is more appropriate to reflect the nature of the chemical entities patterns.
 
@@
 
-The initial dynamic topic modelling motivation suggests the application of time-series into the generative process of data sets.
-However, the way we capture the ions is done simultaneously for the whole corpus.
+The initial dynamic topic modelling paper suggests to induce the time-series into the generative process of data sets.
+However, the way we capture the ions is done instantaneously for the whole corpus.
 Therefore, we are relaxing the assumption of time-series.
@@
 For example, if we assume that the data is generated in a 1-dimensional manner, then time-slice $t$ would depends on its adjacent time-slices $t-1$ and $t+1$.
-However, since the data sets we are working on are 2-dimensional, the induced relationship considers more adjacent time-slices.
+However, since the image of the sample is 2-dimensional, there are more adjacent time-slices.
 This relationship is shown in Figure \ref{fig:2-d_generative_process} below.
@@
 \end{figure}
-In the boundary case shown on the left, time-slice $t_{0,0}$ relates to  three surrounding time-slices, whereas, in the general case indicated on the right side, time-slice $t_{1,1}$ relates to eight surrounding time-slices.
-Thereby, the notion of dynamic topic modelling is preserved even upon relaxing the assumption of time-series.
+In the boundary case shown on the left, time-slice $t_{0,0}$ is related to three surrounding time-slices, whereas, in the general case indicated on the right side, time-slice $t_{1,1}$ relates to eight surrounding time-slices.
+Ultimately, the notion of dynamic topic modelling can be  preserved even upon relaxing the assumption of time-series.
 
@@
 \par In order to address the scalability of the model in terms of the large MSI data sets, we will address two techniques for the dimensionality reduction: utilisation of a dictionary and pre-processing of the data.
-Both of these techniques reduces the sparsity of the data sets.
-To start with, the use of dictionary will eliminate the irrelevant use of memory in storing the words which do not occur in particular documents.
-For example, let's assume that the masses of the ions in our data can be represented as the set $\{1, 2, 3, 4, 5\}$, also assume that the ions can be represented in the form of (mass, intensity) tuple.
-Then, if our sample contained two following documents: $\mathbf{w_0} = \{(1, 0.1), (2, 200), (3, 500)\}$ and $\mathbf{w_1} = \{(5, 300)\}$, the sample representation in matrix and dictionary forms would be interchangeable with the expression given in Table \ref{tab:m_and_d} below.
+Both of these techniques reduce the sparsity of the data sets.
+To start with, the use of dictionary will eliminate the irrelevant use of memory by not storing the words which do not occur in the documents.
+For example, let's assume that the masses of the ions in our data can be represented as the set $\{1, 2, 3, 4, 5\}$, also assume that the ions can be represented in the form of a (mass, intensity) tuple.
+Then, if our sample contained two following documents: $\mathbf{w_0} = \{(1, 0.1), (2, 200), (3, 500)\}$ and $\mathbf{w_1} = \{(5, 300)\}$, the sample representation in matrix and dictionary forms would be interchangeable with the expressions given in Table \ref{tab:m_and_d} below.
 \begin{table}[H]
@@
 For this reason, we are applying another pre-processing step.
-By setting the threshold which would indicate the maximum fluctuation between ions masses, we would group the ions which are likely to represent the same time, also, we would reduce the data set even further.
-By taking the example of previous paragraph and setting the fluctuation tolerance to $1$, we obtain the data structure given in Table \ref{tab:d_f_threshhold} below.
+By setting the threshold which indicated the maximum fluctuation between the masses of the ions, we would group the ions which are likely to represent the same type; also, we would reduce the dimensionality of the data set.
+By taking the example of the previous paragraph and setting the fluctuation tolerance to $1$, we obtain the data structure given in Table \ref{tab:d_f_threshhold} below.
 \begin{table}[H]
@@
 \end{figure}
-The time-slice indicated by the green colour, which is one of the boundary cases, is related to eleven other documents indicated by the red colour.
+The time-slice indicated by the green colour, which is one of the boundary cases, is related to other elven time-slices indicated by the red colour.
 Note that in the general case, the time-slice would have twenty-six relating documents.
-Effectively, we are intending to establish a 3-dimensional topic modelling.
+Effectively, we have introduced the application settings of the dynamic topic modelling in the 3-dimensional space.
 % - To-ask: 4-dimensions (time-series)
@@
 % Inference
-% - Metropolis--Hasting
-\par
+% - Metropolis--Hastings
+\par In Section 3, we have addressed a sampling-based inference method -- Gibbs sampling.
+However, for the dynamic topic models, the latter method has to be reconsidered due to the change from Dirichlet to Gaussian distributions.
+Therefore, we will look into another MCMC based algorithm -- Metropolis--Hastings.
+% - To-do: Analyse the details of the MH.
 
@@
 
+% Intro
+% - Spatial Smoothing
+% - Two inference methods
+% - Risk management
+\par The developed technique addressing the spatial smoothing of the MSI data sets will evaluated by the model's performance.
+The performance will be assessed by investigating the inferred patterns of the captured ions.
+Also, we will utilise two different inference methods: sampling-based and variational.
+By having two model variations, we will be able to compare the time performance.
+Hence, we will be able to deduce which variation scales better with the size of the data sets.
+Finally, the section will be finished by investigating the possible risks.
+Also, we will provide a plan to mitigate these risks.
+
 % Evaluation/Experiments
+\par
+
+% Risks
+\par
+
+% To-do:
+% - Ask about experiments
+% - Ask about risks
 \par

### 2016-12-12 docs/MSci_proposal/mprop.tex
Work plan: schedule. Initial draft.

@@
 \usepackage{cite}
+\usepackage{pgfgantt}
+
 \graphicspath{ {images/} }
@@
 By utilising MSI, we can visualise the distributions of the ions (or ion sets) of different masses.
-For instance, Figure \ref{fig:b_and_e} below represents the whole sample as an image, where each picture's pixel is a small region within the sample.
+For instance, Figure \ref{fig:b_and_e} below represents a captured sample as an image, where each picture's pixel is a small region within the sample.
 \begin{figure}[H]
@@
 An open problem in metabolomics is to find the sets of chemical entities  inducing valid visual patterns.
-Ultimately, the discovered biological patterns could be used to suggest and measure the likelihood of particular chemical process occurrence.
+Ultimately, the discovered biological patterns could be used to suggest and measure the likelihood of a particular chemical process occurrence.
 For a rigorous discussion on machine learning applications in metabolomics, the reader can consult the recent survey conducted by Alonso et al.
@@
 % - Sections
-\par The structure of this proposal is given as follows: in Section 2 we present the issue of noise in MSI data sets; in section 3 we cover the literature in topic modelling which focuses on the variations of LDA-like models; in Section 4 we provide the metric of success of the proposed research; finally, in Section 5, we go into details on how the project will be executed.
+\par The structure of this proposal is given as follows: in Section 2 we present the issue of noise in MSI data sets; in section 3 we cover the literature in topic modelling which focuses on the variations of LDA-like models; in Section 4 we discuss the approach to tackle the problem (we discuss the reviewed literature, address design and evaluation details); finally, in Section 5, we set and schedule the tasks required to complete the project.
 % To-do: rewrite the outline
@@
 % Intro
-\par
+% - Tasks.
+% - Schedule
+% - Deliverables
+\par In this section, we will outline the project milestones and deliverables.
+That is, we will outline the most important steps and provide the estimations in a form of Gantt chart.
+Further, the deliverables will be directly induced from the project's objectives and hypothesis (these are addressed in Section 2).
+The deliverables will contain both successful and successful outcomes of the set hypothesis.
 
-% Schedule
-% - Dates
-% - Tasks
 \subsection{Schedule}
+
+% Tasks
+% - Design
+% -- Investigate the literature on the spatial smoothing application in MSI data sets
+% -- Design the dynamic topic model
+% - Implementation
+% -- Implement the generative process of the dynamic topic modelling with applied spatial smoothing
+% -- Implement both sampling-based and variational inference to the dynamic topic model
+% - Evaluation
+% -- Test the performance of the applied spatial smoothing
+% - Documentation/Presentation
+% -- Dissertation
+% -- Reports
+% -- Presentation
+\par The high-level plan of the project's execution include the following milestones: design, implementation, evaluation, and documentation/presentation.
+To establish a basis for the design, we will continuously review the literature on spatial smoothing, MSI, and dynamic topic models.
+Then, the implementation will require the utilisation of spatial smoothing in the generative process of the dynamic topic model; also, for the inference, we will implement both: sampling-based and variational methods.
+Further, the evaluation will be conducted by assessing the performance of the developed models.
+Finally, the results will be documented in a form of weekly reports, dissertation, and presentation.
+
+% Dates
+% - Non-continuous
+% - Estimations
+\par The outlined tasks will not be necessarily executed in sequential manner. For some of the tasks, we will use parallel approach as given in Figure \ref{fig:gantt} below.
+
+\begin{figure}[H]
+\begin{ganttchart}[
+hgrid, 
+vgrid,
+y unit chart=.8cm
+]{1}{15}
+
+\gantttitle{Weeks}{15} \\
+\gantttitlelist{1,...,15}{1} \\
+
+\ganttgroup{Design}{1}{7} \\
+\ganttbar{Dynamic Generative Process}{1}{5} \\
+\ganttbar{Spatial Smoothing}{3}{7} \ganttnewline
+
+\ganttgroup{Implementation}{3}{10} \\
+\ganttbar{Dynamic Generative Process}{3}{6} \\
+\ganttbar{Inference Methods}{4}{7} \\
+\ganttbar{Spatial Smoothing}{7}{10} \\
+
+\ganttgroup{Evaluation}{8}{11} \\
+\ganttbar{Usability}{8}{10} \\
+\ganttbar{Performance}{9}{11} \ganttnewline
+
+\ganttgroup{Documentation/Presentation}{2}{15} \\
+\ganttbar{Reports}{2}{11} \\
+\ganttbar{Dissertation}{10}{15} \ganttnewline
+\ganttbar[
+=TBA
+]{Presentation}{12}{14}
+
+\end{ganttchart}
+\caption{Gantt chart of the project's execution.}
+\label{fig:gantt}
+\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% it is fine to change the bibliography style if you want
+
 \bibliography{mprop}
 \bibliographystyle{plain}
+% To-do: list, in alphabetical order by author and date
 \end{document}
-
-% To-do: relaxing the matrix dimension assumptions, i.e., introducing the dictionary variation

### 2016-12-12 docs/MSci_proposal/mprop.tex
Work plan: deliverables. Initial draft.

@@
 % - Estimations
-\par The outlined tasks will not be necessarily executed in sequential manner. For some of the tasks, we will use parallel approach as given in Figure \ref{fig:gantt} below.
+\par The outlined tasks will not be necessarily executed in sequential manner. For some of the tasks, we will use the parallel approach as given in Figure \ref{fig:gantt} below.
 
@@
 \ganttbar{Dissertation}{10}{15} \ganttnewline
-\ganttbar[
-=TBA
-]{Presentation}{12}{14}
+\ganttbar{Presentation}{12}{14}
 
@@
 
-% Code
-% Dissertation
 \subsection{Deliverables}
 
-% Conclusion
-\par
+% - Successful/Unsuccessful
+% - Model in Python 
+% -- 2/3 objectives
+% - Dissertation
+% - Presentation
+\par The deliverables of the project will depend on the outcome of the initial hypothesis -- `the degeneracies of MSI data sets can be detected using LDA-like models'.
+Nevertheless, in any of the project's outcomes, we will provide the following deliverables: the presentation; the dissertation; and the spatial smoothing model, which will be written in Python.
+However, the main project's contribution to the research in MSI will depend on the results: if the model displays an improvement in performance, then we could utilise the model in pre-processing MSI data sets; in the case of unsuccessful outcome, we would assess the techniques used and provide details on why the model does not work.
+Both outcomes would bring the motivation for further research in improving MSI data sets: the successful outcome would bring attention to the applicability of LDA-like models; the unsuccessful outcome would suggest the need of a different perspective required to tackle the issue of noise.
 

### 2016-12-17 docs/MSci_proposal/mprop.tex
Finishing the evaluation section.

@@
 % - Spatial Smoothing
-\par This proposal is focused on applying machine learning techniques to improve the analysis of biology data sets.
+\par This proposal is focused on applying machine learning techniques in order to improve the analysis of biology data sets.
 We intend to enhance the quality of \textit{mass spectrometry imaging} (MSI) data.
@@
 \end{enumerate}
-Briefly, the first objective is expected to utilise the general state-of-the-art topic modelling methodology to distinguish the techniques displaying better performance in MSI applications -- we will conduct a survey on LDA-like topic modelling methodology.
-The second objective focuses on optimising those general models for the particular task of spatial smoothing in MSI data sets.
-Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered spatial smoothing techniques might be applicable to other domains.
+Briefly, for the first objective, we will conduct a survey on LDA-like topic modelling methodology.
+This survey will distinguish the potential models for the applications in MSI.
+The second objective focuses on optimising these potential models for the particular task of spatial smoothing in MSI data sets.
+Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered spatial smoothing techniques might be applicable in other domains.
 
@@
 % - Example
-\par Recall that in topic modelling induces a hidden structure within the studied data sets.
-We will provide an example providing an intuition in this process.
-Say, if we were analysing a book, its chapters might address different topics or address the topics at a different extent.
+\par Recall that topic modelling infers hidden patterns from the processed data sets.
+For example, if we were analysing a book, its chapters might address different topics or address the topics at different extent.
 Further, the chapters itself would have unique distributions of the topics.
-Note that we are assuming that we do not know what the topics addressed in the book.
+Note that we are assuming that there is no knowledge about the actual topics covered in the book.
 In order to derive these hidden topics, a probabilistic topic modelling method would analyse the structure of the words used in the book.
@@
 % - A notification that the terminology and the notation will be useful throughout the proposal.
-\par The terminology used in this proposal is equivalent to the terminology used by Blei \cite{blei_2012}.
+\par The terminology used in this proposal is similar to the terminology used by Blei \cite{blei_2012}.
 Figure \ref{fig:lda} below represents the original LDA model in plate notation.
@@
 The circles represent an entity, whereas the plates indicate the number of entities.
-For example, there would $N \times M$ entities of $z$ and $M$ entities of $\theta$.
+For example, there would be $N \times M$ entities of $z$ and $M$ entities of $\theta$.
 Further, LDA is a three-level hierarchical structure: the smallest entity (grey circle) is defined as a \textit{word}; then a collection of words (inner plate) is defined as a \textit{document}; and a collection of documents (outer plate) is a \textit{corpus}.
 Note that Figure \ref{fig:lda} represents $M$ documents and $N$ words per document, it follows that in the corpus there would be $M \times N$ words.
-Also, as suggested by Figure \ref{fig:lda}, let's denote a word by $w$.
-Then, let's denote the sequence of words in document $d$ by
+Also, as suggested by Figure \ref{fig:lda}, let's denote a word by $w$; then, the sequence of words in document $d$ is denoted by
 \begin{equation*}
@@
 \end{equation*}
-then the sequence of words in the corpus would be denoted by
+and the sequence of documents in the corpus is denoted by
 \begin{equation*}
@@
 \par Now, we will introduce the lower level entities.
-The grey circle indicates that the \textit{observable} variables, whereas the white coloured circles indicate the \textit{hidden} variables.
+The grey circle indicates the \textit{observable} variables, whereas the white coloured circles indicate the \textit{hidden} variables.
 Note that in Figure 1 the words are the only observable variables; all other entities are hidden, i.e., they describe the underlying structure of the model.
-Further, recall that topic modelling is applied in order to discover a mixture of topics; or in other words, topic modelling could describe a corpus and its documents in terms of topic distributions.
+Further, recall that topic modelling is applied in order to discover a mixture of topics.
+In other words, topic modelling could describe a corpus and its documents in terms of topic distributions.
 This intuition will simplify the definitions of the remaining entities in Figure 1, these are given as follows:
@@
 % - The derivation of the posterior
-\par The relationship the between observable and hidden variables, effectively, describes the inference of the underlying topic structure.
-We will go through an arbitrary example in order to give an intuition in the inference process and familiarise with the lower level notation of the previous listed variables.
-Figure \ref{fig:lda} shows see that $\alpha_k$ (the prior for the $k$th topic in a document) influences $\theta_{:, k}$ (the latent variable describing the distribution of topic $k$ over all documents).
+\par The relationship of observable and hidden variables, effectively, describes the inference of the underlying topic structure.
+We will go through an arbitrary example in order to give an intuition in the inference process; also, we will familiarise with the lower level notation of the previous listed variables.
+Figure \ref{fig:lda} illustrates that $\alpha_k$ (the prior for the $k$th topic in a document) influences $\theta_{:, k}$ (the latent variable describing the distribution of topic $k$ over all documents).
 It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in document $d$) influences $z_{d, :}$ (the topic assignment over all vocabulary terms for document $d$).
@@
 The approximation of it will be considered during the literature review on LDA-like models.
-The naming of the following subsections corresponds to the reviewed papers.
+Note that the naming of the following subsections corresponds to the reviewed papers.
 
@@
 \cite{blei_2003}.
-Since this paper is the origin of LDA-like models, it will be discussed in more detail to introduce the basis of LDA-like methodology.
+Since this paper is the origin of LDA-like models, it will be discussed in more detail in order to introduce the basis of LDA-like methodology.
 As a result, this coverage will simplify the review on the LDA derivatives discussed in the upcoming subsections.
@@
 % - The generative algorithm
-\par As discussed in the preliminaries subsection, we are also taking an assumption that the corpus is induced by a probabilistic process.
+\par As discussed in the preliminaries subsection, we also set an assumption that the corpus is induced by a probabilistic process.
 The generation of the corpus has three levels: the parameters $\alpha$ and $\beta$ are sampled once; the variable $\theta$ is sampled once per document; and the variables $z$ and $w$ are sampled once per word.
@@
 In order to generate the whole corpus, we would run the algorithm for each document.
-Going back to Algorithm 1, a brief intuition behind it is given as follows: in line 1 we draw the number of words $N$ in a document ($\xi$ is an ancillary variable used as the mean for the Poisson Distribution); in line 2 we draw topic distribution $\theta$ from the Dirichlet distribution on parameter $\alpha$; further, in lines 3--6 we generate the words: in line 4 we draw topic $z_n$ from the multinomial distribution on $\theta$ and, finally, in line 5 we obtain  word $w_n$ from the conditional probability on $z_n$ and $\beta$, recall that $\beta$ is the prior on the vocabulary terms correspondence to topics.
+Going back to Algorithm 1, a brief intuition behind it is given as follows: in line 1 we draw the number of words in a document $N$($\xi$ is an ancillary variable used as the mean for the Poisson Distribution); in line 2 we draw topic distribution $\theta$ from the Dirichlet distribution on parameter $\alpha$; further, in lines 3--6 we generate the words: in line 4 we draw topic $z_n$ from the multinomial distribution on $\theta$ and, finally, in line 5 we obtain  word $w_n$ from the conditional probability on $z_n$ and $\beta$, recall that $\beta$ is the prior on the vocabulary terms correspondence to topics.
 
@@
 % - Variational parameters gamma and fi
-\par As suggested in the ending paragraph of the preliminaries subsection, the inference relates to the computation of evidence $p(w)$ (it is the denominator term given in Equation 2).
-Even though the original paper mentions the inference methods such as Markov chain Monte Carlo, it suggests using the method of variational inference.
+\par As suggested by the last paragraph of the preliminaries subsection, the inference relates to the computation of evidence $p(w)$ (it is the denominator term given in Equation 2).
+Even though the original paper mentions the inference methods such as Markov chain Monte Carlo, the authors suggest using the method of variational inference.
 In order to develop an intuition over the method, the authors introduce the variational parameters $\gamma$ and $\phi$.
@@
 The first experiment addresses document modelling.
-LDA  has displayed balanced hidden topic proportions in the set of tested documents.
+To be more specific, LDA has displayed balanced hidden topic proportions in the set of tested documents.
 That is, the tested documents has replicated the topic proportions of the training documents.
 The second experiment is directed towards the problem of document classification.
-They have addressed the \textit{dimensionality reduction} in order to surpass the performance of the support-vector-machine-like models.
+The authors have addressed the \textit{dimensionality reduction} in order to surpass the performance of the support vector machine (SVM) models.
 To be more specific, the documents with reduced dimensionality would possess a lower amount of features.
 The results of the experiment indicate that the basic LDA model has successfully reduced the number of features.
-Also, it has displayed an improvement in accuracy compared to the support-vector machine-like model.
+Also, it has displayed an improvement in accuracy compared to the SVM-like models.
 
@@
 \par The review on the original LDA model has set up a basis required to understand the model's derivatives.
-That is, the introduced methodology will be useful in considering the following improvements: (1) the  relaxation of the exchangeability assumption; (2) the adaptation of time-series suggesting word correspondence to topic over time; (3) the discussion over alternative inference methods.
-Apart from that, the experiments introduced in this section has reflected the LDA performance upon the model's introduction to the public.
+That is, the introduced methodology will be useful in considering the following improvements: (1) the  relaxation of the exchangeability assumption; (2) the adaptation of time-series suggesting the word-topic correspondence over time; (3) the discussion over alternative inference methods.
+Apart from that, the experiments introduced in this section have reflected the LDA performance during the time of the initial model release.
 The current state-of-the-hard topic modelling techniques overcome the shown results.
-The experiments were addressed in order to suggest the application fields of LDA-like models.
+Nevertheless, the carried experiments suggest the application fields of LDA-like models.
 % To-do: rework the paragraph name to reflect its purpose: remove the discussion (you might think it's done, but it's not)
@@
 % - Sampling as a method of inference
-\par The article titled as `Finding Scientific Topics' \cite{griffiths} suggests an alternative approach to the inference which is suitable for LDA-like models.
+\par The research paper titled as `Finding Scientific Topics' \cite{griffiths} suggests an alternative approach to the inference which is suitable for LDA-like models.
 To be more specific, the article provides a readable introduction on the application of the Markov chain Monte Carlo (MCMC) algorithm.
@@
 \par The MCMC-based inference method is often referred by \textit{Gibbs sampling}.
-The authors claim that the method provides a first-order approximation upon inferring the hidden structure of the data.
+The authors claim that the method provides a first-order approximation for the inference of the hidden structure of the data.
 That is, the performance is sufficient to establish quantitative reasoning in assessing the correlating sections and semantic structure of the data.
@@
 % - Time series
-\par The authors has shown a use of Gibbs sampling on a problem of vision.
-In their experiment settings, it is assumed that a collection of images represents a corpus; an image represent a document; and a pixel of an image is a word.
+\par The authors have shown an application of Gibbs sampling in a problem of vision.
+In their experiment settings, it is assumed that a collection of images represents a corpus; an image represents a document; and a pixel of an image is a word.
 The experiment has shown that the model is able to infer the underlying basis (topics) from which the images could be constructed.
@@
 Note that the performance was assessed in terms of \textit{perplexity} -- the uncertainty in assigning a word assignment to a topic.
-It has been shown that Gibbs sampling displays superior performance in the small data sets, whereas the performance in larger data tends to fall off.
+It has also been shown that Gibbs sampling displays superior performance in the small data sets, whereas the performance in larger data sets tends to fall off.
 Furthermore, the authors have assessed the performance upon varying the number of topics.
-The results of this experiment indicate a trade-off in deriving high-level and low-level structures.
-That is, the low number of topics might result in mashing several structures, whereas the high number of topics is likely to infer irrelevant structures.
+The results of this experiment indicate a trade-off in deriving high-level and low-level patterns.
+That is, the low number of topics might result in mashing significant patterns, whereas the high number of topics is likely to infer irrelevant patterns.
 
@@
 % - PNAS
-\par The main experiment of the article is conducted on the PNAS data set containing abstracts of the articles published from 1991 to 2001.
+\par The main experiment of the research paper is conducted on the PNAS data set containing abstracts of the articles published from 1991 to 2001.
 The authors have noticed that the topic distribution over the vocabulary terms change over time.
@@
 % - Qualitative and quantitative improvement
-\par At this point, we have set up the preliminaries to cover the article on utilising the notion of time-series in LDA-like models.
-We review the initial article on utilising time-series -- `Dynamic Topic Models' \cite{blei_2006}.
-Note that by a \textit{static} model we mean that the assumed generative process does not take the development/change of the vocabulary terms into account, whereas in \textit{dynamic} topic modelling we assume the notion of that the documents were generated in a sequential process.
+\par At this point, we have set up the preliminaries to cover the research paper on utilising the notion of time-series in LDA-like models.
+That is, we review the initial paper on utilising time-series in LDA-like models -- `Dynamic Topic Models' \cite{blei_2006}.
+Note that by a \textit{static} model we mean that the assumed generative process does not take the development/change of the vocabulary terms into account, whereas in \textit{dynamic} topic modelling we assume the notion that the documents were generated in a sequential process.
 That is, in dynamic topic modelling, the topics evolve in terms of their distribution over the vocabulary terms.
-Upon reviewing the dynamic topic modelling article, we emphasise the following aspects: the fundamental dynamic topic modelling assumptions and their differences compared to the basic LDA model; the generative process; the suggested inference methods; and the results displayed by the experiments on the initial dynamic topic model.
+By reviewing this research paper, we emphasise the following aspects: the fundamental dynamic topic modelling assumptions and their differences compared to the basic LDA model; the generative process; the suggested inference methods; and the results displayed by the experiments on the initial dynamic topic model.
 % To-do: qualitative and quantitative improvements
@@
 Note that the documents in each time-slice are exchangeable.
-This treatment of time set ups a basis to tackle the problems which involve the sequential development of the data.
+This treatment of time sets a basis to tackle the problems which involve the sequential development of the data.
 On the other hand, note that another assumption of discrete data remains the same as in the basic LDA model.
@@
 The algorithm is initialised by drawing the parameters $\alpha_t$ and $\beta_t$ from the Gaussian distribution with the means given by the previous parameter values.
-Note that the parameters $\alpha_0$ and $\beta_0$ are the the special case.
-They can be initialised randomly, i.e.
-the mean of the Gaussian distribution is random.
+Note that the parameters $\alpha_0$ and $\beta_0$ are the special case.
+That is, these parameters can be initialised randomly.
 However, for the parameters $\alpha_0$ and $\beta_0$ we take larger variance values compared to $\alpha_t$ and $\beta_t$.
-The larger variance values  induce the initial time-slice with larger fluctuations in the topic and vocabulary term distributions; and the smaller variance values in the following time-slices induce the smoothness.
-Further, in lines 3--9, for every document in time-slice $t$ we execute the following procedure: for document $d$, we draw the topic distribution $\eta_d$ (the meaning of the parameter is equivalent to the previously introduced $\theta_d$); then, in lines 5--7, for every word in document $d$ we draw topic assignment $z_{d, n}$ and, finally, we draw word $w_{t, d, n}$.
+The larger variance values  induce the initial time-slice with larger fluctuations in the topic and vocabulary term distributions; and the smaller variance values in the upcoming time-slices induce the smoothness.
+Further, in lines 3--9, for every document in time-slice $t$ we execute the following procedure: for document $d$, we draw the topic distribution $\eta_d$ (the parameter is equivalent to the previously introduced $\theta_d$); then, in lines 5--7, for every word in document $d$ we draw topic assignment $z_{d, n}$ and, finally, we draw word $w_{t, d, n}$.
 Note that the means for the Multinomial distributions are normalised by the use of \textit{parametrisation}.
-Parametrisation mapping $\pi$ is expressed as
+To be more specific, the parametrisation mapping $\pi$ is expressed as
 \begin{equation}
@@
 The implementation details of these methods can be found in the `Dynamic Topic Modelling' article \cite{blei_2006}.
-Note that the implementation details are provided in a brief manner suggesting the use of additional literature, these are also provided in the article.
+Note that the implementation details are provided in a brief manner suggesting the need of supplementary literature.
 
@@
 % - Evolution of Science
-\par The dynamic topic model is evaluated by conducting an experiment on the journals of `Science'.
+\par The implemented dynamic topic model was evaluated by conducting an experiment on the journals of `Science'.
 The experiment is expected to infer the prevalent science fields and the prevalent vocabulary used in these fields.
@@
 The experiment has successfully established this notion of time-series.
-This is proved by comparing the inferred rises and falls of the scientific fields to the history of science.
-Note that the experiment has been performed both Kalman Filtering and Wavelet Regression methods.
+This can be proved by comparing the inferred rises and falls of the scientific fields to the actual history of science.
+Note that the experiment has been performed using both Kalman Filtering and Wavelet Regression methods.
 
@@
 We have provided the visualisation of the model and introduced the algorithm for the generative process.
-Further, we have outlined the inference methods suggests be the authors.
-Finally, we have familiarised with the experiment settings which were used to infer time series in the articles of `Science'.
+Further, we have outlined the inference methods suggested by the authors.
+Finally, we have familiarised with the experiment settings which were used to infer the time series in the articles of `Science'.
 
@@
 Then, we will go over the high-level design and implementation requirements.
-Finally, we will set the basis for the model's evaluation.
+Finally, we will set the basis for the model's evaluation and assess the potential risks.
 
@@
 \par Throughout the literature review, we have looked into the papers establishing the mathematical underlying of the LDA-like models.
-However, the approach in utilising those models in MSI like data sets have not been reviewed.
+However, the approach in utilising those models in MSI-like data sets have not been reviewed.
 Nevertheless, two sources of relevant research have been provided (the paper by Hooft et al.
@@
 \cite{alonso-et-al}).
-Ultimately, this section is focused to emphasise the superior approaches of the previously reviewed papers; also, we will establish the appropriate techniques for the generative and inference processes.
+Ultimately, this section is focused on emphasising the superior approaches of the previously reviewed papers; also, we will establish the appropriate techniques for the generative and inference processes.
 
@@
 Recall that we have reviewed two generative processes: static and dynamic.
-The static generative process is given by the initial LDA paper -- `Latent Dirichlet Allocation', and the dynamic generative processes is given by the `Dynamic Topic Models' paper.
+The static generative process is addressed in the initial LDA paper -- `Latent Dirichlet Allocation', and the dynamic generative process is addressed in the `Dynamic Topic Models' paper.
 The applications of these generative processes are expressed respectively in Figure \ref{fig:static} and Figure \ref{fig:dynamic} below.
@@
 The static generative process suggests a random distribution of the entities, whereas the dynamic generative process induces a particular pattern.
-Since we expect the ions of same type be in clusters (rather than to be randomly distributed across the sample), the dynamic generative process is more appropriate to reflect the nature of the chemical entities patterns.
+Since we expect the ions of the same type to be gathered in clusters (rather than to be randomly distributed across the sample), the dynamic generative process is more appropriate to reflect the nature of the chemical entities patterns.
 
@@
 \end{figure}
-In the boundary case shown on the left, time-slice $t_{0,0}$ is related to three surrounding time-slices, whereas, in the general case indicated on the right side, time-slice $t_{1,1}$ relates to eight surrounding time-slices.
+In the boundary case shown on the left, time-slice $t_{1,1}$ is related to three surrounding time-slices, whereas, in the general case indicated on the right side, time-slice $t_{2,2}$ relates to eight surrounding time-slices.
 Ultimately, the notion of dynamic topic modelling can be  preserved even upon relaxing the assumption of time-series.
@@
 Since both methods have been utilised in the state-of-the-art topic modelling applications, we will implement both of them in order to set up a discussion for the developed model's evaluation.
-However, since the authors of `Dynamic Topic Models' has utilised only the variational methods, we have to address the compatibility of the sampling-based methods in terms of applicability to the dynamic topic modelling.
+However, since the authors of `Dynamic Topic Models' have utilised only the variational methods, we have to address the compatibility of the sampling-based methods in terms of the applicability to the dynamic topic modelling.
 
@@
 \hline
-$\mathbf{w_0}$&0.1&200&500&$\O$&$\O$\\
-$\mathbf{w_1}$&$\O$&$\O$&$\O$&$\O$&300\\
+$\mathbf{w_0}$&0.1&200&500&\O&\O\\
+$\mathbf{w_1}$&\O&\O&\O&\O&300\\
 \hline
@@
 \end{table}
-Note that, in these settings, empty set symbol $\O$ represents a value which is not assigned to the memory; and the matrix and dictionary forms are respectively given on the left and right hand sides.
+Note that, in these settings, the empty set symbol $\O$ represents a value which is not assigned in the memory; and the matrix and dictionary forms are respectively given on the left and right hand sides.
 It follows that the use of dictionary requires about as half as much memory.
@@
 \hline
-$\mathbf{w_0}$&\O&200&500&$\O$&$\O$\\
-$\mathbf{w_1}$&$\O$&$\O$&$\O$&$\O$&300\\
+$\mathbf{w_0}$&\O&200&500&\O&\O\\
+$\mathbf{w_1}$&\O&\O&\O&\O&300\\
 \hline
@@
 % Ion grouping
-\par As discussion in Section 1, the captured ions of same time are likely to have fluctuations in their masses.
+\par As discussed in Section 1, the captured ions of the same type are likely to have fluctuations in their masses.
 For this reason, we are applying another pre-processing step.
-By setting the threshold which indicated the maximum fluctuation between the masses of the ions, we would group the ions which are likely to represent the same type; also, we would reduce the dimensionality of the data set.
+By setting another threshold for the maximum fluctuation between the masses of the ions, we would group the ions which are likely to represent the same type; also, we would reduce the dimensionality of the data set.
 By taking the example of the previous paragraph and setting the fluctuation tolerance to $1$, we obtain the data structure given in Table \ref{tab:d_f_threshhold} below.
@@
 \hline
-$\mathbf{w_0}$&700&$\O$\\
-$\mathbf{w_1}$&$\O$&300\\
+$\mathbf{w_0}$&700&\O\\
+$\mathbf{w_1}$&\O&300\\
 \hline
@@
 Notice that the masses are now represented by intervals, and the intensities below the threshold are summed together.
-Ultimately, by utilising the latter dimensionality reduction and ion grouping techniques, we establish data structure which would scale better compared to the initial data capturing settings.
+Ultimately, by utilising the latter dimensionality reduction and ion grouping techniques, we establish a data structure which would scale better compared to the naive data capturing settings.
 
@@
 % 3-D data sets
-\par The generative process of the MSI data set discussed in the discussion over the reviewed literature has utilised 2-dimensional data set.
+\par The generative process of the MSI data set, which was discussed in the discussion over the reviewed literature, has utilised 2-dimensional data sets.
 However, we are expecting to work on 3-dimensional data sets.
@@
 % - Risk management
-\par The developed technique addressing the spatial smoothing of the MSI data sets will evaluated by the model's performance.
-The performance will be assessed by investigating the inferred patterns of the captured ions.
+\par The developed technique addressing the spatial smoothing of the MSI data sets will evaluated by measuring the improvement in performance.
+That is, the performance will be assessed by investigating the quality of the inferred patterns.
 Also, we will utilise two different inference methods: sampling-based and variational.
@@
 Hence, we will be able to deduce which variation scales better with the size of the data sets.
-Finally, the section will be finished by investigating the possible risks.
-Also, we will provide a plan to mitigate these risks.
+Finally, the section will be finished by investigating the possible risks; also, we will provide potential strategies mitigating the risks.
+
+% 
+% - Generating synthetic data sets
+% -- The improvement in the inference
+% 
+\par The implemented model will be evaluated using synthetic and real data sets.
+However, at the start, we will use the synthetic data sets.
+Note that the performance will be measured in the following way: (1) take two equivalent data sets; (2) apply the spatial smoothing on one of the data sets; (3) infer the underlying structure of the data; (4) check whether the applied smoothing has inferred the underlying structures better.
+By working with synthetic data sets, we had the control over inducing the underlying structure of the data.
+Hence, we will able to give a basis in measuring the performance of the applied spatial smoothing.
+
+% Risks
+% - Algorithm complexity (implementation difficulty)
+% - Size of the data sets -> slow computation 
+% - Obtaining real data
+% - Uncertainty whether the model produces an improvement
+\par Even though we have induced a strategy for the project's evaluation, the described process assumes that we can mitigate the potential risks.
+Note that, at the initial stage of the project, we have identified the following potential risks:
+\begin{itemize}
+  \item The implementation complexity of the inference algorithms;
+  \item The access to the real data.
+  \item The model's scalability in large data sets;
+  \item The inability to measure the performance of the applied spatial smoothing.
+\end{itemize}
+The first three risks would slow down the project execution thus raise a concern of finishing the project on time, whereas the fourth one would prevent the confirmation of denial of the initial project's hypothesis -- `the degeneracies of MSI data sets can be detected using LDA-like models'.
+
+% Risk mitigation
+% - Algorithms
+% -- Familiarising with them before the research
+% -- Starting with the simpler one (Gibbs sampling)
+\par The risk of complex (in terms of development) algorithms has been already mitigated by familiarising with basic sampling-based and optimisation-based inference models.
+That is, we have implemented the models utilising Gibbs sampling and variational inference.
+Also, note that the main model will firstly be developed using Gibbs sampling, as this method is easier to implement compared to the variational one.
+
+% - Scalability
+% -- Display the improvement on small data sets
+% -- Leave the optimisation for the future work, if the benefit exists
+\par Further, the risk of model's scalability can be mitigated by  working on the small data sets.
+In case the final model displays an improvement in the inference after the application of the spatial smoothing, this project would bring motivation for another research project on optimising the method for larger data sets.
+
+% - Real data can be obtained from the Web
+\par The risk of the real data access might be partially resolved, as the University has a laboratory for the research in Metabolomics.
+However, in case this options is not possible, we can obtain the data sets from the Web.
+
+% - Uncertainty in the ground truth
+% -- Generate data sets with underlying structure
+% -- Assess the inference by using spatial and non-spatial models
+\par Finally, the risk of the inability to measure the performance of the applied smoothing can impact the outcome of the project the most.
+That is, it would be difficult to reason about the improvement in the inference of the real data sets, because we do not possess the full knowledge of the data set underlying structure.
+We will mitigate this issue by establishing a performance metric from the synthetic data sets.
 
 % Evaluation/Experiments
-\par
-
-% Risks
-\par
-
-% To-do:
-% - Ask about experiments
-% - Ask about risks
-\par
+% - Real data
+\par By having the defined boundaries of the expected improvement in performance, we can attempt to tackle the real data sets.
+That is, the expected performance improvement would serve as an indicator whether the inferred real data deviates in the expected limit.
+To be more specific, the data sets with applied spatial and the data sets without applied smoothing will display some differences in the inferred topics.
+Hence, we will check whether those differences meet the performance boundaries suggested by the inference of the synthetic data sets.
 
@@
 vgrid,
-y unit chart=.8cm
+y unit chart=.7cm
 ]{1}{15}
@@
 \end{ganttchart}
-\caption{Gantt chart of the project's execution.}
+\caption{The Gantt chart estimating the project's execution.}
 \label{fig:gantt}

### 2016-12-18 docs/MSci_proposal/mprop.tex
the final review up to the evaluation.

@@
 \usepackage{amssymb,amsmath}
+\usepackage{textcomp}
 %\usepackage{url}
@@
 \author{Arijus Pleska}
-\date{Date of submission placed here}
 \maketitle
@@
 \par In this project, the application domain of MSI is set to be \textit{metabolomics} -- it is the scientific study of chemical processes which involve metabolites (small molecules) and metabolomes (sets of small molecules).
-In the settings of metabolomics, we utilise \text{mass spectrometry} (MS) to obtain the distribution of chemical entities within a sample.
+In the research settings of metabolomics, we utilise \text{mass spectrometry} (MS) to obtain the distribution of chemical entities within a sample.
 By utilising MSI, we can visualise the distributions of the ions (or ion sets) of different masses.
@@
 In Figure \ref{fig:b_and_e} we have chosen the range of mass to reflect the ink used to inscribe the letters.
-That is, in each pixel we have summed the intensities of the chemical entities meeting the mass criteria.
+That is, for each pixel, we have summed the intensities of the chemical entities meeting the mass criteria.
 Then, each pixel has been assigned an intensity value.
@@
 Also, we are focusing on the \textit{probabilistic} (statistical) models, more specifically, \textit{latent Dirichlet allocation} (LDA).
-First of all, the successor LDA-like models are treated as a state-of-the-art method in tackling semantic-analysis type problems.
-For more detailed review on the LDA applications, the reader can consult the survey conducted by Blei \cite{blei_2012}.
+First of all, the successor LDA-like models are treated as a state-of-the-art method in tackling semantic analysis type problems.
+For more detailed review on LDA applications, the reader can consult the survey conducted by Blei \cite{blei_2012}.
 Also, Hooft et al.
@@
     \item Successfully apply spatial smoothing in MSI data sets;
-    \item Give a basis to a general topic modelling method for error correction.
+    \item Give a basis for the development of a general topic modelling method for error correction.
 \end{enumerate}
@@
 The second objective focuses on optimising these potential models for the particular task of spatial smoothing in MSI data sets.
-Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered spatial smoothing techniques might be applicable in other domains.
+Finally, the third objective can be treated as a general contribution to the field of machine learning -- the discovered spatial smoothing techniques might be applicable to other domains.
 
@@
 % - The negation of the hypothesis would contribute in terms of reassessing the complexity of the problem.
-\par The main project hypothesis -- the LDA-like models can detect and correct the degeneracies of MSI data sets -- can be interpreted as the spatial smoothing of MSI data sets.
+\par The main project hypothesis -- the LDA-like models can detect and correct the degeneracies of MSI data sets -- can be interpreted as a spatial smoothing application in MSI data sets.
 Both cases of proving and disproving this hypothesis would contribute to the research in MSI.
@@
 \par The potential applications of the successful research outcome can be directly induced from the previously listed objectives.
-First of all, we would show that the general models (or the models applied in other domains) can be utilised for MSI data analysis.
+First of all, we would show that the general models (or the models applied in other domains) can be utilised in MSI data analysis.
 Further, the developed spatial smoothing methodology would improve the accuracy of the chemical processes prediction.
@@
 Also, the review assesses the methodology of LDA-like models and proposes several implementation variations.
-The review is structured in progressing order: at the start, we look into the original LDA paper and define the terminology (it will be used throughout the proposal); then, we familiarise with two different techniques used for the inference of LDA-like models; finally, we review LDA derivatives which shows the potential to be utilised in designing the error correction technique.
+The review is structured in progressing order: at the start, we look into the original LDA paper and define the terminology (it will be used throughout the proposal); then, we familiarise with two different techniques used for the inference of LDA-like models; finally, we review LDA derivatives which show the potential to be utilised as a technique of error correction.
 
@@
 Note that we are assuming that there is no knowledge about the actual topics covered in the book.
-In order to derive these hidden topics, a probabilistic topic modelling method would analyse the structure of the words used in the book.
+In order to derive those hidden topics, a probabilistic topic modelling method would analyse the structure of the words used in the book.
 In other words, the model would induce patterns which would refer to a particular topic.
-Regarding the method's performance, it would depend on the method's ability to optimise the inference of the patterns.
-For example, commonly used words (such as and, or, the, etc.) would be relevant in all topics.
+The method's performance would depend on its ability to optimise the inference of the patterns.
+For example, commonly used words (and, or, the, etc.) would be relevant in all topics.
 Therefore, their semantic impact in defining the topics would be negligible.
-For this reason, we could dismiss these to improve the time performance of the method.
+For this reason, we could dismiss those words to improve the time performance of the method.
 
@@
   \includegraphics[width=0.7\textwidth]{lda_model}
-  \caption{The design of the original LDA model}
+  \caption{The design of the original LDA model.}
   \label{fig:lda}
@@
 \par To start with, we will familiarise with the higher level entities of the model.
-The circles represent an entity, whereas the plates indicate the number of entities.
+The circles represent particular entities, whereas the plates indicate the number of entities.
 For example, there would be $N \times M$ entities of $z$ and $M$ entities of $\theta$.
-Further, LDA is a three-level hierarchical structure: the smallest entity (grey circle) is defined as a \textit{word}; then a collection of words (inner plate) is defined as a \textit{document}; and a collection of documents (outer plate) is a \textit{corpus}.
+Further, LDA is a three-level hierarchical structure: the smallest entity (grey circle) is referred by a \textit{word}; then a collection of words (inner plate) is referred by a \textit{document}; and a collection of documents (outer plate) is referred by a \textit{corpus}.
 Note that Figure \ref{fig:lda} represents $M$ documents and $N$ words per document, it follows that in the corpus there would be $M \times N$ words.
-Also, as suggested by Figure \ref{fig:lda}, let's denote a word by $w$; then, the sequence of words in document $d$ is denoted by
+Also, as suggested by Figure \ref{fig:lda}, let's denote a word by $w$; then, the sequence of words in document $d$ would be denoted by
 \begin{equation*}
@@
 \end{equation*}
-and the sequence of documents in the corpus is denoted by
+and the sequence of documents in the corpus would be denoted by
 \begin{equation*}
@@
 \item $\theta$ is the latent variable referring to the topic distribution over documents;
-\item $\phi$ is the latent variable referring to the vocabulary over topics
+\item $\phi$ is the latent variable referring to the vocabulary distribution over topics
 \item $w$ is the observable variable referring to the words in documents;
@@
 \end{itemize}
-Further, note that the listed entities can be perceived as matrices of the following dimensions: $\alpha$ is $1 \times K$; $\beta$ is $K \times V$; $\theta$ is $M \times K$; $\phi$ is $K \times V$; $w$ is $M \times N$; and $z$ is $M \times V$.
-% To-do: check the dimensions
+Further, note that the listed entities can be perceived as matrices of the following dimensions: $\alpha$ is $1 \times K$; $\beta$ is $K \times V$; $\theta$ is $M \times K$; $\phi$ is $K \times V$; $w$ is $M \times N$; and $z$ is $M \times N$.
 
@@
 Figure \ref{fig:lda} illustrates that $\alpha_k$ (the prior for the $k$th topic in a document) influences $\theta_{:, k}$ (the latent variable describing the distribution of topic $k$ over all documents).
-It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in document $d$) influences $z_{d, :}$ (the topic assignment over all vocabulary terms for document $d$).
-Further, $z_{d, v}$ refers to the topic assignment of vocabulary term $v$.
-By taking $z_{d, v}$ and $\beta_{v, k}$ (the prior of topic $k$ on vocabulary term $v$), we obtain word $w_{d, v}$.
+It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in document $d$) influences $z_{d, :}$ (the topic assignments of of the words in document $d$).
+More specifically, $z_{d, v}$ refers to the topic assignment of vocabulary term $v$.
+By taking $z_{d, v}$ and $\beta_{k, v}$ (the prior of topic $k$ on vocabulary term $v$), we obtain word $w_{d, v}$.
 In the complete perspective, the generative process of the corpus is given by the following joint probability distribution
@@
 The problem arises from inefficient analytical computation of evidence $p(w)$.
-The approximation of it will be considered during the literature review on LDA-like models.
+The approximation of $p(w)$ will be considered during the literature review on LDA-like models.
 Note that the naming of the following subsections corresponds to the reviewed papers.
@@
 That is, the order in which these entities are processed does not matter.
-In other words, we can say the model follows the \textit{bag-of-words} principle.
+In other words, the model follows the \textit{bag-of-words} principle.
 The next assumption states that we should use discrete data.
-Effectively, the basic LDA does not cover  continuous features.
+Therefore, note that the basic LDA model does not cover continuous features.
 Another assumption is on the number of topics: it is set to be fixed throughout the complete run of inference.
@@
   \includegraphics[width=0.4\textwidth]{variational_parameters}
-  \caption{Variational parameters}
+  \caption{Variational parameters.}
   \label{fig:var_par}
 \end{figure}
-Note that the authors provide pseudo code to build an intuition in tackling this optimisation problem.
+Note that the authors provide pseudo code to develop an intuition in tackling this optimisation problem.
 
@@
 Note that the authors provide an analytical approach to estimate $\beta$ which is proportional to the changes of $\phi$; and for $\alpha$, they provide an approach utilising the Newton--Raphson method.
-Apart from that, the authors introduce the an issue of \textit{sparsity} in a corpus structure.
+Apart from that, the authors introduce an issue of \textit{sparsity} in the corpus structure.
 A sparse corpus is expected upon using a large vocabulary or a large amount of documents.
@@
 To be more specific, LDA has displayed balanced hidden topic proportions in the set of tested documents.
-That is, the tested documents has replicated the topic proportions of the training documents.
+That is, the tested documents have replicated the topic proportions of the training documents.
 The second experiment is directed towards the problem of document classification.
 The authors have addressed the \textit{dimensionality reduction} in order to surpass the performance of the support vector machine (SVM) models.
-To be more specific, the documents with reduced dimensionality would possess a lower amount of features.
+To be more specific, the documents with reduced dimensionality would possess lower amount of features.
 The results of the experiment indicate that the basic LDA model has successfully reduced the number of features.
@@
 Apart from that, the experiments introduced in this section have reflected the LDA performance during the time of the initial model release.
-The current state-of-the-hard topic modelling techniques overcome the shown results.
+Current state-of-the-art techniques overcome the shown results.
 Nevertheless, the carried experiments suggest the application fields of LDA-like models.
@@
 To be more specific, the article provides a readable introduction on the application of the Markov chain Monte Carlo (MCMC) algorithm.
-That is, the MCMC algorithm is applied as a sampling-based method to infer an underlying topic structure.
+That is, the MCMC algorithm is applied as a sampling-based method to infer an underlying topic mixture.
 Further, the authors discuss the method's application settings for the machine learning problems in vision.
-Also, the authors conduct an experiment on inferring the structure of topics in the abstracts of papers published in `Proceedings of the National Academy of Sciences' (PNAS).
-Finally, the paper is concluded by a discussion on assessing the sampling-based method usability.
+Also, the authors conduct an experiment on the topic inference in the abstracts of the research papers published in `Proceedings of the National Academy of Sciences' (PNAS).
+Finally, the paper is concluded by assessing  the sampling-based method's usability.
 To be more specific, the performance of the sampled-based method is compared to the performance of variational inference methods.
@@
 \end{equation}
-For a brief familiarisation with the notation used in the expression, note the following: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$; recall that $V$ is the number of words in the vocabulary and $K$ is the number of topics.
+For a brief familiarisation with the notation used in the expression, note the following terminology: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$, respectively; recall that $V$ is the number of words in the vocabulary and $K$ is the number of topics.
 Also, note that the counts do not include the last assignment of the topic, this is indicated by $-i$.
@@
 The performance of the model was compared to the variational inference based LDA models.
-Note that the performance was assessed in terms of \textit{perplexity} -- the uncertainty in assigning a word assignment to a topic.
-It has also been shown that Gibbs sampling displays superior performance in the small data sets, whereas the performance in larger data sets tends to fall off.
+Note that the performance was assessed in terms of \textit{perplexity} -- the uncertainty in assigning a word to a topic.
+It has also been shown that Gibbs sampling displays superior performance in small data sets, whereas the performance in larger data sets tends to fall off.
 Furthermore, the authors have assessed the performance upon varying the number of topics.
@@
 However, the model had not yet been optimised to take this notion of time-series into account upon the inference.
-Nevertheless, the authors have emphasised the necessity to assess the dynamic topic modelling.
+Nevertheless, the authors have emphasised the necessity to assess the dynamic topic modelling in the future.
 
@@
 That is, we review the initial paper on utilising time-series in LDA-like models -- `Dynamic Topic Models' \cite{blei_2006}.
-Note that by a \textit{static} model we mean that the assumed generative process does not take the development/change of the vocabulary terms into account, whereas in \textit{dynamic} topic modelling we assume the notion that the documents were generated in a sequential process.
+Note that by a \textit{static} model we mean that the assumed generative process does not take the development/change of the vocabulary terms into account, whereas in \textit{dynamic} topic modelling we assume that the documents were generated in a sequential process.
 That is, in dynamic topic modelling, the topics evolve in terms of their distribution over the vocabulary terms.
@@
 Even though we are inducing time series (which would suggest the use of \textit{continuous} variables), the authors use only \textit{categorical} (discrete) data.
-The distinction between continuous and categorical data is the following: continuous data is strictly numeric and infinite, whereas categorical data takes a fixed value from a finite set of values.
+The distinction between continuous and categorical data is the following: continuous data are strictly numeric and infinite, whereas categorical data take fixed values from a finite set of values.
 
@@
   \includegraphics[width=0.7\textwidth]{dynamic_topic_model}
-  \caption{Dynamic topic model. COPYRIGHT.}
+  \caption{The design of a dynamic topic model. \copyright Dynamic Topic Models [\ref{fig:dtm}].}
   \label{fig:dtm}
@@
 The larger variance values  induce the initial time-slice with larger fluctuations in the topic and vocabulary term distributions; and the smaller variance values in the upcoming time-slices induce the smoothness.
-Further, in lines 3--9, for every document in time-slice $t$ we execute the following procedure: for document $d$, we draw the topic distribution $\eta_d$ (the parameter is equivalent to the previously introduced $\theta_d$); then, in lines 5--7, for every word in document $d$ we draw topic assignment $z_{d, n}$ and, finally, we draw word $w_{t, d, n}$.
+Further, in lines 3--9, for every document in time-slice $t$, we execute the following procedure: for document $d$, we draw the topic distribution $\eta_d$ (the parameter is equivalent to the previously introduced $\theta_d$); then, in lines 5--7, for every word in document $d$ we draw topic assignment $z_{d, n}$ and, finally, we draw word $w_{t, d, n}$.
 Note that the means for the Multinomial distributions are normalised by the use of \textit{parametrisation}.
@@
 % - Evolution of Science
-\par The implemented dynamic topic model was evaluated by conducting an experiment on the journals of `Science'.
-The experiment is expected to infer the prevalent science fields and the prevalent vocabulary used in these fields.
+\par The implemented dynamic topic model has been evaluated by conducting an experiment on the journals of `Science'.
+The experiment was expected to infer the prevalent science fields and the prevalent vocabulary used in these fields.
 Note that the relevance of the science fields can be respectively obtained from the latent variables $\theta$ and $\phi$.
-That is, the authors would compare these variables throughout all time-slices in order to induce the reasoning over the prevalent science fields (by using $\theta$) and the prevalent vocabulary terms (by using $\phi$).
+That is, the authors would compare those latent variables throughout all time-slices in order to induce the reasoning over the prevalent science fields (by using $\theta$) and the prevalent vocabulary terms (by using $\phi$).
 The experiment has successfully established this notion of time-series.
@@
 \par In this section, we will set up the approach of the project's execution.
-To start with, we will go over the literature reviewed in Section 3 to derive the promising techniques in terms of applicability to MSI data sets.
+To start with, we will go over the literature reviewed in Section 3 in order to derive the promising techniques in terms of applicability to MSI data sets.
 Then, we will go over the high-level design and implementation requirements.
@@
 In the previous plots, we have visualised the occurrences of a particular entity within a corpus.
-The static generative process suggests a random distribution of the entities, whereas the dynamic generative process induces a particular pattern.
-Since we expect the ions of the same type to be gathered in clusters (rather than to be randomly distributed across the sample), the dynamic generative process is more appropriate to reflect the nature of the chemical entities patterns.
+The static generative process suggests a random distribution of the entity, whereas the dynamic generative process induces a particular pattern.
+Since we expect the ions of the same type to be gathered in clusters (rather than to be randomly distributed across the sample), the dynamic generative process is more appropriate to reflect the nature of the chemical entity patterns.
 
@@
 For example, if we assume that the data is generated in a 1-dimensional manner, then time-slice $t$ would depends on its adjacent time-slices $t-1$ and $t+1$.
-However, since the image of the sample is 2-dimensional, there are more adjacent time-slices.
+However, since the previous illustration of the sample is 2-dimensional, there are more adjacent time-slices.
 This relationship is shown in Figure \ref{fig:2-d_generative_process} below.
@@
 % -- Storing the ranges of ions
-\par In order to address the scalability of the model in terms of the large MSI data sets, we will address two techniques for the dimensionality reduction: utilisation of a dictionary and pre-processing of the data.
+\par In order to address the scalability of the model in terms of the large MSI data sets, we will present two techniques for the dimensionality reduction: utilisation of a dictionary and pre-processing of the data.
 Both of these techniques reduce the sparsity of the data sets.
@@
 \end{table}
-Note that, in these settings, the empty set symbol $\O$ represents a value which is not assigned in the memory; and the matrix and dictionary forms are respectively given on the left and right hand sides.
+Note that, in these settings, the empty set symbol $\O$ represents a value which is not assigned to the memory; and the matrix and dictionary forms are respectively given on the left and right hand sides.
 It follows that the use of dictionary requires about as half as much memory.
@@
 \par As discussed in Section 1, the captured ions of the same type are likely to have fluctuations in their masses.
-For this reason, we are applying another pre-processing step.
-By setting another threshold for the maximum fluctuation between the masses of the ions, we would group the ions which are likely to represent the same type; also, we would reduce the dimensionality of the data set.
+For this reason, we are applying another pre-processing step: by setting another threshold for the maximum fluctuation between the masses of the ions, we would group the ions which are likely to represent the same type.
+As a result, we would reduce the dimensionality of the data set.
 By taking the example of the previous paragraph and setting the fluctuation tolerance to $1$, we obtain the data structure given in Table \ref{tab:d_f_threshhold} below.
@@
 % 3-D data sets
-\par The generative process of the MSI data set, which was discussed in the discussion over the reviewed literature, has utilised 2-dimensional data sets.
+\par The generative process of MSI data sets, which was addressed in the discussion over the reviewed literature, has utilised 2-dimensional data sets.
 However, we are expecting to work on 3-dimensional data sets.

### 2017-01-13 docs/末legacy/report_generative-process/report_generative-process.tex
Organising the file structure.

--- 2016-12-01 docs/reports/report_generative-process/report_generative-process.tex
@@
 \usepackage{fancyheadings}
+\usepackage{hyperref}
 
@@
 \par During this week I have familiarised with the generative processes of static and dynamic LDA models.
-In this report I will present the results by plotting the generated documents, address the estimation of the parameters $\alpha$ and $\beta$, and raise questions to be considered for the next meeting.
+In this report I will present the results by plotting the generated corpora, address the initialisation of the parameters $\alpha$ and $\beta$, and raise questions for the next meeting.
 
@@
 
-In Figure 1 we can see that the word is present throughout all the corpus, whereas Figure 2 displays a smoother distribution of the word.
+In Figure 1 we can see that the word is present throughout the corpus, whereas Figure 2 displays a smoother distribution of the word.
 That is, the dynamic generative process induces a structure of the word's existence in the corpus.
@@
 
-The implementation of the generative processes follow directly from the original LDA paper \cite{Blei:2003:LDA:944919.944937} and the paper on dynamic topic models \cite{Blei:2006:DTM:1143844.1143859}.
+The implementation of the generative processes follows directly from the original LDA paper \cite{Blei:2003:LDA:944919.944937} and the paper on dynamic topic models \cite{Blei:2006:DTM:1143844.1143859}.
+The Python Notebook containing the implementation of the processes can be found on the following link: \url{https://github.com/perdaug/mlinb/blob/master/notebooks/generating_corpus.ipynb}.
+Note that the initial values of the parameter $\alpha$ are positive and random, whereas the initial values of the parameter $\beta$ are equal and normalised for each topic.
+This formulation of $\beta$ suggests that each word has an equal probability of getting drawn in the first document generation of the dynamic model and throughout all generated documents of the static model.
+Also, note that every generated document in the dynamic model is unique with respect to other documents.
+In other words, each dynamic time-slice contains one document.
+
+During the upcoming meeting I would like to address the following questions:
+\begin{enumerate}
+\item Should we use the variational inference for the implementation of the dynamic topic model? If not, would Gibbs sampling scale with the future datasets?
+\item Why Dirichlet distribution is amenable in the dynamic topic modelling?
+(This is claimed in the last paragraph of the first page of the dynamic topic modelling paper \cite{Blei:2006:DTM:1143844.1143859})
+\item For the dynamic model implementation, should we update $\alpha_{t-1}$ (as discussed in Subsection 5.3 of the original LDA paper \cite{Blei:2003:LDA:944919.944937}) before drawing $\alpha_t$? 
+\end{enumerate}
 

### 2017-01-13 code/create_corpus.py

@@ -1,107 +1,0 @@
-"""
-Creating the vocabulary.
-
-Terms:
-    mni -- mass and intensity.
-"""
-
-import math
-import pymzml
-import pandas as pd
-import numpy as np
-np.set_printoptions(precision=3, threshold=np.inf)
-
-INTENSITY_FLOOR = 10
-TOLERANCE_FACTOR = 7
-DATA_PATH = "../data/"
-DATA_FILENAME = "abcdefgh_1.mzML"
-
-def classify(classes, feature):
-    for class_ in classes:
-        if feature >= class_[0] and feature <= class_[1]:
-            return str(class_)
-
-def generate_coordinates(data_size, number_of_rows, row_length, column_length):
-    single_scan_size = (data_size + 1) / number_of_rows
-    column_size = math.floor(single_scan_size \
-        * (column_length / (row_length + column_length)))
-    row_size = single_scan_size - column_size 
-    coordinates = []
-
-    x_coord = 0
-    y_coord = 0
-    row_scan = True
-    scanning_right = True
-    for _ in range(0, number_of_rows):
-        for i in range(1, single_scan_size + 1):
-            coordinates.append((x_coord, y_coord))
-            # Switching the horizontal scanning to the vertical scanning.
-            if i % row_size == 0:
-                row_scan = not row_scan
-            if row_scan:
-                if scanning_right:
-                    x_coord += 1
-                else:
-                    x_coord -= 1
-            else:
-                y_coord += 1
-        # Reversing scanner's direction.
-        row_scan = not row_scan
-        scanning_right = not scanning_right
-    return coordinates
-
-
-def main():
-    run = pymzml.run.Reader(DATA_PATH + DATA_FILENAME)
-
-    # Iterating through the data and storing into a temporary array (`Pymzml Reader' has no length attribute)
-    preprocessed_mnis = []  
-    for spectrum in run:
-        for mass, intensity in spectrum.peaks:
-            if intensity > INTENSITY_FLOOR:
-                preprocessed_mnis.append((mass, intensity, spectrum['id']))
-    mnis_dtype = [('mass', float), ('intensity', float), ('id', int)]
-
-    # Using numpy to sort the data.
-    mnis = np.array(preprocessed_mnis, dtype=mnis_dtype)
-    mnis.sort(order='mass') 
-    
-    # Defining mass ranges for words.
-    words = []
-    starting_class_mass = mnis[0][0]
-    previous_mass = mnis[0][0]
[…]

### 2017-01-13 code/draw_topic_histograms.py

@@ -1,28 +1,0 @@
-import os
-import pandas as pd
-import matplotlib.pyplot as plt
-import numpy as np
-import plotly.plotly as py
-
-HOME_PATH = os.path.expanduser('~')
-
-
-def main():
-    topic = pd.read_pickle('../pickles/topic_0.pickle')
-    topic = topic.to_dict()
-
-    # print(topic)
-    print np.random.randn(1000)
-    exit()
-
-    plt.hist(topic.items())
-    plt.title("Gaussian Histogram")
-    plt.xlabel("Coordinate")
-    plt.ylabel("% Distribution")
-
-    fig = plt.gcf()
-
-    plot_url = py.plot_mpl(fig, filename='mpl-basic-histogram')
-
-if __name__ == '__main__':
-    main()

### 2017-01-13 code/pre-processing/create_corpus.py

@@ -1,0 +1,101 @@
+
+import math
+import pymzml
+import pandas as pd
+import numpy as np
+import os
+np.set_printoptions(precision=3, threshold=np.inf)
+
+INTENSITY_FLOOR = 10
+TOLERANCE_FACTOR = 7
+HOME_PATH = os.path.expanduser('~')
+DATA_PATH = HOME_PATH + '/Projects/ssmsi/data/'
+PICKLES_PATH = HOME_PATH + '/Projects/ssmsi/heavy_pickles/'
+
+def assign_class(classes, feature):
+	for class_assignment in classes:
+		if feature >= class_assignment[0] and feature <= class_assignment[1]:
+			return str(class_assignment)
+
+def generate_coordinates(data_size, number_of_rows, row_length, column_length):
+	single_scan_size = (data_size + 1) / number_of_rows
+	column_size = math.floor(single_scan_size \
+		* (column_length / (row_length + column_length)))
+	row_size = single_scan_size - column_size
+	coordinates = []
+
+	x_coord = 0
+	y_coord = 0
+	row_scan = True
+	scanning_right = True
+	for _ in range(0, number_of_rows):
+		for i in range(1, single_scan_size + 1):
+			coordinates.append((x_coord, y_coord))
+			# Switching the horizontal scanning to the vertical scanning.
+			if i % row_size == 0:
+				row_scan = not row_scan
+			if row_scan:
+				if scanning_right:
+					x_coord += 1
+				else:
+					x_coord -= 1
+			else:
+				y_coord += 1
+		# Reversing scanner's direction.
+		row_scan = not row_scan
+		scanning_right = not scanning_right
+	return coordinates
+
+def main():
+	run = pymzml.run.Reader(DATA_PATH + 'abcdefgh_1.mzML')
+
+	# Reading the data
+	preprocessed_mnis = []
+	for spectrum in run:
+		for mass, intensity in spectrum.peaks:
+			if intensity > INTENSITY_FLOOR:
+				preprocessed_mnis.append((mass, intensity, spectrum['id']))
+	mnis_dtype = [('mass', float), ('intensity', float), ('id', int)]
+
+	# Sorting the data.
+	mnis = np.array(preprocessed_mnis, dtype=mnis_dtype)
+	mnis.sort(order='mass')
+
+	# Defining mass ranges for words.
+	words = []
+	starting_class_mass = mnis[0][0]
+	previous_mass = mnis[0][0]
[…]

### 2017-01-13 code/pre-processing/create_vocab.py

@@ -1,0 +1,48 @@
+import os
+import pandas as pd
+import numpy as np
+
+HOME_PATH = os.path.expanduser('~')
+PICKLES_PATH = HOME_PATH + '/Projects/ssmsi/heavy_pickles/'
+
+
+def create_vocab(corpus):
+	vocab = []
+	for doc in corpus:
+		for word in corpus[doc]:
+			if word not in vocab:
+				vocab.append(word)
+	vocab_panda = pd.Series(vocab)
+	vocab_panda.to_pickle(HOME_PATH + PICKLES_PATH + 'vocab.pickle')
+	return vocab
+
+# Smaller corpus for Gibbs sampling
+def preprocess_corpus(corpus, vocab, factor):
+	pp_corpus = np.zeros((len(corpus), len(vocab)), dtype=np.int64)
+	for doc in corpus:
+		# Finding maximum intensity.
+		max_intensity = 0
+		for word in corpus[doc]:
+			if corpus[doc][word] > max_intensity:
+				max_intensity = corpus[doc][word]
+
+		# Populating the preprocessed corpus.
+		for w, word in enumerate(vocab):
+			if word in corpus[doc]:
+				d = int(doc) - 1
+				normalised_intensity = int(corpus[doc][word] / \
+					(max_intensity / factor))
+				pp_corpus[d][w] = normalised_intensity
+
+	pp_corpus.dump(HOME_PATH + '/Projects/tminm/heavy_pickles/' + 'preprocessed_corpus.pickle')
+
+def main():
+	corpus_series = pd.read_pickle(HOME_PATH + '/Projects/tminm/heavy_pickles/corpus.pickle')
+	corpus = corpus_series.to_dict()
+
+	vocab = create_vocab(corpus)
+	# preprocess_corpus(corpus, vocab, factor=20)
+
+if __name__ == '__main__':
+	main()
+	

### 2017-01-13 code/preprocessing.py

@@ -1,47 +1,0 @@
-import os
-import pandas as pd
-import numpy as np
-
-HOME_PATH = os.path.expanduser('~')
-
-def create_vocab(corpus):
-    vocab = []
-    for doc in corpus:
-        for word in corpus[doc]:
-            if word not in vocab:
-                vocab.append(word)
-    vocab_panda = pd.Series(vocab)
-    vocab_panda.to_pickle(HOME_PATH + '/Projects/tminm/heavy_pickles/' + 'vocab.pickle')
-    print(vocab)
-    exit()
-    return vocab
-
-def preprocess_corpus(corpus, vocab, factor):
-    pp_corpus = np.zeros((len(corpus), len(vocab)), dtype=np.int64)
-    for doc in corpus:
-        # Finding maximum intensity.
-        max_intensity = 0
-        for word in corpus[doc]:
-            if corpus[doc][word] > max_intensity:
-                max_intensity = corpus[doc][word]
-
-        # Populating the preprocessed corpus.
-        for w, word in enumerate(vocab):
-            if word in corpus[doc]:
-                d = int(doc) - 1
-                normalised_intensity = int(corpus[doc][word] / \
-                    (max_intensity / factor))
-                pp_corpus[d][w] = normalised_intensity
-
-    pp_corpus.dump(HOME_PATH + '/Projects/tminm/heavy_pickles/' + 'preprocessed_corpus.pickle')
-
-def main():
-    corpus_series = pd.read_pickle(HOME_PATH + '/Projects/tminm/heavy_pickles/corpus.pickle')
-    corpus = corpus_series.to_dict()
-
-    vocab = create_vocab(corpus)
-    preprocess_corpus(corpus, vocab, factor=20)
-
-if __name__ == '__main__':
-    main()
-    

### 2017-01-20 docs/msci-proposal/mprop.tex
Renaming the MSci project.

--- 2016-12-18 docs/MSci_proposal/mprop.tex
@@
 Effectively, a metabolite is compounded with an appropriate chemical entity to produce an ion.
-It follows that the mass spectrometers are capable to capture some properties of the ionised metabolites.
+It follows that the mass spectrometers are capable of capturing some properties of the ionised metabolites.
 Since the mass of an ion belongs to these properties, the collection of captured ions can be represented in terms of mass spectra.
@@
 That is, we have no initial settings in how the patterns are expected to look like.
-However, upon appropriate optimisation, the unsupervised machine learning techniques are capable to learn the patterns on their own.
+However, upon appropriate optimisation, the unsupervised machine learning techniques are capable of learning the patterns on their own.
 To be more specific, we will utilise the techniques of \textit{topic modelling} (in this case, a pattern is referred by a topic).
@@
 First of all, the successor LDA-like models are treated as a state-of-the-art method in tackling semantic analysis type problems.
-For more detailed review on LDA applications, the reader can consult the survey conducted by Blei \cite{blei_2012}.
+For a more detailed review of LDA applications, the reader can consult the survey conducted by Blei \cite{blei_2012}.
 Also, Hooft et al.
@@
 % - Sections
-\par The structure of this proposal is given as follows: in Section 2 we present the issue of noise in MSI data sets; in section 3 we cover the literature in topic modelling which focuses on the variations of LDA-like models; in Section 4 we discuss the approach to tackle the problem (we discuss the reviewed literature, address design and evaluation details); finally, in Section 5, we set and schedule the tasks required to complete the project.
+\par The structure of this proposal is given as follows: in Section 2 we present the issue of noise in MSI data sets; in section 3 we cover the literature in topic modelling which focuses on the variations of LDA-like models; in Section 4 we discuss the approach tackling the noise reduction problem (we discuss the reviewed literature, address design and evaluation details); finally, in Section 5, we set and schedule the tasks required to complete the project.
 % To-do: rewrite the outline
@@
 The authors have addressed the \textit{dimensionality reduction} in order to surpass the performance of the support vector machine (SVM) models.
-To be more specific, the documents with reduced dimensionality would possess lower amount of features.
+To be more specific, the documents with reduced dimensionality would possess a lower amount of features.
 The results of the experiment indicate that the basic LDA model has successfully reduced the number of features.
@@
 The implementation details of these methods can be found in the `Dynamic Topic Modelling' article \cite{blei_2006}.
-Note that the implementation details are provided in a brief manner suggesting the need of supplementary literature.
+Note that the implementation details are provided in a brief manner suggesting the need for supplementary literature.
 
@@
 
-The initial dynamic topic modelling paper suggests to induce the time-series into the generative process of data sets.
+The initial dynamic topic modelling paper suggests inducing the time-series into the generative process of data sets.
 However, the way we capture the ions is done instantaneously for the whole corpus.
@@
 Instead of that, we will use the dynamic topic modelling techniques to establish the relationship between the time-slices (even though we are not taking time into account, we will stick to the same terminology to preserve the consistency).
-For example, if we assume that the data is generated in a 1-dimensional manner, then time-slice $t$ would depends on its adjacent time-slices $t-1$ and $t+1$.
+For example, if we assume that the data is generated in a 1-dimensional manner, then time-slice $t$ would depend on its adjacent time-slices $t-1$ and $t+1$.
 However, since the previous illustration of the sample is 2-dimensional, there are more adjacent time-slices.
@@
 However, for the dynamic topic models, the latter method has to be reconsidered due to the change from Dirichlet to Gaussian distributions.
-Therefore, we will look into another MCMC based algorithm -- Metropolis--Hastings.
+Therefore, during the implementation of the model, we will attempt to utilise another MCMC-based algorithm -- Metropolis--Hastings.
 % - To-do: Analyse the details of the MH.
@@
 % - Risk management
-\par The developed technique addressing the spatial smoothing of the MSI data sets will evaluated by measuring the improvement in performance.
+\par The developed technique for the spatial smoothing of the MSI data sets will be evaluated by measuring the improvement in performance.
 That is, the performance will be assessed by investigating the quality of the inferred patterns.
-Also, we will utilise two different inference methods: sampling-based and variational.
-By having two model variations, we will be able to compare the time performance.
+Also, since we plan to develop two models with different inference methods, we will be able to assess the time performance with respect to the inference method used.
 Hence, we will be able to deduce which variation scales better with the size of the data sets.
-Finally, the section will be finished by investigating the possible risks; also, we will provide potential strategies mitigating the risks.
+Furthermore, in this section, we will assess the possible risks; also, we will provide potential strategies for mitigating the risks.
 
@@
 % -- The improvement in the inference
-% 
-\par The implemented model will be evaluated using synthetic and real data sets.
-However, at the start, we will use the synthetic data sets.
-Note that the performance will be measured in the following way: (1) take two equivalent data sets; (2) apply the spatial smoothing on one of the data sets; (3) infer the underlying structure of the data; (4) check whether the applied smoothing has inferred the underlying structures better.
-By working with synthetic data sets, we had the control over inducing the underlying structure of the data.
-Hence, we will able to give a basis in measuring the performance of the applied spatial smoothing.
+\par The model will be evaluated using synthetic and real data sets.
+However, at the start, we will only use the synthetic data sets.
+Note that the performance will be measured using the following algorithm: (1) take two equivalent data sets; (2) apply the spatial smoothing on one of the data sets; (3) infer the underlying structure of the data; (4) check whether the applied smoothing has inferred the underlying structure better.
+By working with synthetic data sets, we have the control over inducing the underlying structure of the data.
+Hence, we will able to give a basis for measuring the performance of the applied spatial smoothing.
 
@@
 % - Uncertainty whether the model produces an improvement
-\par Even though we have induced a strategy for the project's evaluation, the described process assumes that we can mitigate the potential risks.
-Note that, at the initial stage of the project, we have identified the following potential risks:
+\par Even though we have induced a strategy for the project's evaluation, the described process assumes that we can mitigate the following risks:
 \begin{itemize}
@@
 \end{itemize}
-The first three risks would slow down the project execution thus raise a concern of finishing the project on time, whereas the fourth one would prevent the confirmation of denial of the initial project's hypothesis -- `the degeneracies of MSI data sets can be detected using LDA-like models'.
+The first three risks would slow down the project execution thus raise a concern of finishing the project on time, whereas the fourth one would prevent deriving the confirmation or the denial of the initial project's hypothesis -- `the degeneracies of MSI data sets can be detected using LDA-like models'.
 
@@
 Also, note that the main model will firstly be developed using Gibbs sampling, as this method is easier to implement compared to the variational one.
+However, as previously discussed, we intend to exchange the use of the Gibbs sampling method by utilising the Metropolis--Hastings method.
 
@@
 % - Real data can be obtained from the Web
-\par The risk of the real data access might be partially resolved, as the University has a laboratory for the research in Metabolomics.
+\par The risk of the real data access might be partially resolved, as the University has a laboratory for the research in metabolomics.
 However, in case this options is not possible, we can obtain the data sets from the Web.
@@
 \par Finally, the risk of the inability to measure the performance of the applied smoothing can impact the outcome of the project the most.
-That is, it would be difficult to reason about the improvement in the inference of the real data sets, because we do not possess the full knowledge of the data set underlying structure.
+That is, it would be difficult to reason about the improvement in the inference of the real data sets because we do not possess the full knowledge of the data sets' underlying structure.
 We will mitigate this issue by establishing a performance metric from the synthetic data sets.
@@
 \par By having the defined boundaries of the expected improvement in performance, we can attempt to tackle the real data sets.
-That is, the expected performance improvement would serve as an indicator whether the inferred real data deviates in the expected limit.
-To be more specific, the data sets with applied spatial and the data sets without applied smoothing will display some differences in the inferred topics.
+That is, the expected performance improvement would serve as an indicator whether the inferred real data deviates in the expected limit of improvement.
+To be more specific, the data sets with applied spatial smoothing and the data sets without applied spatial smoothing will display some differences in the inferred topics.
 Hence, we will check whether those differences meet the performance boundaries suggested by the inference of the synthetic data sets.
@@
 \par In this section, we will outline the project milestones and deliverables.
-That is, we will outline the most important steps and provide the estimations in a form of Gantt chart.
-Further, the deliverables will be directly induced from the project's objectives and hypothesis (these are addressed in Section 2).
+That is, we will outline the most important steps and provide the estimations in the form of the Gantt chart.
+Further, the deliverables will be induced from the project's objectives and hypothesis (these are addressed in Section 2).
 The deliverables will contain both successful and successful outcomes of the set hypothesis.
@@
 Further, the evaluation will be conducted by assessing the performance of the developed models.
-Finally, the results will be documented in a form of weekly reports, dissertation, and presentation.
+Finally, the results will be documented in the forms of weekly reports, the dissertation, and the presentation.
 
@@
 % - Estimations
-\par The outlined tasks will not be necessarily executed in sequential manner. For some of the tasks, we will use the parallel approach as given in Figure \ref{fig:gantt} below.
+\par The outlined tasks will not be necessarily executed in a sequential manner. For some of the tasks, we will use the parallel approach as given in Figure \ref{fig:gantt} below.
 
@@
 Nevertheless, in any of the project's outcomes, we will provide the following deliverables: the presentation; the dissertation; and the spatial smoothing model, which will be written in Python.
-However, the main project's contribution to the research in MSI will depend on the results: if the model displays an improvement in performance, then we could utilise the model in pre-processing MSI data sets; in the case of unsuccessful outcome, we would assess the techniques used and provide details on why the model does not work.
+However, the main project's contribution to the research in MSI will depend on the results: if the model displays an improvement in performance, then we could utilise the model in the pre-processing of MSI data sets; in the case of unsuccessful outcome, we would assess the techniques used and provide details on why the model does not work.
 Both outcomes would bring the motivation for further research in improving MSI data sets: the successful outcome would bring attention to the applicability of LDA-like models; the unsuccessful outcome would suggest the need of a different perspective required to tackle the issue of noise.

### 2017-01-20 code/ldas/gibbs_lda.py
Fixing the thetas update.

@@ -1,93 +1,99 @@
 import os
 import pandas as pd
 import numpy as np
-
-HOME_PATH = os.path.expanduser('~')
 np.set_printoptions(edgeitems=5)
 
 
-class Gibbs_LDA(object):
+class LDA_Gibbs(object):
 
-    def __init__(self, K, iter_count):
-        self.iter_count = iter_count
-        self.K = K
-        self.thetas_ = []
+	def __init__(self, K, iter_count):
+		self.iter_count = iter_count
+		self.K = K
+		self.thetas = []
 
-    def fit(self, corpus):
-        self._fit(corpus)
+	def fit(self, corpus):
+		self._fit(corpus)
 
-    def dump_thetas(self, path):
-        pickle = np.array(self.thetas_)
-        pickle.dump(path)
+	def dump_thetas(self, path):
+		pickle = np.array(self.thetas)
+		pickle.dump(path)
 
-    def _fit(self, corpus):
-        self._initialise(corpus)
+	def _fit(self, corpus):
+		self._initialise(corpus)
 
-        for it in range(self.iter_count):
-            print("Iteration %s:" % it) 
-            self._sample_topics(corpus)
-            
-            self._store_theta()
-            print("Current theta:")
-            print(self.thetas_[it]) 
+		for it in range(self.iter_count):
+			print("Iteration %s:" % it)
+			self._sample_topics(corpus)
+			self._store_theta()
+			print("Current theta:")
+			print(self.thetas[it])
 
-    def _initialise(self, corpus):
-        self.D, self.W = corpus.shape
-        self._total_frequencies = corpus.sum()
-        self.alpha = 1. / self.K
-        self.beta = 1. / self.W
+	def _initialise(self, corpus):
+		self.D, self.W = corpus.shape
+		self._total_frequencies = corpus.sum()
+		self.alpha = 1. / self.K
+		self.beta = 1. / self.W
 
-        self._zw_counts = np.zeros((self.K, self.W), dtype=np.int)
-        self._dz_counts = np.zeros((self.D, self.K), dtype=np.int)
-        self._z_counts = np.zeros(self.K, dtype=np.int)
-        self._z_history = []
+		self._zw_counts = np.zeros((self.K, self.W), dtype=np.int)
+		self._dz_counts = np.zeros((self.D, self.K), dtype=np.int)
+		self._z_counts = np.zeros(self.K, dtype=np.int)
+		self._z_history = []
 
[…]

### 2017-01-20 code/ldas/lda_gibbs.py
Fixed the plotting of the thetas.

@@ -3,6 +3,10 @@ import pandas as pd
 import numpy as np
 np.set_printoptions(edgeitems=5)
 
+'''
+References:
+- Griffiths and Stevyers (2004)
+'''
 
 class LDA_Gibbs(object):
 
@@ -59,9 +63,7 @@ class LDA_Gibbs(object):
 					self._zw_counts[z, w] -= 1
 					self._z_counts[z] -= 1
 
-					# The second denominator from eq. 5 (Griffiths and Stevyers
-					# 2004) is omitted because it is a constant and we normalise
-					# the vector p_topic anyway.
+					# Eq. 5, the second denominator is  a constant
 					p_topic = ((1.0 * self._zw_counts[:, w] + self.beta) / \
 						(1.0 * self._z_counts + self.W * self.beta)) * \
 						((self._dz_counts[d, :] + self.alpha))
@@ -87,10 +89,11 @@ class LDA_Gibbs(object):
 
 HOME_PATH = os.path.expanduser('~') + '/Projects/ssmsi/'
 DATA_PATH = HOME_PATH + 'pickles/corpora/a-to-h/'
-OUT_PATH = HOME_PATH + 'pickles/corpora/results/lda_gibbs/'
+OUT_PATH = HOME_PATH + 'pickles/results/lda_gibbs/'
 
 def main():
-	lda_gibbs = LDA_Gibbs(K=10, iter_count=10)
+	lda_gibbs = LDA_Gibbs(K=10, iter_count=50)
+	# To-do: Fix the main corpus (the number of col is arbitrary) 
 	corpus = pd.read_pickle(DATA_PATH + 'a-to-h_corpus_r.pickle')
 	lda_gibbs.fit(corpus)
 	lda_gibbs.dump_thetas(OUT_PATH + 'gibbs_thetas.pickle')

### 2017-01-21 code/pre-processing/create_vocab.py
Creating the notebooks to run the models; Creating a separate file for the corpus pre-processing; Redefining the file naming convention.

@@ -1,6 +1,5 @@
 import os
 import pandas as pd
-import numpy as np
 
 HOME_PATH = os.path.expanduser('~')
 PICKLES_PATH = HOME_PATH + '/Projects/ssmsi/heavy_pickles/'
@@ -13,35 +12,15 @@ def create_vocab(corpus):
 			if word not in vocab:
 				vocab.append(word)
 	vocab_panda = pd.Series(vocab)
-	vocab_panda.to_pickle(HOME_PATH + PICKLES_PATH + 'vocab.pickle')
-	return vocab
+	return vocab_panda
 
-# Smaller corpus for Gibbs sampling
-def preprocess_corpus(corpus, vocab, factor):
-	pp_corpus = np.zeros((len(corpus), len(vocab)), dtype=np.int64)
-	for doc in corpus:
-		# Finding maximum intensity.
-		max_intensity = 0
-		for word in corpus[doc]:
-			if corpus[doc][word] > max_intensity:
-				max_intensity = corpus[doc][word]
-
-		# Populating the preprocessed corpus.
-		for w, word in enumerate(vocab):
-			if word in corpus[doc]:
-				d = int(doc) - 1
-				normalised_intensity = int(corpus[doc][word] / \
-					(max_intensity / factor))
-				pp_corpus[d][w] = normalised_intensity
-
-	pp_corpus.dump(HOME_PATH + '/Projects/tminm/heavy_pickles/' + 'preprocessed_corpus.pickle')
 
 def main():
 	corpus_series = pd.read_pickle(HOME_PATH + '/Projects/tminm/heavy_pickles/corpus.pickle')
 	corpus = corpus_series.to_dict()
 
-	vocab = create_vocab(corpus)
-	# preprocess_corpus(corpus, vocab, factor=20)
+	vocab_panda = create_vocab(corpus)
+	vocab_panda.to_pickle(HOME_PATH + PICKLES_PATH + 'vocab.pickle')
 
 if __name__ == '__main__':
 	main()

### 2017-01-21 code/pre-processing/pre-process_corpus.py

@@ -1,0 +1,53 @@
+import os
+import pandas as pd
+import numpy as np
+
+# Smaller corpus for Gibbs sampling
+def preprocess_corpus(corpus, vocab, factor):
+	pp_corpus = np.zeros((len(corpus), len(vocab)), dtype=np.int64)
+	for doc in corpus:
+		# Finding maximum intensity.
+		max_intensity = 0
+		for word in corpus[doc]:
+			if corpus[doc][word] > max_intensity:
+				max_intensity = corpus[doc][word]
+
+		# Populating the preprocessed corpus.
+		for w, word in enumerate(vocab):
+			if word in corpus[doc]:
+				d = int(doc) - 1
+				normalised_intensity = int(corpus[doc][word] / \
+					(max_intensity / factor))
+				pp_corpus[d][w] = normalised_intensity
+	return pp_corpus
+
+
+from optparse import OptionParser
+op = OptionParser()
+op.add_option("--source",
+							action="store", type=str,
+							help="The choice of data origins.")
+op.add_option("--corpus",
+							action="store", type=str,
+							help="The choice of a data set.")
+(opts, args) = op.parse_args()
+
+
+HOME_PATH = os.path.expanduser('~') + '/Projects/ssmsi/'
+DATA_PATH = HOME_PATH + 'pickles/corpora/' + opts.source + '/'
+AUX_DATA_PATH = HOME_PATH + 'pickles/corpora/a-to-h/'
+INPUT_FILE_NAME = opts.source + '_corpus_' + opts.corpus + '.pickle'
+OUT_PATH = DATA_PATH
+OUTPUT_FILE_NAME = opts.source + '_corpus_' + opts.corpus + '_pp.pickle'
+
+def main():
+	corpus_series = pd.read_pickle(DATA_PATH + INPUT_FILE_NAME)
+	corpus = corpus_series.to_dict()
+	vocab_pickle = pd.read_pickle(AUX_DATA_PATH + 'a-to-h_vocab.pickle')
+	vocab = vocab_pickle.tolist()
+
+	pp_corpus = preprocess_corpus(corpus, vocab, factor=20)
+	pp_corpus.dump(OUT_PATH + OUTPUT_FILE_NAME)
+
+if __name__ == '__main__':
+	main()

### 2017-01-21 code/topic-models/dtm_alpha-only.py

@@ -5,10 +5,10 @@ np.set_printoptions(edgeitems=5)
 
 '''
 References:
-- Griffiths and Stevyers (2004)
+- Blei and Lafferty (2006)
 '''
 
-class LDA_Gibbs(object):
+class DTM_alpha(object):
 
 	def __init__(self, K, iter_count):
 		self.iter_count = iter_count
@@ -92,8 +92,8 @@ DATA_PATH = HOME_PATH + 'pickles/corpora/a-to-h/'
 OUT_PATH = HOME_PATH + 'pickles/results/lda_gibbs/'
 
 def main():
-	lda_gibbs = LDA_Gibbs(K=10, iter_count=50)
-	# To-do: Fix the main corpus (the number of col is arbitrary) 
+	lda_gibbs = DTM_alpha(K=10, iter_count=50)
+	# To-do: Fix the main corpus (the number of col is arbitrary)
 	corpus = pd.read_pickle(DATA_PATH + 'a-to-h_corpus_r.pickle')
 	lda_gibbs.fit(corpus)
 	lda_gibbs.dump_thetas(OUT_PATH + 'gibbs_thetas.pickle')

### 2017-01-21 code/topic-models/lda_gibbs.py

@@ -1,0 +1,116 @@
+import os
+import pandas as pd
+import numpy as np
+np.set_printoptions(edgeitems=5)
+
+'''
+References:
+- Griffiths and Stevyers (2004)
+'''
+
+class LDA_Gibbs(object):
+
+	def __init__(self, K, iter_count):
+		self.iter_count = iter_count
+		self.K = K
+		self.thetas = []
+
+	def fit(self, corpus):
+		self._fit(corpus)
+
+	def dump_thetas(self, path):
+		pickle = np.array(self.thetas)
+		pickle.dump(path)
+
+	def _fit(self, corpus):
+		self._initialise(corpus)
+
+		for it in range(self.iter_count):
+			print("Iteration %s:" % it)
+			self._sample_topics(corpus)
+			self._store_theta()
+			print("Current theta:")
+			print(self.thetas[it])
+
+	def _initialise(self, corpus):
+		self.D, self.W = corpus.shape
+		self._total_frequencies = corpus.sum()
+		self.alpha = 1. / self.K
+		self.beta = 1. / self.W
+
+		self._zw_counts = np.zeros((self.K, self.W), dtype=np.int)
+		self._dz_counts = np.zeros((self.D, self.K), dtype=np.int)
+		self._z_counts = np.zeros(self.K, dtype=np.int)
+		self._z_history = []
+
+		for d, doc in enumerate(corpus):
+			self._z_history.append([])
+			for w, word in enumerate(doc):
+				self._z_history[d].append([])
+				word_frequency = int(word)
+				for _ in range(word_frequency):
+					z = np.random.randint(self.K)
+					self._z_history[d][w].append(z)
+					self._dz_counts[d, z] += 1
+					self._zw_counts[z, w] += 1
+					self._z_counts[z] += 1
+
+	def _sample_topics(self, corpus):
+		for d, doc in enumerate(corpus):
+			for w, _ in enumerate(doc):
+				for it, z in enumerate(self._z_history[d][w]):
+					self._dz_counts[d, z] -= 1
+					self._zw_counts[z, w] -= 1
+					self._z_counts[z] -= 1
+
+					# Eq. 5, the second denominator is  a constant
+					p_topic = ((1.0 * self._zw_counts[:, w] + self.beta) / \
+						(1.0 * self._z_counts + self.W * self.beta)) * \
+						((self._dz_counts[d, :] + self.alpha))
[…]

### 2017-01-21 code/末legacy/d_gibbs_lda.py

@@ -8,119 +8,119 @@ np.set_printoptions(edgeitems=5)
 
 class Dynamic_Gibbs_LDA(object):
 
-    def __init__(self, K, iter_count, n_seq):
-        self.iter_count = iter_count
-        self.K = K
-        self.n_seq = n_seq
-        self.thetas_ = []
-        self.phis_ = []
-        self.alphas_ = []
+	def __init__(self, K, iter_count, n_seq):
+		self.iter_count = iter_count
+		self.K = K
+		self.n_seq = n_seq
+		self.thetas_ = []
+		self.phis_ = []
+		self.alphas_ = []
 
-    def fit(self, corpus):
-        self._fit(corpus)
+	def fit(self, corpus):
+		self._fit(corpus)
 
-    def dump(self, array, path):
-        pickle = np.array(array)
-        pickle.dump(path)
+	def dump(self, array, path):
+		pickle = np.array(array)
+		pickle.dump(path)
 
-    def _fit(self, corpus):
-        list_of_seq_corpus = np.array_split(corpus, self.n_seq)
+	def _fit(self, corpus):
+		list_of_seq_corpus = np.array_split(corpus, self.n_seq)
 
-        for seq_c, seq_corpus in enumerate(list_of_seq_corpus):
+		for seq_c, seq_corpus in enumerate(list_of_seq_corpus):
 
-            self._initialise(seq_corpus, seq_c)
-            print(self.alpha.shape)
-            print(self.alpha)
-            print(self.alpha.sum())
-            print(self.beta.shape)
-            print(self.beta[0, :])
-            print(self.beta[0, :].sum())
+			self._initialise(seq_corpus, seq_c)
+			print(self.alpha.shape)
+			print(self.alpha)
+			print(self.alpha.sum())
+			print(self.beta.shape)
+			print(self.beta[0, :])
+			print(self.beta[0, :].sum())
 
-            for it in range(self.iter_count):
-                
-                print("Sequence: %s, Iteration: %s." % (seq_c, it)) 
-                self._sample_topics(seq_corpus)
-                self._store_theta()
-                self._store_phi()
-                print("Current theta:")
-                print(self.thetas_[it]) 
+			for it in range(self.iter_count):
+				
+				print("Sequence: %s, Iteration: %s." % (seq_c, it)) 
+				self._sample_topics(seq_corpus)
+				self._store_theta()
[…]

### 2017-01-22 code/topic-models/dtm_alpha-only.py
Implementing the barebones for the alpha update.

@@ -10,10 +10,16 @@ References:
 
 class DTM_alpha(object):
 
-	def __init__(self, K, iter_count):
+	def __init__(self, K, iter_count, t_slices):
 		self.iter_count = iter_count
 		self.K = K
+		self.t_slices = t_slices
 		self.thetas = []
+		self.alpha_stars = []
+		self.alphas = []
+		self.deviation_a_basic = 0.1
+		self.deviation_a_guess = 3
+		self.r_rates = []
 
 	def fit(self, corpus):
 		self._fit(corpus)
@@ -23,27 +29,37 @@ class DTM_alpha(object):
 		pickle.dump(path)
 
 	def _fit(self, corpus):
-		self._initialise(corpus)
+		corpus_partitions = np.array_split(corpus, self.t_slices)
+		for idx_p, corpus_p in enumerate(corpus_partitions):
+			self._initialise(corpus_p, idx_p)
+			for it in range(self.iter_count):
+				print("Time-slice: %s, Iteration: %s." % (idx_p, it))
+				self._sample_topics(corpus_p)
+				self._store_theta()
+			self._dynamic_step()
+		print self.r_rates
+		print sum(self.r_rates)/5
 
-		for it in range(self.iter_count):
-			print("Iteration %s:" % it)
-			self._sample_topics(corpus)
-			self._store_theta()
-			print("Current theta:")
-			print(self.thetas[it])
 
-	def _initialise(self, corpus):
-		self.D, self.W = corpus.shape
-		self._total_frequencies = corpus.sum()
-		self.alpha = 1. / self.K
-		self.beta = 1. / self.W
+	def _initialise(self, corpus_p, idx_p):
+		self.D, self.W = corpus_p.shape
+		if idx_p == 0:
+			mean_a_init = 1./self.K
+			deviation_a_init = 0.2
+			a_init = np.random.normal(mean_a_init, deviation_a_init, self.K)
+			self.alpha = a_init
+			self.alphas.append(a_init)
+			a_star_init = np.random.normal(a_init, self.deviation_a_guess,
+																		 self.K)
+			self.alpha_stars.append(a_star_init)
+			self.beta = np.full((self.K, self.W), 1./self.W)
 
 		self._zw_counts = np.zeros((self.K, self.W), dtype=np.int)
 		self._dz_counts = np.zeros((self.D, self.K), dtype=np.int)
[…]

### 2017-01-22 docs/reports/report_mh_alpha-update/report_mh_alpha-update.tex
Evening update

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
\setlength{\parskip}{1em}

% \pagestyle{empty}

\begin{document}

\begingroup  
  \centering
  \large Report: The implementation of the Metropolis--Hastings (MH) algorithm to update the alphas \par
  \large Arijus Pleska \par
\endgroup

\par This report is structured in two sections: an introduction to the current model's state; and a follow-up with the faced difficulties and knowledge gaps. Note that I have prioritised to implement the MH algorithm in order to establish a complete work-flow of the model. Even though I am uncertain about some concepts, hopefully, by having a report, it will be easier to discuss the issues during the project meetings.

\subsection*{Current Stage}

% - Data and model settings
% -- Synthetic data set
% -- Parameters initialisation
% --- Number of topics
% --- Initial alphas and betas
% ---- Variance

\par During the experiment, I have used the following settings:
\begin{itemize}
\item The synthetic data has been created by the previously implemented dynamic topic modelling (DNT) generative process:
\begin{itemize}
\item The number of documents: $|\mbox{D}|\approx6000$;
\item The size of the vocabulary: $|\mbox{V}|\approx2000$;
\item The number of words per document: $\mbox{N}_d\approx20$,\quad$\forall d \in \mbox{D}$;
\item Instead of intensity values, it is assumed that the document dictionaries contain word counts. For example, $d_{111}=\{v_{20}:15, v_{40}:5\}$. 
\end{itemize}
\item The number of topics: $K=10$;
\item The number of time-slices: $T=50$;
\item The alpha at $t=0$: $\alpha_0 \sim \mathcal{N}(\mu_0, \sigma^2_0I),\quad \mu_0 = 0.1,\quad\sigma_0^2 = 0.2$;
\item The alphas at $t>0$: $\alpha_t \sim \mathcal{N}(\alpha_{t-1}, \sigma^2I), \quad\sigma^2 = 0.1$;
\item The candidate alphas: $\alpha'_t \sim \mathcal{N}(\alpha_{t}, \delta^2I), \quad\delta^2 = 2$; 
\item The acceptance rate: $r_t=\min(1,p(\alpha'_t)/p(\alpha_t))$;
\item The probability of the state: $p(\alpha_t)=p(\alpha_{t} | \alpha_{t-1})\cdot p(\alpha_{t+1} | \alpha_{t})\cdot \pi(\alpha_t)$, where $\pi$ is a mapping to the mean parameterisation;
\end{itemize}

The rationale of the implementation follows the following principle: $\alpha_t$ is set to $\alpha'_t$ on the successful `toss' based on $r_t$. Also, the variances are tuned to obtain $r_t\approx30\%$.

\subsection*{Issues}
\par My uncertainties with the proposed solution are the following:
\begin{itemize}
\item The estimation of $p(\alpha_t)$:
\begin{itemize}
\item The third term of the expression, $\pi(\alpha_t)$, represents the topic distribution in documents in time-slice $t$;
\item The current model treats the vocabulary term distributions over the topics, $\beta$, to have same values; therefore, this term was omitted -- it cancels out upon the estimation of $r_t$;
\item The first (and second) term $p(\alpha_{t} | \alpha_{t-1})$ is drawn from $\mathcal{N}(\alpha_{t-1}, \sigma^2I)$.
\end{itemize}
\item Since $\alpha_t$ is a vector, the initial $r_t$ is a vector as well.
\end{itemize}

%\bibliography{report_mh_alpha-update}{}
%\bibliographystyle{plain}
\end{document}

### 2017-02-14 presentations/alpha-update/dtm_alpha.py
The barebones for the experiment report.

@@ -14,10 +14,12 @@ class DTM_alpha(object):
 		self.var_alpha_prop = var_prop
 		self.assignments_alpha = 0
 		self.alpha_updates_potential = 0
+		self.beta = None
 
 	def _initialise(self, corpus):
 		self.T, self.V = corpus.shape
-		self.beta = np.full((self.K, self.V), 1./self.V)
+		if self.beta != None:
+			self.beta = np.full((self.K, self.V), 1./self.V)
 		# INIT ALPHAS
 		self.history_alpha = np.zeros(
 				shape=(self.iter_count, self.T, self.K))

### 2017-02-14 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Wrote the experiment settings of the experiments.

--- 2017-01-22 docs/reports/report_mh_alpha-update/report_mh_alpha-update.tex
@@
 \usepackage[english]{babel}
+\usepackage{graphicx}
+\usepackage{float}
 
@@
 % \pagestyle{empty}
+\graphicspath{ {images/} }
+% ___________________________________________________________________
 
@@
   \centering
-  \large Report: The implementation of the Metropolis--Hastings (MH) algorithm to update the alphas \par
+  \large Report: Experiments on $\alpha$ updates \par
   \large Arijus Pleska \par
 \endgroup
+% ___________________________________________________________________
 
-\par This report is structured in two sections: an introduction to the current model's state; and a follow-up with the faced difficulties and knowledge gaps.
-Note that I have prioritised to implement the MH algorithm in order to establish a complete work-flow of the model.
-Even though I am uncertain about some concepts, hopefully, by having a report, it will be easier to discuss the issues during the project meetings.
+\par This report assesses some experiments performed on reproducing $\alpha$ parameters used in a generative data process.
+Note that the report is structures in the following sections: 1) defining the experiment settings; 2) assessing the experiment results; 3) settings some questions to be discussed during next meeting.
+% ___________________________________________________________________
 
+\section*{The Experiment Settings}
+
+\par The intention of the carried experiments is to identify the optimal settings for the Metropolis--Hastings algorithm application. 
+To start with, I have generated a synthetic corpus; the parameters used in the corpus generation will allow to assess the performance achieved in the experiments.
+The corpus generation parameters are set as follows:
+\begin{itemize}
+	\item The number of topics: K = 2;
+	\item The number of documents (time-slices): T = 20;
+	\item The size of vocabulary: V = 10;
+	\item The number of words per document t: $N_t \sim \mbox{Pois}(\lambda),\quad \lambda = 1000$.
+\end{itemize}
+Further, to consider the initial settings of $\alpha_k$ development over documents, $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding \textit{softmax} expressions of the curves are illustrated in Figure 1 below.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\textwidth]{alpha_initial}
+  \caption{The values of $\mu$ used in the generative process.}
+  \label{fig:mu}
+\end{figure}
+Speaking of $\beta$, it was initially predefined and kept constant throughout the dynamic generative process; $\beta$ is illustrated in Figure 2 below.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\textwidth]{beta_initial}
+  \caption{The values of $\beta$ used in the generative process.}
+  \label{fig:beta}
+\end{figure}
+Note that the latter $\beta$ values were applied to the autoregressive topic model for the $\alpha$ update experiments.
+% ___________________________________________________________________
+
+\section*{The experiment results}
+
+\par
+
+% ___________________________________________________________________
 \subsection*{Current Stage}
@@
 \begin{itemize}
-\item The synthetic data has been created by the previously implemented dynamic topic modelling (DNT) generative process:
+\item The synthetic data has been created by inducing the previously implemented dynamic topic modelling (DNT) generative process:
 \begin{itemize}

### 2017-02-14 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Decribing the experiment results (beginning).

@@
 \usepackage{float}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 
-\par
+\par The first experiment is focused on discovering the choice of the variances.
+To be more specific, the alpha update is based on three different variances were used: the `initial' variance $\sigma_0^2I$ to induce $\alpha_t$ at $t=0$, the `basic' variance $\sigma^2I$ to induce $\alpha_t$ at $t>0$, and the `proposed' variance $\delta^2I$ to induce $\alpha'_t$ at $t=0$; also, note that $\alpha'_t$ at $t>0$ were induced using the `basic' variance.
+
+\par The experiment involved $500$ autoregressive iterations and $\sigma^2$ were set to $0.1$.
+The idea was to tune the model with different values of $\sigma_0$ and $\delta^2$; the choice of different values is illustrated in Figures 3, 4, 5 below.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\textwidth]{var-init-10_var-prop-5_iter-500}
+  \caption{The values of $\beta$ used in the generative process.}
+  \label{fig:a}
+\end{figure}
+\begin{figure}[H]
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{var-init-10_var-prop-5_iter-500}
+                \caption{A gull}
+                \label{fig:gull}
+        \end{subfigure}%
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{var-init-10_var-prop-5_iter-500}
+                \caption{A gull2}
+                \label{fig:gull2}
+        \end{subfigure}%
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{var-init-10_var-prop-5_iter-500}
+                \caption{A tiger}
+                \label{fig:tiger}
+        \end{subfigure}
+        \caption{Pictures of animals}\label{fig:animals}
+\end{figure}
 

### 2017-02-15 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Adjusting the model tuning.

@@
 \end{itemize}
-Further, to consider the initial settings of $\alpha_k$ development over documents, $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding \textit{softmax} expressions of the curves are illustrated in Figure 1 below.
+Further, to consider the initial settings of $\alpha_k$ development over documents, $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding softmax expressions of the curves, i.e $\mu = \mbox{softmax}(\alpha)$, are illustrated in Figure 1 below.
 \begin{figure}[H]
@@
 
-\par The experiment involved $500$ autoregressive iterations and $\sigma^2$ were set to $0.1$.
-The idea was to tune the model with different values of $\sigma_0$ and $\delta^2$; the choice of different values is illustrated in Figures 3, 4, 5 below.
+\par For both experiments, the number of autoregressive iterations was set to $500$ and $\sigma^2$ was set to $0.1$.
+The resulting plots of $\mu$ with different values of $\sigma_0$ and $\delta^2$ are illustrated in Figures 3, 4, 5 below.
 
@@
 \end{figure}
+
+\par The second experiment was carried to determine the impact of the $\alpha$ update in recovering the original topic fluctuations in the synthetic corpus.
+For this reason, the autoregressive part of the dynamic topic was disabled.
+The topic assignments to the documents of the last iteration are visualised in Figure 6 below.
+% ___________________________________________________________________
+
+\section*{Observations}
+
+\par For the last section of this report, I have set some questions to be addressed during next meeting; these are listed below:
+\begin{itemize}
+	\item In some cases, the larger number of iterations reduced the performance in reproducing the initial $\alpha$;
+\end{itemize}
 

### 2017-02-15 presentations/alpha-update/dtm_alpha.py

@@ -15,6 +15,8 @@ class DTM_alpha(object):
 		self.assignments_alpha = 0
 		self.alpha_updates_potential = 0
 		self.beta = None
+		self.thetas = []
+
 
 	def _initialise(self, corpus):
 		self.T, self.V = corpus.shape
@@ -60,6 +62,7 @@ class DTM_alpha(object):
 			self._update_alphas()
 			# 3) Sampling
 			self._sample_topics(corpus)
+			self._store_theta()
 
 	def eval_n_dis_log(self, x, mean, var):
 		product = np.dot(np.transpose(x-mean), (x-mean))
@@ -138,3 +141,25 @@ class DTM_alpha(object):
 					self._zw_counts[z_new, w] += 1
 					self._z_counts[z_new] += 1
 
+	def _store_theta(self):
+		thetas_prev = (self._dz_counts + self.alphas)
+		thetas_p_norm = thetas_prev \
+										/ np.sum(thetas_prev, axis=1)[:, np.newaxis]
+
+		thetas_current = np.zeros(shape=thetas_prev.shape)
+		# for doc_idx, theta_p_n in enumerate(thetas_p_norm):
+		# 	theta_current = np.random.dirichlet(theta_p_n)
+		# 	thetas_current[doc_idx, :] = theta_current
+		self.thetas.append(np.array(thetas_current))
+
+
+corpus = pd.read_pickle('corpus_test.pkl')
+iter_count = 100
+var_init = 1
+var_basic = 0.1
+var_prop = 0.15
+K = 2
+dtm_alpha = DTM_alpha(K=K, iter_count=iter_count, var_init=var_init, var_basic=var_basic, var_prop=var_prop)
+beta = np.array([[0, 0, 0, 0.3, 0.7, 0, 0, 0, 0, 0], [0, 0, 0.2, 0, 0, 0, 0, 0.3, 0.3, 0.2]])
+dtm_alpha.beta = beta
+dtm_alpha.fit(corpus)

### 2017-02-16 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Report nr2.

@@
 
-\par This report assesses some experiments performed on reproducing $\alpha$ parameters used in a generative data process.
-Note that the report is structures in the following sections: 1) defining the experiment settings; 2) assessing the experiment results; 3) settings some questions to be discussed during next meeting.
+\par This report assesses some experiments performed on reproducing $\alpha$ parameters used in generating synthetic data.
+Note that the report is structured in the following order: 1) defining the experiment settings; 2) assessing the experiment results.
 % ___________________________________________________________________
@@
   \includegraphics[width=0.5\textwidth]{alpha_initial}
-  \caption{The values of $\mu$ used in the generative process.}
+  \caption{The values of $\mbox{softmax}(\alpha)$ used in the generative process.}
   \label{fig:mu}
@@
 \par The first experiment is focused on discovering the choice of the variances.
-To be more specific, the alpha update is based on three different variances were used: the `initial' variance $\sigma_0^2I$ to induce $\alpha_t$ at $t=0$, the `basic' variance $\sigma^2I$ to induce $\alpha_t$ at $t>0$, and the `proposed' variance $\delta^2I$ to induce $\alpha'_t$ at $t=0$; also, note that $\alpha'_t$ at $t>0$ were induced using the `basic' variance.
+To be more specific, the alpha update is based on three different variances were used: the `initial' variance $\sigma^2_0I$ to induce $\alpha_t$ at $t=0$, the `basic' variance $\sigma^2I$ to induce $\alpha_t$ at $t>0$, and the `proposed' variance $\delta^2I$ to induce $\alpha'_t$ at $t=0$; also, note that $\alpha'_t$ at $t>0$ were induced using the `basic' variance.
 
-\par For both experiments, the number of autoregressive iterations was set to $500$ and $\sigma^2$ was set to $0.1$.
-The resulting plots of $\mu$ with different values of $\sigma_0$ and $\delta^2$ are illustrated in Figures 3, 4, 5 below.
+\par For both experiments, the number of autoregressive iterations was set to $500$ and $\sigma^2$ was set to $0.01$.
+The resulting plots of $\mu$ with different values of $\sigma^2_0$ and $\delta^2$ are illustrated in Figures 3, 4, 5 below.
 
 \begin{figure}[H]
-  \centering
-  \includegraphics[width=0.5\textwidth]{var-init-10_var-prop-5_iter-500}
-  \caption{The values of $\beta$ used in the generative process.}
-  \label{fig:a}
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{init-1_prop-05}
+                \caption{$\sigma^2_0=1,\quad \delta^2=0.5$}
+                \label{fig:gull}
+        \end{subfigure}%
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{init-1_prop-2}
+                \caption{$\sigma^2_0=1,\quad \delta^2=2$}
+                \label{fig:gull2}
+        \end{subfigure}%
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{init-1_prop-5}
+                \caption{$\sigma^2_0=1,\quad \delta^2=5$}
+                \label{fig:tiger}
+        \end{subfigure}
+        \caption{$\mbox{softmax}(\alpha)$ with the lowest value of the initial variance.}\label{fig:animals}
 \end{figure}
@@
         \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{var-init-10_var-prop-5_iter-500}
-                \caption{A gull}
+                \includegraphics[width=\linewidth]{init-5_prop-05}
+                \caption{$\sigma^2_0=5,\quad \delta^2=0.5$}
                 \label{fig:gull}
@@
         \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{var-init-10_var-prop-5_iter-500}
-                \caption{A gull2}
+                \includegraphics[width=\linewidth]{init-5_prop-2}
+                \caption{$\sigma^2_0=5,\quad \delta^2=2$}
                 \label{fig:gull2}
@@
         \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{var-init-10_var-prop-5_iter-500}
-                \caption{A tiger}
+                \includegraphics[width=\linewidth]{init-5_prop-5}
+                \caption{$\sigma^2_0=5,\quad \delta^2=5$}
                 \label{fig:tiger}
         \end{subfigure}
-        \caption{Pictures of animals}\label{fig:animals}
+        \caption{$\mbox{softmax}(\alpha)$ with the average value of the initial variance.}\label{fig:animals}
+\end{figure}
+\begin{figure}[H]
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{init-10_prop-05}
+                \caption{$\sigma^2_0=10,\quad \delta^2=0.5$}
+                \label{fig:gull}
+        \end{subfigure}%
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{init-10_prop-2}
+                \caption{$\sigma^2_0=10,\quad \delta^2=2$}
+                \label{fig:gull2}
+        \end{subfigure}%
+        \begin{subfigure}[b]{0.33\textwidth}
+                \includegraphics[width=\linewidth]{init-10_prop-5}
+                \caption{$\sigma^2_0=10,\quad \delta^2=5$}
+                \label{fig:tiger}
+        \end{subfigure}
+        \caption{$\mbox{softmax}(\alpha)$ with the highest value of the initial variance.}\label{fig:animals}
 \end{figure}
@@
 The topic assignments to the documents of the last iteration are visualised in Figure 6 below.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\textwidth]{thetas}
+  \caption{The topic assignments to the documents.}
+  \label{fig:mu}
+\end{figure}
+
 % ___________________________________________________________________
 
-\section*{Observations}
+% \section*{Observations}
 
-\par For the last section of this report, I have set some questions to be addressed during next meeting; these are listed below:
-\begin{itemize}
-	\item In some cases, the larger number of iterations reduced the performance in reproducing the initial $\alpha$;
-\end{itemize}
-
+% \par For the last section of this report, I have set some questions to be addressed during next meeting; these are listed below:
+% \begin{itemize}
+%   \item In some cases, the larger number of iterations reduced the performance in reproducing the initial $\alpha$;
+% \end{itemize}
 % ___________________________________________________________________
-\subsection*{Current Stage}
-
-% - Data and model settings
-% -- Synthetic data set
-% -- Parameters initialisation
-% --- Number of topics
-% --- Initial alphas and betas
-% ---- Variance
-
-\par During the experiment, I have used the following settings:
-\begin{itemize}
-\item The synthetic data has been created by inducing the previously implemented dynamic topic modelling (DNT) generative process:
-\begin{itemize}
-\item The number of documents: $|\mbox{D}|\approx6000$;
-\item The size of the vocabulary: $|\mbox{V}|\approx2000$;
-\item The number of words per document: $\mbox{N}_d\approx20$,\quad$\forall d \in \mbox{D}$;
-\item Instead of intensity values, it is assumed that the document dictionaries contain word counts. For example, $d_{111}=\{v_{20}:15, v_{40}:5\}$. 
-\end{itemize}
-\item The number of topics: $K=10$;
-\item The number of time-slices: $T=50$;
-\item The alpha at $t=0$: $\alpha_0 \sim \mathcal{N}(\mu_0, \sigma^2_0I),\quad \mu_0 = 0.1,\quad\sigma_0^2 = 0.2$;
-\item The alphas at $t>0$: $\alpha_t \sim \mathcal{N}(\alpha_{t-1}, \sigma^2I), \quad\sigma^2 = 0.1$;
-\item The candidate alphas: $\alpha'_t \sim \mathcal{N}(\alpha_{t}, \delta^2I), \quad\delta^2 = 2$; 
-\item The acceptance rate: $r_t=\min(1,p(\alpha'_t)/p(\alpha_t))$;
-\item The probability of the state: $p(\alpha_t)=p(\alpha_{t} | \alpha_{t-1})\cdot p(\alpha_{t+1} | \alpha_{t})\cdot \pi(\alpha_t)$, where $\pi$ is a mapping to the mean parameterisation;
-\end{itemize}
-
-The rationale of the implementation follows the following principle: $\alpha_t$ is set to $\alpha'_t$ on the successful `toss' based on $r_t$. Also, the variances are tuned to obtain $r_t\approx30\%$.
-
-\subsection*{Issues}
-\par My uncertainties with the proposed solution are the following:
-\begin{itemize}
-\item The estimation of $p(\alpha_t)$:
-\begin{itemize}
-\item The third term of the expression, $\pi(\alpha_t)$, represents the topic distribution in documents in time-slice $t$;
-\item The current model treats the vocabulary term distributions over the topics, $\beta$, to have same values; therefore, this term was omitted -- it cancels out upon the estimation of $r_t$;
-\item The first (and second) term $p(\alpha_{t} | \alpha_{t-1})$ is drawn from $\mathcal{N}(\alpha_{t-1}, \sigma^2I)$.
-\end{itemize}
-\item Since $\alpha_t$ is a vector, the initial $r_t$ is a vector as well.
-\end{itemize}
 

### 2017-02-16 presentations/alpha-update/dtm_alpha.py

@@ -54,12 +54,12 @@ class DTM_alpha(object):
 		for it in range(self.iter_count):
 			if it != 0 and it % 10 == 0:
 				print 'Iteration: %s' % it
-				print 'Total alpha changes: %s' % self.assignments_alpha
-				print 'Alpha update rate: %.4f' % (1.0 \
-						* self.assignments_alpha / self.alpha_updates_potential)
+				# print 'Total alpha changes: %s' % self.assignments_alpha
+				# print 'Alpha update rate: %.4f' % (1.0 \
+				# 		* self.assignments_alpha / self.alpha_updates_potential)
 			# 2) Alpha Update
 			self.history_alpha[it] = copy.deepcopy(self.alphas)
-			self._update_alphas()
+			# self._update_alphas()
 			# 3) Sampling
 			self._sample_topics(corpus)
 			self._store_theta()
@@ -147,19 +147,21 @@ class DTM_alpha(object):
 										/ np.sum(thetas_prev, axis=1)[:, np.newaxis]
 
 		thetas_current = np.zeros(shape=thetas_prev.shape)
-		# for doc_idx, theta_p_n in enumerate(thetas_p_norm):
-		# 	theta_current = np.random.dirichlet(theta_p_n)
-		# 	thetas_current[doc_idx, :] = theta_current
+		for doc_idx, theta_p_n in enumerate(thetas_p_norm):
+			theta_current = np.random.dirichlet(theta_p_n)
+			thetas_current[doc_idx, :] = theta_current
 		self.thetas.append(np.array(thetas_current))
 
 
 corpus = pd.read_pickle('corpus_test.pkl')
-iter_count = 100
+iter_count = 500
 var_init = 1
-var_basic = 0.1
-var_prop = 0.15
+var_basic = 0.01
+var_prop = 0.5
 K = 2
 dtm_alpha = DTM_alpha(K=K, iter_count=iter_count, var_init=var_init, var_basic=var_basic, var_prop=var_prop)
 beta = np.array([[0, 0, 0, 0.3, 0.7, 0, 0, 0, 0, 0], [0, 0, 0.2, 0, 0, 0, 0, 0.3, 0.3, 0.2]])
 dtm_alpha.beta = beta
 dtm_alpha.fit(corpus)
+thetas = np.array(dtm_alpha.thetas)
+thetas.dump('thetas.pkl')

### 2017-02-18 presentations/alpha-update/dtm_alpha.py
Implemented the auto-regressive variation and fixed the M--H algorithm.@

@@ -20,8 +20,8 @@ class DTM_alpha(object):
 
 	def _initialise(self, corpus):
 		self.T, self.V = corpus.shape
-		if self.beta != None:
-			self.beta = np.full((self.K, self.V), 1./self.V)
+		# if self.beta != None:
+		# 	self.beta = np.full((self.K, self.V), 1./self.V)
 		# INIT ALPHAS
 		self.history_alpha = np.zeros(
 				shape=(self.iter_count, self.T, self.K))
@@ -54,12 +54,12 @@ class DTM_alpha(object):
 		for it in range(self.iter_count):
 			if it != 0 and it % 10 == 0:
 				print 'Iteration: %s' % it
-				# print 'Total alpha changes: %s' % self.assignments_alpha
-				# print 'Alpha update rate: %.4f' % (1.0 \
-				# 		* self.assignments_alpha / self.alpha_updates_potential)
+				print 'Total alpha changes: %s' % self.assignments_alpha
+				print 'Alpha update rate: %.4f' % (1.0 \
+						* self.assignments_alpha / self.alpha_updates_potential)
 			# 2) Alpha Update
 			self.history_alpha[it] = copy.deepcopy(self.alphas)
-			# self._update_alphas()
+			self._update_alphas()
 			# 3) Sampling
 			self._sample_topics(corpus)
 			self._store_theta()

### 2017-02-18 presentations/alpha-update/dtm_autoreg.py

@@ -1,0 +1,184 @@
+import os
+import pandas as pd
+import numpy as np
+import copy
+from scipy.stats import norm
+
+np.set_printoptions(threshold=np.nan)
+
+class dtm_autoreg(object):
+
+	'''
+	Parameters:
+	- sigma_0_sq is the initial variance for alpha[0];
+	- sigma_sq is the variance for alpha[t]
+	- delta_sq is the variance for alpha'[t]
+	'''
+	def __init__(self, K, n_it, sigma_0_sq, sigma_sq, delta_sq):
+		self.n_it = n_it
+		self.K = K
+		self.sigma_0_sq = sigma_0_sq
+		self.sigma_sq = sigma_sq
+		self.delta_sq = delta_sq
+
+		self.n_upd_alpha = 0
+		self.n_it_alpha = 0
+		self.beta = None
+		self.thetas = []
+
+	def fit(self, corpus):
+		self._initialise(corpus)
+		for it in range(self.n_it):
+			if it != 0 and it % 10 == 0:
+				self._print_update(it)
+			self.hist_alpha[it] = copy.deepcopy(self.alpha)
+			self._update_alpha()
+			self._sample_topics(corpus)
+			self._store_theta()
+
+	def _print_update(self, it):
+		print 'Iteration: %s' % it
+		print 'Total alpha changes: %s' % self.n_upd_alpha
+		print 'Alpha update rate: %.4f' \
+				% (1.0 * self.n_upd_alpha / self.n_it_alpha)
+# __________________________________________________________________
+
+	def _initialise(self, corpus):
+		# Initialise the parameters
+		self.T, self.V = corpus.shape
+		self.hist_alpha = np.zeros(shape=(self.n_it,self.T,self.K))
+		self.alpha = np.zeros(shape=(self.T,self.K))
+		self.alpha[0] = np.random.normal(0,self.sigma_0_sq,self.K)
+		for t in xrange(1, self.T):
+			self.alpha[t] = np.random.normal(self.alpha[t-1], \
+					self.sigma_sq)
+		self.n_zw = np.zeros((self.K,self.V), dtype=np.int)
+		self.n_dz = np.zeros((self.T,self.K), dtype=np.int)
+		self.n_z = np.zeros(self.K, dtype=np.int)
+		self.hist_z = []
+
+		# Run the initial assignments
[…]

### 2017-02-18 presentations/alpha-update/dtm_autoreg.py
Implementing the snapshot requirement.

@@ -3,10 +3,11 @@ import pandas as pd
 import numpy as np
 import copy
 from scipy.stats import norm
+import pickle as pkl
 
 np.set_printoptions(threshold=np.nan)
 
-class dtm_autoreg(object):
+class DTM_Alpha(object):
 
 	'''
 	Parameters:
@@ -14,27 +15,46 @@ class dtm_autoreg(object):
 	- sigma_sq is the variance for alpha[t]
 	- delta_sq is the variance for alpha'[t]
 	'''
-	def __init__(self, K, n_it, sigma_0_sq, sigma_sq, delta_sq):
-		self.n_it = n_it
+	def __init__(self, K, sigma_0_sq, sigma_sq, delta_sq, 
+							 autoreg=False, snapshot=False):
+		# Input
 		self.K = K
 		self.sigma_0_sq = sigma_0_sq
 		self.sigma_sq = sigma_sq
 		self.delta_sq = delta_sq
-
+		self.autoreg = autoreg
+		# Counters and array initialisation
+		self.n_it_sampl_last = 0
 		self.n_upd_alpha = 0
 		self.n_it_alpha = 0
+		self.hist_theta = []
+		self.hist_alpha = []
+		self.hist_z = []
+		# Parameters initialised later in the work-flow
+		self.snapshot = snapshot
 		self.beta = None
-		self.thetas = []
+		self.T = None
+		self.V = None
+		self.alpha = None
+		self.n_zw = None
+		self.n_dz = None
+		self.n_z = None
+		self.corpus = None
+# __________________________________________________________________
 
-	def fit(self, corpus):
-		self._initialise(corpus)
-		for it in range(self.n_it):
+	def fit(self, n_it, corpus=None):
+		if not self.snapshot:
+			self._initialise(corpus)
+		print self.n_it_sampl_last
+		for it in range(self.n_it_sampl_last,
+									  self.n_it_sampl_last + n_it):
 			if it != 0 and it % 10 == 0:
 				self._print_update(it)
-			self.hist_alpha[it] = copy.deepcopy(self.alpha)
+			self.hist_alpha.append(copy.deepcopy(self.alpha))
 			self._update_alpha()
-			self._sample_topics(corpus)
[…]

### 2017-02-18 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Adjusting the mac for the latex.@
Adjusting the plotting, implementing the snapshotting, implementing the non auto-regressive variation.

@@
 
-\par This report assesses some experiments performed on reproducing $\alpha$ parameters used in generating synthetic data.
-Note that the report is structured in the following order: 1) defining the experiment settings; 2) assessing the experiment results.
+\par This report assesses experiments on reproducing $\alpha$ parameters used in the synthetic data generation process.
+Note that the report is structured in the following order: at the start, the rationale of the used Bayesian methods are covered; then, the experiment settings are defined; in the following stage, we assess the experiment results; finally, the identified issues are outlined to be discussed during the following meeting.
+% ___________________________________________________________________
+
+\section*{Preliminaries}
+
+\par In order to understand the experiment experiment settings, it is necessary to be familiar with the auto-regressive and non auto-regressive $\alpha$ priors as well as the Metropolis--Hastings (M--H) algorithm.
+
+\par The previously introduced priors can be expressed as follows:
+
 % ___________________________________________________________________
@@
 \end{figure}
-
 % ___________________________________________________________________

### 2017-02-18 presentations/alpha-update/dtm_alpha.py

@@ -2,147 +2,204 @@ import os
 import pandas as pd
 import numpy as np
 import copy
+from scipy.stats import norm
+import pickle as pkl
+
 np.set_printoptions(threshold=np.nan)
 
-class DTM_alpha(object):
+class DTM_Alpha(object):
 
-	def __init__(self, K, iter_count, var_init, var_basic, var_prop):
-		self.iter_count = iter_count
+	'''
+	Parameters:
+	- sigma_0_sq is the initial variance for alpha[0];
+	- sigma_sq is the variance for alpha[t]
+	- delta_sq is the variance for alpha'[t]
+	'''
+	def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None,
+	 						 delta_sq=None, autoreg=False, snapshot=False):
+		# Input
 		self.K = K
-		self.var_alpha_init = var_init
-		self.var_alpha = var_basic
-		self.var_alpha_prop = var_prop
-		self.assignments_alpha = 0
-		self.alpha_updates_potential = 0
+		self.sigma_0_sq = sigma_0_sq
+		self.sigma_sq = sigma_sq
+		self.delta_sq = delta_sq
+		self.autoreg = autoreg
+		# Counters and array initialisation
+		self.n_it_sampl_last = 0
+		self.n_upd_alpha = 0
+		self.n_it_alpha = 0
+		self.hist_theta = []
+		self.hist_alpha = []
+		self.hist_z = []
+		# Parameters initialised later in the work-flow
+		self.snapshot = snapshot
 		self.beta = None
-		self.thetas = []
+		self.T = None
+		self.V = None
+		self.alpha = None
+		self.n_zw = None
+		self.n_dz = None
+		self.n_z = None
+		self.corpus = None
+# __________________________________________________________________
 
+	def fit(self, n_it, corpus=None):
+		if not self.snapshot:
+			self._initialise(corpus)
+		for it in range(self.n_it_sampl_last,
+									  self.n_it_sampl_last + n_it):
+			if it != 0 and it % 10 == 0:
+				self._print_update(it)
+			self.hist_alpha.append(copy.deepcopy(self.alpha))
+			self._update_alpha()
+			self._sample_topics()
+			self._store_theta()
[…]

### 2017-02-18 presentations/alpha-update/dtm_autoreg.py

@@ -15,8 +15,8 @@ class DTM_Alpha(object):
 	- sigma_sq is the variance for alpha[t]
 	- delta_sq is the variance for alpha'[t]
 	'''
-	def __init__(self, K, sigma_0_sq, sigma_sq, delta_sq, 
-							 autoreg=False, snapshot=False):
+	def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None,
+	 						 delta_sq=None, autoreg=False, snapshot=False):
 		# Input
 		self.K = K
 		self.sigma_0_sq = sigma_0_sq
@@ -45,7 +45,6 @@ class DTM_Alpha(object):
 	def fit(self, n_it, corpus=None):
 		if not self.snapshot:
 			self._initialise(corpus)
-		print self.n_it_sampl_last
 		for it in range(self.n_it_sampl_last,
 									  self.n_it_sampl_last + n_it):
 			if it != 0 and it % 10 == 0:
@@ -63,7 +62,7 @@ class DTM_Alpha(object):
 				% (1.0 * self.n_upd_alpha / self.n_it_alpha)
 # __________________________________________________________________
 
-	def load_fit(self, path_file, n_it):
+	def load_fit(self, path_file, n_it=0):
 		vars_loaded = pd.read_pickle(path_file)
 		for var in vars(self):
 			if var is 'snapshot':
@@ -71,6 +70,7 @@ class DTM_Alpha(object):
 			else:
 				setattr(self, var, vars_loaded[var])
 		self.fit(n_it=n_it)
+
 # __________________________________________________________________
 
 	def _initialise(self, corpus):
@@ -220,23 +220,21 @@ def main():
 	dtm_alpha = DTM_Alpha(K=K, sigma_0_sq=sigma_0_sq,
 												sigma_sq=sigma_sq, delta_sq=delta_sq,
 												autoreg=False)
-	dtm_alpha.load_fit('model.pkl', n_it=n_it)
-	print vars(dtm_alpha).keys()
-	vars_dtm = vars(dtm_alpha)
-	with open('model.pkl', 'wb') as f:
-		pkl.dump(vars_dtm, f)
-	return
-	beta = np.array([[0, 0, 0, 0.3, 0.7, 0, 0, 0, 0, 0],
[…]

### 2017-02-19 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Finishing to set up the report.@

--- 2017-02-16 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
@@
 \usepackage{subcaption}
+\usepackage{amsmath}
 
@@
 
-\par This report assesses some experiments performed on reproducing $\alpha$ parameters used in generating synthetic data.
-Note that the report is structured in the following order: 1) defining the experiment settings; 2) assessing the experiment results.
+\par The purpose of this report is to assesses whether an auto-regressive topic model is capable to recover the $\alpha$ parameter used in the synthetic data generation process.
+This report is structured in the following order: at the start, the rationale of the used Bayesian methods are covered; then, the experiment settings are defined; in the next stage, we assess the experiment results; finally, the identified issues are outlined to be discussed during the following meeting.
+% ___________________________________________________________________
+
+\section*{Preliminaries}
+
+\par In order to understand the experiment settings, it is necessary to be familiar with the auto-regressive and non auto-regressive $\alpha$ priors as well as the Metropolis--Hastings (M--H) algorithm; the expressions of the latter concepts are listed below:
+\begin{enumerate}
+  \item The auto-regressive $\alpha$ prior:
+    \begin{align*}
+    p(\alpha_0,\ldots,\alpha_T)=p(\alpha_0)\prod_{t=1}^{T}{p(\alpha_t|\alpha_{t-1})}; \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t=0;\\
+    & p(\alpha_t)=f(\alpha_t; \alpha_{t-1}, \sigma^2I), \quad t>0.
+    \end{align*}
+  \item The non auto-regressive $\alpha$ prior:
+    \begin{align*}
+    p(\alpha_0,\ldots,\alpha_T)=\prod_{t=0}^{T}{p(\alpha_t}); \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t \geq 0.
+\end{align*}
+  \item The rationale of the utilised M--H algorithm variation:\\\\
+    For the start, familiarise with the following expressions:
+    \begin{align*}
+      \dfrac{A(x'|x)}{A(x|x')} = \dfrac{p(x'|x)}{p(x|x')}\cdot \dfrac{q(x'|x)}{q(x|x')} = \dfrac{p(x',x)}{p(x)}\cdot \dfrac{p(x')}{p(x,x')} = \dfrac{p(x')}{p(x)}; \qquad &x' \sim q(x,\delta^2I);\\
+      & q = \mathcal{N} \Rightarrow q(x'|x) = q(x|x');
+    \end{align*}
+    where $x$ is the current state, $x'$ is the proposed state, $A$ is the acceptance distribution, and $q$ is the proposal distribution.
+Taking the previous results into account, the acceptance rate $r$ is expressed as follows:
+    \begin{align*}
+      r = \min{\bigg(1, \dfrac{p(x')}{p(x)}\bigg)}.
+    \end{align*}
+  \item The application of the M--H algorithm to update $\alpha$:\\\\
+    Note that the $\alpha$ values are updated independently; that is, the expression for the acceptance rate below is for a single $\alpha$ entry.
+    \begin{align*}
+      \dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)} &= \dfrac{p(X|z,\alpha^{-tk},\alpha_{t, k}')\cdot p(z|\alpha^{-tk},\alpha_{t, k}')\cdot p(\alpha^{-tk},\alpha_{t, k}')}{p(X)}\cdot \dfrac{p(X)}{p(X|z,\alpha)\cdot p(z|\alpha)\cdot p(\alpha)}\\
+      &= \dfrac{\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}; \qquad t > 0, \quad t \neq T;
+    \end{align*}
+    where $\pi$ is the softmax function, $\alpha_{t,k}' \sim \mathcal{N}(\alpha_{t, k}, \delta^2)$, and $\alpha^{-tk}$ denotes $\alpha$ without $\alpha_{t,k}'$. It follows that,
+    \begin{align*}
+      r_{t,k} = \min{\bigg(1, \dfrac{\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}\bigg)}; \qquad t > 0, \quad t \neq T.
+    \end{align*}
+    For the boundary cases $t=0$ and $t=T$, the $p(\alpha)$ term is proportional to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively.
+
+    % \begin{align*}
+    %   \dfrac{p(\alpha_{t,k}'|\alpha_{t,k})}{p(\alpha_{t,k}|\alpha_{t,k}')}\cdot \dfrac{q(\alpha_{t,k}'|\alpha_{t,k})}{q(\alpha_{t,k}|\alpha_{t,k}')} = \dfrac{p(\alpha_{t,k}',\alpha_{t,k})}{p(\alpha_{t,k})}\cdot \dfrac{p(\alpha_{t,k}')}{p(\alpha_{t,k},\alpha_{t,k}')}, \qquad &\alpha_{t,k}' \sim \mathcal{N}(\alpha_{t,k},\delta^2),\\
+    %   &q(\alpha_{t,k}'|\alpha_{t,k}) = q(\alpha_{t,k}|\alpha_{t,k})'.
+    % \end{align*}
+    %   \Rightarrow \dfrac{p(\alpha_{t,k}')}{p(\alpha_{t,k})}
+
+\end{enumerate}
+
 % ___________________________________________________________________
@@
 \par The intention of the carried experiments is to identify the optimal settings for the Metropolis--Hastings algorithm application. 
-To start with, I have generated a synthetic corpus; the parameters used in the corpus generation will allow to assess the performance achieved in the experiments.
-The corpus generation parameters are set as follows:
+The rationale of the carried experiments is based on generating a corpus with pre-defined $\alpha$ changes.
+Based on the experiments, we will determine which techniques display higher performance in reproducing the pre-defined $\alpha$ changes.
+To expand on the corpus generation settings, the parameters used are listed below:
 \begin{itemize}
   \item The number of topics: K = 2;
-  \item The number of documents (time-slices): T = 20;
+  \item The number of documents (time slices): T = 20;
   \item The size of vocabulary: V = 10;
@@
 \end{itemize}
-Further, to consider the initial settings of $\alpha_k$ development over documents, $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding softmax expressions of the curves, i.e $\mu = \mbox{softmax}(\alpha)$, are illustrated in Figure 1 below.
+Speaking of $\alpha_k$ development over time (documents), $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding softmax expressions of the curves (i.e, $\theta = \mbox{softmax}(\alpha)$) are illustrated in Figure 1 below.
 \begin{figure}[H]
@@
 \end{figure}
-Speaking of $\beta$, it was initially predefined and kept constant throughout the dynamic generative process; $\beta$ is illustrated in Figure 2 below.
+Speaking of $\beta$, it was initially pre-defined and kept constant throughout the dynamic generative process; $\beta$ is illustrated in Figure 2 below.
 \begin{figure}[H]
@@
 \end{figure}
-Note that the latter $\beta$ values were applied to the autoregressive topic model for the $\alpha$ update experiments.
 % ___________________________________________________________________
@@
 
-\par The first experiment is focused on discovering the choice of the variances.
-To be more specific, the alpha update is based on three different variances were used: the `initial' variance $\sigma^2_0I$ to induce $\alpha_t$ at $t=0$, the `basic' variance $\sigma^2I$ to induce $\alpha_t$ at $t>0$, and the `proposed' variance $\delta^2I$ to induce $\alpha'_t$ at $t=0$; also, note that $\alpha'_t$ at $t>0$ were induced using the `basic' variance.
+\par The first experiment is focused on discovering an optimal choice of the variances.
+Note that the first experiment is carried using the auto-regressive model; therefore, three different types of variances were considered: the `basic' variance $\sigma^2_0$, the `auto-regressive' variance $\sigma^2$, and the `proposed' variance $\delta^2$.
 
-\par For both experiments, the number of autoregressive iterations was set to $500$ and $\sigma^2$ was set to $0.01$.
+\par For the first experiment, $\delta^2$ was kept constant and set to $1$.
+Effectively, low $\delta^2$ values suggest that the convergence of $\alpha$ is slow and stable, whereas for high values of $\delta^2$ the convergence is faster and less stable.
+In both cases, with a high number of iterations, a low-error $\alpha$ value will be found.
+Therefore, we are focusing to tune only the $\sigma^2_0$ and $\sigma^2$ variances
+
+% \par For both experiments, the number of autoregressive iterations was set to $500$ and $\sigma^2$ was set to $0.01$.
 The resulting plots of $\mu$ with different values of $\sigma^2_0$ and $\delta^2$ are illustrated in Figures 3, 4, 5 below.
@@
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-1_prop-05}
-                \caption{$\sigma^2_0=1,\quad \delta^2=0.5$}
-                \label{fig:gull}
+    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-02|prop_1E+00|it_500}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-1_prop-2}
-                \caption{$\sigma^2_0=1,\quad \delta^2=2$}
-                \label{fig:gull2}
+    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-02|prop_1E+00|it_500}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-1_prop-5}
-                \caption{$\sigma^2_0=1,\quad \delta^2=5$}
-                \label{fig:tiger}
+    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-02|prop_1E+00|it_500}
   \end{subfigure}
@@
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-5_prop-05}
-                \caption{$\sigma^2_0=5,\quad \delta^2=0.5$}
-                \label{fig:gull}
+    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-03|prop_1E+00|it_500}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-5_prop-2}
-                \caption{$\sigma^2_0=5,\quad \delta^2=2$}
-                \label{fig:gull2}
+    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-03|prop_1E+00|it_500}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-5_prop-5}
-                \caption{$\sigma^2_0=5,\quad \delta^2=5$}
-                \label{fig:tiger}
+    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-03|prop_1E+00|it_500}
   \end{subfigure}
-        \caption{$\mbox{softmax}(\alpha)$ with the average value of the initial variance.}\label{fig:animals}
+  \caption{$\mbox{softmax}(\alpha)$ with the lowest value of the initial variance.}\label{fig:animals}
 \end{figure}
@@
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-10_prop-05}
-                \caption{$\sigma^2_0=10,\quad \delta^2=0.5$}
-                \label{fig:gull}
+    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-04|prop_1E+00|it_500}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-10_prop-2}
-                \caption{$\sigma^2_0=10,\quad \delta^2=2$}
-                \label{fig:gull2}
+    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-04|prop_1E+00|it_500}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-                \includegraphics[width=\linewidth]{init-10_prop-5}
-                \caption{$\sigma^2_0=10,\quad \delta^2=5$}
-                \label{fig:tiger}
+    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-04|prop_1E+00|it_500}
   \end{subfigure}
-        \caption{$\mbox{softmax}(\alpha)$ with the highest value of the initial variance.}\label{fig:animals}
+  \caption{$\mbox{softmax}(\alpha)$ with the lowest value of the initial variance.}\label{fig:animals}
 \end{figure}
 
-\par The second experiment was carried to determine the impact of the $\alpha$ update in recovering the original topic fluctuations in the synthetic corpus.
-For this reason, the autoregressive part of the dynamic topic was disabled.
-The topic assignments to the documents of the last iteration are visualised in Figure 6 below.
+\par The second experiment assesses the impact of the auto-regressive $\alpha$ update.
+For this reason, the $\alpha$ prior was switched to the non auto-regressive one.
+The resulting $\theta$ values are illustrated in Figure X below.
 

### 2017-02-19 presentations/alpha-update/dtm_alpha.py

@@ -81,8 +81,11 @@ class DTM_Alpha(object):
 		self.alpha = np.zeros(shape=(self.T,self.K))
 		self.alpha[0] = np.random.normal(0,self.sigma_0_sq,self.K)
 		for t in xrange(1, self.T):
-			self.alpha[t] = np.random.normal(self.alpha[t-1], \
-					self.sigma_sq)
+			if self.autoreg:
+				self.alpha[t] = np.random.normal(self.alpha[t-1], \
+																				 self.sigma_sq)
+			else:
+				self.alpha[t] = np.random.normal(0, self.sigma_0_sq)
 		self.n_zw = np.zeros((self.K,self.V), dtype=np.int)
 		self.n_dz = np.zeros((self.T,self.K), dtype=np.int)
 		self.n_z = np.zeros(self.K, dtype=np.int)

### 2017-02-20 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Migrating the updated model to the code directory.

@@
 \par The purpose of this report is to assesses whether an auto-regressive topic model is capable to recover the $\alpha$ parameter used in the synthetic data generation process.
-This report is structured in the following order: at the start, the rationale of the used Bayesian methods are covered; then, the experiment settings are defined; in the next stage, we assess the experiment results; finally, the identified issues are outlined to be discussed during the following meeting.
+This report is structured in the following order: at the start, the rationale of the used Bayesian methods are covered; then, the experiment settings are defined; in the next stage, we assess the experiment results; finally, the identified issues are outlined to be discussed during next meeting.
 % ___________________________________________________________________
@@
     \end{align*}
-    For the boundary cases $t=0$ and $t=T$, the $p(\alpha)$ term is proportional to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively.
+    For the boundary cases $t=0$ and $t=T$, the $p(\alpha)$ term corresponds to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively.
 
@@
 The rationale of the carried experiments is based on generating a corpus with pre-defined $\alpha$ changes.
-Based on the experiments, we will determine which techniques display higher performance in reproducing the pre-defined $\alpha$ changes.
+Based on the experiments, we will determine which techniques display higher performance in reproducing the pre-defined $\alpha$ fluctuations over time.
 To expand on the corpus generation settings, the parameters used are listed below:
@@
 \end{itemize}
-Speaking of $\alpha_k$ development over time (documents), $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding softmax expressions of the curves (i.e, $\theta = \mbox{softmax}(\alpha)$) are illustrated in Figure 1 below.
+Speaking of $\alpha_k$ development over time (documents), $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding topic distributions over documents (i.e., $\theta = \mbox{softmax}(\alpha)$)) are illustrated in Figure 1 below.
 \begin{figure}[H]
   \centering
-  \includegraphics[width=0.5\textwidth]{alpha_initial}
+  \includegraphics[width=0.5\textwidth]{theta_initial}
   \caption{The values of $\mbox{softmax}(\alpha)$ used in the generative process.}
-  \label{fig:mu}
 \end{figure}
-Speaking of $\beta$, it was initially pre-defined and kept constant throughout the dynamic generative process; $\beta$ is illustrated in Figure 2 below.
+Speaking of $\beta$, the parameter was initially pre-defined and kept constant throughout the dynamic generative process; $\beta$ is illustrated in Figure 2 below.
 \begin{figure}[H]
@@
   \caption{The values of $\beta$ used in the generative process.}
-  \label{fig:beta}
 \end{figure}
@@
 In both cases, with a high number of iterations, a low-error $\alpha$ value will be found.
-Therefore, we are focusing to tune only the $\sigma^2_0$ and $\sigma^2$ variances
-
-% \par For both experiments, the number of autoregressive iterations was set to $500$ and $\sigma^2$ was set to $0.01$.
-The resulting plots of $\mu$ with different values of $\sigma^2_0$ and $\delta^2$ are illustrated in Figures 3, 4, 5 below.
+Therefore, we are focusing to tune only the $\sigma^2_0$ and $\sigma^2$ variances.
+Also, note that two visualisation of topic distribution over documents are provided: at 1000 iterations; and at 2000 iterations.
 
@@
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-02|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-02|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-02|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-02|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-02|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-02|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}
-  \caption{$\mbox{softmax}(\alpha)$ with the lowest value of the initial variance.}\label{fig:animals}
+  \caption{Fixed $\sigma^2=0.01$.}
 \end{figure}
@@
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-03|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-03|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-03|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}
-  \caption{$\mbox{softmax}(\alpha)$ with the lowest value of the initial variance.}\label{fig:animals}
+  \caption{Fixed $\sigma^2=0.001$.}
 \end{figure}
@@
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-04|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-04|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-04|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}%
   \begin{subfigure}[b]{0.33\textwidth}
-    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-04|prop_1E+00|it_500}
+    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-04|prop_1E+00|it_2000|autoreg_True}
   \end{subfigure}
-  \caption{$\mbox{softmax}(\alpha)$ with the lowest value of the initial variance.}\label{fig:animals}
+  \caption{Fixed $\sigma^2=0.0001$.}
 \end{figure}
@@
 For this reason, the $\alpha$ prior was switched to the non auto-regressive one.
-The resulting $\theta$ values are illustrated in Figure X below.
+Note that in this case $\sigma^2$ has no impact; therefore, we provide illustrations by varying the $\sigma^2_0$ term.
+Again, $\delta^2=1$ was kept constant throughout the experiments.
+Also, note that we ran the model for $200$ iterations.
+The resulting $\mbox{softmax}(\alpha)$ values are illustrated in Figure 6 below.
 
 \begin{figure}[H]
-  \centering
-  \includegraphics[width=0.5\textwidth]{thetas}
-  \caption{The topic assignments to the documents.}
-  \label{fig:mu}
+  \begin{subfigure}[b]{0.33\textwidth}
+    \includegraphics[width=\linewidth]{init_1E-01|basic_1E-02|prop_1E+00|it_200|autoreg_False}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.33\textwidth}
+    \includegraphics[width=\linewidth]{init_1E+00|basic_1E-02|prop_1E+00|it_200|autoreg_False}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.33\textwidth}
+    \includegraphics[width=\linewidth]{init_5E+00|basic_1E-02|prop_1E+00|it_200|autoreg_False}
+  \end{subfigure}
+  \caption{The non auto-regressive $\alpha$ update.}
 \end{figure}
@@
 
-% \section*{Observations}
+\section*{Questions}
 
-% \par For the last section of this report, I have set some questions to be addressed during next meeting; these are listed below:
-% \begin{itemize}
-%   \item In some cases, the larger number of iterations reduced the performance in reproducing the initial $\alpha$;
-% \end{itemize}
+\par For the last section of this report, I have set some questions to be addressed during next meeting; these are listed below:
+\begin{itemize}
+  \item During the derivation of the posterior given in Preliminaries Section, is the $\pi(\alpha_{t}')_k$ term, i.e.
+$p(z|\alpha^{-tk},\alpha_{t, k}')$, derived correctly?
+We did not have it explicitly expressed before.
+  \item Should we consider the impact of tuning the model with the pre-defined $\beta$ (the one used during the synthetic corpus generation)?
+The model is able to recover the topic assignments to documents even if $\beta$ is initialised randomly.
+Also, do we have pre-defined $\beta$ for the non synthetic data sets?
+  \item What are the indications of the faster $\alpha$ converge displayed in the non auto-regressive model? Could it mean that the model implementation is faulty?
+\end{itemize}
 % ___________________________________________________________________

### 2017-02-20 code/topic-models/alpha-update/dtm_alpha.py

@@ -1,0 +1,279 @@
+import os
+import pandas as pd
+import numpy as np
+import copy
+from scipy.stats import norm
+import pickle as pkl
+
+np.set_printoptions(threshold=np.nan)
+
+
+class DTM_Alpha(object):
+
+    '''
+    Parameters:
+    - sigma_0_sq is the initial variance for alpha[0];
+    - sigma_sq is the variance for alpha[t]
+    - delta_sq is the variance for alpha'[t]
+    '''
+    def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None, delta_sq=None,
+                 autoreg=False, snapshot=False):
+        # Input
+        self.K = K
+        self.sigma_0_sq = sigma_0_sq
+        self.sigma_sq = sigma_sq
+        self.delta_sq = delta_sq
+        self.autoreg = autoreg
+        # Counters and array initialisation
+        self.n_it_sampl_last = 0
+        self.n_upd_alpha = 0
+        self.n_it_alpha = 0
+        self.hist_theta = []
+        self.hist_alpha = []
+        self.hist_z = []
+        # Parameters initialised later in the work-flow
+        self.snapshot = snapshot
+        self.beta = None
+        self.T = None
+        self.V = None
+        self.alpha = None
+        self.n_zw = None
+        self.n_dz = None
+        self.n_z = None
+        self.corpus = None
+# ___________________________________________________________________________
+
+    def fit(self, n_it, corpus=None):
+        if not self.snapshot:
+            self._initialise(corpus)
+        for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
+            if it != 0 and it % 10 == 0:
+                self._print_update(it)
+            self.hist_alpha.append(copy.deepcopy(self.alpha))
+            self._update_alpha()
+            self._sample_topics()
+            self._store_theta()
+        self.n_it_sampl_last += n_it
+
+    def _print_update(self, it):
+        print 'Iteration: %s' % self.n_upd_alpha
+        print 'Total alpha changes: %s' % self.n_upd_alpha
+        print 'Alpha update rate: %.4f' \
+            % (1.0 * self.n_upd_alpha / self.n_it_alpha)
+# ___________________________________________________________________________
+
+    def load_fit(self, path_file, n_it=0):
[…]

### 2017-02-20 code/topic-models/dtm_alpha-only.py

@@ -1,176 +1,0 @@
-import os
-import pandas as pd
-import numpy as np
-import copy
-
-np.set_printoptions(threshold=np.nan)
-
-'''
-References:
-- Blei and Lafferty (2006)
-'''
-
-class DTM_alpha(object):
-
-	def __init__(self, K, iter_count):
-		self.iter_count = iter_count
-		self.K = K
-		self.history_alphas = []
-
-	def _initialise(self, corpus):
-		self.T, self.V = corpus.shape
-		self.beta = np.full((self.K, self.V), 1./self.V)
-		# INIT ALPHAS
-		self.alphas = np.zeros(shape=(self.T, self.K))
-		self.mean_alpha_init = 0
-		self.var_alpha_init = 0.2
-		self.var_alpha_prop = 0.05
-		self.var_alpha = 0.01
-
-		self.alphas[0] = np.random.normal(self.mean_alpha_init, 
-																			self.var_alpha_init, 
-																			self.K)
-		for t in xrange(1, self.T):
-			self.alphas[t] = np.random.normal(self.alphas[t-1], 
-																				self.var_alpha)
-		# INIT TOPIC ASSIGNMENTS
-		self._zw_counts = np.zeros((self.K, self.V), dtype=np.int)
-		self._dz_counts = np.zeros((self.T, self.K), dtype=np.int)
-		self._z_counts = np.zeros(self.K, dtype=np.int)
-		self._z_history = []
-
-		for d, doc in enumerate(corpus):
-			self._z_history.append([])
-			for w, word in enumerate(doc):
-				self._z_history[d].append([])
-				word_frequency = int(word)
-				for _ in range(word_frequency):
-					z = np.random.randint(self.K)
-					self._z_history[d][w].append(z)
-					self._dz_counts[d, z] += 1
-					self._zw_counts[z, w] += 1
-					self._z_counts[z] += 1
-
-	def fit(self, corpus):
-		# 1) INITIALISATION
-		self._initialise(corpus)
-		for it in range(self.iter_count):
-			print 'Iteration: %s.' % it
-			# 2) Alpha Update
-			self.history_alphas.append(copy.deepcopy(self.alphas))
-			self._update_alphas()
-			print self.changed
-			# 3) Sampling
-			self._sample_topics(corpus)
-
[…]

### 2017-02-20 code/末legacy/dtm_alpha-only.py

@@ -1,0 +1,176 @@
+import os
+import pandas as pd
+import numpy as np
+import copy
+
+np.set_printoptions(threshold=np.nan)
+
+'''
+References:
+- Blei and Lafferty (2006)
+'''
+
+class DTM_alpha(object):
+
+	def __init__(self, K, iter_count):
+		self.iter_count = iter_count
+		self.K = K
+		self.history_alphas = []
+
+	def _initialise(self, corpus):
+		self.T, self.V = corpus.shape
+		self.beta = np.full((self.K, self.V), 1./self.V)
+		# INIT ALPHAS
+		self.alphas = np.zeros(shape=(self.T, self.K))
+		self.mean_alpha_init = 0
+		self.var_alpha_init = 0.2
+		self.var_alpha_prop = 0.05
+		self.var_alpha = 0.01
+
+		self.alphas[0] = np.random.normal(self.mean_alpha_init, 
+																			self.var_alpha_init, 
+																			self.K)
+		for t in xrange(1, self.T):
+			self.alphas[t] = np.random.normal(self.alphas[t-1], 
+																				self.var_alpha)
+		# INIT TOPIC ASSIGNMENTS
+		self._zw_counts = np.zeros((self.K, self.V), dtype=np.int)
+		self._dz_counts = np.zeros((self.T, self.K), dtype=np.int)
+		self._z_counts = np.zeros(self.K, dtype=np.int)
+		self._z_history = []
+
+		for d, doc in enumerate(corpus):
+			self._z_history.append([])
+			for w, word in enumerate(doc):
+				self._z_history[d].append([])
+				word_frequency = int(word)
+				for _ in range(word_frequency):
+					z = np.random.randint(self.K)
+					self._z_history[d][w].append(z)
+					self._dz_counts[d, z] += 1
+					self._zw_counts[z, w] += 1
+					self._z_counts[z] += 1
+
+	def fit(self, corpus):
+		# 1) INITIALISATION
+		self._initialise(corpus)
+		for it in range(self.iter_count):
+			print 'Iteration: %s.' % it
+			# 2) Alpha Update
+			self.history_alphas.append(copy.deepcopy(self.alphas))
+			self._update_alphas()
+			print self.changed
+			# 3) Sampling
+			self._sample_topics(corpus)
+
[…]

### 2017-02-20 presentations/alpha-update/dtm_alpha.py

@@ -1,243 +1,0 @@
-import os
-import pandas as pd
-import numpy as np
-import copy
-from scipy.stats import norm
-import pickle as pkl
-
-np.set_printoptions(threshold=np.nan)
-
-class DTM_Alpha(object):
-
-	'''
-	Parameters:
-	- sigma_0_sq is the initial variance for alpha[0];
-	- sigma_sq is the variance for alpha[t]
-	- delta_sq is the variance for alpha'[t]
-	'''
-	def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None,
-	 						 delta_sq=None, autoreg=False, snapshot=False):
-		# Input
-		self.K = K
-		self.sigma_0_sq = sigma_0_sq
-		self.sigma_sq = sigma_sq
-		self.delta_sq = delta_sq
-		self.autoreg = autoreg
-		# Counters and array initialisation
-		self.n_it_sampl_last = 0
-		self.n_upd_alpha = 0
-		self.n_it_alpha = 0
-		self.hist_theta = []
-		self.hist_alpha = []
-		self.hist_z = []
-		# Parameters initialised later in the work-flow
-		self.snapshot = snapshot
-		self.beta = None
-		self.T = None
-		self.V = None
-		self.alpha = None
-		self.n_zw = None
-		self.n_dz = None
-		self.n_z = None
-		self.corpus = None
-# __________________________________________________________________
-
-	def fit(self, n_it, corpus=None):
-		if not self.snapshot:
-			self._initialise(corpus)
-		for it in range(self.n_it_sampl_last,
-									  self.n_it_sampl_last + n_it):
-			if it != 0 and it % 10 == 0:
-				self._print_update(it)
-			self.hist_alpha.append(copy.deepcopy(self.alpha))
-			self._update_alpha()
-			self._sample_topics()
-			self._store_theta()
-		self.n_it_sampl_last += n_it
-
-	def _print_update(self, it):
-		print 'Iteration: %s' % it
-		print 'Total alpha changes: %s' % self.n_upd_alpha
-		print 'Alpha update rate: %.4f' \
-				% (1.0 * self.n_upd_alpha / self.n_it_alpha)
-# __________________________________________________________________
-
[…]

### 2017-02-20 presentations/alpha-update/dtm_autoreg.py

@@ -1,240 +1,0 @@
-import os
-import pandas as pd
-import numpy as np
-import copy
-from scipy.stats import norm
-import pickle as pkl
-
-np.set_printoptions(threshold=np.nan)
-
-class DTM_Alpha(object):
-
-	'''
-	Parameters:
-	- sigma_0_sq is the initial variance for alpha[0];
-	- sigma_sq is the variance for alpha[t]
-	- delta_sq is the variance for alpha'[t]
-	'''
-	def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None,
-	 						 delta_sq=None, autoreg=False, snapshot=False):
-		# Input
-		self.K = K
-		self.sigma_0_sq = sigma_0_sq
-		self.sigma_sq = sigma_sq
-		self.delta_sq = delta_sq
-		self.autoreg = autoreg
-		# Counters and array initialisation
-		self.n_it_sampl_last = 0
-		self.n_upd_alpha = 0
-		self.n_it_alpha = 0
-		self.hist_theta = []
-		self.hist_alpha = []
-		self.hist_z = []
-		# Parameters initialised later in the work-flow
-		self.snapshot = snapshot
-		self.beta = None
-		self.T = None
-		self.V = None
-		self.alpha = None
-		self.n_zw = None
-		self.n_dz = None
-		self.n_z = None
-		self.corpus = None
-# __________________________________________________________________
-
-	def fit(self, n_it, corpus=None):
-		if not self.snapshot:
-			self._initialise(corpus)
-		for it in range(self.n_it_sampl_last,
-									  self.n_it_sampl_last + n_it):
-			if it != 0 and it % 10 == 0:
-				self._print_update(it)
-			self.hist_alpha.append(copy.deepcopy(self.alpha))
-			self._update_alpha()
-			self._sample_topics()
-			self._store_theta()
-		self.n_it_sampl_last += n_it
-
-	def _print_update(self, it):
-		print 'Iteration: %s' % it
-		print 'Total alpha changes: %s' % self.n_upd_alpha
-		print 'Alpha update rate: %.4f' \
-				% (1.0 * self.n_upd_alpha / self.n_it_alpha)
-# __________________________________________________________________
-
[…]

### 2017-02-20 presentations/alpha-update/preprocessing.py

@@ -1,23 +1,0 @@
-import numpy as np
-np.set_printoptions(threshold=np.nan)
-
-# Smaller corpus for Gibbs sampling
-def preprocess_corpus(corpus, vocab, threshold, normalise):
-	pp_corpus = np.zeros((len(corpus), len(vocab)), dtype=np.float)
-	factor_norm = 1
-	for doc in corpus:
-		# Finding the maximum intensity.
-		if normalise:
-			max_ND = 0
-			for word in corpus[doc]:
-				if corpus[doc][word] > max_ND:
-					max_ND = corpus[doc][word]
-			factor_norm = threshold / max_ND
-
-		# Populating the preprocessed corpus.
-		for w, word in enumerate(vocab):
-			if word in corpus[doc]:
-				count_norm = factor_norm * corpus[doc][word]
-				pp_corpus[doc][w] = int(count_norm)
-	return pp_corpus
-	

### 2017-02-21 code/topic-models/alpha-update/dtm_alpha.py
"git test."

@@ -224,6 +224,7 @@ class DTM_Alpha(object):
 
 
 def main():
+    print 'gittest'
     corpus = pd.read_pickle('corpus_test.pkl')
     n_it = 250
     sigma_0_sq = 1

### 2017-02-23 docs/reports/experiments_MH-DTM/report_mh-dtm.tex
Summing alpha_k for the acceptance rate calculation

@@
 \section*{Preliminaries}
+% TODO: Fix the pi expression (remove P(X|...))
 
@@
       \dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)} &= \dfrac{p(X|z,\alpha^{-tk},\alpha_{t, k}')\cdot p(z|\alpha^{-tk},\alpha_{t, k}')\cdot p(\alpha^{-tk},\alpha_{t, k}')}{p(X)}\cdot \dfrac{p(X)}{p(X|z,\alpha)\cdot p(z|\alpha)\cdot p(\alpha)}\\
-      &= \dfrac{\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}; \qquad t > 0, \quad t \neq T;
+      &= \dfrac{\prod_{k=0}^K\big[\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\big]\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\prod_{k=0}^K\big[\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\big]\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}; \qquad t > 0, \quad t \neq T;
     \end{align*}
-    where $\pi$ is the softmax function, $\alpha_{t,k}' \sim \mathcal{N}(\alpha_{t, k}, \delta^2)$, and $\alpha^{-tk}$ denotes $\alpha$ without $\alpha_{t,k}'$. It follows that,
+    where $\pi$ is the softmax function, $\alpha_{t,k}' \sim \mathcal{N}(\alpha_{t, k}, \delta^2)$, and $\alpha^{-tk}$ denotes $\alpha$ without $\alpha_{t,k}'$; also, for the boundary cases $t=0$ and $t=T$, note that the $p(\alpha)$ term corresponds to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively.
+Further, for computational stability, the previous ratio would be computed in the log space as follows:
     \begin{align*}
-      r_{t,k} = \min{\bigg(1, \dfrac{\pi(\alpha_{t}')_k^{z_{t,k}}\cdot \pi(\alpha_{t}')_k\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\pi(\alpha_t)_k^{z_{t,k}}\cdot \pi(\alpha_{t})_k\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}\bigg)}; \qquad t > 0, \quad t \neq T.
+      \log{\bigg[\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)}\bigg]} &= \log{\big[p(z,\alpha^{-tk},\alpha_{t, k}'|X)\big]} - \log{\big[p(z,\alpha|X)\big]}; \quad \mbox{where}\\
+      \log{\big[p(z,\alpha^{-tk},\alpha_{t, k}'|X)\big]} &=\sum_{k=0}^K\bigg[(z_{t,k}+1)\cdot\log\big[\pi(\alpha_{t}')_k\big]\bigg] + \log\big[p(\alpha_t'|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t')\big],\\
+      \log\big[p(z,\alpha|X)\big] &= \sum_{k=0}^K\bigg[(z_{t,k}+1)\cdot\log\big[\pi(\alpha_{t})_k\big]\bigg] + \log\big[p(\alpha_t|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t)\big].
     \end{align*}
-    For the boundary cases $t=0$ and $t=T$, the $p(\alpha)$ term corresponds to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively.
+    Finally, the acceptance rate is calculated using the formula below.
+    \begin{align*}
+      r_{t,k} = \exp\bigg[\min{\bigg(0, \log{\bigg[\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)}\bigg]}\bigg)}\bigg].
+    \end{align*}
 

### 2017-02-23 code/topic-models/alpha-update/dtm_alpha.py

@@ -1,3 +1,12 @@
+
+"""
+VERSION
+- Python 2
+
+FUNCTION
+- Infer topics by updating the alphas.
+"""
+
 import os
 import pandas as pd
 import numpy as np
@@ -11,28 +20,31 @@ np.set_printoptions(threshold=np.nan)
 class DTM_Alpha(object):
 
     '''
-    Parameters:
+    PARAMETERS:
     - sigma_0_sq is the initial variance for alpha[0];
     - sigma_sq is the variance for alpha[t]
     - delta_sq is the variance for alpha'[t]
     '''
     def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None, delta_sq=None,
                  autoreg=False, snapshot=False):
-        # Input
         self.K = K
         self.sigma_0_sq = sigma_0_sq
         self.sigma_sq = sigma_sq
         self.delta_sq = delta_sq
         self.autoreg = autoreg
-        # Counters and array initialisation
+        self.snapshot = snapshot
+        '''
+        Counter and array initialisation
+        '''
         self.n_it_sampl_last = 0
         self.n_upd_alpha = 0
         self.n_it_alpha = 0
         self.hist_theta = []
         self.hist_alpha = []
         self.hist_z = []
-        # Parameters initialised later in the work-flow
-        self.snapshot = snapshot
+        '''
+        The parameters initialised later in the work-flow
+        '''
         self.beta = None
         self.T = None
         self.V = None
@@ -48,18 +60,16 @@ class DTM_Alpha(object):
             self._initialise(corpus)
         for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
             if it != 0 and it % 10 == 0:
-                self._print_update(it)
+                print('Iteration: %s' % it)
+                print('Total alpha changes: %s' % self.n_upd_alpha)
+                print('Alpha update rate: %.4f'
+                      % (1.0 * self.n_upd_alpha / self.n_it_alpha))
             self.hist_alpha.append(copy.deepcopy(self.alpha))
             self._update_alpha()
+            # break
             self._sample_topics()
             self._store_theta()
         self.n_it_sampl_last += n_it
-
-    def _print_update(self, it):
-        print 'Iteration: %s' % self.n_upd_alpha
-        print 'Total alpha changes: %s' % self.n_upd_alpha
[…]

### 2017-03-14 code/models/dtm_alpha.py
Repackaging; Establishing an experiment-based work-flow.

@@ -1,0 +1,346 @@
+
+"""
+VERSION
+- Python 2
+
+FUNCTION
+- Infer topics by updating the alphas.
+"""
+
+import os
+import pandas as pd
+import numpy as np
+import copy
+from time import time
+
+np.set_printoptions(threshold=np.nan)
+
+# ___________________________________________________________________________
+
+
+class DTM_Alpha(object):
+    '''
+    PARAMETERS:
+    - sigma_0_sq is the initial variance for alpha[0];
+    - sigma_sq is the variance for alpha[t]
+    - delta_sq is the variance for alpha'[t]
+    '''
+    def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None, delta_sq=None,
+                 autoreg=False, snapshot=False, beta=None,
+                 verbose_init=True):
+        self.K = K
+        self.sigma_0_sq = sigma_0_sq
+        self.sigma_sq = sigma_sq
+        self.delta_sq = delta_sq
+        self.autoreg = autoreg
+        self.snapshot = snapshot
+        self.beta = beta
+        self.verbose_init = verbose_init
+        '''
+        Counter and array initialisation
+        '''
+        self.n_it_sampl_last = 0
+        self.n_upd_alpha = 0
+        self.n_it_alpha = 0
+        self.hist_theta = []
+        self.hist_alpha = []
+        self.hist_z = []
+        '''
+        The parameters initialised later in the work-flow
+        '''
+        self.beta = None
+        self.T = None
+        self.V = None
+        self.alpha = None
+        self.n_zw = None
+        self.n_dz = None
+        self.n_z = None
+        self.corpus = None
+
+        self.phi_norm = []
+        self.phi_softmax = []
+# ===========================================================================
+
+    def load_fit(self, path_file, n_it=0):
+        vars_loaded = pd.read_pickle(path_file)
+        for var in vars(self):
+            if var is 'snapshot':
+                setattr(self, 'snapshot', True)
+            else:
+                setattr(self, var, vars_loaded[var])
+        self.fit(n_it=n_it)
+# ===========================================================================
+
+    def fit(self, n_it, corpus=None):
+        if not self.snapshot:
+            self._initialise(corpus)
+        for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
+            if (it - 1) % 5 == 0:
[…]

### 2017-03-14 code/models/lda_gibbs.py

@@ -1,0 +1,142 @@
+
+import os
+import pandas as pd
+import numpy as np
+from time import time
+
+
+# ___________________________________________________________________________
+
+
+class LDA_Gibbs(object):
+
+    def __init__(self, K, snapshot=False, verbose_init=True):
+        self.n_it_sampl_last = 0
+        self.K = K
+        self.hist_thetas = []
+        self.hist_phis = []
+        self.snapshot = snapshot
+        self.corpus = None
+        self.verbose_init = verbose_init
+
+# ___________________________________________________________________________
+    def fit(self, n_it, corpus=None):
+        if not self.snapshot:
+            self._initialise(corpus)
+        for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
+            if (it % 5 == 0):
+                print("Iteration %s:" % it)
+            self._sample_topics()
+            self._store_theta()
+            self._store_phi()
+            if self.verbose_init:
+                self.verbose_init = False
+        self.n_it_sampl_last += n_it
+# ___________________________________________________________________________
+
+    def load_fit(self, path_file, n_it=0):
+        vars_loaded = pd.read_pickle(path_file)
+        for var in vars(self):
+            if var is 'snapshot':
+                setattr(self, 'snapshot', True)
+            else:
+                setattr(self, var, vars_loaded[var])
+        self.fit(n_it=n_it)
+# ___________________________________________________________________________
+
+    def _initialise(self, corpus):
+        t0 = time()
+        print('{} has started.'.format(self.__class__.__name__))
+        self.corpus = corpus
+        self.D, self.W = self.corpus.shape
+        '''
+        Initialise hyper-parameters
+        '''
+        self.alpha = 1. / self.K
+        self.beta = 1. / self.W
+        '''
+        Run the initial assignment.
+        '''
+        self._zw_counts = np.zeros((self.K, self.W), dtype=np.int)
+        self._dz_counts = np.zeros((self.D, self.K), dtype=np.int)
+        self._z_counts = np.zeros(self.K, dtype=np.int)
+        self._z_history = []
+        for d, doc in enumerate(corpus):
+            self._z_history.append([])
+            for w, word in enumerate(doc):
+                self._z_history[d].append([])
+                word_frequency = int(word)
+                for _ in range(word_frequency):
[…]

### 2017-03-14 code/pre-processing/create_bne-corpus.py

@@ -1,0 +1,114 @@
+
+import math
+import pymzml
+import pandas as pd
+import numpy as np
+import os
+import pickle as pkl
+
+np.set_printoptions(precision=3, threshold=np.inf)
+
+INTENSITY_FLOOR = 10
+TOLERANCE_FACTOR = 7
+HOME_PATH = os.path.expanduser('~')
+PATH_DATA = HOME_PATH + '/Projects/ssmsi/data/corpora_raw/'
+PATH_OUT = HOME_PATH + '/Projects/ssmsi/data/corpora_processed/'
+# ___________________________________________________________________________
+
+
+def generate_coordinates(data_size, number_of_rows, row_length,
+                         column_length):
+    single_scan_size = (data_size + 1) / number_of_rows
+    column_size = math.floor(single_scan_size *
+                             (column_length / (row_length + column_length)))
+    row_size = single_scan_size - column_size
+    coordinates = []
+
+    x_coord = 0
+    y_coord = 0
+    row_scan = True
+    scanning_right = True
+    for _ in range(0, number_of_rows):
+        for i in range(1, single_scan_size + 1):
+            coordinates.append((x_coord, y_coord))
+            # Switching the horizontal scanning to the vertical scanning.
+            if i % row_size == 0:
+                row_scan = not row_scan
+            if row_scan:
+                if scanning_right:
+                    x_coord += 1
+                else:
+                    x_coord -= 1
+            else:
+                y_coord += 1
+        # Reversing scanner's direction.
+        row_scan = not row_scan
+        scanning_right = not scanning_right
+    return coordinates
+# ___________________________________________________________________________
+
+
+def assign_class(classes, feature):
+    for class_assignment in classes:
+        if feature >= class_assignment[0] and feature <= class_assignment[1]:
+            return str(class_assignment)
+
+
+# ___________________________________________________________________________
+def main():
+    '''
+    Pre-processing the data
+    '''
+    run = pymzml.run.Reader(PATH_DATA + 'abcdefgh_1.mzML')
+    preprocessed_mnis = []
+    for spectrum in run:
+        for mass, intensity in spectrum.peaks:
+            if intensity > INTENSITY_FLOOR:
+                preprocessed_mnis.append((mass, intensity, spectrum['id']))
+    mnis_dtype = [('mass', float), ('intensity', float), ('id', int)]
[…]

### 2017-03-14 code/pre-processing/create_corpus.py

@@ -1,101 +1,0 @@
-
-import math
-import pymzml
-import pandas as pd
-import numpy as np
-import os
-np.set_printoptions(precision=3, threshold=np.inf)
-
-INTENSITY_FLOOR = 10
-TOLERANCE_FACTOR = 7
-HOME_PATH = os.path.expanduser('~')
-DATA_PATH = HOME_PATH + '/Projects/ssmsi/data/'
-PICKLES_PATH = HOME_PATH + '/Projects/ssmsi/heavy_pickles/'
-
-def assign_class(classes, feature):
-	for class_assignment in classes:
-		if feature >= class_assignment[0] and feature <= class_assignment[1]:
-			return str(class_assignment)
-
-def generate_coordinates(data_size, number_of_rows, row_length, column_length):
-	single_scan_size = (data_size + 1) / number_of_rows
-	column_size = math.floor(single_scan_size \
-		* (column_length / (row_length + column_length)))
-	row_size = single_scan_size - column_size
-	coordinates = []
-
-	x_coord = 0
-	y_coord = 0
-	row_scan = True
-	scanning_right = True
-	for _ in range(0, number_of_rows):
-		for i in range(1, single_scan_size + 1):
-			coordinates.append((x_coord, y_coord))
-			# Switching the horizontal scanning to the vertical scanning.
-			if i % row_size == 0:
-				row_scan = not row_scan
-			if row_scan:
-				if scanning_right:
-					x_coord += 1
-				else:
-					x_coord -= 1
-			else:
-				y_coord += 1
-		# Reversing scanner's direction.
-		row_scan = not row_scan
-		scanning_right = not scanning_right
-	return coordinates
-
-def main():
-	run = pymzml.run.Reader(DATA_PATH + 'abcdefgh_1.mzML')
-
-	# Reading the data
-	preprocessed_mnis = []
-	for spectrum in run:
-		for mass, intensity in spectrum.peaks:
-			if intensity > INTENSITY_FLOOR:
-				preprocessed_mnis.append((mass, intensity, spectrum['id']))
-	mnis_dtype = [('mass', float), ('intensity', float), ('id', int)]
-
-	# Sorting the data.
-	mnis = np.array(preprocessed_mnis, dtype=mnis_dtype)
-	mnis.sort(order='mass')
-
-	# Defining mass ranges for words.
-	words = []
-	starting_class_mass = mnis[0][0]
-	previous_mass = mnis[0][0]
[…]

### 2017-03-14 code/pre-processing/create_vocab.py

@@ -1,27 +1,0 @@
-import os
-import pandas as pd
-
-HOME_PATH = os.path.expanduser('~')
-PICKLES_PATH = HOME_PATH + '/Projects/ssmsi/heavy_pickles/'
-
-
-def create_vocab(corpus):
-	vocab = []
-	for doc in corpus:
-		for word in corpus[doc]:
-			if word not in vocab:
-				vocab.append(word)
-	vocab_panda = pd.Series(vocab)
-	return vocab_panda
-
-
-def main():
-	corpus_series = pd.read_pickle(HOME_PATH + '/Projects/tminm/heavy_pickles/corpus.pickle')
-	corpus = corpus_series.to_dict()
-
-	vocab_panda = create_vocab(corpus)
-	vocab_panda.to_pickle(HOME_PATH + PICKLES_PATH + 'vocab.pickle')
-
-if __name__ == '__main__':
-	main()
-	

### 2017-03-14 code/pre-processing/generator_corpus.py

@@ -1,0 +1,48 @@
+
+"""
+VERSION
+- Python 2
+
+FUNCTION
+- Corpus generation
+"""
+
+import numpy as np
+import string
+
+
+class Generator_Corpus(object):
+    '''
+    Terms:
+    - pp: pre-processed
+    '''
+    def __init__(self, alphas, beta, xi, T):
+        self.alphas = alphas
+        self.beta = beta
+        self.xi = xi
+        self.K, self.V = self.beta.shape
+        self.vocab = self._generate_vocab()
+        self.T = T
+
+    def _generate_vocab(self):
+        vocab = []
+        for w in range(self.V):
+            vocab.append(string.ascii_lowercase[w])
+        return np.array(vocab)
+# ___________________________________________________________________________
+
+    def generate_corpus(self):
+        corpus = {}
+        for d in range(self.T):
+            corpus[d] = {}
+            N = np.random.poisson(self.xi)
+            for w in range(N):
+                topic_distrib = np.random.multinomial(1, self.alphas[d])
+                z = np.where(topic_distrib == 1)[0][0]
+                word_distrib = np.random.multinomial(1, self.beta[z])
+                w = np.where(word_distrib == 1)[0][0]
+                w_key = self.vocab[w]
+                if w_key not in corpus[d]:
+                    corpus[d][w_key] = 0
+                corpus[d][w_key] += 1
+        return corpus

### 2017-03-14 code/pre-processing/plotter_corpus.py

@@ -1,0 +1,82 @@
+
+import numpy as np
+import math
+import plotly.graph_objs as go
+
+
+class Visualiser_Corpus(object):
+
+    def __init__(self, corpus, n_rows, l_row, l_column):
+        self.corpus = corpus
+        self.vocab = self._extract_vocab()
+        self.T = len(self.corpus)
+        self.V = len(self.vocab)
+        self.n_rows = n_rows
+        self.l_row = l_row
+        self.l_column = l_column
+        self.n_columns, self.map_coord = self._generate_coordinates()
+#____________________________________________________________________________
+
+    def _extract_vocab(self):
+        vocab = []
+        for doc in self.corpus:
+            for word in doc:
+                if word not in vocab:
+                    vocab.append(word)
+        return vocab
+#____________________________________________________________________________
+
+    def _generate_coordinates(self):
+        s_column, s_row = self._calc_sizes_map()
+        x_coord = 0
+        y_coord = self.n_rows - 1
+        scanning_right = True
+        map_coord = {}
+        for j in range(0, n_rows):
+            for i in range(0, s_row):
+                doc_id = i + s_column * j + s_row * j
+                map_coord[doc_id] = {'x': x_coord, 'y': y_coord}
+                # Switching horizontal scanning to vertical.
+                if (i % (s_row - 1) == 0 and i != 0):
+                    y_coord -= 1
+                    continue
+                if scanning_right:
+                    x_coord += 1
+                else:
+                    x_coord -= 1
+            # Reversing the scanner's direction.
+            scanning_right = not scanning_right
+        return s_row, map_coord
+
+    def _calc_sizes_map(self):
+        s_scan = (self.n_docs + 1) / self.n_rows
+        s_column_r = s_scan * (self.l_column / (self.l_row + self.l_column))
+        s_column = int(math.floor(s_column_r))
+        s_row = s_scan - s_column
+        return s_column, s_row
+#____________________________________________________________________________
+
+    def locate_word(self):
+        # Mapping to the z values.
+        z_matrix = np.zeros(shape=(self.n_rows, self.n_columns))
+        for i, doc in corpus.iteritems():
+            if word_key in doc and int(i) in self.map_coord:
+                doc_coord = map_coord[int(i)]
+                z_matrix[doc_coord['y']][doc_coord['x']] = doc[word_key]
+        # Creating the x and y axes.
+        x_vector = np.zeros(shape=(self.n_columns, 1))
[…]

### 2017-03-14 code/pre-processing/pre-process_corpus.py

@@ -1,56 +1,0 @@
-import os
-import pandas as pd
-import numpy as np
-np.set_printoptions(threshold=np.nan)
-from optparse import OptionParser
-
-op = OptionParser()
-op.add_option("--source",
-							action="store", type=str,
-							help="The choice of data origins.")
-op.add_option("--name",
-							action="store", type=str,
-							help="The choice of a data set.")
-op.add_option("--th",
-							action="store", type=float,
-							help="The max count threshold.")
-(opts, args) = op.parse_args()
-
-
-HOME_PATH = os.path.expanduser('~') + '/Projects/ssmsi/'
-DATA_PATH = HOME_PATH + 'pickles/corpora/' + opts.source + '/'
-AUX_DATA_PATH = HOME_PATH + 'pickles/corpora/a-to-h/'
-OUT_PATH = DATA_PATH
-INPUT_FILE_NAME = opts.name + '.pkl'
-OUTPUT_FILE_NAME = 'pp_' + INPUT_FILE_NAME
-
-THRESHOLD_ND = opts.th
-
-# Smaller corpus for Gibbs sampling
-def preprocess_corpus(corpus, vocab):
-	pp_corpus = np.zeros((len(corpus), len(vocab)), dtype=np.float)
-	for doc in corpus:
-		# Finding the maximum intensity.
-		max_ND = 0
-		for word in corpus[doc]:
-			if corpus[doc][word] > max_ND:
-				max_ND = corpus[doc][word]
-		factor_norm = THRESHOLD_ND / max_ND
-
-		# Populating the preprocessed corpus.
-		for w, word in enumerate(vocab):
-			if word in corpus[doc]:
-				count_norm = factor_norm * corpus[doc][word]
-				pp_corpus[doc][w] = int(count_norm)
-	return pp_corpus
-
-def main():
-	corpus = pd.read_pickle(DATA_PATH + INPUT_FILE_NAME)
-	vocab = pd.read_pickle(DATA_PATH + 'vocab_' + INPUT_FILE_NAME)
-	pp_corpus = preprocess_corpus(corpus, vocab)
-	pp_corpus.dump(OUT_PATH + OUTPUT_FILE_NAME)
-	print('Number of words per document: ')
-	print(np.sum(pp_corpus, axis=1))
-
-if __name__ == '__main__':
-	main()

### 2017-03-14 code/pre-processing/processor_corpus.py

@@ -1,0 +1,92 @@
+
+"""
+VERSION
+- Python 2
+
+FUNCTION
+- Corpus pre-prodcessing
+"""
+
+import numpy as np
+import math
+
+
+class Processor_Corpus(object):
+    '''
+    Terms:
+    - pp: pre-processed
+    '''
+    def __init__(self, corpus, vocab=None):
+        self.corpus = corpus
+        self.vocab = vocab
+        if self.vocab is None:
+            self.vocab = self._extract_vocab()
+        self.T = len(self.corpus)
+        self.V = len(self.vocab)
+
+    def _extract_vocab(self):
+        vocab = []
+        for doc in self.corpus.values():
+            for word in doc.keys():
+                if word not in vocab:
+                    vocab.append(word)
+        return np.array(vocab)
+# ___________________________________________________________________________
+
+    '''
+    Terms:
+    - pp: pre-processed
+    '''
+    def process_corpus(self, threshold, normalise):
+        corpus_pp = np.zeros((self.T, self.V), dtype=np.float)
+        for t, doc in self.corpus.items():
+            t = int(t)
+            '''
+            Assign the normalisation factor based on the max intensity
+            '''
+            if normalise:
+                max_ND = 0
+                for word in doc:
+                    if doc[word] > max_ND:
+                        max_ND = doc[word]
+                factor_norm = threshold / max_ND
+            else:
+                factor_norm = 1.
+            '''
+            Transforming into a numpy array
+            '''
+            for w, word in enumerate(self.vocab):
+                if word in doc:
+                    count_norm = factor_norm * doc[word]
+                    corpus_pp[t][w] = int(count_norm)
+        return corpus_pp
+# ___________________________________________________________________________
+
+    def generate_coordinates(self, n_rows, l_row, l_column):
+        s_column, s_row = self.calc_sizes(self.T, n_rows, l_row, l_column)
+        x_coord = 0
+        y_coord = n_rows - 1
+        scanning_right = True
+        map_coord = {}
+        for j in range(0, n_rows):
+            for i in range(0, s_row):
+                doc_id = i + s_column * j + s_row * j
+                map_coord[doc_id] = {'x': x_coord, 'y': y_coord}
+                # Switching horizontal scanning to vertical.
+                if (i % (s_row - 1) == 0 and i != 0):
+                    y_coord -= 1
+                    continue
+                if scanning_right:
+                    x_coord += 1
+                else:
[…]

### 2017-03-14 code/pre-processing/visualiser_corpus.py

@@ -1,0 +1,162 @@
+
+import numpy as np
+import math
+import plotly.graph_objs as go
+import matplotlib.pyplot as plt
+# ___________________________________________________________________________
+
+
+class Visualiser_Corpus(object):
+
+    def __init__(self, corpus, vocab, n_rows, l_row, l_column):
+        self.corpus = corpus
+        # self.vocab = self._extract_vocab()
+        self.vocab = vocab
+        self.T = len(self.corpus)
+        self.V = len(self.vocab)
+        self.n_rows = n_rows
+        self.l_row = l_row
+        self.l_column = l_column
+        self.n_columns, self.map_coord = self._generate_coordinates()
+# ___________________________________________________________________________
+
+    def _extract_vocab(self):
+        vocab = []
+        for doc in self.corpus:
+            for word in doc:
+                if word not in vocab:
+                    vocab.append(word)
+        return vocab
+# ___________________________________________________________________________
+
+    def _generate_coordinates(self):
+        s_column, s_row = self._calc_sizes_map()
+        x_coord = 0
+        y_coord = self.n_rows - 1
+        scanning_right = True
+        map_coord = {}
+        # print(s_row)
+        for j in range(0, self.n_rows):
+            for i in range(0, s_row):
+                doc_id = i + s_column * j + s_row * j
+                map_coord[doc_id] = {'x': x_coord, 'y': y_coord}
+                # Switching horizontal scanning to vertical.
+                if (i % (s_row - 1) == 0 and i != 0):
+                    y_coord -= 1
+                    continue
+                if scanning_right:
+                    x_coord += 1
+                else:
+                    x_coord -= 1
+            # Reversing the scanner's direction.
+            scanning_right = not scanning_right
+        return s_row, map_coord
+
+    def _calc_sizes_map(self):
+        s_scan = (self.T + 1) / self.n_rows
+        s_column_r = s_scan * (self.l_column / (self.l_row + self.l_column))
+        s_column = int(math.floor(s_column_r))
+        s_row = s_scan - s_column
+        return s_column, int(s_row)
+# ___________________________________________________________________________
+
+    '''
+    Used on a pre-processed corpus
+    '''
+    def locate_word(self, word):
+        # Mapping to the z values.
+        z_matrix = np.zeros(shape=(self.n_rows, self.n_columns))
[…]

### 2017-03-14 code/topic-models/alpha-update/dtm_alpha.py

@@ -1,308 +1,0 @@
-
-"""
-VERSION
-- Python 2
-
-FUNCTION
-- Infer topics by updating the alphas.
-"""
-
-import os
-import pandas as pd
-import numpy as np
-import copy
-from scipy.stats import norm
-import pickle as pkl
-
-np.set_printoptions(threshold=np.nan)
-
-
-class DTM_Alpha(object):
-
-    '''
-    PARAMETERS:
-    - sigma_0_sq is the initial variance for alpha[0];
-    - sigma_sq is the variance for alpha[t]
-    - delta_sq is the variance for alpha'[t]
-    '''
-    def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None, delta_sq=None,
-                 autoreg=False, snapshot=False):
-        self.K = K
-        self.sigma_0_sq = sigma_0_sq
-        self.sigma_sq = sigma_sq
-        self.delta_sq = delta_sq
-        self.autoreg = autoreg
-        self.snapshot = snapshot
-        '''
-        Counter and array initialisation
-        '''
-        self.n_it_sampl_last = 0
-        self.n_upd_alpha = 0
-        self.n_it_alpha = 0
-        self.hist_theta = []
-        self.hist_alpha = []
-        self.hist_z = []
-        '''
-        The parameters initialised later in the work-flow
-        '''
-        self.beta = None
-        self.T = None
-        self.V = None
-        self.alpha = None
-        self.n_zw = None
-        self.n_dz = None
-        self.n_z = None
-        self.corpus = None
-# ___________________________________________________________________________
-
-    def fit(self, n_it, corpus=None):
-        if not self.snapshot:
-            self._initialise(corpus)
-        for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
-            if it != 0 and it % 10 == 0:
-                print('Iteration: %s' % it)
-                print('Total alpha changes: %s' % self.n_upd_alpha)
-                print('Alpha update rate: %.4f'
-                      % (1.0 * self.n_upd_alpha / self.n_it_alpha))
-            self.hist_alpha.append(copy.deepcopy(self.alpha))
-            self._update_alpha()
-            # break
-            self._sample_topics()
-            self._store_theta()
-        self.n_it_sampl_last += n_it
-# ___________________________________________________________________________
-
[…]

### 2017-03-14 code/topic-models/lda_gibbs.py

@@ -1,116 +1,0 @@
-import os
-import pandas as pd
-import numpy as np
-np.set_printoptions(edgeitems=5)
-
-'''
-References:
-- Griffiths and Stevyers (2004)
-'''
-
-class LDA_Gibbs(object):
-
-	def __init__(self, K, iter_count):
-		self.iter_count = iter_count
-		self.K = K
-		self.thetas = []
-
-	def fit(self, corpus):
-		self._fit(corpus)
-
-	def dump_thetas(self, path):
-		pickle = np.array(self.thetas)
-		pickle.dump(path)
-
-	def _fit(self, corpus):
-		self._initialise(corpus)
-
-		for it in range(self.iter_count):
-			print("Iteration %s:" % it)
-			self._sample_topics(corpus)
-			self._store_theta()
-			print("Current theta:")
-			print(self.thetas[it])
-
-	def _initialise(self, corpus):
-		self.D, self.W = corpus.shape
-		self._total_frequencies = corpus.sum()
-		self.alpha = 1. / self.K
-		self.beta = 1. / self.W
-
-		self._zw_counts = np.zeros((self.K, self.W), dtype=np.int)
-		self._dz_counts = np.zeros((self.D, self.K), dtype=np.int)
-		self._z_counts = np.zeros(self.K, dtype=np.int)
-		self._z_history = []
-
-		for d, doc in enumerate(corpus):
-			self._z_history.append([])
-			for w, word in enumerate(doc):
-				self._z_history[d].append([])
-				word_frequency = int(word)
-				for _ in range(word_frequency):
-					z = np.random.randint(self.K)
-					self._z_history[d][w].append(z)
-					self._dz_counts[d, z] += 1
-					self._zw_counts[z, w] += 1
-					self._z_counts[z] += 1
-
-	def _sample_topics(self, corpus):
-		for d, doc in enumerate(corpus):
-			for w, _ in enumerate(doc):
-				for it, z in enumerate(self._z_history[d][w]):
-					self._dz_counts[d, z] -= 1
-					self._zw_counts[z, w] -= 1
-					self._z_counts[z] -= 1
-
-					# Eq. 5, the second denominator is  a constant
-					p_topic = ((1.0 * self._zw_counts[:, w] + self.beta) / \
-						(1.0 * self._z_counts + self.W * self.beta)) * \
-						((self._dz_counts[d, :] + self.alpha))
[…]

### 2017-03-15 code/models/dtm_alpha.py
Experiment: obtaining the latent betas.

@@ -39,12 +39,12 @@ class DTM_Alpha(object):
         '''
         Counter and array initialisation
         '''
-        self.n_it_sampl_last = 0
-        self.n_upd_alpha = 0
-        self.n_it_alpha = 0
-        self.hist_theta = []
-        self.hist_alpha = []
-        self.hist_z = []
+        self.n_it_sampl_last = None
+        self.n_upd_alpha = None
+        self.n_it_alpha = None
+        self.hist_theta = None
+        self.hist_alpha = None
+        self.hist_phi = None
         '''
         The parameters initialised later in the work-flow
         '''
@@ -55,10 +55,8 @@ class DTM_Alpha(object):
         self.n_zw = None
         self.n_dz = None
         self.n_z = None
+        self.hist_z = None
         self.corpus = None
-
-        self.phi_norm = []
-        self.phi_softmax = []
 # ===========================================================================
 
     def load_fit(self, path_file, n_it=0):
@@ -75,7 +73,7 @@ class DTM_Alpha(object):
         if not self.snapshot:
             self._initialise(corpus)
         for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
-            if (it - 1) % 5 == 0:
+            if it != 0 and it % 25 == 0:
                 self._print_status(it)
             self._update_alpha()
             self._sample_topics()
@@ -83,6 +81,9 @@ class DTM_Alpha(object):
             self._store_theta()
             if self.verbose_init:
                 self.verbose_init = False
+            if it == 50:
+                # print(self.hist_theta[0])
+                print(self.hist_phi[it])
         self.n_it_sampl_last += n_it
 
     def _print_status(self, it):
@@ -94,6 +95,15 @@ class DTM_Alpha(object):
     def _initialise(self, corpus):
         t0 = time()
         print('{} has started.'.format(self.__class__.__name__))
+        '''
+        Initialise auxiliary variables
+        '''
+        self.n_it_sampl_last = 0
+        self.n_upd_alpha = 0
+        self.n_it_alpha = 0
+        self.hist_theta = []
+        self.hist_alpha = []
+        self.hist_phi = []
         self.corpus = corpus
         self.T, self.V = corpus.shape
         '''
[…]

### 2017-03-15 code/pre-processing/processor_corpus.py

@@ -4,7 +4,7 @@ VERSION
 - Python 2
 
 FUNCTION
-- Corpus pre-prodcessing
+- Corpus pre-processing
 """
 
 import numpy as np

### 2017-03-15 code/pre-processing/visualiser_corpus.py

@@ -18,7 +18,13 @@ class Visualiser_Corpus(object):
         self.l_row = l_row
         self.l_column = l_column
         self.n_columns, self.map_coord = self._generate_coordinates()
+        self.linspace_t = np.linspace(0, self.T - 1, num=self.T)
+        self.linspace_V = np.linspace(0, self.V - 1, num=self.V)
+
 # ___________________________________________________________________________
+    '''
+    TODO: Enable to vocab creation
+    '''
 
     def _extract_vocab(self):
         vocab = []
@@ -58,6 +64,11 @@ class Visualiser_Corpus(object):
         s_column = int(math.floor(s_column_r))
         s_row = s_scan - s_column
         return s_column, int(s_row)
+
+# ___________________________________________________________________________
+    def softmax(self, arr):
+        return np.exp(arr) / np.sum(np.exp(arr))
+
 # ___________________________________________________________________________
 
     '''
@@ -122,7 +133,7 @@ class Visualiser_Corpus(object):
 
 # ___________________________________________________________________________
 
-    def plot_latent_alpha(self, history_alpha, n_it, T, var_init, var_basic,
+    def plot_latent_alpha(self, history_alpha, n_it, var_init, var_basic,
                           var_prop):
         '''
         Calculate the softmax values
@@ -148,15 +159,43 @@ class Visualiser_Corpus(object):
         from matplotlib.ticker import MaxNLocator
         ax = plt.figure().gca()
         ax.xaxis.set_major_locator(MaxNLocator(integer=True))
-        linspace_t = np.linspace(0, T - 1, num=T)
         for y_arr, label, setting in zip(curves, labels, settings):
-            plt.plot(linspace_t, y_arr, label=label, color=setting[0],
+            plt.plot(self.linspace_t, y_arr, label=label, color=setting[0],
                      alpha=setting[1])
         title_fig = '$\sigma_0^2=%.2f,\quad \sigma^2=%.4f,\quad \delta^2=%.1f$' % (var_init, var_basic, var_prop)
         plt.title(title_fig)
         plt.legend(loc=1)
         plt.xlabel('t')
-        plt.ylabel('$\theta$')
+        plt.ylabel(r'$\theta$')
 
-    def softmax(self, arr):
-        return np.exp(arr) / np.sum(np.exp(arr))
+# ___________________________________________________________________________
+
[…]

### 2017-03-17 code/models/dtm_alpha.py
Establishing a standard for the corpus generation, generalising the plotting functions.

@@ -26,15 +26,13 @@ class DTM_Alpha(object):
     - delta_sq is the variance for alpha'[t]
     '''
     def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None, delta_sq=None,
-                 autoreg=False, snapshot=False, beta=None,
-                 verbose_init=True):
+                 autoreg=False, snapshot=False, verbose_init=True):
         self.K = K
         self.sigma_0_sq = sigma_0_sq
         self.sigma_sq = sigma_sq
         self.delta_sq = delta_sq
         self.autoreg = autoreg
         self.snapshot = snapshot
-        self.beta = beta
         self.verbose_init = verbose_init
         '''
         Counter and array initialisation
@@ -48,6 +46,7 @@ class DTM_Alpha(object):
         '''
         The parameters initialised later in the work-flow
         '''
+        self.burnin = None
         self.beta = None
         self.T = None
         self.V = None
@@ -59,45 +58,52 @@ class DTM_Alpha(object):
         self.corpus = None
 # ===========================================================================
 
-    def load_fit(self, path_file, n_it=0):
+    def load_fit(self, path_file, n_it_add=0):
         vars_loaded = pd.read_pickle(path_file)
         for var in vars(self):
             if var is 'snapshot':
                 setattr(self, 'snapshot', True)
             else:
                 setattr(self, var, vars_loaded[var])
-        self.fit(n_it=n_it)
+        self.fit(n_it=n_it_add)
 # ===========================================================================
 
-    def fit(self, n_it, corpus=None):
+    def fit(self, n_it, n_burn_it=0, corpus=None):
+        print('{} has started.'.format(self.__class__.__name__))
         if not self.snapshot:
             self._initialise(corpus)
+        if self.burnin:
+            print('The burn-in started.')
         for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
-            if it != 0 and it % 25 == 0:
-                self._print_status(it)
+            self._update_status(it, n_burn_it)
             self._update_alpha()
             self._sample_topics()
-            self._store_phi()
-            self._store_theta()
+            if not self.burnin:
+                self._store_phi()
+                self._store_theta()
+                # self._calculate_perplexity()
[…]

### 2017-03-17 code/pre-processing/generator_corpus.py

@@ -17,8 +17,10 @@ class Generator_Corpus(object):
     - pp: pre-processed
     '''
     def __init__(self, alphas, beta, xi, T):
-        self.alphas = alphas
+        self.alphas = self.softmax(alphas)
+        # TODO: apply softmax for non-trivial betas
         self.beta = beta
+        # self.beta = self.softmax(beta)
         self.xi = xi
         self.K, self.V = self.beta.shape
         self.vocab = self._generate_vocab()
@@ -45,4 +47,11 @@ class Generator_Corpus(object):
                 if w_key not in corpus[d]:
                     corpus[d][w_key] = 0
                 corpus[d][w_key] += 1
+
         return corpus
+
+    def softmax(self, arr):
+        arr_softmax = np.array(arr)
+        for idx_r, row in enumerate(arr):
+            arr_softmax[idx_r] = np.exp(row) / np.sum(np.exp(row))
+        return arr_softmax

### 2017-03-17 code/pre-processing/helper_experiment.py

@@ -1,0 +1,84 @@
+
+"""
+VERSION
+- Python 2
+
+FUNCTION
+- Functions to be used during the experiments
+"""
+
+import numpy as np
+
+
+class Helper_Experiment(object):
+
+    def calculate_performance(self, alpha_init, beta_init, theta_init,
+                              clf, s_batch):
+        hist_theta = np.array(clf.hist_theta)
+        generator_batch = self._chunks(hist_theta, s_batch)
+        performance_batch_theta = []
+        performance_batch_phi = []
+        '''
+        Calculating the diff in thetas
+        '''
+        for hist_theta_batch in generator_batch:
+            theta_latent = np.average(hist_theta_batch, axis=0)
+            # TODO: Fix theta tuning with non pre-set alphas
+            matchings_topic = self._tune_thetas(alpha_init, theta_latent)
+            theta_init = self._softmax_matrix(alpha_init).T
+            theta_latent = theta_latent.T
+            diffs_theta = []
+            for idx_init, row_init in enumerate(theta_init):
+                idx_latent = matchings_topic[idx_init]
+                diff_theta = np.abs(row_init - theta_latent[idx_latent, :]).sum()
+                diffs_theta.append(diff_theta)
+            sum_theta = np.sum(diffs_theta)
+            performance_batch_theta.append(sum_theta)
+        '''
+        Calculating the diff in phis
+        '''
+        hist_phi = np.array(clf.hist_phi)
+        generator_batch = self._chunks(hist_phi, s_batch)
+        for hist_phi_batch in generator_batch:
+
+            hist_phi = np.array(clf.hist_phi)
+            phi_latent = np.average(hist_phi_batch, axis=0)
+            diffs_phi = []
+            for idx_init, row_init in enumerate(beta_init):
+                idx_latent = matchings_topic[idx_init]
+                diff_phi = np.abs(row_init - phi_latent[idx_latent, :]).sum()
+                diffs_phi.append(diff_phi)
+            sum_phi = np.sum(diffs_phi)
+            performance_batch_phi.append(sum_phi)
+        return (performance_batch_theta, performance_batch_phi)
+
+    def _chunks(self, l, n):
+        """Yield successive n-sized chunks from l."""
+        for i in range(0, len(l), n):
+            yield l[i:i + n]
+
+    def _tune_thetas(self, alpha_init, theta_latent):
+        theta_init = self._softmax_matrix(alpha_init)
+        theta_init = theta_init.T
[…]

### 2017-03-17 code/pre-processing/visualiser_corpus.py

@@ -69,6 +69,12 @@ class Visualiser_Corpus(object):
     def softmax(self, arr):
         return np.exp(arr) / np.sum(np.exp(arr))
 
+    def softmax_matrix(self, arr):
+        arr_softmax = np.array(arr)
+        for idx_r, row in enumerate(arr):
+            arr_softmax[idx_r] = np.exp(row) / np.sum(np.exp(row))
+        return arr_softmax
+
 # ___________________________________________________________________________
 
     '''
@@ -131,9 +137,159 @@ class Visualiser_Corpus(object):
         fig = go.Figure(data=data, layout=layout)
         return fig
 
-# ___________________________________________________________________________
 
-    def plot_latent_alpha(self, history_alpha, n_it, var_init, var_basic,
+# __________________________________________________________________________
+
+    def compare_performances(self, perf_ar, perf_nonar):
+        theta_ar, phi_ar = perf_ar
+        theta_nonar, phi_nonar = perf_nonar
+        plt.figure()
+        plt.title('theta')
+        plt.plot(theta_ar, label='AR')
+        plt.plot(theta_nonar, label='Non-AR')
+        plt.legend(loc=2)
+        plt.figure()
+        plt.title('phi')
+        plt.plot(phi_ar, label='AR')
+        plt.plot(phi_nonar, label='Non-AR')
+        plt.legend(loc=2)
+
+
+
+        # print(theta_ar)
+        # plt.bar(self.linspace_V, phi[0], label='k=0')
+        # plt.bar(self.linspace_V, phi[1], label='k=1')
+        # plt.xticks(np.arange(0, 10, 1.0))
+        # plt.legend(loc=1)
+        # plt.xlabel('V')
+        # plt.ylabel(r'$\phi$')
+
+# __________________________________________________________________________
+# LATENT PLOTS
+
+    def plot_latent_thetas(self, thetas):
+        '''
+        Axis formatting
+        '''
+        from matplotlib.ticker import MaxNLocator
+        ax = plt.figure().gca()
+        ax.xaxis.set_major_locator(MaxNLocator(integer=True))
+        '''
+        Pre-processing
+        '''
+        thetas = thetas.T
+        '''
+        Plotting
+        '''
+        for theta in thetas:
+            plt.plot(self.linspace_t, theta)
+        plt.xlabel('t')
+        plt.ylabel(r'$\theta$')
+
+    def plot_latent_phis(self, phis, title):
+        '''
+        Axis formatting
[…]

### 2017-03-21 code/pre-processing/generator_corpus.py
Defined the preliminary evaluation settings.

@@ -16,11 +16,12 @@ class Generator_Corpus(object):
     Terms:
     - pp: pre-processed
     '''
-    def __init__(self, alphas, beta, xi, T):
-        self.alphas = self.softmax(alphas)
-        # TODO: apply softmax for non-trivial betas
-        self.beta = beta
-        # self.beta = self.softmax(beta)
+    def __init__(self, alpha_init, theta_init, beta_init, xi, T):
+        if alpha_init is not None:
+            self.theta = self._softmax_matrix(alpha_init)
+        elif theta_init is not None:
+            self.theta = theta_init
+        self.beta = beta_init
         self.xi = xi
         self.K, self.V = self.beta.shape
         self.vocab = self._generate_vocab()
@@ -39,7 +40,7 @@ class Generator_Corpus(object):
             corpus[d] = {}
             N = np.random.poisson(self.xi)
             for w in range(N):
-                topic_distrib = np.random.multinomial(1, self.alphas[d])
+                topic_distrib = np.random.multinomial(1, self.theta[d])
                 z = np.where(topic_distrib == 1)[0][0]
                 word_distrib = np.random.multinomial(1, self.beta[z])
                 w = np.where(word_distrib == 1)[0][0]
@@ -50,7 +51,7 @@ class Generator_Corpus(object):
 
         return corpus
 
-    def softmax(self, arr):
+    def _softmax_matrix(self, arr):
         arr_softmax = np.array(arr)
         for idx_r, row in enumerate(arr):
             arr_softmax[idx_r] = np.exp(row) / np.sum(np.exp(row))

### 2017-03-21 code/pre-processing/helper_experiment.py

@@ -15,64 +15,86 @@ class Helper_Experiment(object):
     def calculate_performance(self, alpha_init, beta_init, theta_init,
                               clf, s_batch):
         hist_theta = np.array(clf.hist_theta)
-        generator_batch = self._chunks(hist_theta, s_batch)
+        theta_init = theta_init.T
+        generator_batch = self._split_to_batches(hist_theta, s_batch)
         performance_batch_theta = []
         performance_batch_phi = []
         '''
-        Calculating the diff in thetas
+        Iterate through the theta batches
         '''
         for hist_theta_batch in generator_batch:
+            '''
+            Pre-process and tune the thetas
+            '''
             theta_latent = np.average(hist_theta_batch, axis=0)
-            # TODO: Fix theta tuning with non pre-set alphas
-            matchings_topic = self._tune_thetas(alpha_init, theta_latent)
-            theta_init = self._softmax_matrix(alpha_init).T
+            if alpha_init is not None:
+                theta_init = self._softmax_matrix(alpha_init)
             theta_latent = theta_latent.T
+            matchings_topic = self._tune_thetas(theta_init, theta_latent)
+            '''
+            Calculate the theta similarity
+            '''
             diffs_theta = []
             for idx_init, row_init in enumerate(theta_init):
                 idx_latent = matchings_topic[idx_init]
-                diff_theta = np.abs(row_init - theta_latent[idx_latent, :]).sum()
+                diff_theta = np.abs(row_init - theta_latent[idx_latent, :])
+                diff_theta = diff_theta.sum()
                 diffs_theta.append(diff_theta)
             sum_theta = np.sum(diffs_theta)
             performance_batch_theta.append(sum_theta)
         '''
-        Calculating the diff in phis
+        Iterate through the phi batches
         '''
         hist_phi = np.array(clf.hist_phi)
-        generator_batch = self._chunks(hist_phi, s_batch)
+        generator_batch = self._split_to_batches(hist_phi, s_batch)
         for hist_phi_batch in generator_batch:
-
-            hist_phi = np.array(clf.hist_phi)
             phi_latent = np.average(hist_phi_batch, axis=0)
+            '''
+            Calculate the phi similarity
+            '''
             diffs_phi = []
             for idx_init, row_init in enumerate(beta_init):
                 idx_latent = matchings_topic[idx_init]
[…]

### 2017-03-21 code/pre-processing/plotter_corpus.py

@@ -15,7 +15,7 @@ class Visualiser_Corpus(object):
         self.l_row = l_row
         self.l_column = l_column
         self.n_columns, self.map_coord = self._generate_coordinates()
-#____________________________________________________________________________
+# ___________________________________________________________________________
 
     def _extract_vocab(self):
         vocab = []
@@ -24,7 +24,7 @@ class Visualiser_Corpus(object):
                 if word not in vocab:
                     vocab.append(word)
         return vocab
-#____________________________________________________________________________
+# ___________________________________________________________________________
 
     def _generate_coordinates(self):
         s_column, s_row = self._calc_sizes_map()
@@ -54,7 +54,7 @@ class Visualiser_Corpus(object):
         s_column = int(math.floor(s_column_r))
         s_row = s_scan - s_column
         return s_column, s_row
-#____________________________________________________________________________
+# ___________________________________________________________________________
 
     def locate_word(self):
         # Mapping to the z values.

### 2017-03-21 code/pre-processing/visualiser_corpus.py

@@ -3,14 +3,17 @@ import numpy as np
 import math
 import plotly.graph_objs as go
 import matplotlib.pyplot as plt
+import os
 # ___________________________________________________________________________
 
 
 class Visualiser_Corpus(object):
 
-    def __init__(self, corpus, vocab, n_rows, l_row, l_column):
+    def __init__(self, corpus, vocab, n_rows, l_row, l_column, path_plots,
+                 save=True):
         self.corpus = corpus
         # self.vocab = self._extract_vocab()
+        self.path_plots = path_plots
         self.vocab = vocab
         self.T = len(self.corpus)
         self.V = len(self.vocab)
@@ -20,6 +23,10 @@ class Visualiser_Corpus(object):
         self.n_columns, self.map_coord = self._generate_coordinates()
         self.linspace_t = np.linspace(0, self.T - 1, num=self.T)
         self.linspace_V = np.linspace(0, self.V - 1, num=self.V)
+        self.save = save
+
+        if not os.path.exists(self.path_plots):
+            os.makedirs(self.path_plots)
 
 # ___________________________________________________________________________
     '''
@@ -140,34 +147,44 @@ class Visualiser_Corpus(object):
 
 # __________________________________________________________________________
 
-    def compare_performances(self, perf_ar, perf_nonar):
+    def compare_performances(self, perf_ar, perf_nonar, s_batch):
         theta_ar, phi_ar = perf_ar
         theta_nonar, phi_nonar = perf_nonar
-        plt.figure()
-        plt.title('theta')
+        '''
+        Axis formatting
+        '''
+        from matplotlib.ticker import MaxNLocator
+        '''
+        Plotting thetas
+        '''
+        ax = plt.figure().gca()
+        ax.xaxis.set_major_locator(MaxNLocator(integer=True))
+        plt.title(r'$\theta$')
         plt.plot(theta_ar, label='AR')
         plt.plot(theta_nonar, label='Non-AR')
         plt.legend(loc=2)
-        plt.figure()
-        plt.title('phi')
+        plt.ylabel('Total error')
+        plt.xlabel('Batch no. (s_batch = {})'.format(s_batch))
+        if self.save:
+            plt.savefig(self.path_plots + 'performance_thetas.png', dpi=300)
+        '''
+        Plotting phis
+        '''
+        ax = plt.figure().gca()
[…]

### 2017-03-23 code/models/dtm_alpha.py
Impl: fixing the topic assignment; paper: 2/3 of the introduction.

@@ -25,8 +25,9 @@ class DTM_Alpha(object):
     - sigma_sq is the variance for alpha[t]
     - delta_sq is the variance for alpha'[t]
     '''
-    def __init__(self, K=None, sigma_0_sq=None, sigma_sq=None, delta_sq=None,
-                 autoreg=False, snapshot=False, verbose_init=True):
+    def __init__(self, K=None, beta=None, sigma_0_sq=None, sigma_sq=None,
+                 delta_sq=None, autoreg=False, snapshot=False,
+                 verbose_init=True):
         self.K = K
         self.sigma_0_sq = sigma_0_sq
         self.sigma_sq = sigma_sq
@@ -34,6 +35,7 @@ class DTM_Alpha(object):
         self.autoreg = autoreg
         self.snapshot = snapshot
         self.verbose_init = verbose_init
+        self.beta = beta
         '''
         Counter and array initialisation
         '''
@@ -47,7 +49,6 @@ class DTM_Alpha(object):
         The parameters initialised later in the work-flow
         '''
         self.burnin = None
-        self.beta = None
         self.T = None
         self.V = None
         self.alpha = None
@@ -77,16 +78,15 @@ class DTM_Alpha(object):
         for it in range(self.n_it_sampl_last, self.n_it_sampl_last + n_it):
             self._update_status(it, n_burn_it)
             self._update_alpha()
-            self._sample_topics()
+            self._assign_topics()
             if not self.burnin:
-                self._store_phi()
-                self._store_theta()
+                self._sample_phi()
+                self._sample_theta()
                 # self._calculate_perplexity()
             if self.verbose_init:
                 self.verbose_init = False
         self.n_it_sampl_last += n_it
         print('{} has finished.'.format(self.__class__.__name__))
-
 
     def _update_status(self, it, n_burn_it):
         if it != 0 and it % 50 == 0:
@@ -115,8 +115,6 @@ class DTM_Alpha(object):
         '''
         Initialise hyper-parameters
         '''
-        if self.beta is None:
-            self._init_beta()
         self.alpha = np.zeros(shape=(self.T, self.K))
         self.alpha[0] = np.random.normal(0, self.sigma_0_sq, self.K)
[…]

### 2017-03-23 code/pre-processing/visualiser_corpus.py

@@ -9,8 +9,8 @@ import os
 
 class Visualiser_Corpus(object):
 
-    def __init__(self, corpus, vocab, n_rows, l_row, l_column, path_plots,
-                 save=True):
+    def __init__(self, corpus, vocab, n_rows, l_row, l_column,
+                 path_plots=None, save=True):
         self.corpus = corpus
         # self.vocab = self._extract_vocab()
         self.path_plots = path_plots
@@ -25,8 +25,9 @@ class Visualiser_Corpus(object):
         self.linspace_V = np.linspace(0, self.V - 1, num=self.V)
         self.save = save
 
-        if not os.path.exists(self.path_plots):
-            os.makedirs(self.path_plots)
+        if self.path_plots is not None:
+            if not os.path.exists(self.path_plots):
+                os.makedirs(self.path_plots)
 
 # ___________________________________________________________________________
     '''

### 2017-03-29 docs/reports/experiments_AR-performance/report_AR-performance.tex
Experiment: wrote a report; Paper: started the lit. review

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
\setlength{\parskip}{1em}

\usepackage[a4paper,top=1in,bottom=1in,left=1in,right=1in]{geometry}

% \pagestyle{empty}
\graphicspath{ {images/} }
% ___________________________________________________________________

\begin{document}

\begingroup  
  \centering
  \large Report: AR and Non-AR model performance comparison \par
  \large Arijus Pleska \par
\endgroup
% ___________________________________________________________________

\section*{Experiment I}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\linewidth]{init_theta-1.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.4\textwidth}
    \includegraphics[width=\linewidth]{init_phi-1.png}
  \end{subfigure}%
  \caption{$\theta$ and $\phi$ of the Experiment I corpora.}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas-1.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_1.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_1.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis-1.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_1.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_1.png}
  \end{subfigure}%
  \caption{Corpus no. $1$.}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-2.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-2.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-2.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-2.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-2.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-2.png}
  \end{subfigure}%
  \caption{Corpus no. $2$.}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-3.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-3.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-3.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-3.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-3.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-3.png}
  \end{subfigure}%
  \caption{Corpus no. $3$.}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-4.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-4.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-4.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-4.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-4.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-4.png}
  \end{subfigure}%
  \caption{Corpus no. $4$.}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-5.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-5.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-5.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-5.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-5.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-5.png}
  \end{subfigure}%
  \caption{Corpus no. $5$.}

\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-6.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-6.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-6.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-6.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-6.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-6.png}
  \end{subfigure}%
  \caption{Corpus no. $6$.}

\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-7.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-7.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-7.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-7.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-7.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-7.png}
  \end{subfigure}%
  \caption{Corpus no. $7$.}

\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-8.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-8.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-8.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-8.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-8.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-8.png}
  \end{subfigure}%
  \caption{Corpus no. $8$.}

\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-9.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-9.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-9.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-9.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-9.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-9.png}
  \end{subfigure}%
  \caption{Corpus no. $9$.}

\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_thetas_experiment-1|overlapping-yes_dataset-10.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_AR_experiment-1|overlapping-yes_dataset-10.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_AR_experiment-1|overlapping-yes_dataset-10.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{performance_phis_experiment-1|overlapping-yes_dataset-10.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_thetas_Non-AR_experiment-1|overlapping-yes_dataset-10.png}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\linewidth]{latent_phis_Non-AR_experiment-1|overlapping-yes_dataset-10.png}
  \end{subfigure}%
  \caption{Corpus no. $10$.}

\end{figure}

% \par The purpose of this report is to assesses whether an auto-regressive topic model is capable to recover the $\alpha$ parameter used in the synthetic data generation process. This report is structured in the following order: at the start, the rationale of the used Bayesian methods are covered; then, the experiment settings are defined; in the next stage, we assess the experiment results; finally, the identified issues are outlined to be discussed during next meeting.
% ___________________________________________________________________

% \section*{Preliminaries}
% % TODO: Fix the pi expression (remove P(X|...))§

% \par In order to understand the experiment settings, it is necessary to be familiar with the auto-regressive and non auto-regressive $\alpha$ priors as well as the Metropolis--Hastings (M--H) algorithm; the expressions of the latter concepts are listed below:
% \begin{enumerate}
%   \item The auto-regressive $\alpha$ prior:
%     \begin{align*}
%     p(\alpha_0,\ldots,\alpha_T)=p(\alpha_0)\prod_{t=1}^{T}{p(\alpha_t|\alpha_{t-1})}; \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t=0;\\
%     & p(\alpha_t)=f(\alpha_t; \alpha_{t-1}, \sigma^2I), \quad t>0.
%     \end{align*}
%   \item The non auto-regressive $\alpha$ prior:
%     \begin{align*}
%     p(\alpha_0,\ldots,\alpha_T)=\prod_{t=0}^{T}{p(\alpha_t}); \qquad &p(\alpha_t)=f(\alpha_0; 0, \sigma_0^2I), \quad t \geq 0.
% \end{align*}
%   \item The rationale of the utilised M--H algorithm variation:\\\\
%     For the start, familiarise with the following expressions:
%     \begin{align*}
%       \dfrac{A(x'|x)}{A(x|x')} = \dfrac{p(x'|x)}{p(x|x')}\cdot \dfrac{q(x'|x)}{q(x|x')} = \dfrac{p(x',x)}{p(x)}\cdot \dfrac{p(x')}{p(x,x')} = \dfrac{p(x')}{p(x)}; \qquad &x' \sim q(x,\delta^2I);\\
%       & q = \mathcal{N} \Rightarrow q(x'|x) = q(x|x');
%     \end{align*}
%     where $x$ is the current state, $x'$ is the proposed state, $A$ is the acceptance distribution, and $q$ is the proposal distribution. Taking the previous results into account, the acceptance rate $r$ is expressed as follows:
%     \begin{align*}
%       r = \min{\bigg(1, \dfrac{p(x')}{p(x)}\bigg)}.
%     \end{align*}
%   \item The application of the M--H algorithm to update $\alpha$:\\\\
%     Note that the $\alpha$ values are updated independently; that is, the expression for the acceptance rate below is for a single $\alpha$ entry.
%     \begin{align*}
%       \dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)} &= \dfrac{p(X|z,\alpha^{-tk},\alpha_{t, k}')\cdot p(z|\alpha^{-tk},\alpha_{t, k}')\cdot p(\alpha^{-tk},\alpha_{t, k}')}{p(X)}\cdot \dfrac{p(X)}{p(X|z,\alpha)\cdot p(z|\alpha)\cdot p(\alpha)}\\
%       &= \dfrac{\prod_{k=0}^K\big[\pi(\alpha_{t}')_k^{z_{t,k}}\big]\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\prod_{k=0}^K\big[\pi(\alpha_t)_k^{z_{t,k}}\big]\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)}; \qquad t > 0, \quad t \neq T;
%     \end{align*}
%     where $\pi$ is the softmax function, $\alpha_{t,k}' \sim \mathcal{N}(\alpha_{t, k}, \delta^2)$, and $\alpha^{-tk}$ denotes $\alpha$ without $\alpha_{t,k}'$; also, for the boundary cases $t=0$ and $t=T$, note that the $p(\alpha)$ term corresponds to $p(\alpha_{0})\cdot p(\alpha_{1}|\alpha_{0})$ and $p(\alpha_{T}|\alpha_{T-1})$ respectively. Further, for computational stability, the previous ratio would be computed in the log space as follows: 
%     \begin{align*}
%       \log{\bigg[\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)}\bigg]} &= \log{\big[p(z,\alpha^{-tk},\alpha_{t, k}'|X)\big]} - \log{\big[p(z,\alpha|X)\big]}; \quad \mbox{where}\\
%       \log{\big[p(z,\alpha^{-tk},\alpha_{t, k}'|X)\big]} &=\sum_{k=0}^K\bigg[z_{t,k}\cdot\log\big[\pi(\alpha_{t}')_k\big]\bigg] + \log\big[p(\alpha_t'|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t')\big],\\
%       \log\big[p(z,\alpha|X)\big] &= \sum_{k=0}^K\bigg[z_{t,k}\cdot\log\big[\pi(\alpha_{t})_k\big]\bigg] + \log\big[p(\alpha_t|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t)\big].
%     \end{align*}
%     Finally, the acceptance rate is calculated using the formula below.
%     \begin{align*}
%       r_{t,k} = \exp\bigg[\min{\bigg(0, \log{\bigg[\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)}\bigg]}\bigg)}\bigg].
%     \end{align*}

% \end{enumerate}

% ___________________________________________________________________

% \section*{The Experiment Settings}

% \par The intention of the carried experiments is to identify the optimal settings for the Metropolis--Hastings algorithm application. 
% The rationale of the carried experiments is based on generating a corpus with pre-defined $\alpha$ changes. Based on the experiments, we will determine which techniques display higher performance in reproducing the pre-defined $\alpha$ fluctuations over time. To expand on the corpus generation settings, the parameters used are listed below:
% \begin{itemize}
%   \item The number of topics: K = 2;
%   \item The number of documents (time slices): T = 20;
%   \item The size of vocabulary: V = 10;
%   \item The number of words per document t: $N_t \sim \mbox{Pois}(\lambda),\quad \lambda = 1000$.
% \end{itemize}
% Speaking of $\alpha_k$ development over time (documents), $\alpha_0$ is a sine curve and $\alpha_1$ is a cosine curve; the corresponding topic distributions over documents (i.e., $\theta = \mbox{softmax}(\alpha)$)) are illustrated in Figure 1 below.
% \begin{figure}[H]
%   \centering
%   \includegraphics[width=0.5\textwidth]{theta_initial}
%   \caption{The values of $\mbox{softmax}(\alpha)$ used in the generative process.}
% \end{figure}
% Speaking of $\beta$, the parameter was initially pre-defined and kept constant throughout the dynamic generative process; $\beta$ is illustrated in Figure 2 below.
% \begin{figure}[H]
%   \centering
%   \includegraphics[width=0.5\textwidth]{beta_initial}
%   \caption{The values of $\beta$ used in the generative process.}
% \end{figure}
% ___________________________________________________________________

% \section*{The experiment results}

% \par The first experiment is focused on discovering an optimal choice of the variances. Note that the first experiment is carried using the auto-regressive model; therefore, three different types of variances were considered: the `basic' variance $\sigma^2_0$, the `auto-regressive' variance $\sigma^2$, and the `proposed' variance $\delta^2$.

% \par For the first experiment, $\delta^2$ was kept constant and set to $1$. Effectively, low $\delta^2$ values suggest that the convergence of $\alpha$ is slow and stable, whereas for high values of $\delta^2$ the convergence is faster and less stable. In both cases, with a high number of iterations, a low-error $\alpha$ value will be found. Therefore, we are focusing to tune only the $\sigma^2_0$ and $\sigma^2$ variances. Also, note that two visualisation of topic distribution over documents are provided: at 1000 iterations; and at 2000 iterations.

% \begin{figure}[H]
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E-01|basic_1E-02|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E+00|basic_1E-02|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_5E+00|basic_1E-02|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}
%   \caption{Fixed $\sigma^2=0.01$.}
% \end{figure}
% \begin{figure}[H]
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E-01|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E+00|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_5E+00|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}
%   \caption{Fixed $\sigma^2=0.001$.}
% \end{figure}
% \begin{figure}[H]
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E-01|basic_1E-04|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E+00|basic_1E-03|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_5E+00|basic_1E-04|prop_1E+00|it_2000|autoreg_True}
%   \end{subfigure}
%   \caption{Fixed $\sigma^2=0.0001$.}
% \end{figure}

% \par The second experiment assesses the impact of the auto-regressive $\alpha$ update. For this reason, the $\alpha$ prior was switched to the non auto-regressive one. Note that in this case $\sigma^2$ has no impact; therefore, we provide illustrations by varying the $\sigma^2_0$ term. Again, $\delta^2=1$ was kept constant throughout the experiments. Also, note that we ran the model for $200$ iterations. The resulting $\mbox{softmax}(\alpha)$ values are illustrated in Figure 6 below.

% \begin{figure}[H]
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E-01|basic_1E-02|prop_1E+00|it_200|autoreg_False}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_1E+00|basic_1E-02|prop_1E+00|it_200|autoreg_False}
%   \end{subfigure}%
%   \begin{subfigure}[b]{0.33\textwidth}
%     \includegraphics[width=\linewidth]{init_5E+00|basic_1E-02|prop_1E+00|it_200|autoreg_False}
%   \end{subfigure}
%   \caption{The non auto-regressive $\alpha$ update.}
% \end{figure}
% ___________________________________________________________________

% \section*{Questions}

% \par For the last section of this report, I have set some questions to be addressed during next meeting; these are listed below:
% \begin{itemize}
%   \item During the derivation of the posterior given in Preliminaries Section, is the $\pi(\alpha_{t}')_k$ term, i.e. $p(z|\alpha^{-tk},\alpha_{t, k}')$, derived correctly? We did not have it explicitly expressed before.
%   \item Should we consider the impact of tuning the model with the pre-defined $\beta$ (the one used during the synthetic corpus generation)? The model is able to recover the topic assignments to documents even if $\beta$ is initialised randomly. Also, do we have pre-defined $\beta$ for the non synthetic data sets?
%   \item What are the indications of the faster $\alpha$ converge displayed in the non auto-regressive model? Could it mean that the model implementation is faulty?
% \end{itemize}
% ___________________________________________________________________

%\bibliography{report_mh_alpha-update}{}
%\bibliographystyle{plain}
\end{document}

### 2017-03-29 code/pre-processing/create_bne-corpus.py

@@ -1,7 +1,15 @@
 
 import math
+
+"""
+VERSION
+- Python 2
+
+FUNCTION
+- Pre-process the 'bne' data
+"""
+
 import pymzml
-import pandas as pd
 import numpy as np
 import os
 import pickle as pkl
@@ -64,6 +72,7 @@ def main():
     for spectrum in run:
         for mass, intensity in spectrum.peaks:
             if intensity > INTENSITY_FLOOR:
+                print(intensity)
                 preprocessed_mnis.append((mass, intensity, spectrum['id']))
     mnis_dtype = [('mass', float), ('intensity', float), ('id', int)]
     mnis = np.array(preprocessed_mnis, dtype=mnis_dtype)

### 2017-03-29 code/pre-processing/helper_experiment.py

@@ -15,7 +15,8 @@ class Helper_Experiment(object):
     def calculate_performance(self, alpha_init, beta_init, theta_init,
                               clf, s_batch):
         hist_theta = np.array(clf.hist_theta)
-        theta_init = theta_init.T
+        if alpha_init is None:
+            theta_init = theta_init.T
         generator_batch = self._split_to_batches(hist_theta, s_batch)
         performance_batch_theta = []
         performance_batch_phi = []
@@ -29,6 +30,7 @@ class Helper_Experiment(object):
             theta_latent = np.average(hist_theta_batch, axis=0)
             if alpha_init is not None:
                 theta_init = self._softmax_matrix(alpha_init)
+                theta_init = theta_init.T
             theta_latent = theta_latent.T
             matchings_topic = self._tune_thetas(theta_init, theta_latent)
             '''
@@ -74,6 +76,8 @@ class Helper_Experiment(object):
         for idx_init, row_init in enumerate(theta_init):
             row = []
             for idx_latent, row_latent in enumerate(theta_latent):
+                # print(row_init)
+                # print(row_latent)
                 diff = np.abs(row_latent - row_init).sum()
                 row.append(diff)
             matrix_dist.append(row)

### 2017-03-29 code/pre-processing/visualiser_corpus.py

@@ -9,9 +9,11 @@ import os
 
 class Visualiser_Corpus(object):
 
-    def __init__(self, corpus, vocab, n_rows, l_row, l_column,
-                 path_plots=None, save=True):
+    def __init__(self, corpus, no_corpus, no_experiment, vocab, n_rows, l_row,
+                 l_column, path_plots=None, save=True):
         self.corpus = corpus
+        self.no_corpus = no_corpus
+        self.no_experiment = no_experiment
         # self.vocab = self._extract_vocab()
         self.path_plots = path_plots
         self.vocab = vocab
@@ -167,7 +169,8 @@ class Visualiser_Corpus(object):
         plt.ylabel('Total error')
         plt.xlabel('Batch no. (s_batch = {})'.format(s_batch))
         if self.save:
-            plt.savefig(self.path_plots + 'performance_thetas.png', dpi=300)
+            name_file = 'performance_thetas_{}_dataset-{}.png'.format(self.no_experiment, self.no_corpus)
+            plt.savefig(self.path_plots + name_file, dpi=300)
         '''
         Plotting phis
         '''
@@ -180,7 +183,8 @@ class Visualiser_Corpus(object):
         plt.ylabel('Total error')
         plt.xlabel('Batch no. (s_batch = {})'.format(s_batch))
         if self.save:
-            plt.savefig(self.path_plots + 'performance_phis.png', dpi=300)
+            name_file = 'performance_phis_{}_dataset-{}.png'.format(self.no_experiment, self.no_corpus)
+            plt.savefig(self.path_plots + name_file, dpi=300)
 
 # __________________________________________________________________________
 # LATENT PLOTS
@@ -205,7 +209,7 @@ class Visualiser_Corpus(object):
         plt.ylabel(r'$\theta$')
         plt.title(title)
         if self.save:
-            name_file = 'latent_thetas_{}.png'.format(title)
+            name_file = 'latent_thetas_{}_{}_dataset-{}.png'.format(title, self.no_experiment, self.no_corpus)
             plt.savefig(self.path_plots + name_file, dpi=300)
 
     def plot_latent_phis(self, phis, title):
@@ -226,7 +230,7 @@ class Visualiser_Corpus(object):
         plt.ylabel(r'$\phi$')
         plt.title(title)
         if self.save:
-            name_file = 'latent_phis_{}.png'.format(title)
[…]

### 2017-04-01 code/models/dtm_alpha.py
Paper: finished the methodology

@@ -89,7 +89,7 @@ class DTM_Alpha(object):
         print('{} has finished.'.format(self.__class__.__name__))
 
     def _update_status(self, it, n_burn_it):
-        if it != 0 and it % 50 == 0:
+        if it != 0 and it % 5 == 0:
             print('Iteration: {}'.format(it))
             rate_alpha_update = 1.0 * self.n_upd_alpha / self.n_it_alpha
             print('Alpha update rate: {:.2f}'.format(rate_alpha_update))

### 2017-04-01 code/pre-processing/generator_corpus.py

@@ -33,6 +33,7 @@ class Generator_Corpus(object):
             vocab.append(string.ascii_lowercase[w])
         return np.array(vocab)
 # ___________________________________________________________________________
+# TODO: Beta is phi
 
     def generate_corpus(self):
         corpus = {}

### 2017-04-01 code/pre-processing/visualiser_corpus.py

@@ -9,8 +9,7 @@ import os
 
 class Visualiser_Corpus(object):
 
-    def __init__(self, corpus, no_corpus, no_experiment, vocab, n_rows, l_row,
-                 l_column, path_plots=None, save=True):
+    def __init__(self, corpus, vocab, n_rows, l_row, l_column, path_plots=None, save=True, no_corpus=None, no_experiment=None):
         self.corpus = corpus
         self.no_corpus = no_corpus
         self.no_experiment = no_experiment
@@ -93,7 +92,8 @@ class Visualiser_Corpus(object):
     def locate_word(self, word):
         # Mapping to the z values.
         z_matrix = np.zeros(shape=(self.n_rows, self.n_columns))
-        word_idx = self.vocab.index(word)
+        word_idx = np.where(self.vocab == word)
+        # word_idx = self.vocab.index(word)
         for i, doc in enumerate(self.corpus):
             if int(i) in self.map_coord:
                 doc_coord = self.map_coord[int(i)]
@@ -105,7 +105,8 @@ class Visualiser_Corpus(object):
         y_vector = np.zeros(shape=(self.n_rows, 1))
         for i in range(0, self.n_rows):
             y_vector[i] = i
-        layout = go.Layout(height=350,
+        layout = go.Layout(title="The term's occurrences after the normalisation",
+                           height=350,
                            xaxis=dict(title='x'),
                            yaxis=dict(title='y'))
         data = [go.Heatmap(z=z_matrix,

### 2017-04-13 code/models/dtm_alpha.py
paper: experiment settings.

@@ -89,7 +89,7 @@ class DTM_Alpha(object):
         print('{} has finished.'.format(self.__class__.__name__))
 
     def _update_status(self, it, n_burn_it):
-        if it != 0 and it % 5 == 0:
+        if it != 0 and it % 25 == 0:
             print('Iteration: {}'.format(it))
             rate_alpha_update = 1.0 * self.n_upd_alpha / self.n_it_alpha
             print('Alpha update rate: {:.2f}'.format(rate_alpha_update))
@@ -267,20 +267,29 @@ class DTM_Alpha(object):
 # ___________________________________________________________________________
 
     def _sample_phi(self):
-        alphas_dir = (1.0 * self.n_zw + self.beta)
+        phi_pre = (1.0 * self.n_zw + self.beta)
+        phi_pre /= (1.0 * self.n_z[:, None] + self.V * self.beta)
         phi = []
-        for alpha_dir in alphas_dir:
-            phi_topic = np.random.dirichlet(alpha_dir)
+        for row_phi in phi_pre:
+            phi_topic = np.random.dirichlet(row_phi)
             phi.append(phi_topic)
+        # TOFIX
+        # phi = (1.0 * self.n_zw + self.beta)
+        # phi = self._softmax(phi)
+        # phi /= np.sum(phi, axis=1)[:, None]
         self.hist_phi.append(phi)
 # ___________________________________________________________________________
 
     def _sample_theta(self):
         alphas_dir = self.alpha_softmax
         theta = []
-        for alpha_dir in alphas_dir:
-            theta_doc = np.random.dirichlet(alpha_dir)
-            theta.append(theta_doc)
+        # for alpha_dir in alphas_dir:
+        #     theta_doc = np.random.dirichlet(alpha_dir)
+        #     theta.append(theta_doc)
+        # theta = (self.n_dz + alphas_dir)
+        # theta = self._softmax(theta)
+        # TOFIX
+        theta = self.alpha_softmax
         self.hist_theta.append(np.array(theta))
 
 # ___________________________________________________________________________

### 2017-04-13 code/pre-processing/helper_experiment.py

@@ -17,13 +17,16 @@ class Helper_Experiment(object):
         hist_theta = np.array(clf.hist_theta)
         if alpha_init is None:
             theta_init = theta_init.T
-        generator_batch = self._split_to_batches(hist_theta, s_batch)
+        # generator_batch = self._split_to_batches(hist_theta, s_batch)
         performance_batch_theta = []
         performance_batch_phi = []
         '''
         Iterate through the theta batches
         '''
-        for hist_theta_batch in generator_batch:
+        n_iter = int(len(hist_theta) / s_batch)
+        # for hist_theta_batch in generator_batch:
+        for i in range(n_iter):
+            hist_theta_batch = hist_theta[0:i * s_batch - 1]
             '''
             Pre-process and tune the thetas
             '''
@@ -48,8 +51,10 @@ class Helper_Experiment(object):
         Iterate through the phi batches
         '''
         hist_phi = np.array(clf.hist_phi)
-        generator_batch = self._split_to_batches(hist_phi, s_batch)
-        for hist_phi_batch in generator_batch:
+        # generator_batch = self._split_to_batches(hist_phi, s_batch)
+        # for hist_phi_batch in generator_batch:
+        for i in range(n_iter):
+            hist_phi_batch = hist_phi[0:i * s_batch - 1]
             phi_latent = np.average(hist_phi_batch, axis=0)
             '''
             Calculate the phi similarity

### 2017-04-13 code/pre-processing/visualiser_corpus.py

@@ -120,7 +120,8 @@ class Visualiser_Corpus(object):
 # ___________________________________________________________________________
 
     def locate_topic(self, id_topic, thetas):
-        theta = thetas[-1, :, :]
+        theta = thetas[-1]
+        # theta = np.average(thetas, axis=0)
         # Mapping to the z values.
         z_matrix = np.zeros(shape=(self.n_rows, self.n_columns))
         for i, doc in enumerate(theta):
@@ -170,7 +171,7 @@ class Visualiser_Corpus(object):
         plt.ylabel('Total error')
         plt.xlabel('Batch no. (s_batch = {})'.format(s_batch))
         if self.save:
-            name_file = 'performance_thetas_{}_dataset-{}.png'.format(self.no_experiment, self.no_corpus)
+            name_file = 'performance-theta_dataset-{}.png'.format(self.no_corpus)
             plt.savefig(self.path_plots + name_file, dpi=300)
         '''
         Plotting phis
@@ -184,7 +185,7 @@ class Visualiser_Corpus(object):
         plt.ylabel('Total error')
         plt.xlabel('Batch no. (s_batch = {})'.format(s_batch))
         if self.save:
-            name_file = 'performance_phis_{}_dataset-{}.png'.format(self.no_experiment, self.no_corpus)
+            name_file = 'performance-phi_dataset-{}.png'.format(self.no_corpus)
             plt.savefig(self.path_plots + name_file, dpi=300)
 
 # __________________________________________________________________________
@@ -210,7 +211,7 @@ class Visualiser_Corpus(object):
         plt.ylabel(r'$\theta$')
         plt.title(title)
         if self.save:
-            name_file = 'latent_thetas_{}_{}_dataset-{}.png'.format(title, self.no_experiment, self.no_corpus)
+            name_file = 'theta-latent_{}_dataset-{}.png'.format(title, self.no_corpus)
             plt.savefig(self.path_plots + name_file, dpi=300)
 
     def plot_latent_phis(self, phis, title):
@@ -231,7 +232,7 @@ class Visualiser_Corpus(object):
         plt.ylabel(r'$\phi$')
         plt.title(title)
         if self.save:
-            name_file = 'latent_phis_{}_{}_dataset-{}.png'.format(title, self.no_experiment, self.no_corpus)
+            name_file = 'phi-latent_{}_dataset-{}.png'.format(title, self.no_corpus)
             plt.savefig(self.path_plots + name_file, dpi=300)
 
[…]

### 2017-04-20 docs/msci-paper/paper_ssmsi.tex
code: a code review; paper: the final draft

%% example.tex
%% Jeremy Singer
%% 16 Oct 12

\documentclass{mpaper}

\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{float}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{subcaption}
\graphicspath{{images/}}

\makeatletter
\newcommand{\thickhline}{%
    \noalign {\ifnum 0=`}\fi \hrule height 1.2pt
    \futurelet \reserved@a \@xhline
}
\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1.2pt\hskip\tabcolsep}}
\makeatother

\begin{document}

\title{Spatial Smoothing in Mass Spectrometry Imaging}
\author{Arijus Pleska}
\matricnum{2019828P}

\maketitle

% ___________________________________________________________________________

% - intro: Metabolomics
% - research problem: Spatial smoothing
% - prelims: Design an appropriate topic model
% - method: Provide the methodology topic model
% - experiments: synthetic data
\begin{abstract}
In this paper, we target a data modelling approach used in computational metabolomics; to be specific, we assess whether spatial smoothing improves the topic term and noise identification. By assessing mass spectrometry imaging data, we design an enhancement for latent Dirichlet allocation-based topic models. For both data pre-processing and model design, we survey relevant research. Further, we focus on presenting our methodology in detail providing the preliminaries to the basic models and guiding through the performed adjustments in the proposed model variations. To assess the application's performance, we run the model comparison on a number of diverse synthetic datasets.
\end{abstract}

% ___________________________________________________________________________
\section{Introduction}

% Purpose:
% - A spatial smoothing application in visual data using unsupervised ML
% - Focus: Metabolomics and topic modelling
\par In this research paper, we assess an application of spatial smoothing in visual data; that is, we induce continuity among data elements. The spatial smoothing application is particularly targeted to be applied for unsupervised pattern recognition. To be more specific, our focus is to model a biomedical application in the field of metabolomics. Furthermore, we limit the scope of the applied unsupervised machine learning techniques to the branch of topic modelling.

% Basis:
% - Visual data: MSI application in metabolomics
% - MSI data structure
\par The characteristics of our utilised metabolomics data are expressed in the form of mass spectrometry imaging (MSI). Effectively, we use MSI to visualise the metabolomics data in the form of spatial distribution. Speaking of the metabolomics data, it contain information about ionised metabolites. Note that metabolites are molecules produced by the chemical process of metabolism, whereas by ionisation we refer to the information capture method. In other words, MSI data is a visualisation of ion (sampled metabolite) distributions: the complete dataset is the whole image; the image's pixel is a particular sampling region; and each region contains intensities of ions with unique mass-to-charge $m/z$ values.

% Basis:
% - Unsupervised machine learning: Topic modelling
% - topic modelling w.r.t. MSI
\par Speaking of our machine learning application, topic modelling is a technique used to infer unlabelled topic distributions based on data's underlying semantic structure. With respect to MSI data, we can model the topic distributions over an image and its every pixel; furthermore, topic modelling can express the types of ions corresponding to particular topics. Since we use a topic model to model metabolomics data in a statistic manner, the utilised topic models are tuned to reflect the metabolomics environment as realistically as possible. Relating to our research targets, spatial smoothing is one of such environment settings.

% Research Problem:
% - Noisiness
% -- Fragmentation
% -- Retention time
% - Loss of information
% -- Overlapping topics
\par The basis of the project's research problems comes from the limitations of current metabolite sampling techniques. The main limitation is the loss of information caused by the metabolite ionisation. As a consequence, the MSI data is noisy. To expand on the noisiness, it is caused by metabolite fragmentation and the limitation to tune different metabolite retention times. By metabolite fragmentation, we refer to a metabolite split; the split would cause the captured ions to possess unexpected values. Speaking of the retention time, the intensity value of each ion type varies with respect to time; therefore, since each ion is captured at a different state of its retention, each ion type would possess some variance with respect to its intensity. Finally, note that the loss of information might be also caused by overlapping ion topics. Since different ion topics can contain same ion types, the topic possessing a lower ion intensity value would be overwhelmed and, thus, not reflected in the MSI data.
% Contributions:
% - A specific and extensive assessment of the SS application
% - A tuned topic model and well-organised experiment settings
\par In this paper, we contribute to the research in MSI by carrying an extensive assessment of the spatial smoothing application. The assessment is carried in both quantitative and qualitative manners: we assess the performance on a number of diverse datasets; also, our experiments are designed to reflect the nature of the MSI data reflecting computational metabolomics. Furthermore, we provide a Python implementation of a tuned topic model; also, we establish maintainable experiment settings. By the tuned topic model, we mean that the model's implementation is particularly designed to meet the characteristics of the MSI data. Speaking of the experiment settings, note that we are carrying the experiments using Jupyter notebooks. Effectively, the use of the notebooks creates a portable and well-documented environment to initialise the experiment settings and execute the topic modelling. As a result, external parties could run the released notebooks and reproduce the experiment results in a swift manner.

% The paper's organisation
\par The paper is organised in the following order: in Section 2, we discuss the background of the research project; in Section 3, we provide a formal definition of the assessed research problems; in Section 4, we review the results of the relevant research; in Section 5, we establish the rationale of the applied methodology; in Section 6, we introduce the experiments; finally, in Section 7, we conclude the findings.

% ___________________________________________________________________________
\section{Background}

% Intro
% - Preliminaries: Introduce the rationale of LDA
% - Terminology: Define the applied topic modelling terminology
% - MSI Data characteristics: Introduce the qualities of the data
\par The background section covers the basis of the concepts used throughout the paper. At the start, we provide a high-level overview of the general topic modelling concepts; then, we define the terminology used throughout the paper; finally, we introduce the characteristic qualities of the MSI data.  
\subsection{Topic Modelling Preliminaries}

% LDA Intro:
% - Hierarchical structure
% - Generative model
% - Bayesian methods
% - Assumptions
\par The research project targets a specific branch of topic models. The branch consists of Latent Dirichlet Allocation (LDA) derivatives. Note that the initial LDA model was introduced by Blei et al. \cite{blei2003latent}. One of the model's key characteristics is the three-level hierarchical treatment of the data. In the context of the utilised MSI data, the hierarchical structure can be perceived as follows: in the highest level, we have an MSI image; in the middle level, we have a pixel of an MSI image; and in the lowest level, we have the intensities of particular ions in a pixel.

% Generatitive model:
% - Generative process
\par Another model's key characteristic is the generative treatment of the data. By a generative model, we mean that the latent data instances are treated as a result of a mixture of underlying parameters drawn from probability distributions. In other words, the generative data treatment induces randomness in the end products of the data; however, note that the source of the data -- the lowest level parameters of the probability distributions -- remain the same. The key aspect of the generative model is the degree of freedom in the connections of random variables; this notion allows modelling more realistic, thus, more complex data relations. Ultimately, the rationale of the generative model is based on recovering the underlying semantic structure. Effectively, in order to recover the generative process of the data, we delve into the applications of Bayesian methods.

% Bayesian methods
% - Inference: sampling
% - Connection to LDA
\par By applying Bayes' rule, we can express the underlying semantic structure in the form of a posterior probability distribution. However, note that we can not analytically compute such posterior expression. Instead, we can estimate the posterior distributions using optimisation or direct sampling. In this project, we particularly focus on the underlying semantic structure's inference using direct sampling. Speaking of the sampling-based inference method applications to LDA-like models, the ground-work has been established by Griffiths and Steyvers \cite{griffiths2004finding}. The authors have utilised the collapsed Gibbs sampling -- a Markov chain Monte Carlo (MCMC) algorithm. Effectively, the method integrates the uncertainty out; so that we sample the entities of interest directly. More specific details about the rationale of the collapsed Gibbs sampling application will be provided in Section 4.

% Assumptions
% - Exchengeability (BoW)
% - Discrete data
\par Relating to the initial LDA model, the authors have set the following assumptions for the utilised data: exchangeability among the inner components of the lower and middle data hierarchy levels; also, a discrete treatment of the lower-level data. By exchangeability, it is meant that the components follow the bag-of-words principle; that is, the order of the data has no correlation with the underlying semantic structure. Speaking of the discrete data treatment, it is assumed that the lower-level components have no spatial connection. With respect to our project, both assumption types -- exchangeability and discreteness -- do not correspond to the characteristics of the MSI data. Therefore, we will establish a methodology to relax the latter assumptions.

\subsection{Topic Modelling Terminology}

% Terminology intro:
% - Introduce the three-level hierarchical structure
\par In this subsection, we introduce our topic modelling terminology. In order to put the MSI data in the topic modelling context, the components of  LDA's three-level hierarchical structure are defined as follows: \textit{a corpus} is the whole image of a sample; \textit{a document} is an MSI image's pixel; and \textit{a word} is an interval of mass-to-charge values corresponding to a particular ion. Starting from now on, we will use the latter topic modelling concepts to introduce the LDA model's architecture and notation.

% LDA design
% - Graphical model
\par Since LDA-like models are also graphical models, the dependence of random variables can be illustrated using graphs. In order to familiarise with the architecture and the variables of LDA-like models, we provide Figure \ref{fig:arch_LDA-init} illustrating the initial LDA model in the plate notation.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.25\textwidth]{LDA-initial.png}
  \caption{The initial LDA model's architecture.}
  \label{fig:arch_LDA-init}
\end{figure}
The circles indicate the model's variables: the coloured circle corresponds to the observable variable, whereas the uncoloured circles correspond to the hidden variables. Effectively, the hidden variables define the model's underlying structure. Speaking of the plates, the plate denoted by $K$ corresponds to the number of topics; the plate denoted by $T$ corresponds to the number of documents, and the plate denoted by $N$ corresponds to the number of words. Effectively, the letters in the bottom right corners indicate the total number of variables. Therefore, a corpus has a $T$ number of documents, and each document has an $N$ number of words.

% Terminology
% - Parameter meaning
% - Listing
\par Before providing a listing with the definitions of the LDA variables, we introduce the purpose of the variables. The variables denoted by $\theta$ and $\phi$ correspond to the underlying probability distributions (e.g., in the case of the initial LDA model, we use Dirichlet distributions). It follows that the variables denoted by $\alpha$ and $\beta$ act as the parameters of the latter probability distributions; note that in the context of machine learning, such auxiliary parameters are called hyper-parameters. Effectively, \textit{a hyper-parameter} allows tuning a machine learning model for a particular dataset application. As a side reference, note that by \textit{a vocabulary} we refer to a collection of terms reflecting a fixed range of the words. At this point, we provide the following list containing the definitions of the initial LDA model's variables: 
\begin{itemize}
	\item $K$ is the number of topics;
	\item $T$ is the number of documents;
	\item $N$ is the number of words per document;
	\item $V$ is the size of a vocabulary;
	\item $w$ is a word;
	\item $z$ is a word's topic assignment;
	\item $\theta_t$ is the topic distribution over the document $t$;
	\item $\phi_k$ is the vocabulary term distribution over the topic $k$;
	\item $\alpha$ is the hyper-parameter for the topic distributions;
	\item $\beta$ is the hyper-parameter for the vocabulary term distributions.
\end{itemize}

\subsection{MSI Data Characteristics}

% MSI dataset intro:
% - Introduce the raw data format
% - Describe the data pre-processing requirements
\par In this subsection, we introduce the qualities of the raw MSI data. Furthermore, we set the requirements for the raw data's pre-processing; note that the pre-processing serves as an auxiliary method making the raw MSI data compatible for a scalable topic modelling application. As a side note, the basis of our applied MSI data characteristics is established from the mzML data format. Conveniently, we parse mzML data using the \texttt{pymzml} Python library.

% Raw data format
% - Data meaning
% - Explain the raw data by looking into it
\par Essentially, raw MSI data contain mass spectra of a sample -- the sample's every pixel is expressed in the form of a mass spectrum. By a mass spectrum, we refer to a map from a mass-to-charge value to an intensity value. Ultimately, we can establish a continuous notion of the MSI data by ordering the mass-to-charge values. Note that since the sampling equipment can detect the mass-to-charge values in the millidalton (mDa) precision, the MSI data is sparse (i.e., a large portion of mass-to-charge values are mapped to the zero intensity).

% Data pre-processing requirements
% - Vocabulary generation
% - Intensity normalisation
% - Dismissing inferior information
\par In the context of topic modelling, the raw MSI data possess a large vocabulary (above $5000$ terms). Note that we consider every mass-to-charge value as a word; whereas every intensity value is perceived as a word's occurrence count. Further, since the intensity values could spike up above $1000$, the time complexity of the latent topic inference would require pro-longed runs. To overcome the introduced scalability issues, we carry data pre-processing -- the applied techniques are discussed in the upcoming methodology section.

% ___________________________________________________________________________
\section{Statement of Problem}

% Hypothesis
% - A topic is continuous throughout nearby regions
\par To start with, we set the hypothesis of this research project to \textit{`The spatial smoothing application induces more realistic representation of the visual computational metabolomics data -- MSI'.} By spatial smoothing, it is meant that the topic model would have an auto-regressive treatment among the nearby MSI instances. As an example, we assume that adjacent pixels would have similar latent topic distributions. This assumption corresponds to the nature of our datasets -- a metabolite construction (i.e., a topic) is continuous throughout nearby regions (i.e., sets of adjacent pixels).

% Contributions
% - Naive AR model application
% - Motivating for the state-of-the-art AR model application
\par The impact of proving the hypothesis would bring the following contributions:
\begin{itemize}
	\item Improving the detection of overlapping topics and vocabulary terms;
	\item Reducing the noisiness of the MSI data;
	\item Motivating a further research in the spatial smoothing application in MSI data.
\end{itemize} 
To expand on the overlapping topic detection, the issue arises when two underlying topics are made of similar vocabulary terms: instead of a separate representation, the topics are merged. We intend to identify the flow of distinct topics by the application of spatial smoothing; as a consequence, the spatial smoothing would also impact the data noisiness reduction. Ultimately, if a naive spatial smoothing application displayed a performance improvement, the contribution would set a basis to utilise state-of-the-art auto-regression techniques on the MSI data.

% Approach
% - Clearly define the experiment settings
% - Tune the models by studying synthetic datasets.
\par To my knowledge, the impact of the spatial smoothing application to the MSI domain has not yet been thoroughly studied. For the latter reason, this research project serves as an exploratory assessment of the spatial smoothing application: we introduce the rationale behind the applied methodology; also, we clearly define the range of the experiment settings. To give a brief intuition about the methodology, the study assesses the domain-specific parameter tuning and its impact on a diverse range of synthetic datasets.

% ___________________________________________________________________________
\section{Relevant Research}

% Intro (separate into three)
% - MSI DATA:
% -- alonso2015analytical
% -- Review the metabolomics feature pre-processing
% -- Review the metabolomics alternative approaches
% - Review the implementation approaches
% -- blei2012probabilistic
% -- Gibbs sampling
% -- DTM
\par In this section, we review the ground-work carried on the following aspects: the pre-processing of MSI data; the rationale of the utilised topic models; and novel approaches exploiting the characteristics of MSI data. Note that the covered groundwork is selected to reflect the rationale of the utilised techniques. In order to establish the basis of the state-of-the-art computational metabolomics methodology, we consult the survey carried by Alonso et al. \cite{alonso2015analytical}; speaking of the key probabilistic topic modelling branches, we consult the survey carried by Blei \cite{blei2012probabilistic}.

\subsection{MSI Data Pre-processing}

% Metabolomics: feature pre-processing
% - Normalisation:
% -- kohl2012state
% -- bolstad2003comparison

\par In order to establish a scalable topic inference, we review the following MSI data pre-processing techniques: data normalisation; feature binning; and noise reduction. To start with, data normalisation serves as a method establishing an adaptable data structure. Bolstat et al. \cite{bolstad2003comparison} have proposed a data normalisation method called linear baseline scaling. Note that the method is particularly targeted at sparse datasets. Effectively, the method is applicable to numerical features; note that the method's work-flow is carried as follows: we find the largest numerical feature of all data instances; then, we calculate the scaling factors by aligning the largest values to a pre-set upper threshold; finally, we align the remaining numerical features based on the established scaling factors. Relating linear baseline scaling to applications on MSI data, Kohl et al. \cite{kohl2012state} have shown that the method's application does not produce a significant loss of information to establish a well-performing MSI data inference.

% - Binning
% -- Evenly spaced 
% -- craig2006scaling
% -- Spectral peaks
% -- de2008nmr
\par Speaking of feature binning, the technique is used to merge the instances of a feature space; as a result, feature binning reduces the data dimensionality. Note that by the MSI data feature space, we refer to distinct $m/z$ values. Since raw MSI data is sparse, a successful application of feature binning is based on identifying appropriate $m/z$ boundaries reflecting unique metabolite types. To introduce some examples, the research carried by Craig et al. \cite{craig2006scaling} have utilised an equally spaced feature binning. Even though the authors have succeeded to reduce the data dimensionality, they have encountered a loss of information by splitting the metabolite topics into arbitrary regions. As an alternative approach, De et al. \cite{de2008nmr} have performed a feature binning based on the spectral peak identification. Effectively, the technique induces a dynamic notion of bin boundaries which are based on the identified intensity peak regions. As a result, the merged $m/z$ values serve as a better representation of metabolite terms.

% - Denoising (removing small intensities)
% -- smith2014proteomics
\par The last reviewed MSI data pre-processing techniques addresses the MSI data noisiness. Even though the MSI data noise reduction is an open research problem going well beyond data pre-processing, Smith et al. \cite{smith2014proteomics} have shown that the application of a general data pre-processing routine displays performance improvements. For example, the authors have induced a lower intensity bound. Effectively, the intensity values below the intensity threshold would be treated as insignificant and/or as a product of data capturing device imperfections. By applying this procedure, the authors have successfully reduced the data dimensionality and, thus, increased the data scalability.

\subsection{Prospective Topic Models}

% Intro
% - Inference
% - Derivatives
% - Gibbs sampling
% -- griffiths2004finding
\par In this subsection, we review the key characteristics of the prospective topic modelling inference techniques and model variations. Speaking of the inference techniques, we employ the rationale of a sampling-based LDA model. Griffiths and Steyvers \cite{griffiths2004finding} have displayed a successful application of the collapsed Gibbs sampler for the topic inference of both textual and visual data. To expand on the collapsed sampling, the technique allows skipping the estimation of $\theta$ and $\phi$ values; instead, the inference is based on the notion of the assignment counts. In order to estimate a topic assignment's probability, the authors suggest using the following expression:
\begin{equation*}
P(z_i = k | z_{-i}, w) \propto \dfrac{n_{-i, k}^{(w_i)} + \beta}{n_{-i, k}^{(\cdot)} + V\beta}\cdot \dfrac{n_{-i, k}^{(t_i)} + \alpha}{n_{-i, \cdot}^{(t_i)} + K\alpha}.
\end{equation*}
note that $i$ refers to the current iteration ($-i$ refers to the previous iteration); $k$ refers to a particular topic; and $n$ refers to the count of the instances indicated by the term's superscript. If required, $\theta$ and $\phi$ can be sampled from the following distributions:
\begin{align*}
\phi_{k,w} & \sim \mbox{Dirichlet}\bigg(\dfrac{n_{k}^{(w)} + \beta}{n_{k}^{(\cdot)} + W\beta}\bigg),\\
\theta_{t,k} & \sim \mbox{Dirichlet}\bigg(\dfrac{n_{k}^{(t)} + \alpha}{n^{(t)} + K\alpha}\bigg).
\end{align*}
Essentially, more accurate representations of $\theta$ and $\phi$ are obtained by running the Gibbs sampler until a sufficient convergence: we preserve the $\theta$ and $\phi$ values of each iteration; when the sampling is finished, the average of the preserved values is an accurate approximation of the underlying $\theta$ and $\phi$ values.

% Topic modelling:
% - DTM
% -- blei2006dynamic
% - SeqLDA
% -- du2012sequential
\par In order to relax the topic modelling assumptions of the initial LDA model, we review the model's derivatives. One of the assumptions -- word exchangeability -- is addressed by Blei and Lafferty \cite{blei2006dynamic}. 
The authors have proposed dynamic topic model (DTM) inducing the notion of change in the topic and vocabulary distributions; the model's architecture is illustrated in Figure \ref{fig:arch_dtm}.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.4\textwidth]{dtm-architecture.png}
  \caption{The DTM architecture.}
  \label{fig:arch_dtm}
\end{figure}
Ultimately, the documents of a corpus are assigned to segments with unique $\theta$ and $\phi$ values. Note that the changes in $\theta$ and $\phi$ values are impacted by their hyper-parameter updates. Since in the DTM paper the authors use variational inference, we provide a reference for an alternative dynamic topic modelling variation -- sequential LDA -- proposed by Du et al. \cite{du2012sequential}. Note that in the sequential LDA paper, the authors display a rigorous application of the collapsed Gibbs sampler.

\subsection{Prospective Characteristics of MSI Data}
% Metabolomics: spatial smoothing
% - TM application
% -- van2016topic
\par At this point, we look into the prospective topic modelling applications exploiting the characteristics of MSI data and employing spatial smoothing. To start with, Hooft et al. \cite{van2016topic} have utilised an LDA-like model to infer metabolite substructures from MSI data. The authors have established a novel approach utilising a favourable LDA's property -- the option to assign a unique vocabulary term to multiple topics. Effectively, this approach allows identifying metabolite substructures which are made of overlapping elements.

% - Spatial Chaos
% -- palmer2016fdr
\par Speaking of the spatial smoothing application, to my knowledge, the idea has not yet been widely spread among the computational metabolomics community. Nevertheless, a recent study by Palmer et al. \cite{palmer2016fdr} have attempted to quantify spatial chaos among the partitions of MSI data. The authors have reported that the established notion of spatial chaos has improved the speed and accuracy of continuous metabolite pattern identification.

% ___________________________________________________________________________
\section{Methodology}

% Intro
\par In this section, we cover the rationale of a topic model tuned for the spatial smoothing application. To start with, we provide details on how to establish the auto-regressive notion among MSI data; then, we show how to apply the auto-regression to a topic inference based on the collapsed Gibbs sampling. Further, we provide a list of the applied data pre-processing techniques for MSI data. Finally, by considering the data format induced by the data pre-preprocessing, we introduce the generative MSI data process. Effectively, we apply the generative process to generate synthetic data for our experiments.

\subsection{Spatial Smoothing}

% Introduce AR
% - Define the first two variances
% - Point out that we are using the terms corresponding to the MSI data
\par We establish the spatial smoothing among MSI data by inducing the auto-regressiveness among the pixels (documents). Before going into details, note that we cover the established methods using the previously introduced topic modelling notation. To start with, recall that by auto-regressiveness we refer to the smooth topic development among the nearby instances of an MSI corpus. In our settings, the auto-regressiveness is established by assuming that the joint probability distribution of the $\alpha$ priors is given by the following expression:
\begin{align*}
&p(\alpha_1,\ldots,\alpha_T) =p(\alpha_1)\prod_{t=2}^{T}{p(\alpha_t|\alpha_{t-1})}, \quad \mbox{where}\\
&p(\alpha_0) =\mathcal{N} (\alpha_0; 0, \sigma_0^2I), \quad \mbox{and}\\
&p(\alpha_t) =\mathcal{N} (\alpha_t; \alpha_{t-1}, \sigma^2I), \quad t>0.
\end{align*}
To introduce the previous expression, note that the index $t$ refers to a particular pixel (a document). This means that every pixel of an MSI corpus has a unique underlying topic distribution induced by a unique $\alpha$. Further, the variances $\sigma_0^2$ and $\sigma^2$ correspond to the initialisation variance and the smoothness variance, respectively. Effectively, $\sigma_0^2$ is used to create larger gaps among different topics, whereas $\sigma^2$ preserves the smoothness. Therefore, we set $\sigma_0^2$ to possess a higher value compared to $\sigma^2$.

% Introduce the MH algorithm
\par The previously introduced $\alpha$ priors serve as the initial point in estimating the true $\alpha$ values. To estimate the true $\alpha$ values, we utilise the Metropolis--Hastings (MH) algorithm. The work-flow of the MH algorithm is started by drawing the proposed state:
\begin{align*}
&x' \sim q(x,\delta^2I).
\end{align*}
Note that $x$ denotes the current state, $q$ denotes the proposal distribution, and $\delta^2$ denotes the proposal variance; also, note that if $\delta^2$ is large, the proposal state converges to the true posterior in larger yet random increments; alternatively, if $\delta^2$ is small, the convergence is performed in small yet uniform increments. Note that the settings for an optimal convergence are unique for diverse datasets; as an example, we can find the optimal values using cross validation. Going back to the work-flow, for the second step, we consider the acceptance distributions (these are denoted by $A$) and derive the formula for the acceptance rate:
\begin{align*}
  \dfrac{A(x'|x)}{A(x|x')} = \dfrac{p(x'|x)}{p(x|x')}\cdot \dfrac{q(x'|x)}{q(x|x')} = \dfrac{p(x',x)}{p(x)}\cdot \dfrac{p(x')}{p(x,x')} = \dfrac{p(x')}{p(x)}.
\end{align*}
Note that the proposal distributions cancel out as
\begin{align*}
q = \mathcal{N} \quad \Longrightarrow \quad q(x'|x) = q(x|x').
\end{align*}
For the MH algorithm's final step, we make sure that the acceptance rate does not overflow the probability boundaries; that is, we obtain the acceptance rate denoted by $r$ using the following procedure:
\begin{align*}
  r = \min{\bigg(1, \dfrac{p(x')}{p(x)}\bigg)}.
\end{align*}

% Apply the MH algorithm to the MSI data
% TODO: Softmax
\par At this point, we apply the rationale of the introduced MH algorithm to the topic modelling context. Since we utilise the MH algorithm to update a single value at a time (i.e., we update $\alpha_{t,k}$), we make use of the following notation:
\begin{align*}
  \alpha^{-tk} = \alpha \setminus \alpha_{t,k}.
\end{align*} 
Having the previous notation in mind, the MH algorithm's application to update $\alpha$ is given as follows:
\begin{align*}
  &\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)} = \ldots\\
  &\ldots = \dfrac{p(X|z,\alpha^{-tk},\alpha_{t, k}')}{p(X|z,\alpha)}\cdot \dfrac{p(z|\alpha^{-tk},\alpha_{t, k}')}{p(z|\alpha)}\cdot \dfrac{p(\alpha^{-tk},\alpha_{t, k}')}{p(\alpha)}\\
  &\dots = \dfrac{\prod_{k=1}^K\pi(\alpha_{t,k}')^{z_{t,k}}\cdot p(\alpha_t'|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t')}{\prod_{k=1}^K\pi(\alpha_{t,k})^{z_{t,k}}\cdot p(\alpha_t|\alpha_{t-1})\cdot p(\alpha_{t+1}|\alpha_t)};\quad t \not\in \{1,T\}.
\end{align*}
For completeness, the expressions at the boundaries take the following form:
\begin{align*}
t=1 \quad \Longrightarrow \quad &\dfrac{p(\alpha^{-1k},\alpha_{1, k}')}{p(\alpha)} = \dfrac{p(\alpha_{1}')\cdot p(\alpha_{2}|\alpha_{1}')}{p(\alpha_{1})\cdot p(\alpha_{2}|\alpha_{1})},\\
t=T \quad \Longrightarrow \quad &\dfrac{p(\alpha^{-Tk},\alpha_{T, k}')}{p(\alpha)} = \dfrac{p(\alpha_{T}'|\alpha_{T-1})}{p(\alpha_{T}|\alpha_{T-1})}.\\
\end{align*}
Also, note that by $\pi$ we denote the softmax function which is expressed as follows:
\begin{align*}
  \pi(\alpha_{t,k}) = \dfrac{\exp(\alpha_{t,k})}{\sum_{k'=1}^K \exp(\alpha_{t,k'})}
\end{align*}

\subsection{Auto-regressive Dynamic Topic Model}

% Intro
% - Model modifications
\par Our auto-regressive topic model is based on the rationale of the reviewed dynamic topic model. However, based on the MSI data characteristics and the application of spatial smoothing, the auto-regressive model possesses the following aspects:
\begin{itemize}
	\item The static treatment of the $\beta$ hyper-parameter;
	\item The Gibbs sampler utilising spatial smoothing;
	\item The application of logarithmic space to perform calculations.
\end{itemize}
In the following paragraphs, we introduce each of the previous listings.

\par Even though we utilise the rationale of DTM in order to establish the dynamic notion of the topic development, we preserve a static $\beta$ hyper-parameter. This assumption comes from the characteristics of the metabolomics-based MSI data: we expect the metabolite patterns (i.e., a topic's vocabulary) remain constant. Therefore, in our model, we utilise the architecture illustrated in Figure \ref{fig:arch_ar} below.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\textwidth]{ar-architecture.png}
  \caption{The auto-regressive topic model architecture.}
  \label{fig:arch_ar}
\end{figure}

% Gibbs Sampler changes
\par Speaking of the Gibbs sampler's enhancement, the updated topic assignment formula takes the following form:
\begin{equation*}
P(z_i = k | z_{-i}, w, t) \propto \dfrac{n_{-i, k}^{(w_i)} + \beta}{n_{-i, k}^{(\cdot)} + V\beta}\cdot \pi(\alpha_{t,k}).
\end{equation*}
As a consequence, the sampling of $\theta$ also changes; now, we obtain the topic distribution as follows:
\begin{align*}
\theta_{t,k} = \pi(\alpha_{t,k}).
\end{align*}

% Calculations in the log space
\par Finally, we address the computational stability by performing calculations in logarithmic space. Effectively, the application of logarithmic space mitigates the susceptibility to numerical underflow. Note that numerical underflow is especially relevant in the context of probabilistic models: calculations involve large products of probabilities. In logarithmic space, however, the products are transformed into sums. Speaking of our model, we apply logarithmic space for both the auto-regressive $\alpha$ update and the sampling-based inference. The updated expression for the auto-regressive $\alpha$ update is given as follows:
\begin{alignat*}{2}
  &\log&&\bigg[{\dfrac{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}{p(z,\alpha|X)}}\bigg] = \ldots \\
  & \ldots && = \log\big[{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}\big] - \log\big[{p(z,\alpha|X)}\big]\\
  & \ldots && = z_{t}\sum_{k=1}^K\log\big[\pi(\alpha_{t,k}')\big] + \log\big[p(\alpha_t'|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t')\big]\\
  & && - z_{t}\sum_{k=1}^K\log\big[\pi(\alpha_{t, k})\big] + \log\big[p(\alpha_t|\alpha_{t-1})\big] + \log\big[p(\alpha_{t+1}|\alpha_t)\big].
\end{alignat*}
As a result, the acceptance rate takes the following expression:
\begin{align*}
	r_{t,k} = \exp\big[\min{\big(0, \log\big[{p(z,\alpha^{-tk},\alpha_{t, k}'|X)}\big] - \log\big[{p(z,\alpha|X)}\big]}\big)\big].
\end{align*}
Speaking of the updated expression for the inference, it is updated in the following way:
\begin{align*}
&P(z_i = k | z_{-i}, w, t) \propto \ldots\\
&\ldots \propto \log\big[n_{-i, k}^{(w_i)} + \beta\big] - \log\big[n_{-i, k}^{(\cdot)} + V\beta\big] + \log\big[\pi(\alpha_{t,k})\big].
\end{align*}

\subsection{Data Pre-processing and Generative Process}

% Intro
% - The purpose of the real data
% - Introducing the applied generative process
\par In this subsection, we introduce the MSI data format used in the experiments. At the start, we introduce an example of real data. Effectively, the example displays an application of the pre-processing techniques presented in the literature review section. Further, we transfer the qualities of real MSI data into our synthetic data generation module. To be more specific, we provide an algorithm for the generative data process.

% Looking into real data
% - Applying the reviewed pre-processing techniques
% -- low intensity
% -- bucketisation

\par Before carrying the experiments, we familiarise with the raw MSI data characteristics and assess their scalability. To be more specific, we define the characteristics of our synthetic data by pre-processing a real MSI data sample. To introduce the pre-processing details, we dismiss the words below the intensity threshold of $10$; then, we apply the following bucketisation strategy: merge adjacent vocabulary terms which differ less than $7$ mDA. Effectively, the bucketisation strategy is based on the spectral peak identification. Finally, we apply linear baseline scaling to align the highest intensities to $25$. Most importantly, note that these settings are unique with every dataset; however, the provided values allow carrying experiments in a scalable manner (i.e., a single experiment run on one dataset would take approximately $60$ minutes).

% Plots of pre- and past-normalisation
% - Preserving the metabolotite patterns
% - Improving the scalability (inference iteration time)
\par At this point, we take an exemplary sample. Note that, in the sample, there are two letters inscribed with the ink corresponding to a particular mass-to-charge value. In Figure \ref{fig:comparison}, we compare the visualisation of the sample with and without the applied pre-processing.

\begin{figure}[H]
  \centering

  \begin{subfigure}[b]{0.5\textwidth}
    \caption{The term's occurrences before linear baseline scaling.}
    \includegraphics[width=\linewidth]{pre.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.5\textwidth}
    \caption{The term's occurrences after linear baseline scaling.}
    \includegraphics[width=\linewidth]{post.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.5\textwidth}
    \caption{An inferred topic corresponding to the pattern.}
    \includegraphics[width=\linewidth]{topic.png}
  \end{subfigure}%
  \caption{The term's visual comparison.}
  \label{fig:comparison}
\end{figure}

% The generative process
% - The underlying assumption for the inference
% - synthetic data
% -- Large amounts of data
% - The algorithm of the generative process
% - 1 document is 1 segment
\par Having a basis for a scalable inference, we transfer the identified data properties into the synthetic corpus generation. Before introducing the generative process, recall that our dynamic topic treatment is unique with respect to every document. Therefore, contrary to the reviewed dynamic topic models, our dynamic segment consists of only one document. Applying the latter aspects, we establish our utilised generative process given in Algorithm \ref{alg:dynamic_document_generation} below:
\begin{algorithm}[H]
\caption{The generative process for a synthetic corpus.}
\label{alg:dynamic_document_generation}
\begin{algorithmic}[2]
% \State $\alpha_0 \sim \mathcal{N}(0, \sigma_0^2I)$
\For {$t \leftarrow 1, T$}
% \State $\alpha_t \sim \mathcal{N}(\alpha_{t-1}, \sigma^2I)$
\State $N \sim \mbox{Poisson}(\xi)$
\For {$n \leftarrow 1, N$}
\State $z_{t, n} \sim \mbox{Multinomial}(\pi(\alpha_t))$
\State $k = \{i : z_{t, n, i} = 1\}$
\State $w_{t, n} \sim \mbox{Multinomial}(\phi_k)$
\EndFor
\EndFor
\end{algorithmic}
\end{algorithm}
In practical settings, the rationale of the generative process is defined as follows: the $\xi$ term represents an approximate number of words per document; $\alpha_t$ is the pre-defined auto-regressive hyper-parameter for document $t$; and $\phi_k$ is the pre-defined vocabulary term distribution for topic $k$.

% ___________________________________________________________________________
 \section{Experiments}

% Intro
% - Objective: Research problem
% -- Vocabulary term detection
% -- Noise identification
% - The settings
% - The execution
\par In this section, we assess the research problems introduced in Section 3:
\begin{itemize}
	\item The spatial smoothing application for recovering the underlying vocabulary term distributions;
	\item The auto-regressive model's performance in terms of identifying the noise topic.
\end{itemize}
At the start of the section, we define the settings for tuning the topic models; then, we look into the settings for generating the synthetic datasets; afterwards, we introduce the scope of our experiments. Having defined the settings, we provide several illustrative examples of the experiment execution; and finally, we show the results of all experiments.

\subsection{Pre-experiment Settings}

% Topic model settings
% - Two models: AR & non-AR
% - Remark about the variances
\par The performance of the experiments is assessed by running the auto-regressive and non-auto-regressive topic models in parallel. That is, we run the topic models with and without the pre-set assumption of spatial smoothing. For both models, we tune the variances corresponding to the $\alpha$ update introduced in Section 5. Recall that the variance $\delta^2$ is used to propose a new $\alpha$ hyper-parameter's state; the variance $\sigma_0^2$ is used to initialise $\alpha_0$; and the variance $\sigma^2$ controls spatial smoothing. Effectively, in the non-auto-regressive model, we do not have the $\sigma^2$ term as all $\alpha$ terms are initialised using $\sigma_0^2$ (this notion relaxes the assumption of spatial smoothing).

% Topic model settings
% - AR Tuning
\par In order to run the experiments in an efficient manner, we need to identify optimal values of the previously noted variances. One reason behind the variance tuning corresponds to the rate of convergence upon the application of the MH algorithm. Based on the algorithm's rationale -- a low acceptance rate indicates a slow and stable convergence, whereas a high acceptance rate indicates a random and unstable convergence -- we would find the variance $\delta^2$ inducing the acceptance rate of around $30\%$. Another reason behind the variance tuning is related to the spatial smoothing application. Most importantly, we want to keep the variance $\sigma^2$ in tact with the rate of change of the topic smoothing throughout our data. Furthermore, since our topic development is captured in discrete space, we want to make sure that the discretisation step induced by the generative data process is smaller than the $\sigma^2$ variance; otherwise, we would fail to capture the high rate of change induced by steep topic changes.
% MSI data settings
% - Synthetic corpora
% -- Fixed theta and phi
% -- The size of the synthetic corpus 
\par Speaking of the datasets generated for the experiments, these are designed to reflect three following aspects: the effect of overlapping topics; the effect of overlapping vocabulary terms; and the effect of noise. Note that our assessment is based on an intuitive $3$ topic scenario: $2$ topics model distinct metabolite entities, and the remaining topic models the noise topic. Speaking of the dataset size, we set $T=50$ for the number of documents per corpus and $\xi=100$ for the number of words per documents: the choice of $T$ surpasses the discretisation concern; also, as suggested by the generative algorithm given in Subsection 5.3, the use of the $\xi$ parameter establishes a slightly varying number of words in each document. However, in order to speed up the inference, we normalise the number of words per document to possess the maximum value of $50$.

% Experiment settings
% - Overlapping topics
% - Overlapping vocabulary terms
% - Overlapping noise vocabulary terms
\par In the preceding figures, we illustrate the variations of the data generation settings: in Figure \ref{fig:ol_topics}, we display the setting  controlling the topic overlap; in Figure \ref{fig:ol_terms}, we display the setting controlling the topic term overlap; and in Figure \ref{fig:ol_error}, we display the setting controlling the error overlap. Note that the red and green colourings indicate the synthetic metabolite topics; the green colouring indicates the noise topic; and, for the term names set in the horizontal axes of Figures \ref{fig:ol_terms} and \ref{fig:ol_error}, we use arbitrary, unique numbers.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.48\textwidth}
    % \caption{Enabled topic overlap.}
    \includegraphics[width=\linewidth]{theta_topic-ol.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.48\textwidth}
    % \caption{Disabled topic overlap.}
    \includegraphics[width=\linewidth]{theta_topic-nonol.png}
  \end{subfigure}%
  \caption{Controlling the topic overlap.}
  \label{fig:ol_topics}

  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\linewidth]{phi_topic-ol_error-nonol.png}
  \end{subfigure}%
   \caption{Controlling the topic term overlap.}
  \label{fig:ol_terms}

  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\linewidth]{phi_topic-nonol_error-ol.png}
  \end{subfigure}%
   \caption{Controlling the error overlap.}
  \label{fig:ol_error}
\end{figure}

% Experiment settings
% - Illustration overview: example
% - preset theta and phi
\par Relating the latter settings to our experiments, we assessed their all eight possible permutations. To give an example of a permutation, one of the experiments would assess the ability to recover the underlying topic term distributions with the enabled topic overlap, the disabled topic term overlap, and the enabled error overlap. Speaking of the data generation, note that the latter settings directly reflect the $\theta$ and $\phi$ values of the synthetic datasets (we do not use the $\alpha$ and $\beta$ hyper-parameters). By following the latter principle, we establish a clearer representation of the synthetic data; thus, simplify the performance assessment.

\subsection{Experiment execution}

% Evaluation
% - Performance: topic and term recreation error

\par Before going into the experiment execution, note that we assess the performance based on the models' ability to recover the underlying synthetic corpus generation settings. To wrap this assessment into a more concise terminology, the true solution corresponds to the underlying synthetic corpus generation settings; and the approximate solution corresponds to the inference results. As a result, the performance is measured by taking the difference between the true and approximate solutions.

% Specific examples
% - # of iter & burn-in
\par In order to introduce the rationale behind the performance assessment, we look into one of the eight experiments in more detail. Just like for all our experiments, we run both auto-regressive and non-auto-regressive models for $5000$ Gibbs sampling iterations, $1000$ of which are dedicated to the burn-in process. For each of the remaining $4000$ iterations, we sample the corresponding $\theta$ and $\phi$ values; afterwards, in every $100$th iteration, we average the stored $\theta$ and $\phi$ values, respectively; then, this average is compared to the true solution. As an example, in the $1100$th iteration, we would take the average of $100$ samples; in the $1200$th iteration, we would take the average of $200$ samples. In a single experiment, we would have $40$ of such batches indicating the performance -- this is illustrated in Figure \ref{fig:perf_theta} and Figure \ref{fig:perf_phi}.
\begin{figure}[h]
  \centering
  \begin{subfigure}[b]{0.37\textwidth}
    \includegraphics[width=\linewidth]{performance-theta.png}
  \end{subfigure}%
  \caption{The $\theta$ recovery performance.}
  \label{fig:perf_theta}

  \begin{subfigure}[b]{0.37\textwidth}
    \includegraphics[width=\linewidth]{performance-phi.png}
  \end{subfigure}%
  \caption{The $\phi$ recovery performance.}
  \label{fig:perf_phi}
\end{figure}

% The images of theta and phi recovery
\par Relating to the previous example, the $\theta$ and $\phi$ values corresponding to the last iteration are illustrated in Figure \ref{fig:comp_theta} and Figure \ref{fig:comp_phi}, respectively. Also, note that we relax the colour coding of our figures as the topics are inferred in unsupervised manner.

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\linewidth]{theta-ar.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\linewidth]{theta-nonar.png}
  \end{subfigure}%
  \caption{The comparison of the $\theta$ values.}
  \label{fig:comp_theta}

  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\linewidth]{phi-ar.png}
  \end{subfigure}%

  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\linewidth]{phi-nonar.png}
  \end{subfigure}%
  \caption{The comparison of the $\phi$ values.}
  \label{fig:comp_phi}
\end{figure}

% General assessment
% - The table with the paired t-test scores
% - Paired t-test
% -- 10 datasets each
\par To assess the auto-regressive model's performance, we have generated $10$ distinct datasets for each previously introduced corpus generation setting. Recall that we assess the following conditions: the topic overlap; the vocabulary term overlap; and the error term overlap. For every setting permutation, we perform the two-tailed t-test: the t-statistic suggests the difference in performance; and the p-value suggests whether the result is significant. To be more specific, a negative t-statistic indicates the auto-regressive model's superior performance; whereas the results are determined to be significant if the p-value is below $0.05$. The t-statistics and p-values of all eight performed experiments are provided in Table \ref{tab:ttest} below.

\begin{table}[H]
\begin{center}
\begin{tabular}{|>{\centering\arraybackslash}p{0.75cm}|>{\centering\arraybackslash}p{0.75cm}|>{\centering\arraybackslash}p{0.75cm}">{\centering\arraybackslash}p{0.75cm}|>{\centering\arraybackslash}p{0.75cm}|>{\centering\arraybackslash}p{0.75cm}|>{\centering\arraybackslash}p{0.75cm}|} 
\hline
\multicolumn{3}{|c"}{Overlap} &
\multicolumn{2}{c|}{t-statistic} &
\multicolumn{2}{c|}{p-value} \\
\hline
% \hline
Topic & Term & Error & $\theta$ & $\phi$ & $\theta$ & $\phi$ \\
\thickhline
False & True & True & $-5.41$ & $0.47$ & $0.00$ & $0.65$ \\
False & True & False & $-9.46$ & $-0.38$ & $0.00$ & $0.71$ \\
False & False & True & $4.74$ & $3.93$ & $0.00$ & $0.00$ \\
False & False & False & $-2.91$ & $-3.24$ & $0.02$ & $0.01$ \\
True & True & True & $-5.99$ & $-1.71$ & $0.00$ & $0.12$ \\
True & True & False & $-1.78$ & $-1.53$ & $0.11$ & $0.16$ \\
True & False & True & $1.12$ & $1.17$ & $0.29$ & $0.27$ \\
True & False & False & $0.52$ & $1.10$ & $0.61$ & $0.30$ \\
\hline
\end{tabular}
\end{center}
  \caption{The results of the t-test.}
  \label{tab:ttest}
\end{table}

% - Commenting on the significant results
\par Based on the obtained results, the most significant changes (the p-values vary from $0.00$ to $0.01$) occur upon only switching the error overlap setting (the remaining settings are disabled). If the error overlap is disabled, the spatial smoothing application displays an improved performance (i.e., $3.24$ lower error in recovering $\phi$); however, if the setting is enabled, the auto-regressive model performs poorly (i.e., $3.93$ higher error in recovering $\phi$). Further, by relaxing the significance threshold, we can also consider the experiment instances where the p-values vary from $0.12$ to $0.16$. Conveniently, in this experiment pair, we again consider only the switch of the error overlap setting; however, in this case, all other settings are enabled. To comment on the respective performance, the spatial smoothing application is superior in both cases: $1.71$ and $1.53$ lower error rates in recovering $\phi$.

% - Result pairing
% - Notes why some results are insignificant
\par Interestingly, we can group the experiment listings in four pairs: the pairs are centred around the $\phi$ p-values of $0.01$, $0.16$, $0.30$, or $0.71$. The first are pairs presented in the previous paragraph; however, for the last two pairs, the p-values are well beyond the significance threshold. By noting that, in every pair, only the error setting varies, the insignificant results occur when one of the topic and term settings is disabled and another is enabled. By looking into the $\phi$ recovery plots related to the insignificant results, we noticed that both auto-regressive and non-auto-regressive models infer similar latent $\phi$ distributions. Effectively, in the case of the enabled error setting, both models simplify the dataset complexity; that is, the models assign the overlapping error terms to the main topics. Alternatively, in the case when the error setting is disabled, the problem is too simple -- both models recover the $\phi$ values equally well. To give an example of a similar performance, the reader can consider the previously introduced Figure \ref{fig:perf_phi} and Figure \ref{fig:comp_phi}. However, by considering Figure \ref{fig:perf_phi}, note that the $\phi$ value of the auto-regressive model converges faster.

% ___________________________________________________________________________
\section{Conclusion}

% Summary
\par In this research paper, we reviewed an attempt to induce spatial smoothing in MSI data: the research problematic was supported and inspired by covering the relevant literature; the model's design was introduced by providing the preliminary knowledge covering LDA, spatial smoothing, and MSI data pre-processing; finally, the experiment settings were designed to identify both superior and inferior spatial smoothing application prospects.

% Results
% - Comment on the ability to recover phis
\par Our main objectives were to identify the spatial smoothing application's prospect in recovering the $\phi$ values used upon the generative data process and the ability to separate the noise topic. Speaking of the $\phi$ values, only a half of the carried experiments displayed significant performance compared to the topic model without the spatial smoothing application. We report an improved $\phi$ recovery performance on the synthetic datasets with the enabled topic and terms overlap settings; alternatively, when both topic and term overlap settings are disabled, the performance is superior if the error overlap is disabled and inferior if the error overlap is enabled.

% Results
% - Comments on the ability to find the noise
\par Speaking of the overlapping noise topic's identification, the auto-regressive model -- just like the non-auto-regressive model -- assigns the overlapping error terms to the main topics. For this reason, we conclude that the spatial smoothing application's impact in improving the overlapping noise topic terms is negligible. However, looking into the statistical test on the $\theta$ values, $6$ out $8$ experiments display a significant performance in recovering the synthetic corpus generation settings. In $5$ out $6$ cases, the $\theta$ values are recovered with a lower error; this result is mostly impacted by the model's ability to reflect the shape of the noise topic with a better accuracy.

% Results
% - Comment on the further prospects
% -- Spatial smoothing multi-neighbour
\par Since some of the experiment settings display a performance improvement, the spatial smoothing application can be considered for a further research. We would recommend looking into the application of the undirected graphical model -- Markov random field. Effectively, the application would establish more complex spatial smoothing settings: the spatial treatment of neighbouring entities could be improved from 1-dimensional to 2- or 3-dimensional. Effectively, the spatial dimensionality escalation would reflect the visual aspect of MSI data better.

% Future prospects
% - bucketisation
\par Speaking of the alternative research directions in assessing the spatial smoothing application, we propose a research problem on investigating bucketisation enhancements. In other words, the spatial smoothing application might have an impact upon the feature extraction from MSI data. To be more specific, spatial smoothing might establish more appropriate bucket size ranges in concatenating raw mass-to-charge values. If successful, this application would improve the quality of MSI data features and, consequently, improve the performance of the research problems addressed in this research project.

% - Comment on identifying appropriate settings
% -- All disabled settings do not display a metabolomics-like environment
\par To give a final verdict on the spatial smoothing application's performance, we consider the synthetic data settings which reflect the metabolomics-like environment the best. Effectively, if more overlap settings are enabled, then the synthetic corpus reflects the environment better. As shown by Table \ref{tab:ttest}, the auto-regressive model tends to perform better than the non-auto-regressive model when most of the overlap settings are enabled. Ultimately, spatial smoothing is effective upon considering more complex data.

% ___________________________________________________________________________
\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}

### 2017-04-20 code/models/dtm_alpha.py

@@ -3,27 +3,24 @@
 VERSION
 - Python 2
 
-FUNCTION
-- Infer topics by updating the alphas.
+PURPOSE
+- A topic model with spatial smoothing
 """
 
-import os
 import pandas as pd
 import numpy as np
 import copy
 from time import time
-
-np.set_printoptions(threshold=np.nan)
 
 # ___________________________________________________________________________
 
 
 class DTM_Alpha(object):
     '''
-    PARAMETERS:
-    - sigma_0_sq is the initial variance for alpha[0];
-    - sigma_sq is the variance for alpha[t]
-    - delta_sq is the variance for alpha'[t]
+    TERMS:
+    - sigma_0_sq is the initialisation variance for alpha_0;
+    - sigma_sq is the spatial smoothing variance for alpha_t
+    - delta_sq is the variance for the proposed alpha
     '''
     def __init__(self, K=None, beta=None, sigma_0_sq=None, sigma_sq=None,
                  delta_sq=None, autoreg=False, snapshot=False,
@@ -59,6 +56,9 @@ class DTM_Alpha(object):
         self.corpus = None
 # ===========================================================================
 
+    '''
+    Run a model from a captured snapshot.
+    '''
     def load_fit(self, path_file, n_it_add=0):
         vars_loaded = pd.read_pickle(path_file)
         for var in vars(self):
@@ -69,6 +69,9 @@ class DTM_Alpha(object):
         self.fit(n_it=n_it_add)
 # ===========================================================================
 
+    '''
+    Run a model.
+    '''
     def fit(self, n_it, n_burn_it=0, corpus=None):
         print('{} has started.'.format(self.__class__.__name__))
         if not self.snapshot:
@@ -82,12 +85,16 @@ class DTM_Alpha(object):
             if not self.burnin:
                 self._sample_phi()
                 self._sample_theta()
-                # self._calculate_perplexity()
             if self.verbose_init:
                 self.verbose_init = False
         self.n_it_sampl_last += n_it
         print('{} has finished.'.format(self.__class__.__name__))
 
+    '''
+    Print the status of the following:
+    - Alpha update;
+    - Burn-in.
+    '''
     def _update_status(self, it, n_burn_it):
         if it != 0 and it % 25 == 0:
[…]

### 2017-04-20 code/pre-processing/create_bne-corpus.py

@@ -5,8 +5,10 @@ import math
 VERSION
 - Python 2
 
-FUNCTION
-- Pre-process the 'bne' data
+PURPOSE
+- Pre-process the 'b and e' data
+-- 2D Mapping
+-- Vocabulary's bucketisation
 """
 
 import pymzml

### 2017-04-20 code/pre-processing/generator_corpus.py

@@ -3,8 +3,8 @@
 VERSION
 - Python 2
 
-FUNCTION
-- Corpus generation
+PURPOSE
+- A corpus generation for the notebooks
 """
 
 import numpy as np
@@ -12,10 +12,6 @@ import string
 
 
 class Generator_Corpus(object):
-    '''
-    Terms:
-    - pp: pre-processed
-    '''
     def __init__(self, alpha_init, theta_init, beta_init, xi, T):
         if alpha_init is not None:
             self.theta = self._softmax_matrix(alpha_init)
@@ -33,7 +29,6 @@ class Generator_Corpus(object):
             vocab.append(string.ascii_lowercase[w])
         return np.array(vocab)
 # ___________________________________________________________________________
-# TODO: Beta is phi
 
     def generate_corpus(self):
         corpus = {}
@@ -49,7 +44,6 @@ class Generator_Corpus(object):
                 if w_key not in corpus[d]:
                     corpus[d][w_key] = 0
                 corpus[d][w_key] += 1
-
         return corpus
 
     def _softmax_matrix(self, arr):

### 2017-04-20 code/pre-processing/helper_experiment.py

@@ -3,8 +3,8 @@
 VERSION
 - Python 2
 
-FUNCTION
-- Functions to be used during the experiments
+PURPOSE
+- A experiment execution helper for the notebooks
 """
 
 import numpy as np
@@ -12,19 +12,21 @@ import numpy as np
 
 class Helper_Experiment(object):
 
+    '''
+    Calculating the error difference between the true approximated
+    values of theta and phi.
+    '''
     def calculate_performance(self, alpha_init, beta_init, theta_init,
                               clf, s_batch):
         hist_theta = np.array(clf.hist_theta)
         if alpha_init is None:
             theta_init = theta_init.T
-        # generator_batch = self._split_to_batches(hist_theta, s_batch)
         performance_batch_theta = []
         performance_batch_phi = []
         '''
         Iterate through the theta batches
         '''
         n_iter = int(len(hist_theta) / s_batch)
-        # for hist_theta_batch in generator_batch:
         for i in range(n_iter):
             hist_theta_batch = hist_theta[0:i * s_batch - 1]
             '''
@@ -35,7 +37,7 @@ class Helper_Experiment(object):
                 theta_init = self._softmax_matrix(alpha_init)
                 theta_init = theta_init.T
             theta_latent = theta_latent.T
-            matchings_topic = self._tune_thetas(theta_init, theta_latent)
+            matchings_topic = self._match_thetas(theta_init, theta_latent)
             '''
             Calculate the theta similarity
             '''
@@ -51,8 +53,6 @@ class Helper_Experiment(object):
         Iterate through the phi batches
         '''
         hist_phi = np.array(clf.hist_phi)
-        # generator_batch = self._split_to_batches(hist_phi, s_batch)
-        # for hist_phi_batch in generator_batch:
         for i in range(n_iter):
             hist_phi_batch = hist_phi[0:i * s_batch - 1]
             phi_latent = np.average(hist_phi_batch, axis=0)
@@ -69,11 +69,10 @@ class Helper_Experiment(object):
             performance_batch_phi.append(sum_phi)
         return (performance_batch_theta, performance_batch_phi)
 
-    def _split_to_batches(self, arr, s_batch):
-        for i in range(0, len(arr), s_batch):
-            yield arr[i:i + s_batch]
-
-    def _tune_thetas(self, theta_init, theta_latent):
+    '''
+    Matching the real and inferred topics
+    '''
+    def _match_thetas(self, theta_init, theta_latent):
         '''
         Create the matrix of all distances
         '''
[…]

### 2017-04-20 code/pre-processing/plotter_corpus.py

@@ -1,82 +1,0 @@
-
-import numpy as np
-import math
-import plotly.graph_objs as go
-
-
-class Visualiser_Corpus(object):
-
-    def __init__(self, corpus, n_rows, l_row, l_column):
-        self.corpus = corpus
-        self.vocab = self._extract_vocab()
-        self.T = len(self.corpus)
-        self.V = len(self.vocab)
-        self.n_rows = n_rows
-        self.l_row = l_row
-        self.l_column = l_column
-        self.n_columns, self.map_coord = self._generate_coordinates()
-# ___________________________________________________________________________
-
-    def _extract_vocab(self):
-        vocab = []
-        for doc in self.corpus:
-            for word in doc:
-                if word not in vocab:
-                    vocab.append(word)
-        return vocab
-# ___________________________________________________________________________
-
-    def _generate_coordinates(self):
-        s_column, s_row = self._calc_sizes_map()
-        x_coord = 0
-        y_coord = self.n_rows - 1
-        scanning_right = True
-        map_coord = {}
-        for j in range(0, n_rows):
-            for i in range(0, s_row):
-                doc_id = i + s_column * j + s_row * j
-                map_coord[doc_id] = {'x': x_coord, 'y': y_coord}
-                # Switching horizontal scanning to vertical.
-                if (i % (s_row - 1) == 0 and i != 0):
-                    y_coord -= 1
-                    continue
-                if scanning_right:
-                    x_coord += 1
-                else:
-                    x_coord -= 1
-            # Reversing the scanner's direction.
-            scanning_right = not scanning_right
-        return s_row, map_coord
-
-    def _calc_sizes_map(self):
-        s_scan = (self.n_docs + 1) / self.n_rows
-        s_column_r = s_scan * (self.l_column / (self.l_row + self.l_column))
-        s_column = int(math.floor(s_column_r))
-        s_row = s_scan - s_column
-        return s_column, s_row
-# ___________________________________________________________________________
-
-    def locate_word(self):
-        # Mapping to the z values.
-        z_matrix = np.zeros(shape=(self.n_rows, self.n_columns))
-        for i, doc in corpus.iteritems():
-            if word_key in doc and int(i) in self.map_coord:
-                doc_coord = map_coord[int(i)]
-                z_matrix[doc_coord['y']][doc_coord['x']] = doc[word_key]
-        # Creating the x and y axes.
-        x_vector = np.zeros(shape=(self.n_columns, 1))
[…]

### 2017-04-20 code/pre-processing/processor_corpus.py

@@ -12,10 +12,7 @@ import math
 
 
 class Processor_Corpus(object):
-    '''
-    Terms:
-    - pp: pre-processed
-    '''
+
     def __init__(self, corpus, vocab=None):
         self.corpus = corpus
         self.vocab = vocab
@@ -53,7 +50,7 @@ class Processor_Corpus(object):
             else:
                 factor_norm = 1.
             '''
-            Transforming into a numpy array
+            Transforming the corpus into a numpy array
             '''
             for w, word in enumerate(self.vocab):
                 if word in doc:
@@ -62,6 +59,9 @@ class Processor_Corpus(object):
         return corpus_pp
 # ___________________________________________________________________________
 
+    '''
+    Generate the coordinates for the 2-dimensional bne plotting.
+    '''
     def generate_coordinates(self, n_rows, l_row, l_column):
         s_column, s_row = self.calc_sizes(self.T, n_rows, l_row, l_column)
         x_coord = 0
@@ -89,4 +89,3 @@ class Processor_Corpus(object):
         s_column = int(math.floor(s_scan * (l_column / (l_row + l_column))))
         s_row = s_scan - s_column
         return s_column, s_row
-

### 2017-04-20 code/pre-processing/visualiser_corpus.py

@@ -1,19 +1,26 @@
 
+"""
+VERSION
+- Python 2
+
+PURPOSE
+- A plotting helper for the notebooks
+"""
+
 import numpy as np
 import math
 import plotly.graph_objs as go
 import matplotlib.pyplot as plt
 import os
-# ___________________________________________________________________________
 
 
 class Visualiser_Corpus(object):
 
-    def __init__(self, corpus, vocab, n_rows, l_row, l_column, path_plots=None, save=True, no_corpus=None, no_experiment=None):
+    def __init__(self, corpus, vocab, n_rows, l_row, l_column, path_plots=None,
+                 save=True, no_corpus=None, no_experiment=None):
         self.corpus = corpus
         self.no_corpus = no_corpus
         self.no_experiment = no_experiment
-        # self.vocab = self._extract_vocab()
         self.path_plots = path_plots
         self.vocab = vocab
         self.T = len(self.corpus)
@@ -31,18 +38,7 @@ class Visualiser_Corpus(object):
                 os.makedirs(self.path_plots)
 
 # ___________________________________________________________________________
-    '''
-    TODO: Enable to vocab creation
-    '''
-
-    def _extract_vocab(self):
-        vocab = []
-        for doc in self.corpus:
-            for word in doc:
-                if word not in vocab:
-                    vocab.append(word)
-        return vocab
-# ___________________________________________________________________________
+# The class's initialisation
 
     def _generate_coordinates(self):
         s_column, s_row = self._calc_sizes_map()
@@ -50,7 +46,6 @@ class Visualiser_Corpus(object):
         y_coord = self.n_rows - 1
         scanning_right = True
         map_coord = {}
-        # print(s_row)
         for j in range(0, self.n_rows):
             for i in range(0, s_row):
                 doc_id = i + s_column * j + s_row * j
@@ -75,6 +70,8 @@ class Visualiser_Corpus(object):
         return s_column, int(s_row)
 
 # ___________________________________________________________________________
+# The class's helper functions
+
     def softmax(self, arr):
         return np.exp(arr) / np.sum(np.exp(arr))
 
@@ -87,7 +84,7 @@ class Visualiser_Corpus(object):
 # ___________________________________________________________________________
 
     '''
-    Used on a pre-processed corpus
+    Plot a vocabulary term of a pre-processed corpus
     '''
     def locate_word(self, word):
         # Mapping to the z values.
[…]

### 2017-04-20 docs/msci-paper/paper_ssmsi.tex
The final project submission.

@@
 % - experiments: synthetic data
+
 \begin{abstract}
@@
 By assessing mass spectrometry imaging data, we design an enhancement for latent Dirichlet allocation-based topic models.
-For both data pre-processing and model design, we survey relevant research.
-Further, we focus on presenting our methodology in detail providing the preliminaries to the basic models and guiding through the performed adjustments in the proposed model variations.
-To assess the application's performance, we run the model comparison on a number of diverse synthetic datasets.
+For both data pre-processing and topic model design, we survey relevant research.
+Further, we present the proposed methodology in detail providing the preliminaries and guiding through the performed topic model enhancements.
+To assess the performance, we evaluate the spatial smoothing application on a number of diverse synthetic datasets.
 \end{abstract}
@@
 % - Focus: Metabolomics and topic modelling
-\par In this research paper, we assess an application of spatial smoothing in visual data; that is, we induce continuity among data elements.
+\par In the research project, we assess an application of spatial smoothing in visual data; that is, we induce continuity among data elements.
 The spatial smoothing application is particularly targeted to be applied for unsupervised pattern recognition.
-To be more specific, our focus is to model a biomedical application in the field of metabolomics.
-Furthermore, we limit the scope of the applied unsupervised machine learning techniques to the branch of topic modelling.
+To be more specific, our focus is to model visual metabolomics data using a particular branch of unsupervised machine learning -- topic modelling.
 
@@
 % - MSI data structure
-\par The characteristics of our utilised metabolomics data are expressed in the form of mass spectrometry imaging (MSI).
-Effectively, we use MSI to visualise the metabolomics data in the form of spatial distribution.
-Speaking of the metabolomics data, it contain information about ionised metabolites.
-Note that metabolites are molecules produced by the chemical process of metabolism, whereas by ionisation we refer to the information capture method.
-In other words, MSI data is a visualisation of ion (sampled metabolite) distributions: the complete dataset is the whole image; the image's pixel is a particular sampling region; and each region contains intensities of ions with unique mass-to-charge $m/z$ values.
+\par The characteristics of the utilised metabolomics data are expressed in the form of mass spectrometry imaging (MSI).
+To briefly introduce mass spectrometry, the method captures a biological tissue in the form of mass spectrum: at the start, the metabolites of a biological tissue are ionised (by metabolites, we refer to the tissue's contents -- the product of a metabolism chemical process); after the initial step, a mass spectrometer captures the intensity of each ionised metabolite; as a result, we obtain information about the tissue's contents and their concentrations.
+
+% Basis:
+% - MSI data structure
+\par Relating the latter procedure to MSI, note that we can partition the tissue into small regions.
+Effectively, the tissue's partitioning corresponds to a higher granularity of the data.
+As a result, we can identify the contents of the tissue's particular regions.
+As an example, we provide Figure \ref{fig:msi} illustrating the rationale behind MSI: the image is the whole tissue; the image's pixel is the tissue's particular region; and each pixel contains the intensity values of ions (a different ion type has a unique mass-to-charge $m/z$ value).
+In order to visualise MSI data, we would look into a specific mass-to-charge value.
+Effectively, the intensity of a pixel would be set to the intensity value of the respective ion type.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.42\textwidth]{msi3.png}
+  \caption{The MSI data structure.}
+  \label{fig:msi}
+\end{figure}
 
@@
 % - topic modelling w.r.t. MSI
-\par Speaking of our machine learning application, topic modelling is a technique used to infer unlabelled topic distributions based on data's underlying semantic structure.
-With respect to MSI data, we can model the topic distributions over an image and its every pixel; furthermore, topic modelling can express the types of ions corresponding to particular topics.
-Since we use a topic model to model metabolomics data in a statistic manner, the utilised topic models are tuned to reflect the metabolomics environment as realistically as possible.
-Relating to our research targets, spatial smoothing is one of such environment settings.
+\par Going into the machine learning application, note that we use topic modelling for the inference of topic distributions; effectively, the inferred topic distributions would be treated as the underlying semantic structure of a sampled tissue.
+Since our data is MSI-based, we model the topic distributions over an image and its every pixel; also, we model the types of ions corresponding to particular topics.
+Since our machine learning application is based on Bayesian methods, we tune the model parameters to reflect the metabolomics environment as realistically as possible.
+Ultimately, spatial smoothing is one of such environment settings.
 
@@
 \par The basis of the project's research problems comes from the limitations of current metabolite sampling techniques.
-The main limitation is the loss of information caused by the metabolite ionisation.
-As a consequence, the MSI data is noisy.
-To expand on the noisiness, it is caused by metabolite fragmentation and the limitation to tune different metabolite retention times.
-By metabolite fragmentation, we refer to a metabolite split; the split would cause the captured ions to possess unexpected values.
-Speaking of the retention time, the intensity value of each ion type varies with respect to time; therefore, since each ion is captured at a different state of its retention, each ion type would possess some variance with respect to its intensity.
-Finally, note that the loss of information might be also caused by overlapping ion topics.
-Since different ion topics can contain same ion types, the topic possessing a lower ion intensity value would be overwhelmed and, thus, not reflected in the MSI data.
+One of the limitations is the presence of noise.
+For example, MSI data could be distorted by noise as a result of metabolite fragmentation -- the ionisation would split the metabolite's structure causing a faulty mass-to-charge value capture.
+Another limitation is the overlap of the molecules with similar structures.
+Effectively, some of the produced ions could have same mass-to-charge values; as a result, the ion possessing a lower intensity value would be overwhelmed and, thus, not reflected in MSI data.
+A possible approach mitigating the issue is direct infusion; however, in this project, we investigate whether the latter issue could be mitigated using a computational approach.
+
 % Contributions:
@@
 % - A tuned topic model and well-organised experiment settings
-\par In this paper, we contribute to the research in MSI by carrying an extensive assessment of the spatial smoothing application.
-The assessment is carried in both quantitative and qualitative manners: we assess the performance on a number of diverse datasets; also, our experiments are designed to reflect the nature of the MSI data reflecting computational metabolomics.
-Furthermore, we provide a Python implementation of a tuned topic model; also, we establish maintainable experiment settings.
-By the tuned topic model, we mean that the model's implementation is particularly designed to meet the characteristics of the MSI data.
-Speaking of the experiment settings, note that we are carrying the experiments using Jupyter notebooks.
-Effectively, the use of the notebooks creates a portable and well-documented environment to initialise the experiment settings and execute the topic modelling.
-As a result, external parties could run the released notebooks and reproduce the experiment results in a swift manner.
+\par We contribute to the research in MSI by carrying an extensive assessment of the spatial smoothing application.
+The assessment is performed in both quantitative and qualitative manners: we assess the performance on a number of diverse datasets; also, our experiments are designed to reflect the environment settings of computational metabolomics.
+Furthermore, we provide Python implementations of the proposed topic model and the experiment settings.
+Note that the experiments are performed in Jupyter notebooks; effectively, the practice induces a portable and well-documented environment for initialising and running the experiments.
+The notebooks are also useful in validating the experiment results: external parties would be able to re-run the experiments in a swift manner.
 
@@
 One of the model's key characteristics is the three-level hierarchical treatment of the data.
-In the context of the utilised MSI data, the hierarchical structure can be perceived as follows: in the highest level, we have an MSI image; in the middle level, we have a pixel of an MSI image; and in the lowest level, we have the intensities of particular ions in a pixel.
+In the context of the utilised MSI data, the hierarchical structure can be perceived as follows: in the highest level, we have an MSI image; in the middle level, we have the pixels of an MSI image; and in the lowest level, we have the intensities of particular ions in a pixel.
 
@@
 \par Another model's key characteristic is the generative treatment of the data.
-By a generative model, we mean that the latent data instances are treated as a result of a mixture of underlying parameters drawn from probability distributions.
-In other words, the generative data treatment induces randomness in the end products of the data; however, note that the source of the data -- the lowest level parameters of the probability distributions -- remain the same.
-The key aspect of the generative model is the degree of freedom in the connections of random variables; this notion allows modelling more realistic, thus, more complex data relations.
-Ultimately, the rationale of the generative model is based on recovering the underlying semantic structure.
-Effectively, in order to recover the generative process of the data, we delve into the applications of Bayesian methods.
+By a generative model, we mean that the latent data instances are treated as a result of a mixture of underlying parameters.
+In other words, the generative data treatment induces randomness in the end product of the data generation; however, note that the source of the data -- the lowest level parameters -- remain the same.
+The key aspect of the generative model is the degree of freedom in the connections of random variables; this notion allows modelling more realistic, thus, more complex data settings.
+Ultimately, the rationale of a generative machine learning model is based on recovering the underlying semantic structure.
+To do this, we employ the rationale of Bayesian methods.
 
@@
 \par By applying Bayes' rule, we can express the underlying semantic structure in the form of a posterior probability distribution.
-However, note that we can not analytically compute such posterior expression.
-Instead, we can estimate the posterior distributions using optimisation or direct sampling.
-In this project, we particularly focus on the underlying semantic structure's inference using direct sampling.
-Speaking of the sampling-based inference method applications to LDA-like models, the ground-work has been established by Griffiths and Steyvers \cite{griffiths2004finding}.
-The authors have utilised the collapsed Gibbs sampling -- a Markov chain Monte Carlo (MCMC) algorithm.
-Effectively, the method integrates the uncertainty out; so that we sample the entities of interest directly.
-More specific details about the rationale of the collapsed Gibbs sampling application will be provided in Section 4.
+Since the posterior expression can not always be analytically computed, we estimate it using optimisation or direct sampling.
+Note that, in this project, we particularly focus on the topic inference using direct sampling.
+Relating to the attempts for applying the sampling-based inference to LDA-like models, the ground-work was established by Griffiths and Steyvers \cite{griffiths2004finding}.
+The authors display an application of the collapsed Gibbs sampling -- a Markov chain Monte Carlo (MCMC) algorithm.
+Since the method integrates the uncertainty out, we can sample the data entities without the explicit notion of the underlying parameters.
 
@@
 % - Discrete data
-\par Relating to the initial LDA model, the authors have set the following assumptions for the utilised data: exchangeability among the inner components of the lower and middle data hierarchy levels; also, a discrete treatment of the lower-level data.
+\par Going back to the initial LDA model, the model's authors report on setting the following assumptions: exchangeability among the inner components of the lower and middle data hierarchy levels; and a discrete treatment of the lower-level data.
 By exchangeability, it is meant that the components follow the bag-of-words principle; that is, the order of the data has no correlation with the underlying semantic structure.
-Speaking of the discrete data treatment, it is assumed that the lower-level components have no spatial connection.
-With respect to our project, both assumption types -- exchangeability and discreteness -- do not correspond to the characteristics of the MSI data.
-Therefore, we will establish a methodology to relax the latter assumptions.
+Relating to the rationale behind the discrete data treatment, it is assumed that the lower-level components have no spatial connection.
+With respect to our project, both assumptions -- the exchangeability and the discreteness -- do not correspond to the characteristics of the metabolomics environment.
+Therefore, we establish a methodology to relaxing the latter assumptions.
 
@@
 \par In this subsection, we introduce our topic modelling terminology.
-In order to put the MSI data in the topic modelling context, the components of  LDA's three-level hierarchical structure are defined as follows: \textit{a corpus} is the whole image of a sample; \textit{a document} is an MSI image's pixel; and \textit{a word} is an interval of mass-to-charge values corresponding to a particular ion.
-Starting from now on, we will use the latter topic modelling concepts to introduce the LDA model's architecture and notation.
+In order to put the MSI data structure in the topic modelling context, the components of LDA's three-level hierarchical structure are defined as follows: \textit{a corpus} is the whole image of a sample; \textit{a document} is an MSI image's pixel; and \textit{a word} is an interval of mass-to-charge values corresponding to a particular ion.
+Note that from now on, we use the latter topic modelling concepts to introduce the LDA model's architecture and notation.
 
@@
 % - Graphical model
-\par Since LDA-like models are also graphical models, the dependence of random variables can be illustrated using graphs.
+\par Since LDA-like models correspond to the branch of graphical models, the dependence of random variables can be illustrated using graphs.
 In order to familiarise with the architecture and the variables of LDA-like models, we provide Figure \ref{fig:arch_LDA-init} illustrating the initial LDA model in the plate notation.
@@
   \centering
-  \includegraphics[width=0.25\textwidth]{LDA-initial.png}
+  \includegraphics[width=0.20\textwidth]{LDA-initial.png}
   \caption{The initial LDA model's architecture.}
@@
 \end{figure}
-The circles indicate the model's variables: the coloured circle corresponds to the observable variable, whereas the uncoloured circles correspond to the hidden variables.
+To start with, the circles indicate the model's variables: the coloured circle corresponds to the observable variable; and the uncoloured circles correspond to the hidden variables.
 Effectively, the hidden variables define the model's underlying structure.
-Speaking of the plates, the plate denoted by $K$ corresponds to the number of topics; the plate denoted by $T$ corresponds to the number of documents, and the plate denoted by $N$ corresponds to the number of words.
+To introduce the plates, the plate denoted by $K$ corresponds to the number of topics; the plate denoted by $T$ corresponds to the number of documents, and the plate denoted by $N$ corresponds to the number of words.
 Effectively, the letters in the bottom right corners indicate the total number of variables.
@@
 % - Listing
-\par Before providing a listing with the definitions of the LDA variables, we introduce the purpose of the variables.
+\par Before providing a listing with the definitions of the LDA variables, we introduce their purpose.
 The variables denoted by $\theta$ and $\phi$ correspond to the underlying probability distributions (e.g., in the case of the initial LDA model, we use Dirichlet distributions).
@@
 Effectively, \textit{a hyper-parameter} allows tuning a machine learning model for a particular dataset application.
-As a side reference, note that by \textit{a vocabulary} we refer to a collection of terms reflecting a fixed range of the words.
+As a side reference, note that by \textit{a vocabulary} we refer to a collection of terms reflecting a fixed range of words.
 At this point, we provide the following list containing the definitions of the initial LDA model's variables:
@@
 % - Describe the data pre-processing requirements
-\par In this subsection, we introduce the qualities of the raw MSI data.
-Furthermore, we set the requirements for the raw data's pre-processing; note that the pre-processing serves as an auxiliary method making the raw MSI data compatible for a scalable topic modelling application.
+\par In this subsection, we introduce the qualities of the MSI data.
+Furthermore, we set the requirements for the pre-processing of raw MSI data; note that the pre-processing serves as an auxiliary method making raw MSI data compatible for a scalable topic modelling application.
 As a side note, the basis of our applied MSI data characteristics is established from the mzML data format.
@@
 % - Explain the raw data by looking into it
-\par Essentially, raw MSI data contain mass spectra of a sample -- the sample's every pixel is expressed in the form of a mass spectrum.
-By a mass spectrum, we refer to a map from a mass-to-charge value to an intensity value.
-Ultimately, we can establish a continuous notion of the MSI data by ordering the mass-to-charge values.
-Note that since the sampling equipment can detect the mass-to-charge values in the millidalton (mDa) precision, the MSI data is sparse (i.e., a large portion of mass-to-charge values are mapped to the zero intensity).
+\par Recall that raw MSI data is mass spectra of a tissue's sample (i.e., the sample's every pixel is expressed in the form of a mass spectrum); also, every mass spectrum term conveys a particular intensity value.
+To help organise the data, we establish a continuous notion by ordering the mass-to-charge values.
+Further, since the sampling equipment can detect the mass-to-charge values in the millidalton (\textit{mDa}) precision, the MSI data is sparse (i.e., a large portion of mass-to-charge values are mapped to the zero intensity).
 
@@
 % - Dismissing inferior information
-\par In the context of topic modelling, the raw MSI data possess a large vocabulary (above $5000$ terms).
+\par In the context of topic modelling, raw MSI data possess a large vocabulary (above $5000$ terms).
 Note that we consider every mass-to-charge value as a word; whereas every intensity value is perceived as a word's occurrence count.
 Further, since the intensity values could spike up above $1000$, the time complexity of the latent topic inference would require pro-longed runs.
-To overcome the introduced scalability issues, we carry data pre-processing -- the applied techniques are discussed in the upcoming methodology section.
+To overcome the introduced scalability issues, we carry data pre-processing (the applied techniques are discussed in the upcoming methodology section).
 
@@
 \begin{itemize}
-	\item Improving the detection of overlapping topics and vocabulary terms;
-	\item Reducing the noisiness of the MSI data;
-	\item Motivating a further research in the spatial smoothing application in MSI data.
+	\item Improve the detection of overlapping topics and vocabulary terms;
+	\item Reduce the noisiness of the MSI data;
+	\item Motivate a further research in the spatial smoothing application in MSI data.
 \end{itemize} 
 To expand on the overlapping topic detection, the issue arises when two underlying topics are made of similar vocabulary terms: instead of a separate representation, the topics are merged.
-We intend to identify the flow of distinct topics by the application of spatial smoothing; as a consequence, the spatial smoothing would also impact the data noisiness reduction.
-Ultimately, if a naive spatial smoothing application displayed a performance improvement, the contribution would set a basis to utilise state-of-the-art auto-regression techniques on the MSI data.
+We intend to identify the flow of distinct topics by applying spatial smoothing; as a consequence, the spatial smoothing would also impact the data noisiness reduction.
+Ultimately, if a naive spatial smoothing application displayed a performance improvement, the contribution would set a basis to utilise state-of-the-art auto-regression techniques in MSI data.
 
@@
 % - Tune the models by studying synthetic datasets.
-\par To my knowledge, the impact of the spatial smoothing application to the MSI domain has not yet been thoroughly studied.
-For the latter reason, this research project serves as an exploratory assessment of the spatial smoothing application: we introduce the rationale behind the applied methodology; also, we clearly define the range of the experiment settings.
-To give a brief intuition about the methodology, the study assesses the domain-specific parameter tuning and its impact on a diverse range of synthetic datasets.
+\par To my knowledge, the impact of applying spatial smoothing to MSI data has not yet been thoroughly studied.
+For the latter reason, this research project serves as an exploratory assessment -- we introduce the rationale behind the applied methodology; also, we clearly define the range of the experiment settings.
+To give a brief intuition about the methodology, the study assesses the domain-specific parameter tuning and its impact on a range of synthetic datasets.
 
@@
 In order to establish the basis of the state-of-the-art computational metabolomics methodology, we consult the survey carried by Alonso et al.
-\cite{alonso2015analytical}; speaking of the key probabilistic topic modelling branches, we consult the survey carried by Blei \cite{blei2012probabilistic}.
+\cite{alonso2015analytical}; to introduce the key probabilistic topic modelling branches, we consult the survey carried by Blei \cite{blei2012probabilistic}.
 
@@
 To start with, data normalisation serves as a method establishing an adaptable data structure.
-Bolstat et al.
-\cite{bolstad2003comparison} have proposed a data normalisation method called linear baseline scaling.
+Bolstad et al.
+\cite{bolstad2003comparison} have proposed the data normalisation method called linear baseline scaling.
 Note that the method is particularly targeted at sparse datasets.
@@
 % -- de2008nmr
-\par Speaking of feature binning, the technique is used to merge the instances of a feature space; as a result, feature binning reduces the data dimensionality.
+\par To introduce feature binning, the technique is used to merge the instances of a feature space; as a result, feature binning reduces the data dimensionality.
 Note that by the MSI data feature space, we refer to distinct $m/z$ values.
@@
 As an alternative approach, De et al.
-\cite{de2008nmr} have performed a feature binning based on the spectral peak identification.
+\cite{de2008nmr} have performed a feature binning based on the identification of spectral peaks.
 Effectively, the technique induces a dynamic notion of bin boundaries which are based on the identified intensity peak regions.
@@
 \par The last reviewed MSI data pre-processing techniques addresses the MSI data noisiness.
-Even though the MSI data noise reduction is an open research problem going well beyond data pre-processing, Smith et al.
+Even though the MSI data noise reduction is an open research problem utilising techniques beyond data pre-processing, Smith et al.
 \cite{smith2014proteomics} have shown that the application of a general data pre-processing routine displays performance improvements.
@@
 \par In this subsection, we review the key characteristics of the prospective topic modelling inference techniques and model variations.
-Speaking of the inference techniques, we employ the rationale of a sampling-based LDA model.
+To apply the inference techniques, we employ the rationale of a sampling-based LDA model.
 Griffiths and Steyvers \cite{griffiths2004finding} have displayed a successful application of the collapsed Gibbs sampler for the topic inference of both textual and visual data.
-To expand on the collapsed sampling, the technique allows skipping the estimation of $\theta$ and $\phi$ values; instead, the inference is based on the notion of the assignment counts.
+To expand on the collapsed sampling, the technique allows skipping the estimation of the $\theta$ and $\phi$ values; instead, the inference is based on the notion of the assignment counts.
 In order to estimate a topic assignment's probability, the authors suggest using the following expression:
@@
 \end{equation*}
-note that $i$ refers to the current iteration ($-i$ refers to the previous iteration); $k$ refers to a particular topic; and $n$ refers to the count of the instances indicated by the term's superscript.
+Note that $i$ refers to the current iteration ($-i$ refers to the previous iteration); $k$ refers to a particular topic; and $n$ refers to the count of the instances indicated by the term's superscript.
 If required, $\theta$ and $\phi$ can be sampled from the following distributions:
@@
 Ultimately, the documents of a corpus are assigned to segments with unique $\theta$ and $\phi$ values.
-Note that the changes in $\theta$ and $\phi$ values are impacted by their hyper-parameter updates.
-Since in the DTM paper the authors use variational inference, we provide a reference for an alternative dynamic topic modelling variation -- sequential LDA -- proposed by Du et al.
+Note that the changes in the $\theta$ and $\phi$ values are impacted by their hyper-parameter updates.
+Since, in the DTM paper, the authors use variational inference, we provide a reference for an alternative dynamic topic modelling variation -- sequential LDA -- proposed by Du et al.
 \cite{du2012sequential}.
@@
 % -- palmer2016fdr
-\par Speaking of the spatial smoothing application, to my knowledge, the idea has not yet been widely spread among the computational metabolomics community.
+\par Relating to the spatial smoothing application, to my knowledge, the idea has not yet been widely spread among the computational metabolomics community.
 Nevertheless, a recent study by Palmer et al.
@@
 % Intro
-\par In this section, we cover the rationale of a topic model tuned for the spatial smoothing application.
-To start with, we provide details on how to establish the auto-regressive notion among MSI data; then, we show how to apply the auto-regression to a topic inference based on the collapsed Gibbs sampling.
+\par In this section, we cover the rationale of the topic model tuned for the spatial smoothing application.
+To start with, we provide details on how to establish the auto-regressive notion among MSI data; then, we show how to apply the auto-regression to the topic inference based on the collapsed Gibbs sampling.
 Further, we provide a list of the applied data pre-processing techniques for MSI data.
-Finally, by considering the data format induced by the data pre-preprocessing, we introduce the generative MSI data process.
+Finally, considering the data format induced by data pre-preprocessing, we introduce the generative MSI data process.
 Effectively, we apply the generative process to generate synthetic data for our experiments.
@@
 % - Point out that we are using the terms corresponding to the MSI data
-\par We establish the spatial smoothing among MSI data by inducing the auto-regressiveness among the pixels (documents).
+\par We establish the spatial smoothing among MSI data by inducing the auto-regressiveness among pixels (documents).
 Before going into details, note that we cover the established methods using the previously introduced topic modelling notation.
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{ar-architecture.png}
+  \includegraphics[width=0.33\textwidth]{ar-architecture.png}
   \caption{The auto-regressive topic model architecture.}
@@
 % Gibbs Sampler changes
-\par Speaking of the Gibbs sampler's enhancement, the updated topic assignment formula takes the following form:
+\par In order to enhance the Gibbs sampler, we modify the topic assignment formula to take the following form:
 \begin{equation*}
@@
 In logarithmic space, however, the products are transformed into sums.
-Speaking of our model, we apply logarithmic space for both the auto-regressive $\alpha$ update and the sampling-based inference.
+Relating to our model, we apply logarithmic space for both the auto-regressive $\alpha$ update and the sampling-based inference.
 The updated expression for the auto-regressive $\alpha$ update is given as follows:
@@
 \end{align*}
-Speaking of the updated expression for the inference, it is updated in the following way:
+To introduce the updated expression for the inference, it is expressed as follows:
 \begin{align*}
@@
 
-\par Before carrying the experiments, we familiarise with the raw MSI data characteristics and assess their scalability.
+\par Before carrying the experiments, we familiarise with raw MSI data characteristics and assess their scalability.
 To be more specific, we define the characteristics of our synthetic data by pre-processing a real MSI data sample.
-To introduce the pre-processing details, we dismiss the words below the intensity threshold of $10$; then, we apply the following bucketisation strategy: merge adjacent vocabulary terms which differ less than $7$ mDA.
+To introduce the pre-processing details, we dismiss the words below the intensity threshold of $10$; then, we apply the following bucketisation strategy: merge adjacent vocabulary terms which differ less than $7$ \textit{mDa}.
 Effectively, the bucketisation strategy is based on the spectral peak identification.
@@
   \begin{subfigure}[b]{0.5\textwidth}
-    \caption{The term's occurrences before linear baseline scaling.}
+    \caption{The `b and e' term's occurrences before linear baseline scaling.}
     \includegraphics[width=\linewidth]{pre.png}
@@
   \begin{subfigure}[b]{0.5\textwidth}
-    \caption{The term's occurrences after linear baseline scaling.}
+    \caption{The `b and e' term's occurrences after linear baseline scaling.}
     \includegraphics[width=\linewidth]{post.png}
@@
   \begin{subfigure}[b]{0.5\textwidth}
-    \caption{An inferred topic corresponding to the pattern.}
+    \caption{An inferred topic corresponding to the `b and e' pattern.}
     \includegraphics[width=\linewidth]{topic.png}
   \end{subfigure}%
-  \caption{The term's visual comparison.}
+  \caption{The comparison of the `b and e' term's extraction.}
   \label{fig:comparison}
@@
 % - 1 document is 1 segment
-\par Having a basis for a scalable inference, we transfer the identified data properties into the synthetic corpus generation.
+\par Having the basis for a scalable inference, we transfer the identified data properties into the synthetic corpus generation.
 Before introducing the generative process, recall that our dynamic topic treatment is unique with respect to every document.
 Therefore, contrary to the reviewed dynamic topic models, our dynamic segment consists of only one document.
-Applying the latter aspects, we establish our utilised generative process given in Algorithm \ref{alg:dynamic_document_generation} below:
+Considering the latter aspects, we establish our utilised generative using Algorithm \ref{alg:dynamic_document_generation} given below:
 \begin{algorithm}[H]
@@
 \end{algorithm}
-In practical settings, the rationale of the generative process is defined as follows: the $\xi$ term represents an approximate number of words per document; $\alpha_t$ is the pre-defined auto-regressive hyper-parameter for document $t$; and $\phi_k$ is the pre-defined vocabulary term distribution for topic $k$.
-
+In practical settings, the rationale of the generative process is defined as follows: the $\xi$ term represents an approximate number of words per document; $\alpha_t$ is the pre-defined auto-regressive hyper-parameter for the document $t$; and $\phi_k$ is the pre-defined vocabulary term distribution for the topic $k$.
 % ___________________________________________________________________________
@@
 At the start of the section, we define the settings for tuning the topic models; then, we look into the settings for generating the synthetic datasets; afterwards, we introduce the scope of our experiments.
-Having defined the settings, we provide several illustrative examples of the experiment execution; and finally, we show the results of all experiments.
+Having defined the settings, we provide several illustrative examples of the experiment execution; and finally, we show the results of the performed experiments.
 
@@
 Effectively, in the non-auto-regressive model, we do not have the $\sigma^2$ term as all $\alpha$ terms are initialised using $\sigma_0^2$ (this notion relaxes the assumption of spatial smoothing).
-
-% Topic model settings
-% - AR Tuning
-\par In order to run the experiments in an efficient manner, we need to identify optimal values of the previously noted variances.
-One reason behind the variance tuning corresponds to the rate of convergence upon the application of the MH algorithm.
-Based on the algorithm's rationale -- a low acceptance rate indicates a slow and stable convergence, whereas a high acceptance rate indicates a random and unstable convergence -- we would find the variance $\delta^2$ inducing the acceptance rate of around $30\%$.
-Another reason behind the variance tuning is related to the spatial smoothing application.
-Most importantly, we want to keep the variance $\sigma^2$ in tact with the rate of change of the topic smoothing throughout our data.
-Furthermore, since our topic development is captured in discrete space, we want to make sure that the discretisation step induced by the generative data process is smaller than the $\sigma^2$ variance; otherwise, we would fail to capture the high rate of change induced by steep topic changes.
-% MSI data settings
-% - Synthetic corpora
-% -- Fixed theta and phi
-% -- The size of the synthetic corpus 
-\par Speaking of the datasets generated for the experiments, these are designed to reflect three following aspects: the effect of overlapping topics; the effect of overlapping vocabulary terms; and the effect of noise.
-Note that our assessment is based on an intuitive $3$ topic scenario: $2$ topics model distinct metabolite entities, and the remaining topic models the noise topic.
-Speaking of the dataset size, we set $T=50$ for the number of documents per corpus and $\xi=100$ for the number of words per documents: the choice of $T$ surpasses the discretisation concern; also, as suggested by the generative algorithm given in Subsection 5.3, the use of the $\xi$ parameter establishes a slightly varying number of words in each document.
-However, in order to speed up the inference, we normalise the number of words per document to possess the maximum value of $50$.
-
-% Experiment settings
-% - Overlapping topics
-% - Overlapping vocabulary terms
-% - Overlapping noise vocabulary terms
-\par In the preceding figures, we illustrate the variations of the data generation settings: in Figure \ref{fig:ol_topics}, we display the setting  controlling the topic overlap; in Figure \ref{fig:ol_terms}, we display the setting controlling the topic term overlap; and in Figure \ref{fig:ol_error}, we display the setting controlling the error overlap.
-Note that the red and green colourings indicate the synthetic metabolite topics; the green colouring indicates the noise topic; and, for the term names set in the horizontal axes of Figures \ref{fig:ol_terms} and \ref{fig:ol_error}, we use arbitrary, unique numbers.
 
@@
 
+% Topic model settings
+% - AR Tuning
+\par In order to run the experiments in an efficient manner, we identify optimal values of the previously noted variances.
+One reason behind the variance tuning corresponds to the rate of convergence upon the application of the MH algorithm.
+Based on the algorithm's rationale -- a low acceptance rate indicates a slow and stable convergence, whereas a high acceptance rate indicates a random and unstable convergence -- we would find the variance $\delta^2$ inducing the acceptance rate of around $30\%$.
+Another reason behind the variance tuning is related to the spatial smoothing application.
+Most importantly, we keep the variance $\sigma^2$ in tact with the rate of change of the topic smoothing throughout the data.
+Furthermore, since the topic development is captured in discrete space, we want to make sure that the discretisation step induced by the generative data process is smaller than the $\sigma^2$ variance; otherwise, we would fail to capture the high rate of change induced by steep topic changes.
+% MSI data settings
+% - Synthetic corpora
+% -- Fixed theta and phi
+% -- The size of the synthetic corpus 
+\par To comment on the datasets generated for the experiments, these are designed to reflect the three following aspects: the effect of overlapping topics; the effect of overlapping vocabulary terms; and the effect of noise.
+Note that our assessment is based on an intuitive $3$ topic scenario: $2$ topics model distinct metabolite entities, and the remaining topic models the noise topic.
+To comment on the dataset size, we set $T=50$ for the number of documents per corpus and $\xi=100$ for the number of words per documents: the choice of $T$ surpasses the discretisation concern; also, as suggested by the generative algorithm given in Subsection 5.3, the use of the $\xi$ parameter establishes a slightly varying number of words in each document.
+However, in order to speed up the inference, we normalise the number of words per document to possess the maximum value of $50$.
+
+% Experiment settings
+% - Overlapping topics
+% - Overlapping vocabulary terms
+% - Overlapping noise vocabulary terms
+\par In the following figures, we illustrate the variations of the data generation settings: in Figure \ref{fig:ol_topics}, we display the setting  controlling the topic overlap; in Figure \ref{fig:ol_terms}, we display the setting controlling the topic term overlap; and in Figure \ref{fig:ol_error}, we display the setting controlling the error overlap.
+Note that the red and green colourings indicate the synthetic metabolite topics; the green colouring indicates the noise topic; and, for the term names set in the horizontal axes of Figures \ref{fig:ol_terms} and \ref{fig:ol_error}, we use arbitrary, unique numbers.
+
 % Experiment settings
@@
 % - preset theta and phi
-\par Relating the latter settings to our experiments, we assessed their all eight possible permutations.
+\par Relating the latter settings to our experiments, we assess their all (eight) possible permutations.
 To give an example of a permutation, one of the experiments would assess the ability to recover the underlying topic term distributions with the enabled topic overlap, the disabled topic term overlap, and the enabled error overlap.
-Speaking of the data generation, note that the latter settings directly reflect the $\theta$ and $\phi$ values of the synthetic datasets (we do not use the $\alpha$ and $\beta$ hyper-parameters).
+Note that the latter settings directly reflect the $\theta$ and $\phi$ values of the synthetic datasets (we do not use the $\alpha$ and $\beta$ hyper-parameters).
 By following the latter principle, we establish a clearer representation of the synthetic data; thus, simplify the performance assessment.
@@
 In a single experiment, we would have $40$ of such batches indicating the performance -- this is illustrated in Figure \ref{fig:perf_theta} and Figure \ref{fig:perf_phi}.
+
 \begin{figure}[h]
@@
   \label{fig:perf_theta}
-
   \begin{subfigure}[b]{0.37\textwidth}
@@
   \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\linewidth]{theta-ar.png}
+    \includegraphics[width=\linewidth]{theta-ar_11.png}
   \end{subfigure}%
@@
   \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\linewidth]{theta-nonar.png}
+    \includegraphics[width=\linewidth]{theta-nonar_11.png}
   \end{subfigure}%
@@
   \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\linewidth]{phi-ar.png}
+    \includegraphics[width=\linewidth]{phi-ar_11.png}
   \end{subfigure}%
@@
   \begin{subfigure}[b]{0.48\textwidth}
-    \includegraphics[width=\linewidth]{phi-nonar.png}
+    \includegraphics[width=\linewidth]{phi-nonar_11.png}
   \end{subfigure}%
@@
 \end{center}
-  \caption{The results of the t-test.}
+  \caption{The t-test assessing 8 different overlap settings.}
   \label{tab:ttest}
@@
 \par Interestingly, we can group the experiment listings in four pairs: the pairs are centred around the $\phi$ p-values of $0.01$, $0.16$, $0.30$, or $0.71$.
-The first are pairs presented in the previous paragraph; however, for the last two pairs, the p-values are well beyond the significance threshold.
+The first two pairs are presented in the previous paragraph; however, for the last two pairs, the p-values are well beyond the significance threshold.
 By noting that, in every pair, only the error setting varies, the insignificant results occur when one of the topic and term settings is disabled and another is enabled.
@@
 To give an example of a similar performance, the reader can consider the previously introduced Figure \ref{fig:perf_phi} and Figure \ref{fig:comp_phi}.
-However, by considering Figure \ref{fig:perf_phi}, note that the $\phi$ value of the auto-regressive model converges faster.
+However, by examining Figure \ref{fig:perf_phi}, note that the $\phi$ value of the auto-regressive model converges faster.
 
@@
 \par Our main objectives were to identify the spatial smoothing application's prospect in recovering the $\phi$ values used upon the generative data process and the ability to separate the noise topic.
-Speaking of the $\phi$ values, only a half of the carried experiments displayed significant performance compared to the topic model without the spatial smoothing application.
-We report an improved $\phi$ recovery performance on the synthetic datasets with the enabled topic and terms overlap settings; alternatively, when both topic and term overlap settings are disabled, the performance is superior if the error overlap is disabled and inferior if the error overlap is enabled.
+We report that only a half of the carried experiments displayed a significant performance in recovering the $\phi$ values.
+To be more specific, we observe an improved $\phi$ recovery when the synthetic datasets are generated using enabled topic and terms overlap settings; alternatively, when both topic and term overlap settings are disabled, the performance is superior if the error overlap is disabled and inferior if the error overlap is enabled.
 
@@
 % - Comments on the ability to find the noise
-\par Speaking of the overlapping noise topic's identification, the auto-regressive model -- just like the non-auto-regressive model -- assigns the overlapping error terms to the main topics.
-For this reason, we conclude that the spatial smoothing application's impact in improving the overlapping noise topic terms is negligible.
-However, looking into the statistical test on the $\theta$ values, $6$ out $8$ experiments display a significant performance in recovering the synthetic corpus generation settings.
+\par To expand on the overlapping noise topic's identification (i.e.
+noise detection), the auto-regressive model -- just like the non-auto-regressive model -- assigns the overlapping error terms to the main topics.
+For this reason, we conclude that the spatial smoothing application is negligible in improving the overlapping noise topic term detection.
+However, looking into the statistical test on the $\theta$ values, $6$ out $8$ experiments display a significant performance in recovering the $\theta$ values.
 In $5$ out $6$ cases, the $\theta$ values are recovered with a lower error; this result is mostly impacted by the model's ability to reflect the shape of the noise topic with a better accuracy.
@@
 % -- Spatial smoothing multi-neighbour
-\par Since some of the experiment settings display a performance improvement, the spatial smoothing application can be considered for a further research.
+\par Since some of the experiment settings display a performance improvement, the spatial smoothing application could be considered for a further research.
 We would recommend looking into the application of the undirected graphical model -- Markov random field.
@@
 % - bucketisation
-\par Speaking of the alternative research directions in assessing the spatial smoothing application, we propose a research problem on investigating bucketisation enhancements.
+\par To suggest an alternative research direction in assessing the spatial smoothing application, we propose a research problem on investigating bucketisation enhancements.
 In other words, the spatial smoothing application might have an impact upon the feature extraction from MSI data.
 To be more specific, spatial smoothing might establish more appropriate bucket size ranges in concatenating raw mass-to-charge values.
-If successful, this application would improve the quality of MSI data features and, consequently, improve the performance of the research problems addressed in this research project.
+If successful, this application would improve the quality of MSI data features and, consequently, improve the performance of the research problems addressed by this research project.
 
@@
 % -- All disabled settings do not display a metabolomics-like environment
-\par To give a final verdict on the spatial smoothing application's performance, we consider the synthetic data settings which reflect the metabolomics-like environment the best.
-Effectively, if more overlap settings are enabled, then the synthetic corpus reflects the environment better.
-As shown by Table \ref{tab:ttest}, the auto-regressive model tends to perform better than the non-auto-regressive model when most of the overlap settings are enabled.
-Ultimately, spatial smoothing is effective upon considering more complex data.
+\par To give the final verdict on the research project's outcome, we consider the performance obtained using the settings reflecting the metabolomics environment the best.
+Effectively, the more overlap settings are enabled, the better the metabolomics environment is represented.
+As shown by Table \ref{tab:ttest}, the auto-regressive model tends to perform better when most of the overlap settings are enabled.
+Therefore, we conclude that spatial smoothing can be indeed effective on improving the assessment of the MSI data with metabolomics environment characteristics.
 

### 2017-04-20 submission/code/pre-processing/create_bne-corpus.py

@@ -1,0 +1,125 @@
+
+import math
+
+"""
+VERSION
+- Python 2
+
+PURPOSE
+- Pre-process the 'b and e' data
+-- 2D Mapping
+-- Vocabulary's bucketisation
+"""
+
+import pymzml
+import numpy as np
+import os
+import pickle as pkl
+
+np.set_printoptions(precision=3, threshold=np.inf)
+
+INTENSITY_FLOOR = 10
+TOLERANCE_FACTOR = 7
+HOME_PATH = os.path.expanduser('~')
+PATH_DATA = HOME_PATH + '/Projects/ssmsi/data/corpora_raw/'
+PATH_OUT = HOME_PATH + '/Projects/ssmsi/data/corpora_processed/'
+# ___________________________________________________________________________
+
+
+def generate_coordinates(data_size, number_of_rows, row_length,
+                         column_length):
+    single_scan_size = (data_size + 1) / number_of_rows
+    column_size = math.floor(single_scan_size *
+                             (column_length / (row_length + column_length)))
+    row_size = single_scan_size - column_size
+    coordinates = []
+
+    x_coord = 0
+    y_coord = 0
+    row_scan = True
+    scanning_right = True
+    for _ in range(0, number_of_rows):
+        for i in range(1, single_scan_size + 1):
+            coordinates.append((x_coord, y_coord))
+            # Switching the horizontal scanning to the vertical scanning.
+            if i % row_size == 0:
+                row_scan = not row_scan
+            if row_scan:
+                if scanning_right:
+                    x_coord += 1
+                else:
+                    x_coord -= 1
+            else:
+                y_coord += 1
+        # Reversing scanner's direction.
+        row_scan = not row_scan
+        scanning_right = not scanning_right
+    return coordinates
+# ___________________________________________________________________________
+
+
+def assign_class(classes, feature):
+    for class_assignment in classes:
+        if feature >= class_assignment[0] and feature <= class_assignment[1]:
+            return str(class_assignment)
+
+
+# ___________________________________________________________________________
+def main():
+    '''
+    Pre-processing the data
+    '''
+    run = pymzml.run.Reader(PATH_DATA + 'abcdefgh_1.mzML')
+    preprocessed_mnis = []
+    for spectrum in run:
+        for mass, intensity in spectrum.peaks:
+            if intensity > INTENSITY_FLOOR:
[…]

### 2017-04-20 submission/code/pre-processing/generator_corpus.py

@@ -1,0 +1,53 @@
+
+"""
+VERSION
+- Python 2
+
+PURPOSE
+- A corpus generation for the notebooks
+"""
+
+import numpy as np
+import string
+
+
+class Generator_Corpus(object):
+    def __init__(self, alpha_init, theta_init, beta_init, xi, T):
+        if alpha_init is not None:
+            self.theta = self._softmax_matrix(alpha_init)
+        elif theta_init is not None:
+            self.theta = theta_init
+        self.beta = beta_init
+        self.xi = xi
+        self.K, self.V = self.beta.shape
+        self.vocab = self._generate_vocab()
+        self.T = T
+
+    def _generate_vocab(self):
+        vocab = []
+        for w in range(self.V):
+            vocab.append(string.ascii_lowercase[w])
+        return np.array(vocab)
+# ___________________________________________________________________________
+
+    def generate_corpus(self):
+        corpus = {}
+        for d in range(self.T):
+            corpus[d] = {}
+            N = np.random.poisson(self.xi)
+            for w in range(N):
+                topic_distrib = np.random.multinomial(1, self.theta[d])
+                z = np.where(topic_distrib == 1)[0][0]
+                word_distrib = np.random.multinomial(1, self.beta[z])
+                w = np.where(word_distrib == 1)[0][0]
+                w_key = self.vocab[w]
+                if w_key not in corpus[d]:
+                    corpus[d][w_key] = 0
+                corpus[d][w_key] += 1
+        return corpus
+
+    def _softmax_matrix(self, arr):
+        arr_softmax = np.array(arr)
+        for idx_r, row in enumerate(arr):
+            arr_softmax[idx_r] = np.exp(row) / np.sum(np.exp(row))
+        return arr_softmax

### 2017-04-20 submission/code/pre-processing/helper_experiment.py

@@ -1,0 +1,110 @@
+
+"""
+VERSION
+- Python 2
+
+PURPOSE
+- A experiment execution helper for the notebooks
+"""
+
+import numpy as np
+
+
+class Helper_Experiment(object):
+
+    '''
+    Calculating the error difference between the true approximated
+    values of theta and phi.
+    '''
+    def calculate_performance(self, alpha_init, beta_init, theta_init,
+                              clf, s_batch):
+        hist_theta = np.array(clf.hist_theta)
+        if alpha_init is None:
+            theta_init = theta_init.T
+        performance_batch_theta = []
+        performance_batch_phi = []
+        '''
+        Iterate through the theta batches
+        '''
+        n_iter = int(len(hist_theta) / s_batch)
+        for i in range(n_iter):
+            hist_theta_batch = hist_theta[0:i * s_batch - 1]
+            '''
+            Pre-process and tune the thetas
+            '''
+            theta_latent = np.average(hist_theta_batch, axis=0)
+            if alpha_init is not None:
+                theta_init = self._softmax_matrix(alpha_init)
+                theta_init = theta_init.T
+            theta_latent = theta_latent.T
+            matchings_topic = self._match_thetas(theta_init, theta_latent)
+            '''
+            Calculate the theta similarity
+            '''
+            diffs_theta = []
+            for idx_init, row_init in enumerate(theta_init):
+                idx_latent = matchings_topic[idx_init]
+                diff_theta = np.abs(row_init - theta_latent[idx_latent, :])
+                diff_theta = diff_theta.sum()
+                diffs_theta.append(diff_theta)
+            sum_theta = np.sum(diffs_theta)
+            performance_batch_theta.append(sum_theta)
+        '''
+        Iterate through the phi batches
+        '''
+        hist_phi = np.array(clf.hist_phi)
+        for i in range(n_iter):
+            hist_phi_batch = hist_phi[0:i * s_batch - 1]
+            phi_latent = np.average(hist_phi_batch, axis=0)
+            '''
+            Calculate the phi similarity
+            '''
+            diffs_phi = []
+            for idx_init, row_init in enumerate(beta_init):
+                idx_latent = matchings_topic[idx_init]
+                diff_phi = np.abs(row_init - phi_latent[idx_latent, :])
+                diff_phi = diff_phi.sum()
+                diffs_phi.append(diff_phi)
+            sum_phi = np.sum(diffs_phi)
[…]

### 2017-04-20 submission/code/pre-processing/processor_corpus.py

@@ -1,0 +1,91 @@
+
+"""
+VERSION
+- Python 2
+
+FUNCTION
+- Corpus pre-processing
+"""
+
+import numpy as np
+import math
+
+
+class Processor_Corpus(object):
+
+    def __init__(self, corpus, vocab=None):
+        self.corpus = corpus
+        self.vocab = vocab
+        if self.vocab is None:
+            self.vocab = self._extract_vocab()
+        self.T = len(self.corpus)
+        self.V = len(self.vocab)
+
+    def _extract_vocab(self):
+        vocab = []
+        for doc in self.corpus.values():
+            for word in doc.keys():
+                if word not in vocab:
+                    vocab.append(word)
+        return np.array(vocab)
+# ___________________________________________________________________________
+
+    '''
+    Terms:
+    - pp: pre-processed
+    '''
+    def process_corpus(self, threshold, normalise):
+        corpus_pp = np.zeros((self.T, self.V), dtype=np.float)
+        for t, doc in self.corpus.items():
+            t = int(t)
+            '''
+            Assign the normalisation factor based on the max intensity
+            '''
+            if normalise:
+                max_ND = 0
+                for word in doc:
+                    if doc[word] > max_ND:
+                        max_ND = doc[word]
+                factor_norm = threshold / max_ND
+            else:
+                factor_norm = 1.
+            '''
+            Transforming the corpus into a numpy array
+            '''
+            for w, word in enumerate(self.vocab):
+                if word in doc:
+                    count_norm = factor_norm * doc[word]
+                    corpus_pp[t][w] = int(count_norm)
+        return corpus_pp
+# ___________________________________________________________________________
+
+    '''
+    Generate the coordinates for the 2-dimensional bne plotting.
+    '''
+    def generate_coordinates(self, n_rows, l_row, l_column):
+        s_column, s_row = self.calc_sizes(self.T, n_rows, l_row, l_column)
+        x_coord = 0
+        y_coord = n_rows - 1
+        scanning_right = True
+        map_coord = {}
+        for j in range(0, n_rows):
+            for i in range(0, s_row):
+                doc_id = i + s_column * j + s_row * j
+                map_coord[doc_id] = {'x': x_coord, 'y': y_coord}
+                # Switching horizontal scanning to vertical.
+                if (i % (s_row - 1) == 0 and i != 0):
+                    y_coord -= 1
+                    continue
+                if scanning_right:
[…]
